---
title: "Cross-Sector Innovation Returns Analysis"
subtitle: "Building a combined sector chart from Innovation Strategy Explorer CSV downloads"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 14, fig.height = 8)

library(dplyr)
library(ggplot2)
library(tidyr)
library(scales)
```

# How this works

The [Innovation Strategy Explorer](https://019c97ee-12b9-7776-d9aa-ff62518c2622.share.connect.posit.cloud)
app uses Shiny URL bookmarking. Every filter state is encoded in the URL query
string, so you can link directly to a specific view. This notebook:

1. Generates a **hyperlink for every sector group** that pre-sets the app filters
2. You visit each link and click the **"Data (CSV)"** download button
3. The notebook reads all downloaded CSVs and combines them into a single bar chart

---

# 1. URL parameter format

The app encodes inputs as Shiny bookmark parameters. The relevant ones are:

| Parameter | Description | Example value |
|-----------|-------------|---------------|
| `_inputs_` | Required marker — always first | *(empty)* |
| `navbar_page` | Main tab | `"Country Explorer"` |
| `country-country` | Country selection | `["All countries"]` |
| `country-firm` | Firm or sector group | `["Aerospace %26 Defence"]` |
| `country-toflow` | Return flow measure | `"is_global"` |
| `country-tech_categories_plot1` | Technologies (bar chart) | `["All innovations"]` |
| `country-inner_tabs` | Which sub-tab | `"Returns by Technology"` |
| `country-widthscale` | Bar width mode | `"log"` |
| `country-display_mode` | Error bar style | `"confidence"` |

Values are JSON-encoded, with special characters URL-encoded (`&` → `%26`, space → `%20`).

---

# 2. CSV download format

When you click **"Data (CSV)"** on the "Returns by Technology" tab, the downloaded
file contains one row per technology bar with these columns:

| Column | Description |
|--------|-------------|
| `technology` | Technology label (e.g. "AI", "Green Technology", "All innovations") |
| `mean` | Average return (in %, if `is_*` or `av_*` flow) |
| `innos` | Number of innovations (patents) |
| `sem` | Standard error of the mean |
| `q1`, `q2`, `q3` | 25th, 50th, 75th percentile |
| `top25_bin_mean` | Mean of top 25% of innovations |
| `top50_bin_mean` | Mean of top 50% of innovations |
| `allmean` | Overall mean across all innovations (the dashed reference line) |
| `allinnos` | Total innovation count |
| `top3_ids` | Top patent application IDs |
| `greenclass` | Technology colour category |
| `value_label` | Formatted display string (e.g. "4.2%") |

For a single-sector query with `tech_categories = "All innovations"`, the CSV has
**one row** — that row's `mean` is the sector's average return.

---

# 3. Generate hyperlinks for all sectors

```{r build_urls}
# ── Configuration ──────────────────────────────────────────────────────────
# Set the base URL of the deployed app
BASE_URL <- "https://019c97ee-12b9-7776-d9aa-ff62518c2622.share.connect.posit.cloud"

# The sector groups available in the app (from firm_lookup.parquet → firm_sector)
# Run this once with database access to extract them, or copy from the app's
# "Firm or Sector Group" dropdown under "Sector Groups":
sector_groups <- c(
  "Aerospace & Defence",
  "Alternative Energy",
  "Automobiles & Parts",
  "Chemicals",
  "Construction & Materials",
  "Electronic & Electrical Equipment",
  "Fixed Line Telecommunications",
  "Food & Drug Retailers",
  "General Industrials",
  "Health Care Equipment & Services",
  "Industrial Engineering",
  "Industrial Metals & Mining",
  "Leisure Goods",
  "Oil & Gas Producers",
  "Personal Goods",
  "Software & Computer Services",
  "Technology Hardware & Equipment",
  "Travel & Leisure"
)

# ── Build one URL per sector ──────────────────────────────────────────────
build_ise_url <- function(base_url, sector, toflow = "is_global") {
  # JSON-encode values the way Shiny bookmarking does
  firm_json   <- paste0('["', sector, '"]')
  params <- paste0(
    "?_inputs_",
    "&navbar_page=",                  URLencode('"Country Explorer"', reserved = TRUE),
    "&country-country=",              URLencode('["All countries"]', reserved = TRUE),
    "&country-firm=",                 URLencode(firm_json, reserved = TRUE),
    "&country-toflow=",               URLencode(paste0('"', toflow, '"'), reserved = TRUE),
    "&country-tech_categories_plot1=", URLencode('["All innovations"]', reserved = TRUE),
    "&country-inner_tabs=",           URLencode('"Returns by Technology"', reserved = TRUE),
    "&country-widthscale=",           URLencode('"log"', reserved = TRUE),
    "&country-display_mode=",         URLencode('"confidence"', reserved = TRUE)
  )
  paste0(base_url, params)
}

sector_urls <- setNames(
  sapply(sector_groups, function(s) build_ise_url(BASE_URL, s)),
  sector_groups
)
```

## Clickable links

Click each link below to open the app pre-filtered for that sector group.
Then click **"Data (CSV)"** to download the bar chart data.

```{r print_links, results='asis'}
for (sector in sector_groups) {
  cat(sprintf("- [%s](%s)\n", sector, sector_urls[sector]))
}
```

> **Tip:** Save each downloaded CSV with the sector name, e.g.
> `Aerospace & Defence.csv`, `Alternative Energy.csv`, etc.
> Place them all in a single folder (e.g. `downloads/`).

---

# 4. Read downloaded CSVs

```{r read_csvs}
# ── Point this at your download folder ────────────────────────────────────
csv_folder <- "downloads"

if (!dir.exists(csv_folder)) {
  message(
    "Folder '", csv_folder, "' not found. ",
    "Create it, download all CSVs into it, then re-knit."
  )
}

csv_files <- list.files(csv_folder, pattern = "\\.csv$", full.names = TRUE)
cat("Found", length(csv_files), "CSV files\n")
```

```{r parse_csvs}
all_sectors <- list()

for (f in csv_files) {
  df <- read.csv(f, stringsAsFactors = FALSE)

  # Derive sector name from the filename
  sector_name <- tools::file_path_sans_ext(basename(f))
  # Clean up common filename patterns (e.g. "returns_by_technology_2026-03-24")
  sector_name <- gsub("_\\d{4}-\\d{2}-\\d{2}$", "", sector_name)

  # The CSV may have multiple rows (one per technology).
  # For the cross-sector comparison we want the "All innovations" row,
  # or if there is only one row, use that.
  if ("technology" %in% names(df)) {
    summary_row <- df |> filter(technology == "All innovations")
    if (nrow(summary_row) == 0) summary_row <- df[1, , drop = FALSE]
  } else {
    summary_row <- df[1, , drop = FALSE]
  }

  summary_row$sector <- sector_name
  all_sectors[[sector_name]] <- summary_row
}

combined_df <- bind_rows(all_sectors)

if (nrow(combined_df) > 0) {
  cat("Combined", nrow(combined_df), "sectors\n\n")
  knitr::kable(
    combined_df |>
      select(sector, mean, innos, sem, q1, q2, q3) |>
      arrange(desc(mean)),
    digits = 3,
    caption = "Marginal Global Returns by Sector Group"
  )
} else {
  message("No data yet — download the CSVs first.")
}
```

---

# 5. Combined bar chart

```{r combined_plot, fig.height=10, fig.width=12}
if (nrow(combined_df) > 0) {
  # Overall mean across all sectors (use the allmean from any row — it's the
  # global average that the app shows as its dashed reference line)
  global_mean <- if ("allmean" %in% names(combined_df)) {
    mean(combined_df$allmean, na.rm = TRUE)
  } else {
    NA
  }

  p <- ggplot(combined_df, aes(x = reorder(sector, mean), y = mean)) +
    geom_col(aes(fill = mean), width = 0.7) +
    geom_errorbar(
      aes(ymin = pmax(mean - 1.96 * sem, 0), ymax = mean + 1.96 * sem),
      width = 0.25, linewidth = 0.4, alpha = 0.6
    ) +
    scale_fill_viridis_c(option = "C", name = "Return (%)") +
    coord_flip() +
    labs(
      title = "Marginal Global Returns by Firm Sector Group",
      subtitle = paste0("All innovations, All countries | ",
                        nrow(combined_df), " sectors | ",
                        scales::comma(sum(combined_df$innos, na.rm = TRUE)),
                        " total innovations"),
      x = NULL,
      y = "Return in %",
      caption = paste0("Source: Innovation Strategy Explorer CSV downloads\n",
                       "Error bars = 95% confidence intervals")
    ) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "right",
      plot.title    = element_text(face = "bold", size = 14),
      axis.text.y   = element_text(size = 10),
      plot.caption  = element_text(hjust = 1, size = 9, color = "gray50")
    )

  # Add the app's reference line if available
  if (!is.na(global_mean)) {
    p <- p +
      geom_hline(yintercept = global_mean, linetype = "dashed",
                 color = "black", linewidth = 0.8) +
      annotate("text", y = global_mean,
               x = nrow(combined_df) + 0.4,
               label = "Average", angle = -90,
               vjust = 1.5, size = 3.5, color = "black")
  }

  print(p)
} else {
  message("No data to plot — download the CSVs first.")
}
```

# 6. Innovation volume comparison

```{r lollipop, fig.height=10, fig.width=12}
if (nrow(combined_df) > 0) {
  ggplot(combined_df, aes(x = reorder(sector, mean), y = mean)) +
    geom_point(aes(size = innos, color = mean), alpha = 0.8) +
    geom_segment(aes(xend = reorder(sector, mean), y = 0, yend = mean),
                 color = "grey60", linewidth = 0.3) +
    scale_size_continuous(name = "Innovations", labels = scales::comma,
                          range = c(2, 12)) +
    scale_color_viridis_c(option = "C", name = "Return (%)") +
    coord_flip() +
    labs(
      title = "Returns vs Innovation Volume by Sector",
      subtitle = "Bubble size proportional to number of innovations",
      x = NULL,
      y = "Marginal Global Return (%)"
    ) +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(face = "bold", size = 14))
}
```

# 7. Export combined dataset

```{r export}
if (nrow(combined_df) > 0) {
  out_file <- "all_sectors_combined.csv"
  write.csv(
    combined_df |> select(sector, mean, innos, sem, q1, q2, q3,
                           top25_bin_mean, top50_bin_mean, allmean, allinnos),
    out_file, row.names = FALSE
  )
  cat("Saved combined data to:", out_file, "\n")
}
```
