---
title: "Cross-Sector Innovation Returns Analysis"
subtitle: "Building a combined sector chart from Innovation Strategy Explorer CSV downloads"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 14, fig.height = 8)

library(dplyr)
library(ggplot2)
library(tidyr)
library(scales)
library(chromote)
```

# How this works

The [Innovation Strategy Explorer](https://mondpanther-iseapp2.share.connect.posit.cloud)
app uses Shiny URL bookmarking. Every filter state is encoded in the URL query
string, so you can link directly to a specific view. This notebook:

1. Generates a **bookmarked URL for every sector group** that pre-sets the app filters
2. Uses `chromote` (headless Chrome) to visit each URL and **automatically download**
   the CSV via the app's download handler
3. Reads all downloaded CSVs and combines them into a single bar chart

---

# 1. URL parameter format

The app encodes inputs as Shiny bookmark parameters. The relevant ones are:

| Parameter | Description | Example value |
|-----------|-------------|---------------|
| `_inputs_` | Required marker — always first | *(empty)* |
| `navbar_page` | Main tab | `"Country Explorer"` |
| `country-country` | Country selection | `["All countries"]` |
| `country-firm` | Firm or sector group | `["Aerospace %26 Defence"]` |
| `country-toflow` | Return flow measure | `"is_global"` |
| `country-tech_categories_plot1` | Technologies (bar chart) | `["All innovations"]` |
| `country-inner_tabs` | Which sub-tab | `"Returns by Technology"` |
| `country-widthscale` | Bar width mode | `"log"` |
| `country-display_mode` | Error bar style | `"confidence"` |

Values are JSON-encoded, with special characters URL-encoded (`&` → `%26`, space → `%20`).

---

# 2. CSV download format

When you click **"Data (CSV)"** on the "Returns by Technology" tab, the downloaded
file contains one row per technology bar with these columns:

| Column | Description |
|--------|-------------|
| `technology` | Technology label (e.g. "AI", "Green Technology", "All innovations") |
| `mean` | Average return (in %, if `is_*` or `av_*` flow) |
| `innos` | Number of innovations (patents) |
| `sem` | Standard error of the mean |
| `q1`, `q2`, `q3` | 25th, 50th, 75th percentile |
| `top25_bin_mean` | Mean of top 25% of innovations |
| `top50_bin_mean` | Mean of top 50% of innovations |
| `allmean` | Overall mean across all innovations (the dashed reference line) |
| `allinnos` | Total innovation count |
| `top3_ids` | Top patent application IDs |
| `greenclass` | Technology colour category |
| `value_label` | Formatted display string (e.g. "4.2%") |

For a single-sector query with `tech_categories = "All innovations"`, the CSV has
**one row** — that row's `mean` is the sector's average return.

---

# 3. Generate hyperlinks for all sectors

```{r build_urls}
# ── Configuration ──────────────────────────────────────────────────────────
# Set the base URL of the deployed app
BASE_URL <- "https://mondpanther-iseapp2.share.connect.posit.cloud"

# The sector groups available in the app (from firm_lookup.parquet → firm_sector)
# Run this once with database access to extract them, or copy from the app's
# "Firm or Sector Group" dropdown under "Sector Groups":
sector_groups <- c(
  "Aerospace & Defence",
  "Alternative Energy",
  "Automobiles & Parts",
  "Chemicals",
  "Construction & Materials",
  "Electronic & Electrical Equipment",
  "Fixed Line Telecommunications",
  "Food & Drug Retailers",
  "General Industrials",
  "Health Care Equipment & Services",
  "Industrial Engineering",
  "Industrial Metals & Mining",
  "Leisure Goods",
  "Oil & Gas Producers",
  "Personal Goods",
  "Software & Computer Services",
  "Technology Hardware & Equipment",
  "Travel & Leisure"
)

# ── Build one URL per sector ──────────────────────────────────────────────
build_ise_url <- function(base_url, sector, toflow = "ev_emdenocn",
                          country = "High income countries") {
  # JSON-encode values the way Shiny bookmarking does
  params <- paste0(
    "?_inputs_",
    "&globe-render_map=0",
    "&country-sidebar=true",
    "&region-sidebar=true",
    "&globe-sidebar=true",
    "&navbar_page=",                  URLencode('"Country Explorer"', reserved = TRUE),
    "&country-inner_tabs=",           URLencode('"Returns by Technology"', reserved = TRUE),
    "&region-inner_tabs=",            URLencode('"Returns by Technology"', reserved = TRUE),
    "&country-country=",              URLencode(paste0('"', country, '"'), reserved = TRUE),
    "&country-firm=",                 URLencode(paste0('"', sector, '"'), reserved = TRUE),
    "&country-toflow=",               URLencode(paste0('"', toflow, '"'), reserved = TRUE),
    "&country-tech_categories_plot1=", URLencode('"All innovations"', reserved = TRUE),
    "&country-techs=",               URLencode('"Green Technology"', reserved = TRUE),
    "&region-region=",                URLencode('"All UK regions"', reserved = TRUE),
    "&region-firm=null",
    "&region-toflow_region=",         URLencode('"is_global"', reserved = TRUE),
    "&region-tech_categories_plot1_region=", URLencode('["Green Technology","AI"]', reserved = TRUE),
    "&region-techs_region=",          URLencode('"Green Technology"', reserved = TRUE),
    "&region-techs_comparison_region=null",
    "&globe-sce_country=null",
    "&globe-sce_tech_display=null",
    "&globe-sample_size=",            URLencode('""', reserved = TRUE),
    "&globe-max_wave=1",
    "&country-widthscale=",           URLencode('"log"', reserved = TRUE),
    "&country-display_mode=",         URLencode('"confidence"', reserved = TRUE),
    "&region-widthscale_region=",     URLencode('"log"', reserved = TRUE),
    "&region-display_mode_region=",   URLencode('"confidence"', reserved = TRUE),
    "&country-top_n_ids=10",
    "&country-topn=20",
    "&country-mininno=5",
    "&country-topn_rta=20",
    "&country-bottomn_rta=0",
    "&country-mininno_rta=50",
    "&country-minallinnos_rta=100",
    "&region-top_n_ids_region=10",
    "&region-topn_region=12",
    "&region-mininno_region=100",
    "&region-topn_rta_region=12",
    "&region-bottomn_rta_region=0",
    "&region-mininno_rta_region=50",
    "&region-minallinnos_rta_region=100",
    "&waiter_shown=true",
    "&waiter_hidden=true",
    "&country-avstrax_plot1_hovered=null"
  )
  paste0(base_url, params)
}

sector_urls <- setNames(
  sapply(sector_groups, function(s) build_ise_url(BASE_URL, s)),
  sector_groups
)
```

---

# 4. Fetch CSV data in memory

This section uses `chromote` (headless Chrome) to open each sector's
bookmarked URL, wait for the Shiny app to compute, then uses the browser's
`fetch()` API to read the CSV download endpoint directly into R — no files
are written to disk.

```{r fetch_csvs, results='asis'}
WAIT_APP <- 15   # seconds to wait for the Shiny app to render

# Launch a headless Chrome session
b <- ChromoteSession$new()

all_sectors <- list()

for (sector in sector_groups) {
  cat("- **", sector, "** … ", sep = "")

  # Navigate to the bookmarked URL for this sector
  b$Page$navigate(url = sector_urls[[sector]], wait_ = TRUE)
  Sys.sleep(WAIT_APP)

  # Get the download button's href (set dynamically by Shiny once the
  # session is established and the reactive data has been computed)
  href_res <- b$Runtime$evaluate(
    'var el = document.getElementById("country-dl_csv_avstrax_plot1");
     el ? el.href : null;'
  )
  href <- href_res$result$value

  if (is.null(href)) {
    cat("**FAILED** (download link not found)\n")
    next
  }

  # Use the browser's own fetch() to retrieve the CSV text — this
  # automatically carries the Shiny session cookies / headers
  csv_res <- b$Runtime$evaluate(
    sprintf('fetch("%s").then(r => r.text())', href),
    awaitPromise = TRUE
  )
  csv_text <- csv_res$result$value

  if (is.null(csv_text) || nchar(csv_text) == 0) {
    cat("**FAILED** (empty response)\n")
    next
  }

  df <- read.csv(text = csv_text, stringsAsFactors = FALSE)

  # For the cross-sector comparison we want the "All innovations" row,
  # or if there is only one row, use that.
  if ("technology" %in% names(df)) {
    summary_row <- df |> filter(technology == "All innovations")
    if (nrow(summary_row) == 0) summary_row <- df[1, , drop = FALSE]
  } else {
    summary_row <- df[1, , drop = FALSE]
  }

  summary_row$sector <- sector
  all_sectors[[sector]] <- summary_row
  cat("OK\n")
}

tryCatch(b$close(), error = function(e) NULL)

combined_df <- bind_rows(all_sectors)

if (nrow(combined_df) > 0) {
  cat("\nCombined **", nrow(combined_df), "** sectors\n\n", sep = "")
  knitr::kable(
    combined_df |>
      select(sector, mean, innos, sem, q1, q2, q3) |>
      arrange(desc(mean)),
    digits = 3,
    caption = "Marginal Global Returns by Sector Group"
  )
} else {
  message("No data — check the log above for errors.")
}
```

---

# 5. Combined bar chart

```{r combined_plot, fig.height=10, fig.width=12}
if (nrow(combined_df) > 0) {
  # Overall mean across all sectors (use the allmean from any row — it's the
  # global average that the app shows as its dashed reference line)
  global_mean <- if ("allmean" %in% names(combined_df)) {
    mean(combined_df$allmean, na.rm = TRUE)
  } else {
    NA
  }

  p <- ggplot(combined_df, aes(x = reorder(sector, mean), y = mean)) +
    geom_col(aes(fill = mean), width = 0.7) +
    geom_errorbar(
      aes(ymin = pmax(mean - 1.96 * sem, 0), ymax = mean + 1.96 * sem),
      width = 0.25, linewidth = 0.4, alpha = 0.6
    ) +
    scale_fill_viridis_c(option = "C", name = "Return (%)") +
    coord_flip() +
    labs(
      title = "Marginal Global Returns by Firm Sector Group",
      subtitle = paste0("All innovations, All countries | ",
                        nrow(combined_df), " sectors | ",
                        scales::comma(sum(combined_df$innos, na.rm = TRUE)),
                        " total innovations"),
      x = NULL,
      y = "Return in %",
      caption = paste0("Source: Innovation Strategy Explorer CSV downloads\n",
                       "Error bars = 95% confidence intervals")
    ) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "right",
      plot.title    = element_text(face = "bold", size = 14),
      axis.text.y   = element_text(size = 10),
      plot.caption  = element_text(hjust = 1, size = 9, color = "gray50")
    )

  # Add the app's reference line if available
  if (!is.na(global_mean)) {
    p <- p +
      geom_hline(yintercept = global_mean, linetype = "dashed",
                 color = "black", linewidth = 0.8) +
      annotate("text", y = global_mean,
               x = nrow(combined_df) + 0.4,
               label = "Average", angle = -90,
               vjust = 1.5, size = 3.5, color = "black")
  }

  print(p)
} else {
  message("No data to plot — check the download log above for errors.")
}
```

# 6. Innovation volume comparison

```{r lollipop, fig.height=10, fig.width=12}
if (nrow(combined_df) > 0) {
  ggplot(combined_df, aes(x = reorder(sector, mean), y = mean)) +
    geom_point(aes(size = innos, color = mean), alpha = 0.8) +
    geom_segment(aes(xend = reorder(sector, mean), y = 0, yend = mean),
                 color = "grey60", linewidth = 0.3) +
    scale_size_continuous(name = "Innovations", labels = scales::comma,
                          range = c(2, 12)) +
    scale_color_viridis_c(option = "C", name = "Return (%)") +
    coord_flip() +
    labs(
      title = "Returns vs Innovation Volume by Sector",
      subtitle = "Bubble size proportional to number of innovations",
      x = NULL,
      y = "Marginal Global Return (%)"
    ) +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(face = "bold", size = 14))
}
```

# 7. Export combined dataset

```{r export}
if (nrow(combined_df) > 0) {
  out_file <- "all_sectors_combined.csv"
  write.csv(
    combined_df |> select(sector, mean, innos, sem, q1, q2, q3,
                           top25_bin_mean, top50_bin_mean, allmean, allinnos),
    out_file, row.names = FALSE
  )
  cat("Saved combined data to:", out_file, "\n")
}
```
