---
title: "Prepare various bits of data"
output: html_notebook
---

```{r}

library(dplyr)

```



```{r}
library(dplyr)
if (Sys.info()["nodename"] == "PCACL-9RVSQ74") {
  datad="C:\\Users\\rmartin7\\OneDrive - WBG\\projects\\PRINZGlobal\\imperial_college_visualization\\01_data"
  localbig="C:\\Users\\rmartin7\\OneDrive - Imperial College London\\inglobe"
  ifcone="C:\\Users\\rmartin7\\OneDrive - WBG\\projects\\PRINZGlobal\\KEVCexplorer"
}

```


# cpc data

```{r}
cpcfile=paste0(localbig,"\\data\\cpcs.parquet")
if (!file.exists(cpcfile)) {
  library(bigrquery)
  library(DBI)
  
  # Set your project ID
  project_id <- "patbis"
  
  # Option 1: Direct table download
  # Specify dataset and table name
  dataset <- "fromPATSTAT2021"
  table <- "tls225_docdb_fam_cpc"
  
  # Download the entire table
  cpcs <- bq_table_download(
    bq_table(project_id, dataset, table)
  )
  
  
  cpcs=cpcs %>% select(docdb_family_id,cpc_class_symbol)
  cpcs=cpcs %>% arrange(docdb_family_id)
  library(arrow)
  write_parquet(cpcs,cpcfile)
}




```


  
```{r}

library(arrow)
cpcs=read_parquet(cpcfile)


#test=cpcs %>% summarize(n_distinct(docdb_family_id))

cpcs=cpcs %>% arrange(docdb_family_id)

head(cpcs)
library(stringr)

## get sections...
sectionmap1=cpcs %>% mutate(section=substr(cpc_class_symbol,1,1)) %>%
  select(section,docdb_family_id) %>% 
  distinct()

sectionmap1 %>% distinct(section) %>% nrow()
sectionmap1 %>% distinct(section) %>% pull(section)  %>% cat()


sectionmap2=cpcs %>% mutate(section=substr(cpc_class_symbol,1,3)) %>%
  select(section,docdb_family_id) %>% 
  distinct()

sectionmap2 %>% distinct(section) %>% nrow()
sectionmap2 %>% distinct(section) %>% pull(section)  %>% cat()



cpcs=cpcs %>% gsub(" ", "", cpc_class_symbol)



library(data.table)
library(stringi)

setDT(cpcs)
cpcs[, cpc_class_symbol := stri_replace_all_fixed(cpc_class_symbol, " ", "")]


```


# Battery technology

## mapping file
```{r}


source("extrasectorshelper.r") # definitions of the battery sector

battery_df_expanded <- battery_df %>%
  separate_rows(CPC, sep = ";")%>%
  mutate(cpc_class_symbol = trimws(CPC))

setDT(battery_df_expanded)
battery_df_expanded[, cpc_class_symbol := stri_replace_all_fixed(cpc_class_symbol, " ", "")]


battery_df %>% 
  distinct(technology) %>% 
  pull(technology) %>% 
  paste0('"', ., '"') %>%
  paste(collapse = ", ") %>% cat()



batteryclasses=cpcs %>% inner_join(battery_df_expanded %>% select(technology,cpc_class_symbol)) 

all=batteryclasses %>% distinct(docdb_family_id) %>% mutate(technology="Any battery")
batteryclasses=batteryclasses %>% bind_rows(all)
batteryclasses=batteryclasses %>% select(docdb_family_id,technology)


```



```{r}
library(arrow)
library(fst)
techmap <- read_fst("techmap.fst")
techs=techmap %>% distinct(technology)
if(!"Any battery" %in% techs){
  techmap=techmap %>% bind_rows(batteryclasses)
  
  techmap = techmap %>% mutate(technology=ifelse(technology=="Any battery","Any battery technology",technology))
  techmap = techmap %>% mutate(technology=ifelse(technology=="Any Green","Any Green technology",technology))
  write_parquet(techmap,"techmap.parquet")
}


```

## Make an experimental green class (to check on the disparity in green share)

```{r}
#test=head(cpcs,1000)
head(cpcs)
xgreenclasses=cpcs %>% mutate(technology = substr(cpc_class_symbol, 1, 3)) %>% filter(technology=="Y02") %>% distinct(docdb_family_id,technology)
  
library(arrow)
techmap <- read_parquet("techmap.parquet")


test=techmap %>% filter(technology=="Any Green technology")


techs=techmap %>% distinct(technology)
techs %>% pull(technology)

test2=test %>% select(-technology) %>% inner_join(xgreenclasses)

nrow(test2)
nrow(test)
nrow(xgreenclasses)




```
ok, it seems Any Green is truly Y02



## Analysis of battery...
### Comparative advantage in battery


### tech shares by countries
```{r}


countrymap9918 <- read_fst("countrymap9918.fst")

countrymap9918 %>% distinct(docdb_family_id) %>% nrow()

library(data.table)

  tech_shares <- countrymap9918 %>%
  left_join(techmap, by = "docdb_family_id") %>%
  mutate(innos_total=n_distinct(docdb_family_id)) %>% 
  group_by(ctry_code)  %>% mutate(innos_total_c=n_distinct(docdb_family_id)) %>% 
  group_by(technology) %>% mutate(innos_t=n_distinct(docdb_family_id)) %>% 
  group_by(ctry_code, technology) %>%
    summarise(innos_total_c= mean(innos_total_c),
              innos_total  = mean(innos_total),
              innos_t      = mean(innos_t), test=n(),
              innos_ct     = n_distinct(docdb_family_id), .groups = "drop") %>%
    mutate(
      
      share_ct = innos_ct / innos_total_c,
      share_t=innos_t/innos_total,
      
      RTA=share_ct*2 / (share_ct+share_t)-1,
      innosh_ct=innos_ct/innos_t
      
    ) %>%
  ungroup()


library(WDI)
library(dplyr)

# Download GDP per capita data for 2015
gdp_data <- WDI(
  indicator = c(
    "gdp_pc_current" = "NY.GDP.PCAP.CD",      # GDP per capita at market exchange rates
    "gdp_pc_ppp" = "NY.GDP.PCAP.PP.CD"        # GDP per capita at PPP
  ),
  start = 2015,
  end = 2015,
  extra = FALSE
) %>%
  select(ctry_code = iso2c, gdp_pc_current, gdp_pc_ppp)

# Add to tech_shares
tech_shares <- tech_shares %>%
  left_join(gdp_data, by = "ctry_code")  





# World Bank income classifications (2024/2025)
income_classification <- tibble::tribble(
  ~income_group, ~ctry_code,
  "High income", c("AD", "AE", "AG", "AR", "AU", "AT", "AW", "BE", "BH", "BM", "BS", "BW", "CA", "CH", "CL", "CN", "CW", "CY", "CZ", "DE", "DK", "EE", "ES", "FI", "FR", "GB", "GI", "GR", "GU", "HK", "HR", "HU", "IE", "IL", "IM", "IS", "IT", "JP", "KR", "KW", "KY", "LI", "LT", "LU", "LV", "MC", "MP", "MT", "MO", "NL", "NO", "NZ", "OM", "PA", "PL", "PR", "PT", "QA", "RO", "SA", "SC", "SG", "SI", "SK", "SM", "SE", "TC", "TT", "TW", "UY", "US", "VI", "VG"),
  "Upper middle income", c("AL", "AM", "AZ", "BY", "BZ", "BR", "BG", "CO", "CR", "CU", "DM", "DO", "EC", "FJ", "GA", "GD", "GE", "GQ", "GT", "GY", "ID", "IQ", "IR", "JM", "JO", "KZ", "LB", "LC", "LY", "MK", "MU", "MV", "MX", "MY", "NA", "PE", "PY", "RU", "RS", "SR", "SV", "TH", "TM", "TO", "TR", "VE", "ZA", "ME", "XK", "ST", "VC", "WS"),
  "Lower middle income", c("AO", "BD", "BJ", "BO", "BT", "CI", "CM", "CG", "CV", "DZ", "EG", "GH", "HN", "HT", "IN", "KE", "KG", "KH", "KI", "LA", "LK", "LS", "MA", "MD", "MN", "MM", "MR", "NG", "NI", "NP", "PK", "PG", "PH", "PS", "SB", "SN", "TJ", "TL", "TN", "TZ", "UA", "UZ", "VN", "VU", "ZM", "ZW"),
  "Low income", c("AF", "BF", "BI", "CD", "CF", "ER", "ET", "GM", "GN", "GW", "KM", "LR", "MG", "ML", "MW", "MZ", "NE", "RW", "SD", "SL", "SO", "SS", "SY", "TD", "TG", "UG", "YE")
) %>%
  tidyr::unnest(ctry_code)

# Add income classification to your dataframe
tech_shares <- tech_shares %>%
  left_join(income_classification, by = "ctry_code") %>%  mutate(HIC = if_else(income_group == "High income", 1, 0))


  
  
```

```{r}
library(ggplot2)
library(ggrepel)
ctry="VN"
techs=""



green_classes <- c("Any Green technology","Green Energy", "Green Transport", "Circular Economy", "Green Manufacturing",
                   "Adaptation", "Green Housing", "Green ICT", "Green Agriculture",
                   "GHG Capture")


fclasses=c(green_classes,"AI", "Any battery technology")

ts_VN = tech_shares %>% 
  filter(ctry_code==ctry & technology %in% fclasses) %>% 
  mutate(color=ifelse(technology %in% green_classes, "darkgreen",
                      ifelse(technology=="AI", "orange",
                             ifelse(technology=="Any battery technology","gold", "gray")
                            )
                      )
  )



ts_VN %>% 
  ggplot(aes(x=RTA+1, y=(innosh_ct)*100, color=color, size=innos_ct, label=technology)) +
  geom_point(alpha=.4) +
  geom_text_repel(size=4) +
  geom_vline(xintercept = 1, color = "red", linetype = "dashed") +
  theme_minimal() + xlab("Revealed technological advantage")+
  ylab("Share in global innovation in field in %")+
  theme(legend.position = "none")+ scale_color_identity()
```



```{r}
techmap %>% distinct(technology) %>% pull(technology)

```

```{r}
library(ggplot2)

ctech="Any Hard to Abate technology"  
ctech="Any battery technology"
ctech="Any Green technology"
mini=100

y_var="gdp_pc_ppp"
#y_var="gdp_pc_current"
filtered=tech_shares %>% filter(technology==ctech) %>% filter(innos_total_c>=mini)
  
filtered %>%   
    ggplot(aes(x=RTA,y=log(!!sym(y_var)),size=innos_total_c))+geom_point()+
    geom_smooth(method="lm")
  




```

#### tech shares by country group


```{r}

countrymap9918 <- read_parquet("countrymap9918.parquet")
library(dplyr)


countrymap9918 <- countrymap9918 %>%
  left_join(income_classification, by = "ctry_code") %>%
  mutate(
    cgroup1 = case_when(
      ctry_code == "VN" ~ "Vietnam",
      ctry_code == "CN" ~ "China",
      income_group == "High income" ~ "High income countries",
      TRUE ~ "Other non-high income countries",


    )
  ) %>%
  select(-income_group)  


#countrymap %>% distinct(cgroup1) %>% pull(cgroup1)
#countrymap %>% distinct(ctry_code) %>% arrange(ctry_code) %>%  pull(ctry_code)

```


check what is going on 99-18 & green

```{r}

check9918=countrymap9918 %>% left_join(xgreenclasses) %>% 
  mutate(technology=replace_na(technology,"notY02")) %>% 
  select(cgroup1,docdb_family_id,technology) %>% distinct() %>% 
  group_by(cgroup1) %>% summarise(innos=n(),greeninnos=sum(technology=="Y02")) %>% 
  mutate(share=greeninnos/innos)

check9918$greeninnos %>% sum() 
#if(!"Y02" %in% techs){
#  techmap=techmap %>% bind_rows(batteryclasses)
  
#  write_parquet(techmap,"techmap.parquet")
#}
  



```


# 👌 Let's do this over time...

```{r}

countrymap9918 %>% distinct(ctry_code) %>% nrow()
testb=countrymap9918 %>% distinct(ctry_code)

check9918=countrymap9918 %>% left_join(xgreenclasses) %>% 
  mutate(technology=replace_na(technology,"notY02")) %>% 
  select(cgroup1,docdb_family_id,technology) %>% distinct() %>% 
  group_by(cgroup1) %>% summarise(innos=n(),greeninnos=sum(technology=="Y02")) %>% 
  mutate(share=greeninnos/innos)

check9918$greeninnos %>% sum() 
#if(!"Y02" %in% techs){
#  techmap=techmap %>% bind_rows(batteryclasses)
  
#  write_parquet(techmap,"techmap.parquet")
#}

#*testa %>% rename(ctry_code=person_ctry_code) %>%  anti_join(testb) %>% pull(ctry_code)

```


# 😒 How about taking the income classes from the cs file?


```{r}
ifcone="C:\\Users\\rmartin7\\OneDrive - WBG\\projects\\PRINZGlobal\\KEVCexplorer"

path=paste0(ifcone,"\\","../work packages/patent analysis/Data/combined_data/count_docdb_2025-03-04.csv")

path
new_group_map=read.csv(path) %>% arrange(Economy, earliest_filing_year) %>% 
  select(person_ctry_code,Income.group) %>% distinct() %>% 
  rename(cgroup_new=Income.group,
         ctry_code=person_ctry_code)



countrymap9918=countrymap9918 %>% left_join(new_group_map)

check9918_new=countrymap9918 %>% left_join(xgreenclasses) %>% 
  mutate(technology=replace_na(technology,"notY02")) %>% 
  select(cgroup_new,docdb_family_id,technology) %>% distinct() %>% 
  group_by(cgroup_new) %>% summarise(innos=n(),greeninnos=sum(technology=="Y02")) %>% 
  mutate(share=greeninnos/innos)

check9918_new

```




#### Define functions
```{r}
create_country_groups <- function(df, country_list, new_var_name = "cgroup") {
  
  # Join income classification
  df_with_income <- df %>%
    left_join(income_classification, by = "ctry_code")
  
  # Create the grouping variable dynamically
  for(i in seq_along(country_list)) {
    country_code <- country_list[i]
    country_label <- names(country_list)[i]
    
    if(i == 1) {
      df_with_income <- df_with_income %>%
        mutate(!!new_var_name := case_when(
          ctry_code == country_code ~ country_label,
          TRUE ~ NA_character_
        ))
    } else {
      df_with_income <- df_with_income %>%
        mutate(!!new_var_name := case_when(
          ctry_code == country_code ~ country_label,
          TRUE ~ !!sym(new_var_name)
        ))
    }
  }
  
  # Add high income and other categories
  df_with_income <- df_with_income %>%
    mutate(!!new_var_name := case_when(
      !is.na(!!sym(new_var_name)) ~ !!sym(new_var_name),
      income_group == "High income" ~ "High income countries",
      TRUE ~ "Other non-high income countries"
    )) %>%
    select(-income_group)
  
  return(df_with_income)
}

library(dplyr)
library(rlang)

calculate_tech_shares <- function(countrymap, techmap, group_var) {
  
  # Convert string to symbol for tidy evaluation
  group_var_sym <- sym(group_var)
  
  tech_shares <- countrymap %>%
    left_join(techmap, by = "docdb_family_id") %>%
    mutate(innos_total = n_distinct(docdb_family_id)) %>% 
    group_by(!!group_var_sym) %>% 
    mutate(innos_total_c = n_distinct(docdb_family_id)) %>% 
    group_by(technology) %>% 
    mutate(innos_t = n_distinct(docdb_family_id)) %>% 
    group_by(!!group_var_sym, technology) %>%
    summarise(
      innos_total_c = mean(innos_total_c),
      innos_total   = mean(innos_total),
      innos_t       = mean(innos_t), 
      innos_ct      = n_distinct(docdb_family_id), 
      .groups = "drop"
    ) %>%
    mutate(
      share_ct = innos_ct / innos_total_c,
      share_t  = innos_t / innos_total,
      RTA      = share_ct * 2 / (share_ct + share_t) - 1
    ) %>%
    ungroup()
  
  return(tech_shares)
}
```


```{r}
country_list <- c("VN" = "Vietnam", "CN" = "China")
test=create_country_groups(countrymap9918, country_list, "cgroup1")

tech_shares_country <- calculate_tech_shares(countrymap9918,
                                             techmap, "cgroup1")


ctech="Any Hard to Abate technology"  
ctech="Any battery technology"
#ctech="Any Green technology"

test=tech_shares_country %>% filter(technology==ctech)

```





# Hard to abate sectors


```{r}
library(readxl)
tech_df <- read_excel("hard_to_abate.xlsx") %>% 
  rename(detail=technology,technology=sector) %>% 
  mutate(technology=paste0(technology," Decarbonisation"))


tech_df_expanded <- tech_df %>%
  separate_rows(CPC, sep = ";")%>%
  mutate(cpc_class_symbol = trimws(CPC))

setDT(tech_df_expanded)
tech_df_expanded[, cpc_class_symbol := stri_replace_all_fixed(cpc_class_symbol, " ", "")]


tech_df %>% 
  distinct(technology) %>% 
  pull(technology) %>% 
  paste0('"', ., '"') %>%
  paste(collapse = ", ") %>% cat()



tech_classes=cpcs %>% inner_join(tech_df_expanded %>% select(technology,cpc_class_symbol)) 

all=tech_classes %>% distinct(docdb_family_id) %>% mutate(technology="Any Hard to Abate technology")
tech_classes=tech_classes %>% bind_rows(all)
tech_classes=tech_classes %>% select(docdb_family_id,technology)



```



```{r}


techmap <- read_parquet("techmap.parquet")
techs=techmap %>% distinct(technology)

if(!"Aviation Decarbonisation" %in% techs){
  techmap=techmap %>% bind_rows(tech_classes)
  
  write_parquet(techmap,"techmap.parquet")
}


```




# AI

```{r}
library(readxl)
tech_df <- read_excel("New_Sector_mapping.xlsx",sheet="AI") %>% 
  rename(CPC=`CPC/IPC Codes`) %>% 
  rename(technology=`Sub-Technology`) %>% filter(!is.na(technology))
  
  

tech_df_expanded <- tech_df %>%
  separate_rows(CPC, sep = ",")%>%
  mutate(cpc_class_symbol = trimws(CPC))

setDT(tech_df_expanded)
tech_df_expanded[, cpc_class_symbol := stri_replace_all_fixed(cpc_class_symbol, " ", "")]


tech_df %>% 
  distinct(technology) %>% 
  pull(technology) %>% 
  paste0('"', ., '"') %>%
  paste(collapse = ", ") %>% cat()



tech_classes=cpcs %>% inner_join(tech_df_expanded %>% select(technology,cpc_class_symbol)) 

#all=tech_classes %>% distinct(docdb_family_id) %>% mutate(technology="Any Hard to Abate technology")
#tech_classes=tech_classes %>% bind_rows(all)
tech_classes=tech_classes %>% select(docdb_family_id,technology)



```



```{r}

library(fst)
techmap <- read_parquet("techmap.parquet")
techs=techmap %>% distinct(technology)

if(!"Machine Learning" %in% techs){
  techmap=techmap %>% bind_rows(tech_classes)
  techmap=techmap %>% distinct()
  #write_parquet(techmap,"techmap.parquet")
  write_fst(techmap,"techmap.fst")
}



```


# CPC sections

```{r}

cpc <- c(
  A = "Human Necessities",
  B = "Performing Operations; Transporting",
  C = "Chemistry; Metallurgy",
  D = "Textiles; Paper",
  E = "Fixed Constructions",
  F = "Mechanical Engineering; Lighting; Heating; Weapons; Blasting",
  G = "Physics",
  H = "Electricity",
  Y = "General tagging of new or cross-sectional technology"
)

cpc_df <- data.frame(
  section = names(cpc),
  heading = unname(cpc)
)

cpc_classes <- c(
  A01 = "Agriculture; Forestry; Animal husbandry; Hunting; Trapping; Fishing",
  A21 = "Baking; Edible doughs",
  A22 = "Butchering; Meat treatment; Processing poultry or fish",
  A23 = "Foods or foodstuffs; Treatment thereof",
  A24 = "Tobacco; Cigars; Cigarettes; Smokers' requisites",
  A41 = "Wearing apparel",
  A42 = "Headwear",
  A43 = "Footwear",
  A44 = "Haberdashery; Jewellery",
  A45 = "Hand or travelling articles",
  A46 = "Brushware",
  A47 = "Furniture; Domestic articles or appliances; Coffee mills; Spice mills; Suction cleaners",
  A61 = "Medical or veterinary science; Hygiene",
  A62 = "Life-saving; Fire-fighting",
  A63 = "Sports; Games; Amusements",
  A99 = "Subject matter not otherwise provided for in this section",
  
  B01 = "Physical or chemical processes or apparatus in general",
  B02 = "Crushing, pulverising, or disintegrating; Preparatory treatment of grain for milling",
  B03 = "Separation of solid materials using liquids or using pneumatic tables",
  B04 = "Centrifugal apparatus or machines for carrying-out physical or chemical processes",
  B05 = "Spraying or atomising in general; Applying fluent materials to surfaces",
  B06 = "Generating or transmitting mechanical vibrations in general",
  B07 = "Separating solids from solids; Sorting",
  B08 = "Cleaning",
  B09 = "Disposal of solid waste; Reclamation of contaminated soil",
  B21 = "Mechanical metal-working without essentially removing material; Punching metal",
  B22 = "Casting; Powder metallurgy",
  B23 = "Machine tools; Metal-working not otherwise provided for",
  B24 = "Grinding; Polishing",
  B25 = "Hand tools; Portable power-driven tools; Manipulators",
  B26 = "Hand cutting tools; Cutting; Severing",
  B27 = "Working or preserving wood or similar material; Nailing or stapling machines",
  B28 = "Working cement, clay, or stone",
  B29 = "Working of plastics; Working of substances in a plastic state in general",
  B30 = "Presses",
  B31 = "Making articles of paper, cardboard or material worked in a manner analogous to paper",
  B32 = "Layered products",
  B33 = "Additive manufacturing technology",
  B41 = "Printing; Lining machines; Typewriters; Stamps",
  B42 = "Bookbinding; Albums; Files; Special printed matter",
  B43 = "Writing or drawing implements; Bureau accessories",
  B44 = "Decorative arts",
  B60 = "Vehicles in general",
  B61 = "Railways",
  B62 = "Land vehicles for travelling otherwise than on rails",
  B63 = "Ships or other waterborne vessels; Related equipment",
  B64 = "Aircraft; Aviation; Cosmonautics",
  B65 = "Conveying; Packing; Storing; Handling thin or filamentary material",
  B66 = "Hoisting; Lifting; Hauling; Pushing; Positioning",
  B67 = "Opening, closing or cleaning bottles, jars or similar containers; Liquid handling",
  B68 = "Saddlery; Upholstery",
  B81 = "Microstructural technology",
  B82 = "Nanotechnology",
  
  C01 = "Inorganic chemistry",
  C02 = "Treatment of water, waste water, sewage, or sludge",
  C03 = "Glass; Mineral or slag wool",
  C04 = "Cements; Concrete; Artificial stone; Ceramics; Refractories",
  C05 = "Fertilisers; Manufacture thereof",
  C06 = "Explosives; Matches",
  C07 = "Organic chemistry",
  C08 = "Organic macromolecular compounds; Their preparation or chemical working-up; Compositions",
  C09 = "Dyes; Paints; Polishes; Natural resins; Adhesives; Compositions; Miscellaneous applications",
  C10 = "Petroleum, gas or coke industries; Technical gases containing carbon monoxide",
  C11 = "Animal or vegetable oils, fats, fatty substances or waxes; Fatty acids; Detergents; Candles",
  C12 = "Biochemistry; Beer; Spirits; Wine; Vinegar; Microbiology; Enzymology; Mutation or genetic engineering",
  C13 = "Sugar industry",
  C14 = "Skins; Hides; Pelts; Leather",
  C21 = "Metallurgy of iron",
  C22 = "Metallurgy; Ferrous or non-ferrous alloys; Treatment of alloys or non-ferrous metals",
  C23 = "Coating metallic material; Coating material with metallic material; Surface treatment",
  C25 = "Electrolytic or electrophoretic processes; Apparatus therefor",
  C30 = "Crystal growth",
  C40 = "Combinatorial technology",
  C99 = "Subject matter not otherwise provided for in this section",
  
  D01 = "Natural or man-made threads or fibres; Spinning",
  D02 = "Yarns; Mechanical finishing of yarns or ropes; Warping or beaming",
  D03 = "Weaving",
  D04 = "Braiding; Lace-making; Knitting; Trimmings; Non-woven fabrics",
  D05 = "Sewing; Embroidering; Tufting",
  D06 = "Treatment of textiles or the like; Laundering; Flexible materials not otherwise provided for",
  D07 = "Ropes; Cables other than electric",
  D10 = "Indexing scheme associated with subclasses of section D, relating to textiles",
  D21 = "Paper-making; Production of cellulose",
  
  E01 = "Construction of roads, railways, or bridges",
  E02 = "Hydraulic engineering; Foundations; Soil shifting",
  E03 = "Water supply; Sewerage",
  E04 = "Building",
  E05 = "Locks; Keys; Window or door fittings; Safes",
  E06 = "Doors, windows, shutters, or roller blinds in general; Ladders",
  E21 = "Earth or rock drilling; Mining",
  E99 = "Subject matter not otherwise provided for in this section",
  
  F01 = "Machines or engines in general; Engine plants in general; Steam engines",
  F02 = "Combustion engines; Hot-gas or combustion-product engine plants",
  F03 = "Machines or engines for liquids; Wind, spring, or weight motors; Producing mechanical power",
  F04 = "Positive-displacement machines for liquids; Pumps for liquids or elastic fluids",
  F05 = "Indexing schemes relating to engines or pumps in various subclasses of classes F01-F04",
  F15 = "Fluid-pressure actuators; Hydraulics or pneumatics in general",
  F16 = "Engineering elements or units; General measures for producing and maintaining effective functioning",
  F17 = "Storing or distributing gases or liquids",
  F21 = "Lighting",
  F22 = "Steam generation",
  F23 = "Combustion apparatus; Combustion processes",
  F24 = "Heating; Ranges; Ventilating",
  F25 = "Refrigeration or cooling; Combined heating and refrigeration systems; Heat pump systems",
  F26 = "Drying",
  F27 = "Furnaces; Kilns; Ovens; Retorts",
  F28 = "Heat exchange in general",
  F41 = "Weapons",
  F42 = "Ammunition; Blasting",
  F99 = "Subject matter not otherwise provided for in this section",
  
  G01 = "Measuring; Testing",
  G02 = "Optics",
  G03 = "Photography; Cinematography; Analogous techniques using waves other than optical waves",
  G04 = "Horology",
  G05 = "Controlling; Regulating",
  G06 = "Computing; Calculating or counting",
  G07 = "Checking-devices",
  G08 = "Signalling",
  G09 = "Educating; Cryptography; Display; Advertising; Seals",
  G10 = "Musical instruments; Acoustics",
  G11 = "Information storage",
  G12 = "Instrument details",
  G16 = "Information and communication technology specially adapted for specific application fields",
  G21 = "Nuclear physics; Nuclear engineering",
  G99 = "Subject matter not otherwise provided for in this section",
  
  H01 = "Electric elements",
  H02 = "Generation, conversion, or distribution of electric power",
  H03 = "Electronic circuitry",
  H04 = "Electric communication technique",
  H05 = "Electric techniques not otherwise provided for",
  H99 = "Subject matter not otherwise provided for in this section",
  
  Y02 = "Technologies or applications for mitigation or adaptation against climate change",
  Y04 = "Information or communication technologies having an impact on other technology areas",
  Y10 = "Technical subjects covered by former USPC"
)



```



```{r}
ll=cpc_df %>% pull(heading) %>% unlist()

 paste(ll,",") %>% cat()

```


```{r}

#rm(sectionmap)
library(fst)
countrymap=read_fst("countrymap.fst")

sectionmap1=sectionmap1 %>% inner_join(countrymap %>% distinct(docdb_family_id)) %>% 
  inner_join(cpc_df) %>% 
  
sections=sectionmap1  %>% rename(technology=heading) %>% select(-section)





library(fst)
techmap <- read_fst("techmap.fst")
techs=techmap %>% distinct(technology)

if(!"Physics" %in% techs){
  techmap=techmap %>% bind_rows(sections)
  techmap=techmap %>% distinct()
  
  techmap=techmap %>%  inner_join(countrymap %>% distinct(docdb_family_id))
  #write_parquet(techmap,"techmap.parquet")
  write_fst(techmap,"techmap.fst",compress=100)
}




```



# Write things to dropbox

```{r}



```