---
title: "Prepare various bits of data"
output: html_notebook
---

```{r}

library(dplyr)

```



```{r}
library(dplyr)
if (Sys.info()["nodename"] == "PCACL-9RVSQ74") {
  datad="C:\\Users\\rmartin7\\OneDrive - WBG\\projects\\PRINZGlobal\\imperial_college_visualization\\01_data"
  localbig="C:\\Users\\rmartin7\\OneDrive - Imperial College London\\inglobe"
  ifcone="C:\\Users\\rmartin7\\OneDrive - WBG\\projects\\PRINZGlobal\\KEVCexplorer"
}

```


# cpc data

```{r}
cpcfile=paste0(localbig,"\\data\\cpcs.parquet")
if (!file.exists(cpcfile)) {
  library(bigrquery)
  library(DBI)
  
  # Set your project ID
  project_id <- "patbis"
  
  # Option 1: Direct table download
  # Specify dataset and table name
  dataset <- "fromPATSTAT2021"
  table <- "tls225_docdb_fam_cpc"
  
  # Download the entire table
  cpcs <- bq_table_download(
    bq_table(project_id, dataset, table)
  )
  
  
  cpcs=cpcs %>% select(docdb_family_id,cpc_class_symbol)
  cpcs=cpcs %>% arrange(docdb_family_id)
  library(arrow)
  write_parquet(cpcs,cpcfile)
}




```


  
```{r}

library(arrow)
cpcs=read_parquet(cpcfile)


#test=cpcs %>% summarize(n_distinct(docdb_family_id))

cpcs=cpcs %>% arrange(docdb_family_id)


library(stringr)



cpcs=cpcs %>% gsub(" ", "", cpc_class_symbol)



library(data.table)
library(stringi)

setDT(cpcs)
cpcs[, cpc_class_symbol := stri_replace_all_fixed(cpc_class_symbol, " ", "")]


```


# Battery technology

## mapping file
```{r}


source("extrasectorshelper.r") # definitions of the battery sector

battery_df_expanded <- battery_df %>%
  separate_rows(CPC, sep = ";")%>%
  mutate(cpc_class_symbol = trimws(CPC))

setDT(battery_df_expanded)
battery_df_expanded[, cpc_class_symbol := stri_replace_all_fixed(cpc_class_symbol, " ", "")]


battery_df %>% 
  distinct(technology) %>% 
  pull(technology) %>% 
  paste0('"', ., '"') %>%
  paste(collapse = ", ") %>% cat()



batteryclasses=cpcs %>% inner_join(battery_df_expanded %>% select(technology,cpc_class_symbol)) 

all=batteryclasses %>% distinct(docdb_family_id) %>% mutate(technology="Any battery")
batteryclasses=batteryclasses %>% bind_rows(all)
batteryclasses=batteryclasses %>% select(docdb_family_id,technology)


```



```{r}
library(arrow)
techmap <- read_parquet("techmap.parquet")
techs=techmap %>% distinct(technology)
if(!"Any battery" %in% techs){
  techmap=techmap %>% bind_rows(batteryclasses)
  
  techmap = techmap %>% mutate(technology=ifelse(technology=="Any battery","Any battery technology",technology))
  techmap = techmap %>% mutate(technology=ifelse(technology=="Any Green","Any Green technology",technology))
  write_parquet(techmap,"techmap.parquet")
}


```

## Make an experimental green class (to check on the disparity in green share)

```{r}
#test=head(cpcs,1000)
head(cpcs)
xgreenclasses=cpcs %>% mutate(technology = substr(cpc_class_symbol, 1, 3)) %>% filter(technology=="Y02") %>% distinct(docdb_family_id,technology)
  
library(arrow)
techmap <- read_parquet("techmap.parquet")


test=techmap %>% filter(technology=="Any Green technology")


techs=techmap %>% distinct(technology)
techs %>% pull(technology)

test2=test %>% select(-technology) %>% inner_join(xgreenclasses)

nrow(test2)
nrow(test)
nrow(xgreenclasses)




```
ok, it seems Any Green is truly Y02



## Analysis of battery...
### Comparative advantage in battery


### tech shares by countries
```{r}


countrymap9918 <- read_parquet("countrymap9918.parquet")

countrymap9918 %>% distinct(docdb_family_id) %>% nrow()

library(data.table)

  tech_shares <- countrymap9918 %>%
  left_join(techmap, by = "docdb_family_id") %>%
  mutate(innos_total=n_distinct(docdb_family_id)) %>% 
  group_by(ctry_code)  %>% mutate(innos_total_c=n_distinct(docdb_family_id)) %>% 
  group_by(technology) %>% mutate(innos_t=n_distinct(docdb_family_id)) %>% 
  group_by(ctry_code, technology) %>%
    summarise(innos_total_c= mean(innos_total_c),
              innos_total  = mean(innos_total),
              innos_t      = mean(innos_t), test=n(),
              innos_ct     = n_distinct(docdb_family_id), .groups = "drop") %>%
    mutate(
      
      share_ct = innos_ct / innos_total_c,
      share_t=innos_t/innos_total,
      
      RTA=share_ct*2 / (share_ct+share_t)-1
      
    ) %>%
  ungroup()


library(WDI)
library(dplyr)

# Download GDP per capita data for 2015
gdp_data <- WDI(
  indicator = c(
    "gdp_pc_current" = "NY.GDP.PCAP.CD",      # GDP per capita at market exchange rates
    "gdp_pc_ppp" = "NY.GDP.PCAP.PP.CD"        # GDP per capita at PPP
  ),
  start = 2015,
  end = 2015,
  extra = FALSE
) %>%
  select(ctry_code = iso2c, gdp_pc_current, gdp_pc_ppp)

# Add to tech_shares
tech_shares <- tech_shares %>%
  left_join(gdp_data, by = "ctry_code")  





# World Bank income classifications (2024/2025)
income_classification <- tibble::tribble(
  ~income_group, ~ctry_code,
  "High income", c("AD", "AE", "AG", "AR", "AU", "AT", "AW", "BE", "BH", "BM", "BS", "BW", "CA", "CH", "CL", "CN", "CW", "CY", "CZ", "DE", "DK", "EE", "ES", "FI", "FR", "GB", "GI", "GR", "GU", "HK", "HR", "HU", "IE", "IL", "IM", "IS", "IT", "JP", "KR", "KW", "KY", "LI", "LT", "LU", "LV", "MC", "MP", "MT", "MO", "NL", "NO", "NZ", "OM", "PA", "PL", "PR", "PT", "QA", "RO", "SA", "SC", "SG", "SI", "SK", "SM", "SE", "TC", "TT", "TW", "UY", "US", "VI", "VG"),
  "Upper middle income", c("AL", "AM", "AZ", "BY", "BZ", "BR", "BG", "CO", "CR", "CU", "DM", "DO", "EC", "FJ", "GA", "GD", "GE", "GQ", "GT", "GY", "ID", "IQ", "IR", "JM", "JO", "KZ", "LB", "LC", "LY", "MK", "MU", "MV", "MX", "MY", "NA", "PE", "PY", "RU", "RS", "SR", "SV", "TH", "TM", "TO", "TR", "VE", "ZA", "ME", "XK", "ST", "VC", "WS"),
  "Lower middle income", c("AO", "BD", "BJ", "BO", "BT", "CI", "CM", "CG", "CV", "DZ", "EG", "GH", "HN", "HT", "IN", "KE", "KG", "KH", "KI", "LA", "LK", "LS", "MA", "MD", "MN", "MM", "MR", "NG", "NI", "NP", "PK", "PG", "PH", "PS", "SB", "SN", "TJ", "TL", "TN", "TZ", "UA", "UZ", "VN", "VU", "ZM", "ZW"),
  "Low income", c("AF", "BF", "BI", "CD", "CF", "ER", "ET", "GM", "GN", "GW", "KM", "LR", "MG", "ML", "MW", "MZ", "NE", "RW", "SD", "SL", "SO", "SS", "SY", "TD", "TG", "UG", "YE")
) %>%
  tidyr::unnest(ctry_code)

# Add income classification to your dataframe
tech_shares <- tech_shares %>%
  left_join(income_classification, by = "ctry_code") %>%  mutate(HIC = if_else(income_group == "High income", 1, 0))


  
  
```


```{r}
techmap %>% distinct(technology) %>% pull(technology)

```

```{r}
library(ggplot2)

ctech="Any Hard to Abate technology"  
ctech="Any battery technology"
ctech="Any Green technology"
mini=100

y_var="gdp_pc_ppp"
#y_var="gdp_pc_current"
filtered=tech_shares %>% filter(technology==ctech) %>% filter(innos_total_c>=mini)
  
filtered %>%   
    ggplot(aes(x=RTA,y=log(!!sym(y_var)),size=innos_total_c))+geom_point()+
    geom_smooth(method="lm")
  




```

#### tech shares by country group


```{r}

countrymap9918 <- read_parquet("countrymap9918.parquet")
library(dplyr)


countrymap9918 <- countrymap9918 %>%
  left_join(income_classification, by = "ctry_code") %>%
  mutate(
    cgroup1 = case_when(
      ctry_code == "VN" ~ "Vietnam",
      ctry_code == "CN" ~ "China",
      income_group == "High income" ~ "High income countries",
      TRUE ~ "Other non-high income countries",


    )
  ) %>%
  select(-income_group)  


#countrymap %>% distinct(cgroup1) %>% pull(cgroup1)
#countrymap %>% distinct(ctry_code) %>% arrange(ctry_code) %>%  pull(ctry_code)

```


check what is going on 99-18 & green

```{r}

check9918=countrymap9918 %>% left_join(xgreenclasses) %>% 
  mutate(technology=replace_na(technology,"notY02")) %>% 
  select(cgroup1,docdb_family_id,technology) %>% distinct() %>% 
  group_by(cgroup1) %>% summarise(innos=n(),greeninnos=sum(technology=="Y02")) %>% 
  mutate(share=greeninnos/innos)

check9918$greeninnos %>% sum() 
#if(!"Y02" %in% techs){
#  techmap=techmap %>% bind_rows(batteryclasses)
  
#  write_parquet(techmap,"techmap.parquet")
#}
  



```


# 👌 Let's do this over time...

```{r}

countrymap9918 %>% distinct(ctry_code) %>% nrow()
testb=countrymap9918 %>% distinct(ctry_code)

check9918=countrymap9918 %>% left_join(xgreenclasses) %>% 
  mutate(technology=replace_na(technology,"notY02")) %>% 
  select(cgroup1,docdb_family_id,technology) %>% distinct() %>% 
  group_by(cgroup1) %>% summarise(innos=n(),greeninnos=sum(technology=="Y02")) %>% 
  mutate(share=greeninnos/innos)

check9918$greeninnos %>% sum() 
#if(!"Y02" %in% techs){
#  techmap=techmap %>% bind_rows(batteryclasses)
  
#  write_parquet(techmap,"techmap.parquet")
#}

#*testa %>% rename(ctry_code=person_ctry_code) %>%  anti_join(testb) %>% pull(ctry_code)

```


# 😒 How about taking the income classes from the cs file?


```{r}
ifcone="C:\\Users\\rmartin7\\OneDrive - WBG\\projects\\PRINZGlobal\\KEVCexplorer"

path=paste0(ifcone,"\\","../work packages/patent analysis/Data/combined_data/count_docdb_2025-03-04.csv")

path
new_group_map=read.csv(path) %>% arrange(Economy, earliest_filing_year) %>% 
  select(person_ctry_code,Income.group) %>% distinct() %>% 
  rename(cgroup_new=Income.group,
         ctry_code=person_ctry_code)



countrymap9918=countrymap9918 %>% left_join(new_group_map)

check9918_new=countrymap9918 %>% left_join(xgreenclasses) %>% 
  mutate(technology=replace_na(technology,"notY02")) %>% 
  select(cgroup_new,docdb_family_id,technology) %>% distinct() %>% 
  group_by(cgroup_new) %>% summarise(innos=n(),greeninnos=sum(technology=="Y02")) %>% 
  mutate(share=greeninnos/innos)

check9918_new

```




#### Define functions
```{r}
create_country_groups <- function(df, country_list, new_var_name = "cgroup") {
  
  # Join income classification
  df_with_income <- df %>%
    left_join(income_classification, by = "ctry_code")
  
  # Create the grouping variable dynamically
  for(i in seq_along(country_list)) {
    country_code <- country_list[i]
    country_label <- names(country_list)[i]
    
    if(i == 1) {
      df_with_income <- df_with_income %>%
        mutate(!!new_var_name := case_when(
          ctry_code == country_code ~ country_label,
          TRUE ~ NA_character_
        ))
    } else {
      df_with_income <- df_with_income %>%
        mutate(!!new_var_name := case_when(
          ctry_code == country_code ~ country_label,
          TRUE ~ !!sym(new_var_name)
        ))
    }
  }
  
  # Add high income and other categories
  df_with_income <- df_with_income %>%
    mutate(!!new_var_name := case_when(
      !is.na(!!sym(new_var_name)) ~ !!sym(new_var_name),
      income_group == "High income" ~ "High income countries",
      TRUE ~ "Other non-high income countries"
    )) %>%
    select(-income_group)
  
  return(df_with_income)
}

library(dplyr)
library(rlang)

calculate_tech_shares <- function(countrymap, techmap, group_var) {
  
  # Convert string to symbol for tidy evaluation
  group_var_sym <- sym(group_var)
  
  tech_shares <- countrymap %>%
    left_join(techmap, by = "docdb_family_id") %>%
    mutate(innos_total = n_distinct(docdb_family_id)) %>% 
    group_by(!!group_var_sym) %>% 
    mutate(innos_total_c = n_distinct(docdb_family_id)) %>% 
    group_by(technology) %>% 
    mutate(innos_t = n_distinct(docdb_family_id)) %>% 
    group_by(!!group_var_sym, technology) %>%
    summarise(
      innos_total_c = mean(innos_total_c),
      innos_total   = mean(innos_total),
      innos_t       = mean(innos_t), 
      innos_ct      = n_distinct(docdb_family_id), 
      .groups = "drop"
    ) %>%
    mutate(
      share_ct = innos_ct / innos_total_c,
      share_t  = innos_t / innos_total,
      RTA      = share_ct * 2 / (share_ct + share_t) - 1
    ) %>%
    ungroup()
  
  return(tech_shares)
}
```


```{r}
country_list <- c("VN" = "Vietnam", "CN" = "China")
test=create_country_groups(countrymap9918, country_list, "cgroup1")

tech_shares_country <- calculate_tech_shares(countrymap9918,
                                             techmap, "cgroup1")


ctech="Any Hard to Abate technology"  
ctech="Any battery technology"
#ctech="Any Green technology"

test=tech_shares_country %>% filter(technology==ctech)

```





# Hard to abate sectors


```{r}
library(readxl)
tech_df <- read_excel("hard_to_abate.xlsx") %>% 
  rename(detail=technology,technology=sector) %>% 
  mutate(technology=paste0(technology," Decarbonisation"))


tech_df_expanded <- tech_df %>%
  separate_rows(CPC, sep = ";")%>%
  mutate(cpc_class_symbol = trimws(CPC))

setDT(tech_df_expanded)
tech_df_expanded[, cpc_class_symbol := stri_replace_all_fixed(cpc_class_symbol, " ", "")]


tech_df %>% 
  distinct(technology) %>% 
  pull(technology) %>% 
  paste0('"', ., '"') %>%
  paste(collapse = ", ") %>% cat()



tech_classes=cpcs %>% inner_join(tech_df_expanded %>% select(technology,cpc_class_symbol)) 

all=tech_classes %>% distinct(docdb_family_id) %>% mutate(technology="Any Hard to Abate technology")
tech_classes=tech_classes %>% bind_rows(all)
tech_classes=tech_classes %>% select(docdb_family_id,technology)



```



```{r}


techmap <- read_parquet("techmap.parquet")
techs=techmap %>% distinct(technology)

if(!"Aviation Decarbonisation" %in% techs){
  techmap=techmap %>% bind_rows(tech_classes)
  
  write_parquet(techmap,"techmap.parquet")
}


```




# AI

```{r}
library(readxl)
tech_df <- read_excel("New_Sector_mapping.xlsx",sheet="AI") %>% 
  rename(CPC=`CPC/IPC Codes`) %>% 
  rename(technology=`Sub-Technology`) %>% filter(!is.na(technology))
  
  

tech_df_expanded <- tech_df %>%
  separate_rows(CPC, sep = ",")%>%
  mutate(cpc_class_symbol = trimws(CPC))

setDT(tech_df_expanded)
tech_df_expanded[, cpc_class_symbol := stri_replace_all_fixed(cpc_class_symbol, " ", "")]


tech_df %>% 
  distinct(technology) %>% 
  pull(technology) %>% 
  paste0('"', ., '"') %>%
  paste(collapse = ", ") %>% cat()



tech_classes=cpcs %>% inner_join(tech_df_expanded %>% select(technology,cpc_class_symbol)) 

#all=tech_classes %>% distinct(docdb_family_id) %>% mutate(technology="Any Hard to Abate technology")
#tech_classes=tech_classes %>% bind_rows(all)
tech_classes=tech_classes %>% select(docdb_family_id,technology)



```



```{r}


techmap <- read_parquet("techmap.parquet")
techs=techmap %>% distinct(technology)

if(!"Machine Learning" %in% techs){
  techmap=techmap %>% bind_rows(tech_classes)
  techmap=techmap %>% distinct()
  write_parquet(techmap,"techmap.parquet")
}


```
