---
title: "Patent Abstracts — Bigram Word Cloud"
output:
  html_document:
    df_print: paged
---

This notebook fetches patent abstracts from the EPO Open Patent Services (OPS)
API for a set of application numbers sourced from an Espacenet search, then
draws a word cloud of the most frequent bigrams.

## 0 — Setup

```{r setup, message=FALSE, warning=FALSE}
# Install any missing packages
pkgs <- c("httr2", "xml2", "dplyr", "tidyr", "tidytext", "stringr",
           "ggplot2", "ggwordcloud", "purrr", "tibble")
install.packages(setdiff(pkgs, rownames(installed.packages())), quiet = TRUE)

library(httr2)
library(xml2)
library(dplyr)
library(tidyr)
library(tidytext)
library(stringr)
library(ggplot2)
library(ggwordcloud)
library(purrr)
library(tibble)
```

## 1 — Application numbers

Extracted from the Espacenet search URL:

```{r app_numbers}
app_numbers <- c(
 "CN201810134850", "PT12858932",      "US89428710",       "US201414220990",
 "ARP120101066",   "EP09305250",      "CA3117920",        "SG11201506775V",
 "EP11833120",     "KR20207024776",   "AU2011349446",     "US201213679498",
 "BR112013025572", "KR20127020023",   "CN201380028604",   "US59260509",
 "US201114008192", "US76340110",      "MX2014002289",     "WO2011039386",
 "WO2011021301",   "HUE16193906",     "WO2009007269",     "US201414293546",
 "US201113107352", "US201314425505",  "SG10201504187Y",   "US60981509",
 "US201615193033", "WO2010003130",    "PL13159003",       "BRPI1002602",
 "WO2015033929",   "BRPI1102284",     "AU2015314826",     "US86898910",
 "BR112017009050", "WO2011038946",    "KR20147016656",    "CN201080049139",
 "PL11858867",     "AU2013360086",    "ES11728509",       "WO2010000938",
 "US48126009",     "EP13762719",      "US201313920673",   "NZ61861310",
 "MX2011003856",   "ARP110104677"
)

cat(length(app_numbers), "application numbers to process\n")
```

## 2 — Helper: parse an application number into country + doc-number

The OPS API expects the EPODOC format `CC.number` (country code dot number).

```{r parse_appnum}
parse_app_number <- function(x) {
  # Match leading 2-letter country code (some have 4-letter like BRPI)
  m <- str_match(x, "^([A-Z]{2,4})(\\d.*)$")
  tibble(raw = x, country = m[, 2], docnumber = m[, 3])
}

parsed <- bind_rows(lapply(app_numbers, parse_app_number))
parsed
```

## 3 — Authenticate with EPO OPS

The OPS API requires **free** developer credentials.
Register at <https://developers.epo.org/> to get a Consumer Key and Consumer
Secret, then paste them below.

```{r ops_auth}
# -------------------------------------------------------------------
# PASTE YOUR OPS CREDENTIALS HERE (or set as environment variables)
# -------------------------------------------------------------------
ops_key    <- Sys.getenv("OPS_CONSUMER_KEY",    unset = "YOUR_KEY_HERE")
ops_secret <- Sys.getenv("OPS_CONSUMER_SECRET", unset = "YOUR_SECRET_HERE")

stopifnot("Set your OPS credentials above or via env vars OPS_CONSUMER_KEY / OPS_CONSUMER_SECRET" =
            ops_key != "YOUR_KEY_HERE")

# OAuth2 client-credentials token
token_resp <- request("https://ops.epo.org/3.2/auth/accesstoken") |>
  req_auth_basic(ops_key, ops_secret) |>
  req_body_form(grant_type = "client_credentials") |>
  req_perform()

access_token <- resp_body_json(token_resp)$access_token
cat("OPS access token obtained.\n")
```

## 4a — Debug: inspect raw XML for one patent

Let's see what OPS actually returns so we can write the right XPath.

```{r debug_xml, message=TRUE}
# Pick one that returned "no abstract" — e.g. EP.09305250
test_url <- paste0(
  "https://ops.epo.org/3.2/rest-services/published-data/application/epodoc/",
  "EP.09305250/abstract"
)
test_resp <- request(test_url) |>
  req_headers(Authorization = paste("Bearer", access_token),
              Accept        = "application/xml") |>
  req_error(is_error = \(resp) FALSE) |>
  req_perform()

cat("Status:", resp_status(test_resp), "\n\n")
if (resp_status(test_resp) < 400) {
  raw_xml <- resp_body_string(test_resp)
  cat(raw_xml)
}
```

## 4b — Fetch abstracts from OPS

We query the *published-data* endpoint one patent at a time.
OPS anonymous throttle is ~4 req/s; we add a small delay to be polite.

```{r fetch_abstracts, message=TRUE, warning=TRUE}
ns <- c(ex = "http://www.epo.org/exchange", ops = "http://ops.epo.org")

extract_en_abstract <- function(doc) {
  abs_nodes <- xml_find_all(doc, ".//ex:abstract", ns)
  langs <- xml_attr(abs_nodes, "lang")
  en <- abs_nodes[!is.na(langs) & langs == "en"]
  node <- if (length(en) > 0) en[[1]] else NULL
  if (is.null(node)) return(NA_character_)

  p_nodes <- xml_find_all(node, ".//ex:p", ns)
  text <- if (length(p_nodes) > 0) paste(xml_text(p_nodes), collapse = " ") else xml_text(node)
  text <- str_squish(text)
  if (nchar(text) < 10) NA_character_ else text
}

extract_en_title <- function(doc) {
  title_nodes <- xml_find_all(doc, ".//ex:invention-title", ns)
  langs <- xml_attr(title_nodes, "lang")
  en <- title_nodes[!is.na(langs) & langs == "en"]
  node <- if (length(en) > 0) en[[1]] else NULL
  if (is.null(node)) return(NA_character_)
  str_squish(xml_text(node))
}

fetch_patent <- function(country, docnumber, token) {
  epodoc <- paste0(country, ".", docnumber)
  abstract <- NA_character_
  title    <- NA_character_

  # Helper to try one URL and extract abstract + title
  try_endpoint <- function(url) {
    resp <- request(url) |>
      req_headers(Authorization = paste("Bearer", token),
                  Accept        = "application/xml") |>
      req_error(is_error = \(resp) FALSE) |>
      req_perform()
    if (resp_status(resp) >= 400) return(NULL)
    resp_body_xml(resp)
  }

  # --- Step 1: direct application endpoint (biblio for title + abstract) ---
  tryCatch({
    doc <- try_endpoint(paste0(
      "https://ops.epo.org/3.2/rest-services/published-data/application/epodoc/",
      epodoc, "/biblio,abstract"
    ))
    if (!is.null(doc)) {
      abstract <- extract_en_abstract(doc)
      title    <- extract_en_title(doc)
    }
  }, error = function(e) NULL)

  # --- Step 2: family fallback if no English abstract ---
  if (is.na(abstract)) {
    tryCatch({
      Sys.sleep(0.3)
      doc <- try_endpoint(paste0(
        "https://ops.epo.org/3.2/rest-services/family/application/epodoc/",
        epodoc, "/biblio,abstract"
      ))
      if (!is.null(doc)) {
        abstract <- extract_en_abstract(doc)
        if (is.na(title)) title <- extract_en_title(doc)
      }
    }, error = function(e) NULL)
  }

  src <- if (!is.na(abstract)) "OK" else "miss"
  message("  [", src, "] ", epodoc,
          if (!is.na(title)) paste0(" — ", str_trunc(title, 60)) else "")

  tibble(title = title, abstract = abstract)
}

results <- parsed |>
  mutate(data = map2(country, docnumber, \(cc, dn) {
    Sys.sleep(0.3)
    fetch_patent(cc, dn, access_token)
  })) |>
  unnest(data)

cat("\nFetched", sum(!is.na(results$abstract)), "abstracts and",
    sum(!is.na(results$title)), "titles out of",
    nrow(results), "applications\n")

# Keep backward-compatible name for downstream chunks
abstracts <- results
```

## 5 — AI-derived innovation labels

For each patent we ask OpenAI to produce a short n-gram (2–5 words) that
captures the core innovation, using both the title and the abstract.

```{r ai_labels, message=TRUE}
openai_key <- Sys.getenv("OPENAI_API_KEY", unset = "")
stopifnot("Set OPENAI_API_KEY in your .Renviron" = nchar(openai_key) > 0)

# Build a batch prompt — send all patents in one call for speed & cost
patent_block <- abstracts |>
  filter(!is.na(abstract) | !is.na(title)) |>
  mutate(entry = paste0(
    "[", row_number(), "] ",
    if_else(!is.na(title), paste0("Title: ", title, ". "), ""),
    if_else(!is.na(abstract), paste0("Abstract: ", str_trunc(abstract, 500)), "")
  )) |>
  pull(entry) |>
  paste(collapse = "\n\n")

label_prompt <- paste0(
 "Below are patents, each with a number in brackets, a title and/or abstract.\n\n",
  patent_block,
  "\n\n---\n",
  "For EACH patent, produce a short descriptive n-gram (2 to 5 words, lowercase) ",
  "that captures the core innovation or technology. ",
  "Do NOT use generic phrases like 'novel method' or 'improved system'. ",
  "Be specific and technical, e.g. 'antibody fc region engineering', ",
  "'lithium battery cathode recycling', 'crispr gene editing delivery'.\n\n",
  "Return ONLY a JSON array of objects with keys \"id\" (the bracket number) ",
  "and \"label\" (the n-gram). No other text."
)

label_resp <- request("https://api.openai.com/v1/chat/completions") |>
  req_headers(Authorization = paste("Bearer", openai_key),
              `Content-Type` = "application/json") |>
  req_body_json(list(
    model    = "gpt-4o-mini",
    messages = list(list(role = "user", content = label_prompt)),
    temperature = 0.2,
    max_tokens  = 2000
  )) |>
  req_perform()

raw_json <- resp_body_json(label_resp)$choices[[1]]$message$content

# Strip markdown code fences if present
raw_json <- str_remove_all(raw_json, "```json\\s*|```\\s*")

labels_df <- jsonlite::fromJSON(raw_json) |>
  as_tibble() |>
  mutate(id = as.integer(id),
         label = str_to_lower(str_squish(label)))

# Attach back to the abstracts that were sent
sent <- abstracts |>
  filter(!is.na(abstract) | !is.na(title)) |>
  mutate(id = row_number())

labelled <- sent |>
  left_join(labels_df, by = "id")

cat("Labelled", sum(!is.na(labelled$label)), "patents\n")
labelled |> select(raw, label) |> print(n = 60)
```

## 6 — AI-label word cloud

```{r ai_wordcloud, fig.width=10, fig.height=8, dpi=150}
label_counts <- labelled |>
  filter(!is.na(label)) |>
  count(label, sort = TRUE)

ggplot(label_counts, aes(label = label, size = n, colour = n)) +
  geom_text_wordcloud_area(rm_outside = TRUE, shape = "circle",
                           eccentricity = 0.65) +
  scale_size_area(max_size = 16) +
  scale_colour_gradient(low = "#2c7bb6", high = "#d7191c") +
  theme_minimal() +
  labs(title = "Innovation Word Cloud — AI-derived Labels",
       subtitle = paste(nrow(label_counts), "distinct innovation themes"))
```

## 7 — AI-label bar chart

```{r ai_bar_chart, fig.width=9, fig.height=7}
label_counts |>
  slice_max(n, n = 30, with_ties = FALSE) |>
  mutate(label = reorder(label, n)) |>
  ggplot(aes(n, label, fill = n)) +
  geom_col(show.legend = FALSE) +
  scale_fill_gradient(low = "#2c7bb6", high = "#d7191c") +
  labs(x = "Frequency", y = NULL,
       title = "Top Innovation Themes (AI-labelled)") +
  theme_minimal()
```

## 8 — Statistical bigrams (original)

For reference, the traditional bigram word cloud from raw abstract text.

```{r bigrams}
custom_stop <- tibble(word = c("patent", "invention", "present", "method",
                               "comprises", "comprising", "device", "apparatus",
                               "provided", "relates", "described", "including",
                               "wherein", "thereof", "according"))

bigrams <- abstracts |>
  filter(!is.na(abstract)) |>
  select(raw, abstract) |>
  unnest_tokens(bigram, abstract, token = "ngrams", n = 2) |>
  separate(bigram, into = c("word1", "word2"), sep = " ", remove = FALSE) |>
  anti_join(stop_words, by = c("word1" = "word")) |>
  anti_join(stop_words, by = c("word2" = "word")) |>
  anti_join(custom_stop, by = c("word1" = "word")) |>
  anti_join(custom_stop, by = c("word2" = "word")) |>
  filter(!str_detect(word1, "\\d"), !str_detect(word2, "\\d"))

bigram_counts <- bigrams |>
  count(bigram, sort = TRUE)

head(bigram_counts, 20)
```

```{r stat_wordcloud, fig.width=10, fig.height=8, dpi=150}
top_bigrams <- bigram_counts |>
  slice_max(n, n = 80, with_ties = FALSE)

ggplot(top_bigrams, aes(label = bigram, size = n, colour = n)) +
  geom_text_wordcloud_area(rm_outside = TRUE, shape = "circle",
                           eccentricity = 0.65) +
  scale_size_area(max_size = 14) +
  scale_colour_gradient(low = "#2c7bb6", high = "#d7191c") +
  theme_minimal() +
  labs(title = "Statistical Bigram Word Cloud — Patent Abstracts",
       subtitle = paste(sum(!is.na(abstracts$abstract)), "abstracts"))
```

## 8 — AI Summary of Key Innovations (OpenAI)

Uses the OpenAI API to produce a brief paragraph summarising the most important
innovations across all fetched abstracts.

```{r ai_summary, message=TRUE}
openai_key <- Sys.getenv("OPENAI_API_KEY", unset = "")
stopifnot("Set OPENAI_API_KEY in your .Renviron" = nchar(openai_key) > 0)

# Combine all abstracts into a single block, numbered
abstract_texts <- abstracts |>
  filter(!is.na(abstract)) |>
  mutate(entry = paste0("[", row_number(), "] ", raw, ": ", abstract)) |>
  pull(entry) |>
  paste(collapse = "\n\n")

prompt <- paste0(
  "Below are abstracts from ", sum(!is.na(abstracts$abstract)),
  " patents in the same selection.\n\n",
  abstract_texts,
  "\n\n---\n",
  "Write a single concise paragraph (150-200 words) summarising the most ",
  "important innovations represented by these patents. Focus on the key ",
  "technological themes, what problems they solve, and why they matter. ",
  "Do not list individual patents — synthesise across all of them."
)

resp <- request("https://api.openai.com/v1/chat/completions") |>
  req_headers(Authorization = paste("Bearer", openai_key),
              `Content-Type` = "application/json") |>
  req_body_json(list(
    model    = "gpt-4o-mini",
    messages = list(list(role = "user", content = prompt)),
    temperature = 0.4,
    max_tokens  = 400
  )) |>
  req_perform()

summary_text <- resp_body_json(resp)$choices[[1]]$message$content
```

### Innovation Summary

```{r print_summary, results='asis'}
cat(summary_text)
```
