---
title: "Patent Cluster Comparison — LMICs vs HICs"
output:
  html_document:
    df_print: paged
---

This notebook compares two sets of patents (e.g., LMICs vs HICs) by embedding
their titles and abstracts, clustering them together, and visualizing with
distinct styling for each group.

## 0 — Setup

```{r setup, message=FALSE, warning=FALSE}
pkgs <- c("httr2", "xml2", "dplyr", "tidyr", "tidytext", "stringr",
          "ggplot2", "ggwordcloud", "purrr", "tibble", "digest", "umap",
          "jsonlite", "ggrepel", "DBI", "duckdb", "glue")
install.packages(setdiff(pkgs, rownames(installed.packages())), quiet = TRUE)

library(httr2)
library(xml2)
library(dplyr)
library(tidyr)
library(tidytext)
library(stringr)
library(ggplot2)
library(purrr)
library(tibble)
library(umap)
library(jsonlite)
library(ggrepel)
library(DBI)
library(duckdb)
library(glue)
```

## 1 — Query patent lists from app data

Derive the top Adaptation patents for LMICs and HICs directly from the app's
parquet database, using the same logic as the Shiny app.

```{r query_patents}
# --- Configuration (mirrors the URL parameters) ---
tech_selection <- "Adaptation"
toflow         <- "ev_global"
top_n_ids      <- 500
country_group1 <- "LMICs (excl. China)"
country_group2 <- "High income countries"

# --- Connect to parquet files via DuckDB ---
con <- DBI::dbConnect(duckdb::duckdb())

parquet_dir <- file.path("inst", "extdata")
DBI::dbExecute(con, glue::glue(
  "CREATE OR REPLACE VIEW full_patent_database AS
   SELECT * FROM read_parquet('{file.path(parquet_dir, \"patent_database.parquet\")}')"
))
DBI::dbExecute(con, glue::glue(
  "CREATE OR REPLACE TABLE patents_x_tech AS
   SELECT * FROM read_parquet('{file.path(parquet_dir, \"patents_x_tech.parquet\")}')"
))
DBI::dbExecute(con, glue::glue(
  "CREATE OR REPLACE TABLE tech_lookup AS
   SELECT * FROM read_parquet('{file.path(parquet_dir, \"tech_lookup.parquet\")}')"
))

# --- Load country group definitions from the package's sysdata ---
load("R/sysdata.rda")

# --- Helper: get top patent appln_ids for a country group + tech ---
get_top_patent_ids <- function(con, country_group, tech, toflow, top_n) {
  country_codes <- if (country_group %in% names(group_definitions)) {
    group_definitions[[country_group]]
  } else {
    country_group
  }
  country_sql   <- paste0("'", paste(country_codes, collapse = "', '"), "'")

  sql <- glue::glue("
    WITH filtered_tech AS (
      SELECT DISTINCT t.docdb_family_id, '{tech}' AS technology
      FROM patents_x_tech t
      JOIN tech_lookup tl ON t.technology = tl.technology
      WHERE tl.tech_group = '{tech}' OR t.technology = '{tech}'
    ),

    deduped AS (
      SELECT DISTINCT ON (ft.technology, p.docdb_family_id)
        ft.technology,
        p.docdb_family_id,
        p.appln_id,
        p.{toflow}
      FROM full_patent_database p
      INNER JOIN filtered_tech ft ON p.docdb_family_id = ft.docdb_family_id
      WHERE p.ctry_code IN ({country_sql})
        AND p.{toflow} IS NOT NULL
    ),

    ranked AS (
      SELECT
        appln_id,
        {toflow},
        ROW_NUMBER() OVER (ORDER BY {toflow} DESC) AS rnk
      FROM deduped
    )

    SELECT appln_id
    FROM ranked
    WHERE rnk <= {top_n}
    ORDER BY {toflow} DESC
  ")

  DBI::dbGetQuery(con, sql)$appln_id
}

# --- Fetch patent IDs for both groups ---
apps_lmic <- get_top_patent_ids(con, country_group1, tech_selection, toflow, top_n_ids)
apps_hic  <- get_top_patent_ids(con, country_group2, tech_selection, toflow, top_n_ids)

DBI::dbDisconnect(con, shutdown = TRUE)

cat("LMICs (", country_group1, "):", length(apps_lmic), "applications\n")
cat("HICs (", country_group2, "):", length(apps_hic), "applications\n")

# Union of both lists, tracking source
all_apps <- tibble(
  raw = c(apps_lmic, apps_hic),
  source = c(rep("LMIC", length(apps_lmic)), rep("HIC", length(apps_hic)))
) |>
  distinct(raw, .keep_all = TRUE)

cat("Union (deduplicated):", nrow(all_apps), "applications\n")
table(all_apps$source)
```

## 2 — Parse application numbers

```{r parse_appnum}
parse_app_number <- function(x) {
  m <- str_match(x, "^([A-Z]{2,4})(\\d.*)$")
  tibble(raw = x, country = m[, 2], docnumber = m[, 3])
}

parsed <- all_apps |>
  left_join(bind_rows(lapply(all_apps$raw, parse_app_number)), by = "raw")

parsed
```

## 3 — Authenticate with EPO OPS

```{r ops_auth}
ops_key    <- Sys.getenv("OPS_CONSUMER_KEY",    unset = "YOUR_KEY_HERE")
ops_secret <- Sys.getenv("OPS_CONSUMER_SECRET", unset = "YOUR_SECRET_HERE")

stopifnot("Set your OPS credentials via env vars OPS_CONSUMER_KEY / OPS_CONSUMER_SECRET" =
            ops_key != "YOUR_KEY_HERE")

token_resp <- request("https://ops.epo.org/3.2/auth/accesstoken") |>
  req_auth_basic(ops_key, ops_secret) |>
  req_body_form(grant_type = "client_credentials") |>
  req_perform()

access_token <- resp_body_json(token_resp)$access_token
cat("OPS access token obtained.\n")
```

## 4 — Fetch abstracts (with caching)

```{r fetch_abstracts, message=TRUE, warning=TRUE}
cache_dir <- ".bigdata"
if (!dir.exists(cache_dir)) dir.create(cache_dir, recursive = TRUE)

cache_file <- file.path(cache_dir, paste0(
  "epo_compare_", digest::digest(all_apps$raw, algo = "md5"), ".rds"
))

if (file.exists(cache_file)) {
  cat("Loading cached EPO data from", cache_file, "\n")
  results <- readRDS(cache_file)
  cat("Loaded", sum(!is.na(results$abstract)), "abstracts from cache\n")
} else {
  cat("No cache found, fetching from EPO API...\n")

  ns <- c(ex = "http://www.epo.org/exchange", ops = "http://ops.epo.org")

  extract_en_abstract <- function(doc) {
    abs_nodes <- xml_find_all(doc, ".//ex:abstract", ns)
    if (length(abs_nodes) == 0) return(NA_character_)
    langs <- xml_attr(abs_nodes, "lang")
    # Prefer English; fall back to first available language
    en <- abs_nodes[!is.na(langs) & langs == "en"]
    node <- if (length(en) > 0) en[[1]] else abs_nodes[[1]]

    p_nodes <- xml_find_all(node, ".//ex:p", ns)
    text <- if (length(p_nodes) > 0) paste(xml_text(p_nodes), collapse = " ") else xml_text(node)
    text <- str_squish(text)
    if (nchar(text) < 10) NA_character_ else text
  }

  extract_en_title <- function(doc) {
    title_nodes <- xml_find_all(doc, ".//ex:invention-title", ns)
    if (length(title_nodes) == 0) return(NA_character_)
    langs <- xml_attr(title_nodes, "lang")
    # Prefer English; fall back to first available language
    en <- title_nodes[!is.na(langs) & langs == "en"]
    node <- if (length(en) > 0) en[[1]] else title_nodes[[1]]
    str_squish(xml_text(node))
  }

  extract_applicants <- function(doc) {
    # Find all applicant nodes in docdb format (deduplicated, standardised names)
    app_nodes <- xml_find_all(doc, ".//ex:applicant[@data-format='docdb']", ns)
    if (length(app_nodes) == 0) {
      # Fall back to any applicant nodes
      app_nodes <- xml_find_all(doc, ".//ex:applicant", ns)
    }
    if (length(app_nodes) == 0) {
      return(tibble(
        applicant_names = NA_character_,
        applicant_types = NA_character_,
        is_company      = NA
      ))
    }

    applicants <- lapply(app_nodes, function(node) {
      # Get name — prefer English, fall back to first available
      name_nodes <- xml_find_all(node, ".//ex:applicant-name", ns)
      name_langs <- xml_attr(name_nodes, "lang")
      en_name <- name_nodes[!is.na(name_langs) & name_langs == "en"]
      name_node <- if (length(en_name) > 0) en_name[[1]] else if (length(name_nodes) > 0) name_nodes[[1]] else NULL
      name <- if (!is.null(name_node)) str_squish(xml_text(name_node)) else NA_character_

      # Determine if company: presence of orgname vs first-name/last-name
      has_orgname   <- length(xml_find_all(name_node, ".//ex:orgname", ns)) > 0
      has_firstname <- length(xml_find_all(name_node, ".//ex:first-name", ns)) > 0

      # app-type attribute: "applicant" vs "applicant-inventor"
      app_type <- xml_attr(node, "app-type")

      # Heuristic: it's a company if orgname present, or if no first-name
      # and the name looks corporate (contains common suffixes)
      is_corp <- has_orgname || (!has_firstname && !is.na(name) &&
        str_detect(name, regex("\\b(inc|corp|ltd|llc|gmbh|ag|sa|sas|bv|nv|co|plc|pty|ab|oy|as|se|spa|srl|kk|univ)\\b",
                               ignore_case = TRUE)))

      list(name = name, app_type = app_type %||% NA_character_, is_company = is_corp)
    })

    # Deduplicate by name
    names_vec <- sapply(applicants, \(a) a$name)
    unique_idx <- !duplicated(names_vec) & !is.na(names_vec)
    applicants <- applicants[unique_idx]

    if (length(applicants) == 0) {
      return(tibble(
        applicant_names = NA_character_,
        applicant_types = NA_character_,
        is_company      = NA
      ))
    }

    tibble(
      applicant_names = paste(sapply(applicants, \(a) a$name), collapse = "; "),
      applicant_types = paste(sapply(applicants, \(a) a$app_type), collapse = "; "),
      is_company      = any(sapply(applicants, \(a) a$is_company))
    )
  }

  # Token manager: refreshes the access token when it's close to expiring
  # OPS tokens expire after ~20 minutes, so refresh every 15 minutes to be safe
  token_env <- new.env()
  token_env$token      <- access_token
  token_env$expires_at <- Sys.time() + 15 * 60  # 15 min from now

  refresh_token <- function() {
    message("    [TOKEN] Refreshing OPS access token...")
    resp <- request("https://ops.epo.org/3.2/auth/accesstoken") |>
      req_auth_basic(ops_key, ops_secret) |>
      req_body_form(grant_type = "client_credentials") |>
      req_perform()
    token_env$token      <- resp_body_json(resp)$access_token
    token_env$expires_at <- Sys.time() + 15 * 60
    message("    [TOKEN] Refreshed, valid until ", format(token_env$expires_at, "%H:%M:%S"))
  }

  get_token <- function() {
    if (Sys.time() > token_env$expires_at) refresh_token()
    token_env$token
  }

  # Throttle-aware HTTP caller: waits patiently until throttle clears
  ops_request <- function(url, max_retries = 15) {
    for (attempt in seq_len(max_retries + 1)) {
      resp <- request(url) |>
        req_headers(Authorization = paste("Bearer", get_token()),
                    Accept        = "application/xml") |>
        req_error(is_error = \(resp) FALSE) |>
        req_perform()

      status <- resp_status(resp)

      throttle_hdr <- resp_header(resp, "X-Throttling-Control")
      if (!is.null(throttle_hdr) && grepl("overloaded", throttle_hdr, ignore.case = TRUE)) {
        message("    [THROTTLED] waiting 60s (attempt ", attempt, "/", max_retries, ")...")
        Sys.sleep(60)
        next
      }
      # 401/403 = possibly expired token → force refresh and retry
      if (status %in% c(401, 403)) {
        message("    [", status, "] possibly expired token, refreshing...")
        refresh_token()
        Sys.sleep(5)
        next
      }
      if (status %in% c(429, 500, 503)) {
        message("    [", status, "] waiting 60s (attempt ", attempt, "/", max_retries, ")...")
        Sys.sleep(60)
        next
      }

      if (status >= 400) return(NULL)
      return(resp_body_xml(resp))
    }
    return(NULL)
  }

  fetch_patent <- function(raw_id, country, docnumber) {
    abstract       <- NA_character_
    title          <- NA_character_
    applicant_info <- tibble(applicant_names = NA_character_,
                             applicant_types = NA_character_,
                             is_company = NA)

    # Use OPS SEARCH API (same as Espacenet web search — accepts ap= syntax)
    search_url <- paste0(
      "https://ops.epo.org/3.2/rest-services/published-data/search/biblio,abstract?q=ap%3D",
      raw_id
    )

    doc <- tryCatch(ops_request(search_url), error = function(e) NULL)

    if (!is.null(doc)) {
      abstract <- extract_en_abstract(doc)
      title    <- extract_en_title(doc)
      applicant_info <- extract_applicants(doc)
    }

    src <- if (!is.na(abstract) || !is.na(title)) "OK" else "miss"
    message("  [", src, "] ", raw_id,
            if (!is.na(title)) paste0(" - ", str_trunc(title, 50)) else "",
            if (!is.na(applicant_info$applicant_names))
              paste0(" | ", str_trunc(applicant_info$applicant_names, 40)) else "")

    bind_cols(
      tibble(title = title, abstract = abstract),
      applicant_info
    )
  }

  # Incremental fetching with checkpoint saves every 100 patents
  partial_file <- paste0(cache_file, ".partial.rds")

  # Resume from partial if available
  if (file.exists(partial_file)) {
    done <- readRDS(partial_file)
    done_keys <- done$raw
    cat("Resuming from partial cache:", nrow(done), "already fetched\n")
  } else {
    done <- tibble()
    done_keys <- character(0)
  }

  todo <- parsed |> filter(!raw %in% done_keys)
  n_total <- nrow(parsed)
  n_done  <- nrow(done)
  cat("Fetching", nrow(todo), "remaining patents (", n_done, "already done)...\n")

  for (i in seq_len(nrow(todo))) {
    row <- todo[i, ]
    Sys.sleep(2)  # 2s between patents to stay under per-minute limit

    if ((n_done + i) %% 50 == 0)
      message("--- Progress: ", n_done + i, " / ", n_total, " ---")

    fetched <- tryCatch(
      fetch_patent(row$raw, row$country, row$docnumber),
      error = function(e) {
        message("  [ERROR] ", row$raw, ": ", e$message)
        bind_cols(
          tibble(title = NA_character_, abstract = NA_character_),
          tibble(applicant_names = NA_character_, applicant_types = NA_character_, is_company = NA)
        )
      }
    )

    new_row <- bind_cols(row, fetched)
    done <- bind_rows(done, new_row)

    # Save checkpoint every 100 patents
    if (i %% 100 == 0) {
      saveRDS(done, partial_file)
      message("  [CHECKPOINT] Saved ", nrow(done), " / ", n_total)
    }
  }

  results <- done

  cat("\nFetched", sum(!is.na(results$abstract)), "abstracts,",
      sum(!is.na(results$title)), "titles, and",
      sum(!is.na(results$applicant_names)), "applicant records\n")
  cat("  of which", sum(results$is_company, na.rm = TRUE), "have a company assignee\n")

  saveRDS(results, cache_file)
  if (file.exists(partial_file)) file.remove(partial_file)
  cat("Cached results to", cache_file, "\n")
}
```

## 5 — Create embeddings

```{r create_embeddings, message=TRUE}
openai_key <- Sys.getenv("OPENAI_API_KEY", unset = "")
stopifnot("Set OPENAI_API_KEY in your .Renviron" = nchar(openai_key) > 0)

embed_data <- results |>
  filter(!is.na(abstract) | !is.na(title)) |>
  mutate(
    text = paste(
      if_else(!is.na(title), title, ""),
      if_else(!is.na(abstract), abstract, "")
    ) |> str_squish()
  ) |>
  filter(nchar(text) > 10)

cat("Creating embeddings for", nrow(embed_data), "patents...\n")

get_embeddings <- function(texts, api_key) {
  resp <- request("https://api.openai.com/v1/embeddings") |>
    req_headers(Authorization = paste("Bearer", api_key),
                `Content-Type` = "application/json") |>
    req_body_json(list(
      model = "text-embedding-3-small",
      input = texts
    )) |>
    req_perform()

  result <- resp_body_json(resp)
  embeddings <- lapply(result$data, function(x) unlist(x$embedding))
  do.call(rbind, embeddings)
}

embedding_matrix <- get_embeddings(embed_data$text, openai_key)
cat("Embedding dimensions:", dim(embedding_matrix), "\n")
```

## 6 — Cluster embeddings

```{r cluster_embeddings}
set.seed(42)
k <- min(10, nrow(embedding_matrix))
clusters <- kmeans(embedding_matrix, centers = k, nstart = 25)

embed_data$cluster <- as.factor(clusters$cluster)
cat("Patents per cluster:\n")
table(embed_data$cluster)
```

## 7 — AI-generated cluster labels

```{r cluster_labels, message=TRUE}
cluster_summaries <- embed_data |>
  group_by(cluster) |>
  summarise(
    patents = paste(
      if_else(!is.na(title), title, str_trunc(text, 100)),
      collapse = "\n- "
    ),
    n = n(),
    .groups = "drop"
  ) |>
  mutate(entry = paste0("Cluster ", cluster, " (", n, " patents):\n- ", patents))

label_prompt <- paste0(
  "Below are clusters of patents grouped by semantic similarity.\n\n",
  paste(cluster_summaries$entry, collapse = "\n\n"),
  "\n\n---\n",
  "For EACH cluster, provide a short descriptive label (2-5 words) that captures ",
  "the common technological theme. Be specific and technical.\n\n",
  "Return ONLY a JSON array of objects with keys \"cluster\" (the number) ",
  "and \"label\" (the short description). No other text."
)

label_resp <- request("https://api.openai.com/v1/chat/completions") |>
  req_headers(Authorization = paste("Bearer", openai_key),
              `Content-Type` = "application/json") |>
  req_body_json(list(
    model    = "gpt-4o-mini",
    messages = list(list(role = "user", content = label_prompt)),
    temperature = 0.2,
    max_tokens  = 500
  )) |>
  req_perform()

raw_json <- resp_body_json(label_resp)$choices[[1]]$message$content
raw_json <- str_remove_all(raw_json, "```json\\s*|```\\s*")

cluster_labels <- jsonlite::fromJSON(raw_json) |>
  as_tibble() |>
  mutate(cluster = as.character(cluster))

label_map <- setNames(cluster_labels$label, cluster_labels$cluster)

embed_data <- embed_data |>
  mutate(cluster_label = factor(label_map[as.character(cluster)], levels = label_map))

cat("Cluster labels:\n")
cluster_labels |> print(n = 20)
```

## 8 — UMAP projection and visualization

```{r umap_viz, fig.width=12, fig.height=10, dpi=150}
umap_config <- umap.defaults
umap_config$n_neighbors <- min(15, nrow(embedding_matrix) - 1)
umap_config$min_dist <- 0.1

umap_result <- umap(embedding_matrix, config = umap_config)

embed_data$umap_x <- umap_result$layout[, 1]
embed_data$umap_y <- umap_result$layout[, 2]

embed_data <- embed_data |>
  mutate(plot_label = if_else(!is.na(title), str_trunc(title, 30), raw))

# High-contrast colours for clusters
cluster_colors <- c(
 "#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00",
  "#a65628", "#f781bf", "#1b9e77", "#d95f02", "#7570b3"
)

# Create faded versions for HICs (add transparency via alpha)
# We'll use alpha in the plot instead

# Separate data for layering
lmic_data <- embed_data |> filter(source == "LMIC")
hic_data <- embed_data |> filter(source == "HIC")

# Plot with LMICs in full color and HICs faded with italic labels
ggplot() +
  # HICs: faded points, italic labels
  geom_point(data = hic_data,
             aes(x = umap_x, y = umap_y, colour = cluster_label),
             size = 4, alpha = 0.35, shape = 16) +
  geom_text_repel(data = hic_data,
                  aes(x = umap_x, y = umap_y, label = plot_label, colour = cluster_label),
                  size = 2.2, fontface = "italic", alpha = 0.6,
                  max.overlaps = 15, show.legend = FALSE) +
  # LMICs: full color points, normal labels
  geom_point(data = lmic_data,
             aes(x = umap_x, y = umap_y, colour = cluster_label),
             size = 4, alpha = 0.9, shape = 16) +
  geom_text_repel(data = lmic_data,
                  aes(x = umap_x, y = umap_y, label = plot_label, colour = cluster_label),
                  size = 2.2, fontface = "plain", alpha = 0.9,
                  max.overlaps = 15, show.legend = FALSE) +
  scale_colour_manual(values = cluster_colors) +
  theme_minimal() +
  labs(
    title = "Patent Embedding Clusters — LMICs vs HICs",
    subtitle = paste0(nrow(lmic_data), " LMIC patents (solid) and ",
                      nrow(hic_data), " HIC patents (faded/italic) in ", k, " clusters"),
    x = "UMAP 1",
    y = "UMAP 2",
    colour = "Theme"
  ) +
  theme(
    legend.position = "right",
    legend.text = element_text(size = 9)
  )
```

## 9 — Summary statistics

```{r summary_stats}
src_summary <- embed_data |>
  count(source, cluster_label) |>
  tidyr::pivot_wider(names_from = source, values_from = n, values_fill = 0)
# Ensure both columns exist even if one group has no patents
if (!"LMIC" %in% names(src_summary)) src_summary$LMIC <- 0L
if (!"HIC" %in% names(src_summary)) src_summary$HIC <- 0L
src_summary |>
  mutate(Total = LMIC + HIC) |>
  arrange(desc(Total))
```

## 10 — Simple titles and word cloud

Generate plain-language 3-word titles for each patent that a non-expert can understand.
Similar patents can share the same title.

```{r simple_titles, message=TRUE}
# Build prompt with all patents
patent_entries <- embed_data |>
  mutate(entry = paste0(
    "[", row_number(), "] ",
    if_else(!is.na(title), paste0("Title: ", title, ". "), ""),
    if_else(!is.na(abstract), paste0("Abstract: ", str_trunc(abstract, 300)), "")
  )) |>
  pull(entry) |>
  paste(collapse = "\n\n")

title_prompt <- paste0(
  "Below are patents with their titles and/or abstracts.\n\n",
  patent_entries,
  "\n\n---\n",
  "For EACH patent, create a simple 3-word label (maximum) that a non-expert ",
  "can understand. Use plain language, not technical jargon. ",
  "Examples: 'solar panel cleaning', 'cancer drug delivery', 'water purification system'.\n\n",
  "Similar patents CAN and SHOULD have identical labels if they cover the same topic. ",
  "Do NOT try to make labels unique — group similar innovations together.\n\n",
  "Return ONLY a JSON array of objects with keys \"id\" (the bracket number) ",
  "and \"simple_title\" (the 3-word label, lowercase). No other text."
)

title_resp <- request("https://api.openai.com/v1/chat/completions") |>
  req_headers(Authorization = paste("Bearer", openai_key),
              `Content-Type` = "application/json") |>
  req_body_json(list(
    model    = "gpt-4o-mini",
    messages = list(list(role = "user", content = title_prompt)),
    temperature = 0.3,
    max_tokens  = 3000
  )) |>
  req_perform()

raw_json <- resp_body_json(title_resp)$choices[[1]]$message$content
raw_json <- str_remove_all(raw_json, "```json\\s*|```\\s*")

simple_titles_df <- jsonlite::fromJSON(raw_json) |>
  as_tibble() |>
  mutate(id = as.integer(id),
         simple_title = str_to_lower(str_squish(simple_title)))

# Attach to embed_data
embed_data <- embed_data |>
  mutate(id = row_number()) |>
  left_join(simple_titles_df, by = "id")

cat("Generated", sum(!is.na(embed_data$simple_title)), "simple titles\n")
embed_data |> select(raw, simple_title) |> print(n = 30)
```

```{r simple_title_wordcloud, fig.width=12, fig.height=10, dpi=150}
library(ggwordcloud)

# Count frequencies of simple titles
title_counts <- embed_data |>
  filter(!is.na(simple_title)) |>
  count(simple_title, sort = TRUE)

cat("Top simple titles:\n")
print(title_counts, n = 25)

# Word cloud
ggplot(title_counts, aes(label = simple_title, size = n, colour = n)) +
geom_text_wordcloud_area(rm_outside = TRUE, shape = "circle",
                           eccentricity = 0.65) +
  scale_size_area(max_size = 18) +
  scale_colour_gradient(low = "#2c7bb6", high = "#d7191c") +
  theme_minimal() +
  labs(
    title = "Patent Innovation Themes — Simple Titles",
    subtitle = paste(nrow(title_counts), "distinct themes from", nrow(embed_data), "patents")
  )
```

```{r simple_title_bar, fig.width=10, fig.height=8}
# Bar chart of top titles
title_counts |>
  slice_max(n, n = 25, with_ties = FALSE) |>
  mutate(simple_title = reorder(simple_title, n)) |>
  ggplot(aes(n, simple_title, fill = n)) +
  geom_col(show.legend = FALSE) +
  scale_fill_gradient(low = "#2c7bb6", high = "#d7191c") +
  labs(x = "Number of Patents", y = NULL,
       title = "Top Innovation Themes (Simple Titles)") +
  theme_minimal()
```

```{r simple_title_by_source, fig.width=12, fig.height=8}
# Compare LMIC vs HIC distribution of simple titles
title_source_wide <- embed_data |>
  filter(!is.na(simple_title)) |>
  count(simple_title, source) |>
  tidyr::pivot_wider(names_from = source, values_from = n, values_fill = 0)
if (!"LMIC" %in% names(title_source_wide)) title_source_wide$LMIC <- 0L
if (!"HIC" %in% names(title_source_wide)) title_source_wide$HIC <- 0L
title_source_counts <- title_source_wide |>
  mutate(Total = LMIC + HIC) |>
  slice_max(Total, n = 20, with_ties = FALSE) |>
  tidyr::pivot_longer(cols = c(LMIC, HIC), names_to = "source", values_to = "n")

ggplot(title_source_counts, aes(x = n, y = reorder(simple_title, Total),
                                 fill = source)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("LMIC" = "#e41a1c", "HIC" = "#377eb8")) +
  labs(x = "Number of Patents", y = NULL,
       title = "Top Innovation Themes by Source",
       fill = "Source") +
  theme_minimal() +
  theme(legend.position = "top")
```
