---
title: "Patent Cluster Comparison — LMICs vs HICs"
output:
  html_document:
    df_print: paged
---

This notebook compares two sets of patents (LMICs vs HICs) for each technology
in a configurable list, by embedding their titles and abstracts, clustering
them together, and visualizing with distinct styling for each group.

To analyse a new technology, just add it to `tech_selections` in Section 2.

## 0 — Setup

```{r setup, message=FALSE, warning=FALSE}
pkgs <- c("httr2", "xml2", "dplyr", "tidyr", "tidytext", "stringr",
          "ggplot2", "ggwordcloud", "purrr", "tibble", "digest", "umap",
          "jsonlite", "ggrepel", "DBI", "duckdb", "glue")
install.packages(setdiff(pkgs, rownames(installed.packages())), quiet = TRUE)

library(httr2)
library(xml2)
library(dplyr)
library(tidyr)
library(tidytext)
library(stringr)
library(ggplot2)
library(ggwordcloud)
library(purrr)
library(tibble)
library(umap)
library(jsonlite)
library(ggrepel)
library(DBI)
library(duckdb)
library(glue)
```

## 1 — Helper functions

All reusable machinery lives here. Section 3 defines the pipeline function
that calls these; Section 4 runs the pipeline for each technology.

```{r helpers}
# ---- DuckDB connection & patent query ----------------------------------------

open_patent_db <- function(parquet_dir = file.path("inst", "extdata")) {
  con <- DBI::dbConnect(duckdb::duckdb())
  DBI::dbExecute(con, glue::glue(
    "CREATE OR REPLACE VIEW full_patent_database AS
     SELECT * FROM read_parquet('{file.path(parquet_dir, \"patent_database.parquet\")}')"
  ))
  DBI::dbExecute(con, glue::glue(
    "CREATE OR REPLACE TABLE patents_x_tech AS
     SELECT * FROM read_parquet('{file.path(parquet_dir, \"patents_x_tech.parquet\")}')"
  ))
  DBI::dbExecute(con, glue::glue(
    "CREATE OR REPLACE TABLE tech_lookup AS
     SELECT * FROM read_parquet('{file.path(parquet_dir, \"tech_lookup.parquet\")}')"
  ))
  con
}

get_top_patent_ids <- function(con, country_group, tech, toflow, top_n,
                               group_definitions) {
  country_codes <- if (country_group %in% names(group_definitions)) {
    group_definitions[[country_group]]
  } else {
    country_group
  }
  country_sql <- paste0("'", paste(country_codes, collapse = "', '"), "'")

  sql <- glue::glue("
    WITH filtered_tech AS (
      SELECT DISTINCT t.docdb_family_id, '{tech}' AS technology
      FROM patents_x_tech t
      JOIN tech_lookup tl ON t.technology = tl.technology
      WHERE tl.tech_group = '{tech}' OR t.technology = '{tech}'
    ),

    deduped AS (
      SELECT DISTINCT ON (ft.technology, p.docdb_family_id)
        ft.technology, p.docdb_family_id, p.appln_id, p.{toflow}
      FROM full_patent_database p
      INNER JOIN filtered_tech ft ON p.docdb_family_id = ft.docdb_family_id
      WHERE p.ctry_code IN ({country_sql})
        AND p.{toflow} IS NOT NULL
    ),

    ranked AS (
      SELECT appln_id, {toflow},
        ROW_NUMBER() OVER (ORDER BY {toflow} DESC) AS rnk
      FROM deduped
    )

    SELECT appln_id FROM ranked
    WHERE rnk <= {top_n}
    ORDER BY {toflow} DESC
  ")
  DBI::dbGetQuery(con, sql)$appln_id
}

parse_app_number <- function(x) {
  m <- str_match(x, "^([A-Z]{2,4})(\\d.*)$")
  tibble(raw = x, country = m[, 2], docnumber = m[, 3])
}

# ---- EPO OPS: XML parsing -----------------------------------------------------

ns_epo <- c(ex = "http://www.epo.org/exchange", ops = "http://ops.epo.org")

extract_en_abstract <- function(doc) {
  abs_nodes <- xml_find_all(doc, ".//ex:abstract", ns_epo)
  if (length(abs_nodes) == 0) return(NA_character_)
  langs <- xml_attr(abs_nodes, "lang")
  en <- abs_nodes[!is.na(langs) & langs == "en"]
  node <- if (length(en) > 0) en[[1]] else abs_nodes[[1]]

  p_nodes <- xml_find_all(node, ".//ex:p", ns_epo)
  text <- if (length(p_nodes) > 0) paste(xml_text(p_nodes), collapse = " ") else xml_text(node)
  text <- str_squish(text)
  if (nchar(text) < 10) NA_character_ else text
}

extract_en_title <- function(doc) {
  title_nodes <- xml_find_all(doc, ".//ex:invention-title", ns_epo)
  if (length(title_nodes) == 0) return(NA_character_)
  langs <- xml_attr(title_nodes, "lang")
  en <- title_nodes[!is.na(langs) & langs == "en"]
  node <- if (length(en) > 0) en[[1]] else title_nodes[[1]]
  str_squish(xml_text(node))
}

extract_applicants <- function(doc) {
  app_nodes <- xml_find_all(doc, ".//ex:applicant[@data-format='docdb']", ns_epo)
  if (length(app_nodes) == 0) {
    app_nodes <- xml_find_all(doc, ".//ex:applicant", ns_epo)
  }
  empty <- tibble(applicant_names = NA_character_,
                  applicant_types = NA_character_,
                  is_company = NA)
  if (length(app_nodes) == 0) return(empty)

  applicants <- lapply(app_nodes, function(node) {
    name_nodes <- xml_find_all(node, ".//ex:applicant-name", ns_epo)
    name_langs <- xml_attr(name_nodes, "lang")
    en_name <- name_nodes[!is.na(name_langs) & name_langs == "en"]
    name_node <- if (length(en_name) > 0) en_name[[1]]
                 else if (length(name_nodes) > 0) name_nodes[[1]]
                 else NULL
    name <- if (!is.null(name_node)) str_squish(xml_text(name_node)) else NA_character_

    has_orgname   <- length(xml_find_all(name_node, ".//ex:orgname", ns_epo)) > 0
    has_firstname <- length(xml_find_all(name_node, ".//ex:first-name", ns_epo)) > 0
    app_type <- xml_attr(node, "app-type")

    is_corp <- has_orgname || (!has_firstname && !is.na(name) &&
      str_detect(name, regex("\\b(inc|corp|ltd|llc|gmbh|ag|sa|sas|bv|nv|co|plc|pty|ab|oy|as|se|spa|srl|kk|univ)\\b",
                             ignore_case = TRUE)))

    list(name = name, app_type = app_type %||% NA_character_, is_company = is_corp)
  })

  names_vec <- sapply(applicants, \(a) a$name)
  unique_idx <- !duplicated(names_vec) & !is.na(names_vec)
  applicants <- applicants[unique_idx]
  if (length(applicants) == 0) return(empty)

  tibble(
    applicant_names = paste(sapply(applicants, \(a) a$name), collapse = "; "),
    applicant_types = paste(sapply(applicants, \(a) a$app_type), collapse = "; "),
    is_company      = any(sapply(applicants, \(a) a$is_company))
  )
}

# ---- EPO OPS: token management & HTTP with retries ---------------------------

make_token_manager <- function(ops_key, ops_secret, initial_token) {
  env <- new.env()
  env$token      <- initial_token
  env$expires_at <- Sys.time() + 15 * 60

  env$refresh <- function() {
    message("    [TOKEN] Refreshing OPS access token...")
    resp <- request("https://ops.epo.org/3.2/auth/accesstoken") |>
      req_auth_basic(ops_key, ops_secret) |>
      req_body_form(grant_type = "client_credentials") |>
      req_perform()
    env$token      <- resp_body_json(resp)$access_token
    env$expires_at <- Sys.time() + 15 * 60
    message("    [TOKEN] Refreshed, valid until ",
            format(env$expires_at, "%H:%M:%S"))
  }

  env$get <- function() {
    if (Sys.time() > env$expires_at) env$refresh()
    env$token
  }
  env
}

make_ops_request <- function(token_mgr, max_retries = 15) {
  function(url) {
    for (attempt in seq_len(max_retries + 1)) {
      resp <- request(url) |>
        req_headers(Authorization = paste("Bearer", token_mgr$get()),
                    Accept        = "application/xml") |>
        req_error(is_error = \(resp) FALSE) |>
        req_perform()

      status <- resp_status(resp)
      throttle_hdr <- resp_header(resp, "X-Throttling-Control")

      if (!is.null(throttle_hdr) && grepl("overloaded", throttle_hdr, ignore.case = TRUE)) {
        message("    [THROTTLED] waiting 60s (attempt ", attempt, "/", max_retries, ")...")
        Sys.sleep(60); next
      }
      if (status %in% c(401, 403)) {
        message("    [", status, "] possibly expired token, refreshing...")
        token_mgr$refresh(); Sys.sleep(5); next
      }
      if (status %in% c(429, 500, 503)) {
        message("    [", status, "] waiting 60s (attempt ", attempt, "/", max_retries, ")...")
        Sys.sleep(60); next
      }
      if (status >= 400) return(NULL)
      return(resp_body_xml(resp))
    }
    NULL
  }
}

fetch_one_patent <- function(raw_id, ops_request) {
  abstract <- NA_character_
  title    <- NA_character_
  applicant_info <- tibble(applicant_names = NA_character_,
                           applicant_types = NA_character_,
                           is_company = NA)

  search_url <- paste0(
    "https://ops.epo.org/3.2/rest-services/published-data/search/biblio,abstract?q=ap%3D",
    raw_id
  )
  doc <- tryCatch(ops_request(search_url), error = function(e) NULL)

  if (!is.null(doc)) {
    abstract       <- extract_en_abstract(doc)
    title          <- extract_en_title(doc)
    applicant_info <- extract_applicants(doc)
  }

  src <- if (!is.na(abstract) || !is.na(title)) "OK" else "miss"
  message("  [", src, "] ", raw_id,
          if (!is.na(title)) paste0(" - ", str_trunc(title, 50)) else "",
          if (!is.na(applicant_info$applicant_names))
            paste0(" | ", str_trunc(applicant_info$applicant_names, 40)) else "")

  bind_cols(tibble(title = title, abstract = abstract), applicant_info)
}

fetch_patents <- function(parsed, cache_file, token_mgr) {
  if (file.exists(cache_file)) {
    cat("Loading cached EPO data from", cache_file, "\n")
    results <- readRDS(cache_file)
    cat("Loaded", sum(!is.na(results$abstract)), "abstracts from cache\n")
    return(results)
  }

  cat("No cache found, fetching from EPO API...\n")
  ops_request <- make_ops_request(token_mgr)
  partial_file <- paste0(cache_file, ".partial.rds")

  if (file.exists(partial_file)) {
    done <- readRDS(partial_file)
    done_keys <- done$raw
    cat("Resuming from partial cache:", nrow(done), "already fetched\n")
  } else {
    done <- tibble(); done_keys <- character(0)
  }

  todo <- parsed |> filter(!raw %in% done_keys)
  n_total <- nrow(parsed); n_done <- nrow(done)
  cat("Fetching", nrow(todo), "remaining patents (", n_done, "already done)...\n")

  for (i in seq_len(nrow(todo))) {
    row <- todo[i, ]
    Sys.sleep(2)
    if ((n_done + i) %% 50 == 0)
      message("--- Progress: ", n_done + i, " / ", n_total, " ---")

    fetched <- tryCatch(
      fetch_one_patent(row$raw, ops_request),
      error = function(e) {
        message("  [ERROR] ", row$raw, ": ", e$message)
        bind_cols(
          tibble(title = NA_character_, abstract = NA_character_),
          tibble(applicant_names = NA_character_,
                 applicant_types = NA_character_, is_company = NA)
        )
      }
    )
    done <- bind_rows(done, bind_cols(row, fetched))
    if (i %% 100 == 0) {
      saveRDS(done, partial_file)
      message("  [CHECKPOINT] Saved ", nrow(done), " / ", n_total)
    }
  }

  results <- done
  cat("\nFetched", sum(!is.na(results$abstract)), "abstracts,",
      sum(!is.na(results$title)), "titles, and",
      sum(!is.na(results$applicant_names)), "applicant records\n")
  cat("  of which", sum(results$is_company, na.rm = TRUE), "have a company assignee\n")

  saveRDS(results, cache_file)
  if (file.exists(partial_file)) file.remove(partial_file)
  results
}

# ---- OpenAI: embeddings & chat ----------------------------------------------

get_embeddings <- function(texts, api_key) {
  resp <- request("https://api.openai.com/v1/embeddings") |>
    req_headers(Authorization = paste("Bearer", api_key),
                `Content-Type` = "application/json") |>
    req_body_json(list(model = "text-embedding-3-small", input = texts)) |>
    req_perform()
  result <- resp_body_json(resp)
  do.call(rbind, lapply(result$data, function(x) unlist(x$embedding)))
}

chat_completion <- function(prompt, api_key, model = "gpt-4o-mini",
                            temperature = 0.2, max_tokens = 500) {
  resp <- request("https://api.openai.com/v1/chat/completions") |>
    req_headers(Authorization = paste("Bearer", api_key),
                `Content-Type` = "application/json") |>
    req_body_json(list(
      model = model,
      messages = list(list(role = "user", content = prompt)),
      temperature = temperature,
      max_tokens = max_tokens
    )) |>
    req_perform()
  raw <- resp_body_json(resp)$choices[[1]]$message$content
  str_remove_all(raw, "```json\\s*|```\\s*")
}

label_clusters <- function(embed_data, openai_key) {
  cluster_summaries <- embed_data |>
    group_by(cluster) |>
    summarise(
      patents = paste(
        if_else(!is.na(title), title, str_trunc(text, 100)),
        collapse = "\n- "),
      n = n(), .groups = "drop"
    ) |>
    mutate(entry = paste0("Cluster ", cluster, " (", n, " patents):\n- ", patents))

  prompt <- paste0(
    "Below are clusters of patents grouped by semantic similarity.\n\n",
    paste(cluster_summaries$entry, collapse = "\n\n"),
    "\n\n---\n",
    "For EACH cluster, provide a short descriptive label (2-5 words) that captures ",
    "the common technological theme. Be specific and technical.\n\n",
    "Return ONLY a JSON array of objects with keys \"cluster\" (the number) ",
    "and \"label\" (the short description). No other text."
  )

  raw_json <- chat_completion(prompt, openai_key, max_tokens = 500)
  jsonlite::fromJSON(raw_json) |>
    as_tibble() |>
    mutate(cluster = as.character(cluster))
}

generate_simple_titles <- function(embed_data, openai_key, batch_size = 100) {
  prep <- embed_data |>
    mutate(global_id = row_number()) |>
    mutate(entry_text = paste0(
      if_else(!is.na(title), paste0("Title: ", title, ". "), ""),
      if_else(!is.na(abstract), paste0("Abstract: ", str_trunc(abstract, 300)), "")
    ))

  n <- nrow(prep)
  batches <- split(prep, ceiling(seq_len(n) / batch_size))

  results <- lapply(seq_along(batches), function(bi) {
    batch <- batches[[bi]]
    # Use local indices [1..batch_size] in the prompt, map back afterwards
    entries <- paste0("[", seq_len(nrow(batch)), "] ", batch$entry_text) |>
      paste(collapse = "\n\n")

    prompt <- paste0(
      "Below are patents with their titles and/or abstracts.\n\n",
      entries, "\n\n---\n",
      "For EACH patent, create a simple 3-word label (maximum) that a non-expert ",
      "can understand. Use plain language, not technical jargon. ",
      "Examples: 'solar panel cleaning', 'cancer drug delivery', 'water purification system'.\n\n",
      "Similar patents CAN and SHOULD have identical labels if they cover the same topic. ",
      "Do NOT try to make labels unique — group similar innovations together.\n\n",
      "Return ONLY a JSON array of objects with keys \"id\" (the bracket number) ",
      "and \"simple_title\" (the 3-word label, lowercase). No other text."
    )

    message("    [titles] batch ", bi, "/", length(batches),
            " (", nrow(batch), " patents)")

    raw_json <- chat_completion(prompt, openai_key,
                                temperature = 0.3, max_tokens = 4000)

    parsed <- tryCatch(
      jsonlite::fromJSON(raw_json) |> as_tibble(),
      error = function(e) {
        warning("Failed to parse batch ", bi, ": ", e$message)
        tibble(id = integer(), simple_title = character())
      }
    )
    if (nrow(parsed) == 0) return(parsed)

    parsed |>
      mutate(
        id = as.integer(id),
        simple_title = str_to_lower(str_squish(simple_title)),
        # Map local batch id back to global row_number id
        id = batch$global_id[id]
      )
  })

  bind_rows(results)
}

# ---- Plotting functions ------------------------------------------------------

cluster_colors <- c(
  "#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00",
  "#a65628", "#f781bf", "#1b9e77", "#d95f02", "#7570b3"
)

make_umap_plot <- function(embed_data, k, tech_name) {
  lmic_data <- embed_data |> filter(source == "LMIC")
  hic_data  <- embed_data |> filter(source == "HIC")

  ggplot() +
    geom_point(data = hic_data,
               aes(x = umap_x, y = umap_y, colour = cluster_label),
               size = 4, alpha = 0.35, shape = 16) +
    geom_text_repel(data = hic_data,
                    aes(x = umap_x, y = umap_y, label = plot_label, colour = cluster_label),
                    size = 2.2, fontface = "italic", alpha = 0.6,
                    max.overlaps = 15, show.legend = FALSE) +
    geom_point(data = lmic_data,
               aes(x = umap_x, y = umap_y, colour = cluster_label),
               size = 4, alpha = 0.9, shape = 16) +
    geom_text_repel(data = lmic_data,
                    aes(x = umap_x, y = umap_y, label = plot_label, colour = cluster_label),
                    size = 2.2, fontface = "plain", alpha = 0.9,
                    max.overlaps = 15, show.legend = FALSE) +
    scale_colour_manual(values = cluster_colors) +
    theme_minimal() +
    labs(
      title = paste0("Patent Embedding Clusters — ", tech_name, " (LMICs vs HICs)"),
      subtitle = paste0(nrow(lmic_data), " LMIC patents (solid) and ",
                        nrow(hic_data), " HIC patents (faded/italic) in ", k, " clusters"),
      x = "UMAP 1", y = "UMAP 2", colour = "Theme"
    ) +
    theme(legend.position = "right", legend.text = element_text(size = 9))
}

make_wordcloud_plot <- function(title_counts, tech_name) {
  ggplot(title_counts, aes(label = simple_title, size = n, colour = n)) +
    geom_text_wordcloud_area(rm_outside = TRUE, shape = "circle", eccentricity = 0.65) +
    scale_size_area(max_size = 18) +
    scale_colour_gradient(low = "#2c7bb6", high = "#d7191c") +
    theme_minimal() +
    labs(
      title = paste0("Patent Innovation Themes — ", tech_name),
      subtitle = paste(nrow(title_counts), "distinct themes")
    )
}

make_top_titles_bar <- function(title_counts, tech_name) {
  title_counts |>
    slice_max(n, n = 25, with_ties = FALSE) |>
    mutate(simple_title = reorder(simple_title, n)) |>
    ggplot(aes(n, simple_title, fill = n)) +
    geom_col(show.legend = FALSE) +
    scale_fill_gradient(low = "#2c7bb6", high = "#d7191c") +
    labs(x = "Number of Patents", y = NULL,
         title = paste0("Top Innovation Themes — ", tech_name)) +
    theme_minimal()
}

make_source_comparison_bar <- function(embed_data, tech_name) {
  wide <- embed_data |>
    filter(!is.na(simple_title)) |>
    count(simple_title, source) |>
    tidyr::pivot_wider(names_from = source, values_from = n, values_fill = 0)
  if (!"LMIC" %in% names(wide)) wide$LMIC <- 0L
  if (!"HIC"  %in% names(wide)) wide$HIC  <- 0L

  long <- wide |>
    mutate(Total = LMIC + HIC) |>
    slice_max(Total, n = 20, with_ties = FALSE) |>
    tidyr::pivot_longer(cols = c(LMIC, HIC), names_to = "source", values_to = "n")

  ggplot(long, aes(x = n, y = reorder(simple_title, Total), fill = source)) +
    geom_col(position = "dodge") +
    scale_fill_manual(values = c("LMIC" = "#e41a1c", "HIC" = "#377eb8")) +
    labs(x = "Number of Patents", y = NULL,
         title = paste0("Top Innovation Themes by Source — ", tech_name),
         fill = "Source") +
    theme_minimal() +
    theme(legend.position = "top")
}
```

## 2 — Configuration

Add or remove technology names here to change what gets analysed. Each entry
produces its own full set of plots in Section 5.

```{r config}
tech_selections <- c( "Green Technology")

toflow         <- "ev_global"
top_n_ids      <- 200
country_group1 <- "LMICs (excl. China)"
country_group2 <- "High income countries"

# Load country group definitions from the package's sysdata
load("R/sysdata.rda")

# Open shared DuckDB connection
con <- open_patent_db()

# OPS authentication (done once; token_mgr will auto-refresh)
ops_key    <- Sys.getenv("OPS_CONSUMER_KEY",    unset = "YOUR_KEY_HERE")
ops_secret <- Sys.getenv("OPS_CONSUMER_SECRET", unset = "YOUR_SECRET_HERE")
stopifnot("Set your OPS credentials via env vars OPS_CONSUMER_KEY / OPS_CONSUMER_SECRET" =
            ops_key != "YOUR_KEY_HERE")

token_resp <- request("https://ops.epo.org/3.2/auth/accesstoken") |>
  req_auth_basic(ops_key, ops_secret) |>
  req_body_form(grant_type = "client_credentials") |>
  req_perform()
initial_token <- resp_body_json(token_resp)$access_token
token_mgr <- make_token_manager(ops_key, ops_secret, initial_token)

# OpenAI key
openai_key <- Sys.getenv("OPENAI_API_KEY", unset = "")
stopifnot("Set OPENAI_API_KEY in your .Renviron" = nchar(openai_key) > 0)

cache_dir <- ".bigdata"
if (!dir.exists(cache_dir)) dir.create(cache_dir, recursive = TRUE)
```

## 3 — Pipeline function

`analyze_tech()` runs the full pipeline for a single technology and returns a
list containing the processed data and the plots.

```{r pipeline}
analyze_tech <- function(tech_selection) {
  cat("\n========== ", tech_selection, " ==========\n")

  # --- 1. Query patent IDs ---
  apps_lmic <- get_top_patent_ids(con, country_group1, tech_selection,
                                  toflow, top_n_ids, group_definitions)
  apps_hic  <- get_top_patent_ids(con, country_group2, tech_selection,
                                  toflow, top_n_ids, group_definitions)
  cat("LMICs:", length(apps_lmic), "  HICs:", length(apps_hic), "\n")

  all_apps <- tibble(
    raw = c(apps_lmic, apps_hic),
    source = c(rep("LMIC", length(apps_lmic)), rep("HIC", length(apps_hic)))
  ) |>
    distinct(raw, .keep_all = TRUE)
  cat("Union (deduplicated):", nrow(all_apps), "\n")

  # --- 2. Parse & fetch abstracts ---
  parsed <- all_apps |>
    left_join(bind_rows(lapply(all_apps$raw, parse_app_number)), by = "raw")

  cache_file <- file.path(cache_dir, paste0(
    "epo_compare_", digest::digest(all_apps$raw, algo = "md5"), ".rds"
  ))
  results <- fetch_patents(parsed, cache_file, token_mgr)

  # --- 3. Build embedding dataset ---
  embed_data <- results |>
    filter(!is.na(abstract) | !is.na(title)) |>
    mutate(
      text = paste(
        if_else(!is.na(title), title, ""),
        if_else(!is.na(abstract), abstract, "")
      ) |> str_squish()
    ) |>
    filter(nchar(text) > 10)

  if (nrow(embed_data) < 3) {
    warning("Too few patents with text for ", tech_selection, " — skipping.")
    return(NULL)
  }

  cat("Creating embeddings for", nrow(embed_data), "patents...\n")
  embedding_matrix <- get_embeddings(embed_data$text, openai_key)

  # --- 4. Cluster ---
  set.seed(42)
  k <- min(10, nrow(embedding_matrix))
  clusters <- kmeans(embedding_matrix, centers = k, nstart = 25)
  embed_data$cluster <- as.factor(clusters$cluster)

  # --- 5. AI cluster labels ---
  cluster_labels <- label_clusters(embed_data, openai_key)
  label_map <- setNames(cluster_labels$label, cluster_labels$cluster)
  embed_data <- embed_data |>
    mutate(cluster_label = factor(label_map[as.character(cluster)], levels = label_map))

  # --- 6. UMAP projection ---
  umap_config <- umap.defaults
  umap_config$n_neighbors <- min(15, nrow(embedding_matrix) - 1)
  umap_config$min_dist <- 0.1
  umap_result <- umap(embedding_matrix, config = umap_config)
  embed_data$umap_x <- umap_result$layout[, 1]
  embed_data$umap_y <- umap_result$layout[, 2]
  embed_data <- embed_data |>
    mutate(plot_label = if_else(!is.na(title), str_trunc(title, 30), raw))

  # --- 7. Simple titles ---
  simple_titles_df <- generate_simple_titles(embed_data, openai_key)
  embed_data <- embed_data |>
    mutate(id = row_number()) |>
    left_join(simple_titles_df, by = "id")

  # --- 8. Build plots ---
  title_counts <- embed_data |>
    filter(!is.na(simple_title)) |>
    count(simple_title, sort = TRUE)

  src_summary_wide <- embed_data |>
    count(source, cluster_label) |>
    tidyr::pivot_wider(names_from = source, values_from = n, values_fill = 0)
  if (!"LMIC" %in% names(src_summary_wide)) src_summary_wide$LMIC <- 0L
  if (!"HIC"  %in% names(src_summary_wide)) src_summary_wide$HIC  <- 0L
  src_summary <- src_summary_wide |>
    mutate(Total = LMIC + HIC) |>
    arrange(desc(Total))

  list(
    tech             = tech_selection,
    embed_data       = embed_data,
    cluster_labels   = cluster_labels,
    k                = k,
    title_counts     = title_counts,
    src_summary      = src_summary,
    umap_plot        = make_umap_plot(embed_data, k, tech_selection),
    wordcloud        = make_wordcloud_plot(title_counts, tech_selection),
    top_titles_bar   = make_top_titles_bar(title_counts, tech_selection),
    source_bar       = make_source_comparison_bar(embed_data, tech_selection)
  )
}
```

## 4 — Run pipeline for each technology

```{r run_pipelines, message=TRUE, warning=TRUE}
results_by_tech <- lapply(tech_selections, analyze_tech)
names(results_by_tech) <- tech_selections

# Drop any that were skipped (NULL)
results_by_tech <- results_by_tech[!sapply(results_by_tech, is.null)]
```

## 5 — Results per technology

```{r results_display, results='asis', fig.width=12, fig.height=10, dpi=150}
for (tech in names(results_by_tech)) {
  res <- results_by_tech[[tech]]

  cat("\n\n### ", tech, "\n\n")

  cat("**Cluster labels:**\n\n")
  print(knitr::kable(res$cluster_labels))
  cat("\n\n")

  cat("**Patents per cluster × source:**\n\n")
  print(knitr::kable(res$src_summary))
  cat("\n\n")

  print(res$umap_plot)
  print(res$wordcloud)
  print(res$top_titles_bar)
  print(res$source_bar)
}
```

## 6 — Clean up

```{r cleanup}
DBI::dbDisconnect(con, shutdown = TRUE)
```
