---
title: "Patent Abstracts — Bigram Word Cloud"
output:
  html_document:
    df_print: paged
---

This notebook fetches patent abstracts from the EPO Open Patent Services (OPS)
API for a set of application numbers sourced from an Espacenet search, then
draws a word cloud of the most frequent bigrams.

## 0 — Setup

```{r setup, message=FALSE, warning=FALSE}
# Install any missing packages
pkgs <- c("httr2", "xml2", "dplyr", "tidyr", "tidytext", "stringr",
           "ggplot2", "ggwordcloud", "purrr", "tibble", "digest")
install.packages(setdiff(pkgs, rownames(installed.packages())), quiet = TRUE)

library(httr2)
library(xml2)
library(dplyr)
library(tidyr)
library(tidytext)
library(stringr)
library(ggplot2)
library(ggwordcloud)
library(purrr)
library(tibble)
```

## 1 — Application numbers

Extracted from the Espacenet search URL:

```{r app_numbers}
app_numbers <- c(
 "CN201810134850", "PT12858932",      "US89428710",       "US201414220990",
 "ARP120101066",   "EP09305250",      "CA3117920",        "SG11201506775V",
 "EP11833120",     "KR20207024776",   "AU2011349446",     "US201213679498",
 "BR112013025572", "KR20127020023",   "CN201380028604",   "US59260509",
 "US201114008192", "US76340110",      "MX2014002289",     "WO2011039386",
 "WO2011021301",   "HUE16193906",     "WO2009007269",     "US201414293546",
 "US201113107352", "US201314425505",  "SG10201504187Y",   "US60981509",
 "US201615193033", "WO2010003130",    "PL13159003",       "BRPI1002602",
 "WO2015033929",   "BRPI1102284",     "AU2015314826",     "US86898910",
 "BR112017009050", "WO2011038946",    "KR20147016656",    "CN201080049139",
 "PL11858867",     "AU2013360086",    "ES11728509",       "WO2010000938",
 "US48126009",     "EP13762719",      "US201313920673",   "NZ61861310",
 "MX2011003856",   "ARP110104677"
)

cat(length(app_numbers), "application numbers to process\n")
```

## 2 — Helper: parse an application number into country + doc-number

The OPS API expects the EPODOC format `CC.number` (country code dot number).

```{r parse_appnum}
parse_app_number <- function(x) {
  # Match leading 2-letter country code (some have 4-letter like BRPI)
  m <- str_match(x, "^([A-Z]{2,4})(\\d.*)$")
  tibble(raw = x, country = m[, 2], docnumber = m[, 3])
}

parsed <- bind_rows(lapply(app_numbers, parse_app_number))
parsed
```

## 3 — Authenticate with EPO OPS

The OPS API requires **free** developer credentials.
Register at <https://developers.epo.org/> to get a Consumer Key and Consumer
Secret, then paste them below.

```{r ops_auth}
# -------------------------------------------------------------------
# PASTE YOUR OPS CREDENTIALS HERE (or set as environment variables)
# -------------------------------------------------------------------
ops_key    <- Sys.getenv("OPS_CONSUMER_KEY",    unset = "YOUR_KEY_HERE")
ops_secret <- Sys.getenv("OPS_CONSUMER_SECRET", unset = "YOUR_SECRET_HERE")

stopifnot("Set your OPS credentials above or via env vars OPS_CONSUMER_KEY / OPS_CONSUMER_SECRET" =
            ops_key != "YOUR_KEY_HERE")

# OAuth2 client-credentials token
token_resp <- request("https://ops.epo.org/3.2/auth/accesstoken") |>
  req_auth_basic(ops_key, ops_secret) |>
  req_body_form(grant_type = "client_credentials") |>
  req_perform()

access_token <- resp_body_json(token_resp)$access_token
cat("OPS access token obtained.\n")
```

## 4a — Debug: inspect raw XML for one patent

Let's see what OPS actually returns so we can write the right XPath.

```{r debug_xml, message=TRUE}
# Pick one that returned "no abstract" — e.g. EP.09305250
test_url <- paste0(
  "https://ops.epo.org/3.2/rest-services/published-data/application/epodoc/",
  "EP.09305250/abstract"
)
test_resp <- request(test_url) |>
  req_headers(Authorization = paste("Bearer", access_token),
              Accept        = "application/xml") |>
  req_error(is_error = \(resp) FALSE) |>
  req_perform()

cat("Status:", resp_status(test_resp), "\n\n")
if (resp_status(test_resp) < 400) {
  raw_xml <- resp_body_string(test_resp)
  cat(raw_xml)
}
```

## 4b — Fetch abstracts from OPS (with caching)

We query the *published-data* endpoint one patent at a time.
OPS anonymous throttle is ~4 req/s; we add a small delay to be polite.

Results are cached to `.bigdata/` to avoid repeated API calls.

```{r fetch_abstracts, message=TRUE, warning=TRUE}
# Create cache directory if it doesn't exist
cache_dir <- ".bigdata"
if (!dir.exists(cache_dir)) dir.create(cache_dir, recursive = TRUE)

# Cache file path based on hash of application numbers
cache_file <- file.path(cache_dir, paste0(
  "epo_abstracts_", digest::digest(app_numbers, algo = "md5"), ".rds"
))

# Check if cached data exists
if (file.exists(cache_file)) {
  cat("Loading cached EPO data from", cache_file, "\n")
  results <- readRDS(cache_file)
  abstracts <- results
  cat("Loaded", sum(!is.na(results$abstract)), "abstracts and",
      sum(!is.na(results$title)), "titles from cache\n")
} else {
  cat("No cache found, fetching from EPO API...\n")

ns <- c(ex = "http://www.epo.org/exchange", ops = "http://ops.epo.org")

extract_en_abstract <- function(doc) {
  abs_nodes <- xml_find_all(doc, ".//ex:abstract", ns)
  langs <- xml_attr(abs_nodes, "lang")
  en <- abs_nodes[!is.na(langs) & langs == "en"]
  node <- if (length(en) > 0) en[[1]] else NULL
  if (is.null(node)) return(NA_character_)

  p_nodes <- xml_find_all(node, ".//ex:p", ns)
  text <- if (length(p_nodes) > 0) paste(xml_text(p_nodes), collapse = " ") else xml_text(node)
  text <- str_squish(text)
  if (nchar(text) < 10) NA_character_ else text
}

extract_en_title <- function(doc) {
  title_nodes <- xml_find_all(doc, ".//ex:invention-title", ns)
  langs <- xml_attr(title_nodes, "lang")
  en <- title_nodes[!is.na(langs) & langs == "en"]
  node <- if (length(en) > 0) en[[1]] else NULL
  if (is.null(node)) return(NA_character_)
  str_squish(xml_text(node))
}

fetch_patent <- function(country, docnumber, token) {
  epodoc <- paste0(country, ".", docnumber)
  abstract <- NA_character_
  title    <- NA_character_

  # Helper to try one URL and extract abstract + title
  try_endpoint <- function(url) {
    resp <- request(url) |>
      req_headers(Authorization = paste("Bearer", token),
                  Accept        = "application/xml") |>
      req_error(is_error = \(resp) FALSE) |>
      req_perform()
    if (resp_status(resp) >= 400) return(NULL)
    resp_body_xml(resp)
  }

  # --- Step 1: direct application endpoint (biblio for title + abstract) ---
  tryCatch({
    doc <- try_endpoint(paste0(
      "https://ops.epo.org/3.2/rest-services/published-data/application/epodoc/",
      epodoc, "/biblio,abstract"
    ))
    if (!is.null(doc)) {
      abstract <- extract_en_abstract(doc)
      title    <- extract_en_title(doc)
    }
  }, error = function(e) NULL)

  # --- Step 2: family fallback if no English abstract ---
  if (is.na(abstract)) {
    tryCatch({
      Sys.sleep(0.3)
      doc <- try_endpoint(paste0(
        "https://ops.epo.org/3.2/rest-services/family/application/epodoc/",
        epodoc, "/biblio,abstract"
      ))
      if (!is.null(doc)) {
        abstract <- extract_en_abstract(doc)
        if (is.na(title)) title <- extract_en_title(doc)
      }
    }, error = function(e) NULL)
  }

  src <- if (!is.na(abstract)) "OK" else "miss"
  message("  [", src, "] ", epodoc,
          if (!is.na(title)) paste0(" — ", str_trunc(title, 60)) else "")

  tibble(title = title, abstract = abstract)
}

results <- parsed |>
  mutate(data = map2(country, docnumber, \(cc, dn) {
    Sys.sleep(0.3)
    fetch_patent(cc, dn, access_token)
  })) |>
  unnest(data)

cat("\nFetched", sum(!is.na(results$abstract)), "abstracts and",
    sum(!is.na(results$title)), "titles out of",
    nrow(results), "applications\n")

  # Keep backward-compatible name for downstream chunks
  abstracts <- results

  # Save to cache
  saveRDS(results, cache_file)
  cat("Cached results to", cache_file, "\n")
}
```

## 5 — AI-derived innovation labels

For each patent we ask OpenAI to produce a short n-gram (2–5 words) that
captures the core innovation, using both the title and the abstract.

```{r ai_labels, message=TRUE}
openai_key <- Sys.getenv("OPENAI_API_KEY", unset = "")
stopifnot("Set OPENAI_API_KEY in your .Renviron" = nchar(openai_key) > 0)

# Build a batch prompt — send all patents in one call for speed & cost
patent_block <- abstracts |>
  filter(!is.na(abstract) | !is.na(title)) |>
  mutate(entry = paste0(
    "[", row_number(), "] ",
    if_else(!is.na(title), paste0("Title: ", title, ". "), ""),
    if_else(!is.na(abstract), paste0("Abstract: ", str_trunc(abstract, 500)), "")
  )) |>
  pull(entry) |>
  paste(collapse = "\n\n")

label_prompt <- paste0(
 "Below are patents, each with a number in brackets, a title and/or abstract.\n\n",
  patent_block,
  "\n\n---\n",
  "For EACH patent, produce a short descriptive n-gram (2 to 5 words, lowercase) ",
  "that captures the core innovation or technology. ",
  "Do NOT use generic phrases like 'novel method' or 'improved system'. ",
  "Be specific and technical, e.g. 'antibody fc region engineering', ",
  "'lithium battery cathode recycling', 'crispr gene editing delivery'.\n\n",
  "Return ONLY a JSON array of objects with keys \"id\" (the bracket number) ",
  "and \"label\" (the n-gram). No other text."
)

label_resp <- request("https://api.openai.com/v1/chat/completions") |>
  req_headers(Authorization = paste("Bearer", openai_key),
              `Content-Type` = "application/json") |>
  req_body_json(list(
    model    = "gpt-4o-mini",
    messages = list(list(role = "user", content = label_prompt)),
    temperature = 0.2,
    max_tokens  = 2000
  )) |>
  req_perform()

raw_json <- resp_body_json(label_resp)$choices[[1]]$message$content

# Strip markdown code fences if present
raw_json <- str_remove_all(raw_json, "```json\\s*|```\\s*")

labels_df <- jsonlite::fromJSON(raw_json) |>
  as_tibble() |>
  mutate(id = as.integer(id),
         label = str_to_lower(str_squish(label)))

# Attach back to the abstracts that were sent
sent <- abstracts |>
  filter(!is.na(abstract) | !is.na(title)) |>
  mutate(id = row_number())

labelled <- sent |>
  left_join(labels_df, by = "id")

cat("Labelled", sum(!is.na(labelled$label)), "patents\n")
labelled |> select(raw, label) |> print(n = 60)
```

## 6 — AI-label word cloud

```{r ai_wordcloud, fig.width=10, fig.height=8, dpi=150}
label_counts <- labelled |>
  filter(!is.na(label)) |>
  count(label, sort = TRUE)

ggplot(label_counts, aes(label = label, size = n, colour = n)) +
  geom_text_wordcloud_area(rm_outside = TRUE, shape = "circle",
                           eccentricity = 0.65) +
  scale_size_area(max_size = 16) +
  scale_colour_gradient(low = "#2c7bb6", high = "#d7191c") +
  theme_minimal() +
  labs(title = "Innovation Word Cloud — AI-derived Labels",
       subtitle = paste(nrow(label_counts), "distinct innovation themes"))
```

## 7 — AI-label bar chart

```{r ai_bar_chart, fig.width=9, fig.height=7}
label_counts |>
  slice_max(n, n = 30, with_ties = FALSE) |>
  mutate(label = reorder(label, n)) |>
  ggplot(aes(n, label, fill = n)) +
  geom_col(show.legend = FALSE) +
  scale_fill_gradient(low = "#2c7bb6", high = "#d7191c") +
  labs(x = "Frequency", y = NULL,
       title = "Top Innovation Themes (AI-labelled)") +
  theme_minimal()
```

## 8 — Statistical bigrams (original)

For reference, the traditional bigram word cloud from raw abstract text.

```{r bigrams}
custom_stop <- tibble(word = c("patent", "invention", "present", "method",
                               "comprises", "comprising", "device", "apparatus",
                               "provided", "relates", "described", "including",
                               "wherein", "thereof", "according"))

bigrams <- abstracts |>
  filter(!is.na(abstract)) |>
  select(raw, abstract) |>
  unnest_tokens(bigram, abstract, token = "ngrams", n = 2) |>
  separate(bigram, into = c("word1", "word2"), sep = " ", remove = FALSE) |>
  anti_join(stop_words, by = c("word1" = "word")) |>
  anti_join(stop_words, by = c("word2" = "word")) |>
  anti_join(custom_stop, by = c("word1" = "word")) |>
  anti_join(custom_stop, by = c("word2" = "word")) |>
  filter(!str_detect(word1, "\\d"), !str_detect(word2, "\\d"))

bigram_counts <- bigrams |>
  count(bigram, sort = TRUE)

head(bigram_counts, 20)
```

```{r stat_wordcloud, fig.width=10, fig.height=8, dpi=150}
top_bigrams <- bigram_counts |>
  slice_max(n, n = 80, with_ties = FALSE)

ggplot(top_bigrams, aes(label = bigram, size = n, colour = n)) +
  geom_text_wordcloud_area(rm_outside = TRUE, shape = "circle",
                           eccentricity = 0.65) +
  scale_size_area(max_size = 14) +
  scale_colour_gradient(low = "#2c7bb6", high = "#d7191c") +
  theme_minimal() +
  labs(title = "Statistical Bigram Word Cloud — Patent Abstracts",
       subtitle = paste(sum(!is.na(abstracts$abstract)), "abstracts"))
```

## 8 — AI Summary of Key Innovations (OpenAI)

Uses the OpenAI API to produce a brief paragraph summarising the most important
innovations across all fetched abstracts.

```{r ai_summary, message=TRUE}
openai_key <- Sys.getenv("OPENAI_API_KEY", unset = "")
stopifnot("Set OPENAI_API_KEY in your .Renviron" = nchar(openai_key) > 0)

# Combine all abstracts into a single block, numbered
abstract_texts <- abstracts |>
  filter(!is.na(abstract)) |>
  mutate(entry = paste0("[", row_number(), "] ", raw, ": ", abstract)) |>
  pull(entry) |>
  paste(collapse = "\n\n")

prompt <- paste0(
  "Below are abstracts from ", sum(!is.na(abstracts$abstract)),
  " patents in the same selection.\n\n",
  abstract_texts,
  "\n\n---\n",
  "Write a single concise paragraph (150-200 words) summarising the most ",
  "important innovations represented by these patents. Focus on the key ",
  "technological themes, what problems they solve, and why they matter. ",
  "Do not list individual patents — synthesise across all of them."
)

resp <- request("https://api.openai.com/v1/chat/completions") |>
  req_headers(Authorization = paste("Bearer", openai_key),
              `Content-Type` = "application/json") |>
  req_body_json(list(
    model    = "gpt-4o-mini",
    messages = list(list(role = "user", content = prompt)),
    temperature = 0.4,
    max_tokens  = 400
  )) |>
  req_perform()

summary_text <- resp_body_json(resp)$choices[[1]]$message$content
```

### Innovation Summary

```{r print_summary, results='asis'}
cat(summary_text)
```

## 9 — Embedding Clusters

Uses OpenAI embeddings to create vector representations of each patent's
title + abstract, then clusters them with k-means and visualizes in 2D using UMAP.

```{r embeddings_setup, message=FALSE, warning=FALSE}
# Additional packages for clustering and dimensionality reduction
pkgs_embed <- c("umap")
install.packages(setdiff(pkgs_embed, rownames(installed.packages())), quiet = TRUE)
library(umap)
```

```{r create_embeddings, message=TRUE}
# Combine title and abstract for each patent
embed_data <- abstracts |>

filter(!is.na(abstract) | !is.na(title)) |>
  mutate(
    text = paste(
      if_else(!is.na(title), title, ""),
      if_else(!is.na(abstract), abstract, "")
    ) |> str_squish()
  ) |>
  filter(nchar(text) > 10)

cat("Creating embeddings for", nrow(embed_data), "patents...\n")

# Function to get embeddings from OpenAI
get_embeddings <- function(texts, api_key) {
  resp <- request("https://api.openai.com/v1/embeddings") |>
    req_headers(Authorization = paste("Bearer", api_key),
                `Content-Type` = "application/json") |>
    req_body_json(list(
      model = "text-embedding-3-small",
      input = texts
    )) |>
    req_perform()

  result <- resp_body_json(resp)
  # Extract embedding vectors and convert to matrix
  embeddings <- lapply(result$data, function(x) unlist(x$embedding))
  do.call(rbind, embeddings)  # Return as matrix with rows = documents
}

# Get embeddings (OpenAI allows batching)
embedding_matrix <- get_embeddings(embed_data$text, openai_key)
cat("Embedding dimensions:", dim(embedding_matrix), "\n")
```

```{r cluster_embeddings}
# K-means clustering into 10 clusters (or fewer if not enough documents)
set.seed(42)
k <- min(10, nrow(embedding_matrix))
clusters <- kmeans(embedding_matrix, centers = k, nstart = 25)

embed_data$cluster <- as.factor(clusters$cluster)
cat("Patents per cluster:\n")
table(embed_data$cluster)
```

```{r cluster_labels, message=TRUE}
# Build a summary of each cluster for the AI to label
cluster_summaries <- embed_data |>
  group_by(cluster) |>
  summarise(
    patents = paste(
      if_else(!is.na(title), title, str_trunc(text, 100)),
      collapse = "\n- "
    ),
    n = n(),
    .groups = "drop"
  ) |>
  mutate(entry = paste0(
    "Cluster ", cluster, " (", n, " patents):\n- ", patents
  ))

label_prompt <- paste0(
  "Below are clusters of patents grouped by semantic similarity.\n\n",
  paste(cluster_summaries$entry, collapse = "\n\n"),
  "\n\n---\n",
  "For EACH cluster, provide a short descriptive label (2-5 words) that captures ",
  "the common technological theme. Be specific and technical.\n\n",
  "Return ONLY a JSON array of objects with keys \"cluster\" (the number) ",
  "and \"label\" (the short description). No other text."
)

label_resp <- request("https://api.openai.com/v1/chat/completions") |>
  req_headers(Authorization = paste("Bearer", openai_key),
              `Content-Type` = "application/json") |>
  req_body_json(list(
    model    = "gpt-4o-mini",
    messages = list(list(role = "user", content = label_prompt)),
    temperature = 0.2,
    max_tokens  = 500
  )) |>
  req_perform()

raw_json <- resp_body_json(label_resp)$choices[[1]]$message$content
raw_json <- str_remove_all(raw_json, "```json\\s*|```\\s*")

cluster_labels <- jsonlite::fromJSON(raw_json) |>
  as_tibble() |>
  mutate(cluster = as.character(cluster))

# Create named vector for relabeling
label_map <- setNames(cluster_labels$label, cluster_labels$cluster)

# Update cluster factor with descriptive labels
embed_data <- embed_data |>
  mutate(cluster_label = factor(label_map[as.character(cluster)],
                                levels = label_map))

cat("Cluster labels:\n")
cluster_labels |> print(n = 20)
```

```{r umap_viz, fig.width=10, fig.height=8, dpi=150}
# UMAP dimensionality reduction to 2D
umap_config <- umap.defaults
umap_config$n_neighbors <- min(15, nrow(embedding_matrix) - 1)
umap_config$min_dist <- 0.1

umap_result <- umap(embedding_matrix, config = umap_config)

embed_data$umap_x <- umap_result$layout[, 1]
embed_data$umap_y <- umap_result$layout[, 2]

# Create labels for the plot (truncated title or app number)
embed_data <- embed_data |>
  mutate(plot_label = if_else(!is.na(title), str_trunc(title, 30), raw))

# High-contrast colour palette for clusters
cluster_colors <- c(
  "#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00",
  "#a65628", "#f781bf", "#1b9e77", "#d95f02", "#7570b3"
)

# Plot clusters with AI-generated labels
ggplot(embed_data, aes(x = umap_x, y = umap_y, colour = cluster_label)) +
  geom_point(size = 4, alpha = 0.9) +
  geom_text(aes(label = plot_label), size = 2.2, vjust = -1,
            check_overlap = TRUE, show.legend = FALSE) +
  scale_colour_manual(values = cluster_colors) +
  theme_minimal() +
  labs(
    title = "Patent Embedding Clusters (UMAP projection)",
    subtitle = paste(nrow(embed_data), "patents clustered into", k, "groups"),
    x = "UMAP 1",
    y = "UMAP 2",
    colour = "Theme"
  ) +
  theme(legend.position = "right")
```
