---
title: "Pre-compute avstrax Data for ISE App Performance"
output: html_notebook
---

# Overview

This notebook analyzes the ISE Shiny app to identify all parameter combinations that affect
the `avstrax` computation, and pre-computes the aggregated data frames to speed up the app.

The pre-computed data will be stored in `prepdata/` subfolder in the iseapp Dropbox folder.

## Parameter Analysis

The app has two main visualizations that require expensive computation:

### Plot 1: Returns by Technology (for selected countries)
**Function:** `plot_avstrax_by_country()` calls `compute_avstrax()`

Parameters affecting computation:
- `toflow`: The return flow variable (istrax_*, avstrax_*, ev_*)
- `country_code`: Selected countries or country groups
- `tech_categories`: Technology categories to display

### Plot 2: Returns by Country (for selected technologies)
**Function:** `plot_avstrax_by_technology()` calls `compute_avstrax_for_techs()`

Parameters affecting computation:
- `toflow`: The return flow variable
- `techs`: Selected technology categories
- Country filtering (from Plot 1 selection)

### Region Plots
Same structure as country plots but using regionmap instead of countrymap.

## Pre-computation Strategy

We'll pre-compute for the most common combinations:

1. **By Technology** (for Plot 1): For each `toflow` × `country_group`
2. **By Country** (for Plot 2): For each `toflow` × `technology_category`
3. **By Region** (for Region tab): For each `toflow` × `technology_category`

# Setup

```{r setup}
library(fst)
library(dplyr)
library(tidyr)
library(data.table)

source("dropbox_auth.R")

# Dropbox path detection (same as app.R)
get_dropbox_path <- function() {
  if (.Platform$OS.type == "windows") {
    info_paths <- c(
      file.path(Sys.getenv("APPDATA"), "Dropbox", "info.json"),
      file.path(Sys.getenv("LOCALAPPDATA"), "Dropbox", "info.json")
    )
  } else {
    info_paths <- file.path(Sys.getenv("HOME"), ".dropbox", "info.json")
  }
  for (p in info_paths) {
    if (file.exists(p)) {
      info <- jsonlite::fromJSON(p)
      if (!is.null(info$business)) return(info$business$path)
      else if (!is.null(info$personal)) return(info$personal$path)
    }
  }
  return(NULL)
}

get_local_iseapp_path <- function() {
  env_path <- Sys.getenv("ISEAPP_PATH_LOCAL")
  if (nzchar(env_path) && dir.exists(env_path)) return(env_path)
  dropbox_root <- get_dropbox_path()
  if (!is.null(dropbox_root)) {
    iseapp_path <- file.path(dropbox_root, "Apps", "iseapp")
    if (dir.exists(iseapp_path)) return(iseapp_path)
  }
  return(NULL)
}

iseapp_local_path <- get_local_iseapp_path()
message("Using iseapp path: ", iseapp_local_path)

localpath_fname <- function(fname) {
  if (!is.null(iseapp_local_path)) {
    fname <- sub("^/+", "", fname)
    return(file.path(iseapp_local_path, fname))
  }
  return("")
}

# Create prepdata directory
prepdata_path <- file.path(iseapp_local_path, "prepdata")
if (!dir.exists(prepdata_path)) {
  dir.create(prepdata_path, recursive = TRUE)
  message("Created prepdata directory: ", prepdata_path)
}
```

# Define All Parameter Options

```{r define-parameters}
# ============================================
# TOFLOW OPTIONS (Return flow variables)
# ============================================
# These are the istrax files in the istraxes folder

toflow_vars <- c(
  # Marginal Returns
  "istrax_global",
  "istrax_nationalkey_2009_2018",
  "istrax_EMDE",
  "istrax_EMDENOCN",
  "istrax_EMDENOCNIN",
  "istrax_HIC",
  "istrax_EU",
  "istrax_US",
  "istrax_CN",
  "istrax_GB",
  "istrax_AT",
  "istrax_FR",
  # Average Returns
  "avstrax_global",
  "avstrax_nationalkey_2009_2018",
  "avstrax_EMDE",
  "avstrax_EMDENOCN",
  "avstrax_EMDENOCNIN",
  "avstrax_HIC",
  "avstrax_EU",
  "avstrax_US",
  "avstrax_CN",
  "avstrax_GB",
  "avstrax_AT",
  "avstrax_FR",
  # Spillovers
  "ev_global",
  "ev_nationalkey_2009_2018",
  "ev_EMDE",
  "ev_EMDENOCN",
  "ev_EMDENOCNIN",
  "ev_HIC",
  "ev_EU",
  "ev_US",
  "ev_CN",
  "ev_GB",
  "ev_AT",
  "ev_FR"
)

# ============================================
# COUNTRY GROUP OPTIONS
# ============================================
all_countries <- sort(unique(na.omit(countrycode::codelist$iso2c)))

lmics <- c("AF","AL","DZ","AO","AR","AM","AZ","BD","BJ","BO","BA","BW","BR","BG",
           "BF","BI","KH","CM","CV","CF","TD","CL","CN","CO","KM","CG","CR","CI",
           "CU","DJ","DM","DO","EC","EG","SV","GQ","ER","ET","FJ","GA","GM","GE",
           "GH","GT","GN","GW","GY","HT","HN","IN","ID","IR","IQ","JM","JO","KZ",
           "KE","KI","KP","KG","LA","LB","LS","LR","LY","MG","MW","MY","MV","ML",
           "MR","MU","MX","MD","MN","ME","MA","MZ","MM","NA","NP","NI","NE","NG",
           "MK","PK","PW","PA","PG","PY","PE","PH","RW","WS","ST","SN","RS","SC",
           "SL","SB","SO","ZA","LK","SD","SR","SY","TJ","TZ","TH","TL","TG","TO",
           "TN","TR","TM","TV","UG","UA","UZ","VU","VE","VN","YE","ZM","ZW")

eu_countries <- c("AT","BE","BG","HR","CY","CZ","DK","EE","FI","FR","DE","GR",
                  "HU","IE","IT","LV","LT","LU","MT","NL","PL","PT","RO","SK",
                  "SI","ES","SE")

hic <- setdiff(all_countries, lmics)

country_groups <- list(
  "All_countries" = all_countries,
  "LMICs" = lmics,
  "LMICs_excl_China" = setdiff(lmics, "CN"),
  "EU_countries" = eu_countries,
  "High_income_countries" = hic
)

# ============================================
# TECHNOLOGY CATEGORY OPTIONS
# ============================================
green_classes <- c("Green Technology", "Green Energy", "Green Transport",
                   "Circular Economy", "Green Manufacturing", "Adaptation",
                   "Green Housing", "Green ICT", "Green Agriculture", "GHG Capture")

battery_classes <- c("Battery Technology", "Lithium Extraction & Processing",
                     "Graphite & Carbon Materials", "Cathode Materials", "Anode Materials",
                     "Electrolytes & Additives", "Separators", "Battery Cell Design & Assembly",
                     "Battery Management Systems (BMS)", "Electric Vehicles & Mobility",
                     "Battery Recycling & Recovery")

hard_to_abate_classes <- c("Hard to Abate Sector Decarbonization", "Aviation Decarbonisation",
                           "Cement & Concrete Decarbonisation", "Chemicals & Plastics Decarbonisation",
                           "Shipping Decarbonisation", "Steel & Iron Decarbonisation")

ai_classes <- c("AI", "Machine Learning", "Deep Learning", "Natural Language Processing (NLP)",
                "Computer Vision", "Speech Recognition & Synthesis", "Robotics & Autonomous Systems",
                "Knowledge Representation & Reasoning", "Planning & Decision Making", "Generative AI",
                "Semiconductors", "Cloud & Data Infrastructure", "Data Rettrieval & Processing System",
                "Platform & Frameworks", "Deployment & Support")

cpc_sections <- c("Human Necessities", "Performing Operations; Transporting ",
                  "Chemistry; Metallurgy ", "Textiles; Paper", "Fixed Constructions",
                  "Mechanical Engineering; Lighting; Heating; Weapons; Blasting",
                  "Physics", "Electricity", "General tagging of new or cross-sectional technology")

# Key technology categories for pre-computation
tech_categories <- list(
  "All" = "All",
  "Green_Technology" = "Green Technology",
  "Battery_Technology" = "Battery Technology",
  "Hard_to_Abate" = "Hard to Abate Sector Decarbonization",
  "AI" = "AI",
  "Other" = "Other"
)

# ============================================
# UK REGION OPTIONS
# ============================================
uk_regions <- c("UKC", "UKD", "UKE", "UKF", "UKG", "UKH", "UKI", "UKJ", "UKK", "UKL", "UKM", "UKN")

cat("Parameter summary:\n")
cat("- toflow variables:", length(toflow_vars), "\n")
cat("- Country groups:", length(country_groups), "\n")
cat("- Technology categories:", length(tech_categories), "\n")
cat("- UK regions:", length(uk_regions), "\n")
```

# Load Base Data

```{r load-base-data}
# Load countrymap
pp <- localpath_fname("countrymap.fst")
if (file.exists(pp)) {
  countrymap <- read_fst(pp)
  message("Loaded countrymap: ", nrow(countrymap), " rows")
} else {
  stop("countrymap.fst not found")
}

# Load techmap
pp <- localpath_fname("techmap.fst")
if (file.exists(pp)) {
  techmap <- read_fst(pp)
  message("Loaded techmap: ", nrow(techmap), " rows")
} else {
  stop("techmap.fst not found")
}

# Add "All" technology
techmap <- countrymap %>%
  select(docdb_family_id) %>%
  distinct() %>%
  mutate(technology = "All") %>%
  bind_rows(techmap)

# Standardize technology names
setDT(techmap)
techmap[, technology := fcase(
  technology == "Any Green technology", "Green Technology",
  technology == "Any battery technology", "Battery Technology",
  technology == "Any Hard to Abate technology", "Hard to Abate Sector Decarbonization",
  default = technology
)]

# Load regionmap if available
pp_region <- localpath_fname("regionmap.fst")
if (file.exists(pp_region)) {
  regionmap <- read_fst(pp_region)
  message("Loaded regionmap: ", nrow(regionmap), " rows")
} else {
  regionmap <- NULL
  message("regionmap.fst not found - skipping region pre-computation")
}
```

# Pre-computation Functions

```{r precompute-functions}
source("istraxfunctions.R")

# Function to generate safe filename
safe_filename <- function(...) {
  parts <- c(...)
  name <- paste(parts, collapse = "_")
  name <- gsub("[^a-zA-Z0-9_]", "_", name)
  name <- gsub("_+", "_", name)
  name <- gsub("^_|_$", "", name)
  paste0(name, ".fst")
}

# Pre-compute avstrax by technology for a given toflow and country group
precompute_by_tech <- function(toflow, country_group_name, country_codes) {
  # Load istrax data
  istrax_path <- localpath_fname(paste0("istraxes/", toflow, ".fst"))
  if (!file.exists(istrax_path)) {
    message("  Skipping - istrax file not found: ", toflow)
    return(NULL)
  }

  istrax_data <- read_fst(istrax_path)

  # Replace missing values with 0 in the value column
  # Some files (avstrax, ev) contain NAs that need to be treated as 0
  if (toflow %in% names(istrax_data)) {
    istrax_data[[toflow]][is.na(istrax_data[[toflow]])] <- 0
  }

  # Join with countrymap on both docdb_family_id and ctry_code
  data <- countrymap %>%
    filter(ctry_code %in% country_codes) %>%
    left_join(istrax_data, by = c("docdb_family_id", "ctry_code"))

  if (nrow(data) == 0) {
    message("  Skipping - no data for country group: ", country_group_name)
    return(NULL)
  }

  # Compute avstrax using the existing function
  # We need to pass techmap as classes
  result <- tryCatch({
    compute_avstrax(data, toflow, techmap, colorings = NULL)
  }, error = function(e) {
    message("  Error computing avstrax: ", e$message)
    return(NULL)
  })

  if (!is.null(result)) {
    result$toflow <- toflow
    result$country_group <- country_group_name
  }

  return(result)
}

# Pre-compute avstrax by country for a given toflow and technology
precompute_by_country <- function(toflow, tech_name, tech_filter) {
  # Load istrax data
  istrax_path <- localpath_fname(paste0("istraxes/", toflow, ".fst"))
  if (!file.exists(istrax_path)) {
    message("  Skipping - istrax file not found: ", toflow)
    return(NULL)
  }

  istrax_data <- read_fst(istrax_path)

  # Replace missing values with 0 in the value column
  # Some files (avstrax, ev) contain NAs that need to be treated as 0
  if (toflow %in% names(istrax_data)) {
    istrax_data[[toflow]][is.na(istrax_data[[toflow]])] <- 0
  }

  # Join with countrymap on both docdb_family_id and ctry_code
  data <- countrymap %>%
    left_join(istrax_data, by = c("docdb_family_id", "ctry_code"))

  # Filter techmap for specific technology
  if (tech_filter == "All") {
    classes <- data.frame()  # Empty = all innovations
  } else if (tech_filter == "Other") {
    # Other = not in green, battery, hard_to_abate, ai
    exclude_techs <- c(green_classes, battery_classes, hard_to_abate_classes, ai_classes)
    classes <- techmap %>% filter(!technology %in% exclude_techs)
  } else {
    classes <- techmap %>% filter(technology == tech_filter)
  }

  # Compute avstrax by country
  result <- tryCatch({
    compute_avstrax_for_techs(data, toflow, classes)
  }, error = function(e) {
    message("  Error computing avstrax: ", e$message)
    return(NULL)
  })

  if (!is.null(result)) {
    result$toflow <- toflow
    result$tech_category <- tech_name
  }

  return(result)
}
```

# Run Pre-computation

```{r run-precompute-by-tech, message=TRUE}
# Pre-compute by technology for each toflow × country_group combination
message("=== Pre-computing by Technology ===\n")

for (toflow in toflow_vars) {
  message("\nProcessing toflow: ", toflow)

  for (group_name in names(country_groups)) {
    filename <- safe_filename("by_tech", toflow, group_name)
    filepath <- file.path(prepdata_path, filename)

    # Skip if file already exists
    if (file.exists(filepath)) {
      message("  Country group: ", group_name, " - SKIPPED (file exists)")
      next
    }

    message("  Country group: ", group_name)

    result <- precompute_by_tech(toflow, group_name, country_groups[[group_name]])

    if (!is.null(result) && nrow(result) > 0) {
      write_fst(result, filepath, compress = 50)
      message("    Saved: ", filename, " (", nrow(result), " rows)")
    }
  }
}
```

```{r run-precompute-by-country, message=TRUE}
# Pre-compute by country for each toflow × technology combination
message("\n=== Pre-computing by Country ===\n")

for (toflow in toflow_vars) {
  message("\nProcessing toflow: ", toflow)

  for (tech_name in names(tech_categories)) {
    filename <- safe_filename("by_country", toflow, tech_name)
    filepath <- file.path(prepdata_path, filename)

    # Skip if file already exists
    if (file.exists(filepath)) {
      message("  Technology: ", tech_name, " - SKIPPED (file exists)")
      next
    }

    message("  Technology: ", tech_name)

    result <- precompute_by_country(toflow, tech_name, tech_categories[[tech_name]])

    if (!is.null(result) && nrow(result) > 0) {
      write_fst(result, filepath, compress = 50)
      message("    Saved: ", filename, " (", nrow(result), " rows)")
    }
  }
}
```

```{r run-precompute-regions, message=TRUE}
# Pre-compute for regions (if regionmap available)
if (!is.null(regionmap) && nrow(regionmap) > 0) {
  message("\n=== Pre-computing by Region ===\n")

  for (toflow in toflow_vars) {
    message("\nProcessing toflow: ", toflow)

    # Load istrax data
    istrax_path <- localpath_fname(paste0("istraxes/", toflow, ".fst"))
    if (!file.exists(istrax_path)) {
      message("  Skipping - istrax file not found")
      next
    }

    istrax_data <- read_fst(istrax_path)

    # Replace missing values with 0 in the value column
    # Some files (avstrax, ev) contain NAs that need to be treated as 0
    if (toflow %in% names(istrax_data)) {
      istrax_data[[toflow]][is.na(istrax_data[[toflow]])] <- 0
    }

    # Keep ctry_code in istrax_data for joining (patents have country-specific returns)
    # Remove ctry_name to avoid column conflicts
    istrax_data_for_join <- istrax_data %>% select(-any_of("ctry_name"))

    for (tech_name in names(tech_categories)) {
      filename <- safe_filename("by_region", toflow, tech_name)
      filepath <- file.path(prepdata_path, filename)

      # Skip if file already exists
      if (file.exists(filepath)) {
        message("  Technology: ", tech_name, " - SKIPPED (file exists)")
        next
      }

      message("  Technology: ", tech_name)

      # Prepare region data - join by both docdb_family_id and ctry_code to get GB-specific returns
      # Then rename region_code to ctry_code for compatibility with plotting functions
      region_data <- regionmap %>%
        left_join(istrax_data_for_join, by = c("docdb_family_id", "ctry_code")) %>%
        select(-ctry_code) %>%
        rename(ctry_code = region_code, country_name = region_name)

      # Filter techmap
      if (tech_categories[[tech_name]] == "All") {
        classes <- data.frame()
      } else if (tech_categories[[tech_name]] == "Other") {
        exclude_techs <- c(green_classes, battery_classes, hard_to_abate_classes, ai_classes)
        classes <- techmap %>% filter(!technology %in% exclude_techs)
      } else {
        classes <- techmap %>% filter(technology == tech_categories[[tech_name]])
      }

      result <- tryCatch({
        compute_avstrax_for_techs(region_data, toflow, classes)
      }, error = function(e) {
        message("    Error: ", e$message)
        return(NULL)
      })

      if (!is.null(result) && nrow(result) > 0) {
        result$toflow <- toflow
        result$tech_category <- tech_name

        write_fst(result, filepath, compress = 50)
        message("    Saved: ", filename, " (", nrow(result), " rows)")
      }
    }
  }
} else {
  message("\nSkipping region pre-computation - regionmap not available")
}
```

# Summary

```{r summary}
# List all pre-computed files
files <- list.files(prepdata_path, pattern = "\\.fst$", full.names = TRUE)
message("\n=== Pre-computation Complete ===\n")
message("Total files created: ", length(files))
message("Location: ", prepdata_path)

# Show file sizes
if (length(files) > 0) {
  file_info <- file.info(files)
  total_size <- sum(file_info$size) / 1024 / 1024
  message("Total size: ", round(total_size, 2), " MB")

  cat("\nFiles by type:\n")
  cat("- by_tech files:", sum(grepl("by_tech", files)), "\n")
  cat("- by_country files:", sum(grepl("by_country", files)), "\n")
  cat("- by_region files:", sum(grepl("by_region", files)), "\n")
}
```

# How to Add New Options

When adding new options to the app, update the following:

1. **New toflow variable**: Add to `toflow_vars` list and ensure the .fst file exists in `istraxes/`

2. **New country group**: Add to `country_groups` list with the ISO2 country codes

3. **New technology category**: Add to `tech_categories` list

4. **New region**: Update UK regions list if needed

Then re-run this notebook to regenerate the pre-computed files.

```{r session-info}
sessionInfo()
```
