---
title: "Prepare and Explore ISE Data"
output: html_notebook
---

# Setup

```{r setup}
library(fst)
library(arrow)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggiraph)
library(wbstats)
library(countrycode)
library(jsonlite)

source("dropbox_auth.R")

# Function to detect local Dropbox path from Dropbox's config file
get_dropbox_path <- function() {
  if (.Platform$OS.type == "windows") {
    info_paths <- c(
      file.path(Sys.getenv("APPDATA"), "Dropbox", "info.json"),
      file.path(Sys.getenv("LOCALAPPDATA"), "Dropbox", "info.json")
    )
  } else {
    info_paths <- file.path(Sys.getenv("HOME"), ".dropbox", "info.json")
  }

  for (p in info_paths) {
    if (file.exists(p)) {
      info <- jsonlite::fromJSON(p)
      if (!is.null(info$business)) {
        return(info$business$path)
      } else if (!is.null(info$personal)) {
        return(info$personal$path)
      }
    }
  }
  return(NULL)
}

# Try to find local Dropbox Apps/iseapp folder
get_local_iseapp_path <- function() {
  # First check environment variable
  env_path <- Sys.getenv("ISEAPP_PATH_LOCAL")
  if (nzchar(env_path) && dir.exists(env_path)) {
    return(env_path)
  }
  # Try to detect Dropbox path automatically
  dropbox_root <- get_dropbox_path()
  if (!is.null(dropbox_root)) {
    iseapp_path <- file.path(dropbox_root, "Apps", "iseapp")
    if (dir.exists(iseapp_path)) {
      return(iseapp_path)
    }
  }
  return(NULL)
}

# Get the local path (cached for session)
iseapp_local_path <- get_local_iseapp_path()
if (!is.null(iseapp_local_path)) {
  message("Using local Dropbox path: ", iseapp_local_path)
} else {
  message("Local Dropbox not found, will use online Dropbox API")
}

# Helper function for local file paths
localpath_fname <- function(fname) {
  if (!is.null(iseapp_local_path)) {
    fname <- sub("^/+", "", fname)
    return(file.path(iseapp_local_path, fname))
  }
  return("")
}
```

# Data Loading

Load the core mapping files with Dropbox fallback pattern.

```{r load-countrymap}
pp <- localpath_fname("countrymap.fst")
if (file.exists(pp)) {
  countrymap <- read_fst(pp)
} else {
  countrymap <- dropbox_read_fst("/countrymap.fst")
}

cat("Loaded countrymap:", nrow(countrymap), "rows\n")


countrymap %>% distinct(docdb_family_id) %>% nrow()
test=countrymap %>% filter(ctry_code=="GB") 

test=countrymap %>% filter(ctry_code=="klj")

test=countrymap %>% distinct(ctry_code)

```

```{r load-techmap}
pp <- localpath_fname("techmap.fst")
if (file.exists(pp)) {
  techmap <- read_fst(pp)
} else {
  techmap <- dropbox_read_fst("/techmap.fst")
}

cat("Loaded techmap:", nrow(techmap), "rows\n")

techmap %>% distinct(technology)


```

# Data Overview

## Countrymap Structure

```{r countrymap-overview}
cat("Dimensions:", nrow(countrymap), "rows x", ncol(countrymap), "columns\n\n")
cat("Column names:\n")
print(names(countrymap))
```

```{r countrymap-types}
cat("\nData types:\n")
str(countrymap)
```

```{r countrymap-head}
head(countrymap, 10)
```

## Techmap Structure

```{r techmap-overview}
cat("Dimensions:", nrow(techmap), "rows x", ncol(techmap), "columns\n\n")
cat("Column names:\n")
print(names(techmap))
```

```{r techmap-types}
cat("\nData types:\n")
str(techmap)
```

```{r techmap-head}
head(techmap, 10)
```

# Missing Value Analysis

## Countrymap Completeness

```{r countrymap-missing}
countrymap_missing <- countrymap %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "na_count") %>%
  mutate(
    total = nrow(countrymap),
    na_pct = round(100 * na_count / total, 2)
  ) %>%
  arrange(desc(na_pct))

print(countrymap_missing)



```

## Techmap Completeness

```{r techmap-missing}
techmap_missing <- techmap %>%
  summarise(across(everything(), ~sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "na_count") %>%
  mutate(
    total = nrow(techmap),
    na_pct = round(100 * na_count / total, 2)
  ) %>%
  arrange(desc(na_pct))

print(techmap_missing)
```

# Summary Statistics

## Countrymap Key Metrics

```{r countrymap-summary}
# Numeric columns summary
countrymap %>%
  select(where(is.numeric)) %>%
  summary()
```

## Techmap Key Metrics

```{r techmap-summary}
# Numeric columns summary
techmap %>%
  select(where(is.numeric)) %>%
  summary()
```

# Country Analysis

## Patent Counts by Country

```{r country-patent-counts}
if ("ctry_code" %in% names(countrymap)) {
  country_counts <- countrymap %>%
    count(ctry_code, name = "patent_count") %>%
    arrange(desc(patent_count))

  cat("Total unique countries:", n_distinct(country_counts$ctry_code), "\n\n")
  cat("Top 20 countries by patent count:\n")
  print(head(country_counts, 20))
}
```

## Top Contributors Visualization

```{r country-top-chart, fig.width=10, fig.height=8}
if ("ctry_code" %in% names(countrymap)) {
  top_countries <- country_counts %>%
    slice_max(patent_count, n = 20) %>%
    mutate(ctry_code = fct_reorder(ctry_code, patent_count))

  p <- ggplot(top_countries, aes(x = ctry_code, y = patent_count)) +
    geom_col_interactive(
      aes(tooltip = paste0(ctry_code, ": ", scales::comma(patent_count))),
      fill = "steelblue"
    ) +
    coord_flip() +
    labs(
      title = "Top 20 Countries by Patent Count",
      x = "Country Code",
      y = "Number of Patents"
    ) +
    theme_minimal() +
    scale_y_continuous(labels = scales::comma)

  girafe(ggobj = p)
}
```

## Country Distribution with World Bank Data

```{r country-income-analysis}
if ("ctry_code" %in% names(countrymap)) {
  # Get income classification
  income_class <- wb_countries() %>%
    select(ctry_code = iso2c, income_level, region)

  country_summary <- country_counts %>%
    left_join(income_class, by = "ctry_code") %>%
    filter(!is.na(income_level))

  # Summary by income level
  income_summary <- country_summary %>%
    group_by(income_level) %>%
    summarise(
      n_countries = n(),
      total_patents = sum(patent_count),
      mean_patents = mean(patent_count),
      median_patents = median(patent_count)
    ) %>%
    arrange(desc(total_patents))

  print(income_summary)
}
```

# Technology Analysis

## Technology Category Distribution

```{r tech-distribution}
# Check for technology classification columns in techmap
tech_cols <- names(techmap)[grepl("tech|ipc|cpc|class", names(techmap), ignore.case = TRUE)]
cat("Technology-related columns found:", paste(tech_cols, collapse = ", "), "\n\n")

if (length(tech_cols) > 0) {
  # Show distribution for first tech column found
  first_tech_col <- tech_cols[1]
  cat("Distribution for", first_tech_col, ":\n")

  tech_dist <- techmap %>%
    count(!!sym(first_tech_col), name = "count") %>%
    arrange(desc(count))

  print(head(tech_dist, 20))
}
```

## Technology Categories Visualization

```{r tech-chart, fig.width=10, fig.height=8}
if (length(tech_cols) > 0 && nrow(tech_dist) > 0) {
  top_tech <- tech_dist %>%
    slice_max(count, n = 15) %>%
    mutate(category = fct_reorder(as.character(!!sym(first_tech_col)), count))

  p <- ggplot(top_tech, aes(x = category, y = count)) +
    geom_col_interactive(
      aes(tooltip = paste0(category, ": ", scales::comma(count))),
      fill = "darkgreen"
    ) +
    coord_flip() +
    labs(
      title = paste("Top 15 Technology Categories by", first_tech_col),
      x = "Category",
      y = "Count"
    ) +
    theme_minimal() +
    scale_y_continuous(labels = scales::comma)

  girafe(ggobj = p)
}
```

# Time Series Analysis

## Filing Trends Over Years

```{r time-series-prep}
# Check for year columns
year_cols <- names(countrymap)[grepl("year|date", names(countrymap), ignore.case = TRUE)]
cat("Year-related columns found:", paste(year_cols, collapse = ", "), "\n")
```

```{r filing-trends, fig.width=12, fig.height=6}
if (length(year_cols) > 0) {
  year_col <- year_cols[1]
  cat("Using column:", year_col, "\n\n")

  yearly_counts <- countrymap %>%
    filter(!is.na(!!sym(year_col))) %>%
    count(year = !!sym(year_col), name = "patents") %>%
    arrange(year)

  # Filter to reasonable year range
  yearly_counts <- yearly_counts %>%
    filter(year >= 1980 & year <= 2025)

  print(yearly_counts)

  p <- ggplot(yearly_counts, aes(x = year, y = patents)) +
    geom_line_interactive(
      aes(tooltip = paste0("Year ", year, ": ", scales::comma(patents), " patents")),
      color = "steelblue",
      linewidth = 1
    ) +
    geom_point_interactive(
      aes(tooltip = paste0("Year ", year, ": ", scales::comma(patents), " patents")),
      color = "steelblue",
      size = 2
    ) +
    labs(
      title = "Patent Filing Trends Over Time",
      x = "Year",
      y = "Number of Patents"
    ) +
    theme_minimal() +
    scale_y_continuous(labels = scales::comma)

  girafe(ggobj = p)
}
```

## Trends by Top Countries

```{r trends-by-country, fig.width=12, fig.height=8}
if (length(year_cols) > 0 && "ctry_code" %in% names(countrymap)) {
  # Get top 5 countries
  top5 <- country_counts %>%
    slice_max(patent_count, n = 5) %>%
    pull(ctry_code)

  yearly_by_country <- countrymap %>%
    filter(ctry_code %in% top5, !is.na(!!sym(year_col))) %>%
    count(year = !!sym(year_col), ctry_code, name = "patents") %>%
    filter(year >= 1990 & year <= 2025)

  p <- ggplot(yearly_by_country, aes(x = year, y = patents, color = ctry_code)) +
    geom_line(linewidth = 1) +
    geom_point(size = 1.5) +
    labs(
      title = "Patent Filing Trends by Top 5 Countries",
      x = "Year",
      y = "Number of Patents",
      color = "Country"
    ) +
    theme_minimal() +
    scale_y_continuous(labels = scales::comma)

  print(p)
}
```

# Cross-tabulation Analysis

## Country x Technology Matrix (if applicable)

```{r cross-tab}
# Check if we can join countrymap and techmap
common_cols <- intersect(names(countrymap), names(techmap))
cat("Common columns between countrymap and techmap:\n")
print(common_cols)

if (length(common_cols) > 0) {
  cat("\nThese columns can be used to join the datasets.\n")
}
```

# Data Quality Summary

```{r quality-summary}
cat("=== DATA QUALITY SUMMARY ===\n\n")

cat("COUNTRYMAP:\n")
cat("  - Total records:", scales::comma(nrow(countrymap)), "\n")
cat("  - Total columns:", ncol(countrymap), "\n")
cat("  - Complete cases:", scales::comma(sum(complete.cases(countrymap))), "\n")
cat("  - Complete case rate:", round(100 * sum(complete.cases(countrymap)) / nrow(countrymap), 1), "%\n\n")

cat("TECHMAP:\n")
cat("  - Total records:", scales::comma(nrow(techmap)), "\n")
cat("  - Total columns:", ncol(techmap), "\n")
cat("  - Complete cases:", scales::comma(sum(complete.cases(techmap))), "\n")
cat("  - Complete case rate:", round(100 * sum(complete.cases(techmap)) / nrow(techmap), 1), "%\n")
```

# Session Info

```{r session-info}
sessionInfo()
```
