---
title: "Test: national istrax for developing countries"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fst)
library(arrow)
library(dplyr)
library(readr)
```

## Green share 2009-2018 by country group (HIC / CN / non-HIC excl. China)

Filter each PRINZglobal tagging to docdbs with earliest_filing_year in [2009, 2018],
attach harmonized inventor countries, and compute the share of green docdbs within
each country group.

```{r}
library(data.table)

prinz_dir <- "C:/Users/rmart/OneDrive/PRINZglobal/data"
fromPATSTAT <- "C:/Users/rmart/Dropbox/patbis2021/data/fromPATSTAT"
bigdata_dir <- ".bigdata"

# 1. Family -> earliest_filing_year (from innos.dsv)
fy <- fread(file.path(fromPATSTAT, "innos.dsv"),
            select = c("docdb_family_id", "earliest_filing_year"))
fy <- unique(fy[earliest_filing_year >= 2000 & earliest_filing_year <= 2018,
                .(docdb_family_id)])
cat("Families in [2009, 2018]:", nrow(fy), "\n")

# 2. Harmonized inventor countries -> country group per family
inv_harm <- as.data.table(arrow::read_parquet(
  file.path(bigdata_dir, "inventor_countries_harm.parquet")
))
setnames(inv_harm, "person_ctry_code", "ctry_code")

# Restrict to valid ISO2 country codes (drops "EP", "WO", blanks, typos, etc.)
valid_iso2 <- unique(na.omit(countrycode::codelist$iso2c))
inv_harm  <- inv_harm[ctry_code %in% valid_iso2]

hic_codes <- c("AD","AE","AG","AR","AU","AT","AW","BE","BH","BM","BS","BW","CA","CH","CL","CN","CW","CY","CZ","DE","DK","EE","ES","FI","FR","GB","GI","GR","GU","HK","HR","HU","IE","IL","IM","IS","IT","JP","KR","KW","KY","LI","LT","LU","LV","MC","MP","MT","MO","NL","NO","NZ","OM","PA","PL","PR","PT","QA","RO","SA","SC","SG","SI","SK","SM","SE","TC","TT","TW","UY","US","VI","VG")

inv_harm[, group := fifelse(ctry_code == "CN", "China",
                     fifelse(ctry_code %in% hic_codes, "HIC",
                             "Non-HIC (excl. China)"))]

# For each family, collapse to the set of groups it's attached to
fam_groups <- unique(inv_harm[, .(docdb_family_id, group)])

# 3. Helper: for a tag file, produce the 3-group share table
compute_shares <- function(tag_file, green_values, label) {
  t <- fread(tag_file)
  # second column is the tag; normalise name
  setnames(t, 2, "tag")
  green_fams <- unique(t[tag %in% green_values, .(docdb_family_id)])
  green_fams[, is_green := TRUE]

  # filter to 2009-2018 families, attach group(s), attach green flag
  fg <- fam_groups[fy, on = "docdb_family_id", nomatch = 0L]
  fg <- green_fams[fg, on = "docdb_family_id"]
  fg[is.na(is_green), is_green := FALSE]

  res <- fg[, .(total_fams = uniqueN(docdb_family_id),
                green_fams = uniqueN(docdb_family_id[is_green]),
                share      = uniqueN(docdb_family_id[is_green]) /
                             uniqueN(docdb_family_id)),
            by = group][order(group)]
  res[, source := label]
  setcolorder(res, c("source", "group", "total_fams", "green_fams", "share"))
  res
}

res_green_or_not <- compute_shares(
  file.path(prinz_dir, "docdb_green_or_not.csv"),
  green_values = c("Green Technology"),
  label = "docdb_green_or_not"
)

res_level0 <- compute_shares(
  file.path(prinz_dir, "docdb_level0.csv"),
  # All non-"Trending Technologies", non-"Fossil Fuel Technologies" level0 tags
  # are green sub-categories. Adjust if you want a different set.
  green_values = c("Adaptation","Circular Economy","GHG Capture","Green Agriculture",
                   "Green Buildings","Green Energy","Green ICT","Green Manufacturing",
                   "Green Transport"),
  label = "docdb_level0"
)

res_nolevel <- compute_shares(
  file.path(prinz_dir, "docdb_nolevel.csv"),
  green_values = c("Green technology"),
  label = "docdb_nolevel"
)

rbindlist(list(res_green_or_not, res_level0, res_nolevel))
```

## Load RTA vs GDP data

```{r}
rta_vs_gdp <- read_csv(".bigdata/rta_vs_gdp_2026-04-17.csv")
cat("rta_vs_gdp rows:", nrow(rta_vs_gdp), "\n")
head(rta_vs_gdp)
```

## Load PRINZglobal docdb counts

```{r}
count_docdb <- read_csv("C:/Users/rmart/OneDrive/PRINZglobal/data/count_docdb_2024-11-21.csv")
cat("count_docdb rows:", nrow(count_docdb), "\n")
cat("columns:", paste(names(count_docdb), collapse = ", "), "\n")
head(count_docdb)
count_docdb %>% distinct(technology) %>% pull(technology)
```

## Green technology: pre/post 2010 x income group

```{r}
count_docdb |>
  dplyr::filter(technology == "Green technology") |>
  dplyr::mutate(
    period = dplyr::if_else(earliest_filing_year < 2010, "<2010", ">=2010"),
    income_grp = dplyr::case_when(
      person_ctry_code == "CN"           ~ "China",
      `Income group` == "High income"    ~ "High income",
      TRUE                               ~ "Non-high income (excl. China)"
    )
  ) |>
  dplyr::group_by(period, income_grp) |>
  dplyr::summarise(
    tech_sum = sum(docdb_family_count_tech, na.rm = TRUE),
    all_sum  = sum(docdb_family_count,      na.rm = TRUE),
    share    = tech_sum / all_sum,
    .groups  = "drop"
  ) |>
  dplyr::arrange(period, income_grp)
```

## Compare countries present in count_docdb vs rta_vs_gdp

```{r}
library(countrycode)

code_to_name <- function(codes) {
  nm <- countrycode(codes, origin = "iso2c", destination = "country.name", warn = FALSE)
  paste0(codes, " (", ifelse(is.na(nm), "??", nm), ")")
}

cd_countries  <- sort(unique(na.omit(count_docdb$person_ctry_code)))
rvg_countries <- sort(unique(na.omit(rta_vs_gdp$ctry_code)))

cat("count_docdb countries: ", length(cd_countries), "\n")
cat("rta_vs_gdp countries:  ", length(rvg_countries), "\n")
cat("Overlap:               ", length(intersect(cd_countries, rvg_countries)), "\n\n")

cat("In count_docdb only (not in rta_vs_gdp):\n")
print(code_to_name(setdiff(cd_countries, rvg_countries)))

cat("\nIn rta_vs_gdp only (not in count_docdb):\n")
print(code_to_name(setdiff(rvg_countries, cd_countries)))
```

## Compare Green Technology mapping: PRINZglobal vs Shiny DB

```{r}
# PRINZglobal mapping (all rows are "Green Technology")
green_prinz <- read_csv("C:/Users/rmart/OneDrive/PRINZglobal/data/docdb_green_or_not.csv") |>
  dplyr::filter(green_or_not == "Green Technology") |>
  dplyr::distinct(docdb_family_id)

# Shiny DB mapping: every family in patents_x_tech whose technology belongs
# to tech_group "Green Technology"
patents_x_tech <- arrow::read_parquet("inst/extdata/patents_x_tech.parquet")
tech_lookup    <- arrow::read_parquet("inst/extdata/tech_lookup.parquet")

green_shiny <- patents_x_tech |>
  dplyr::inner_join(tech_lookup, by = "technology") |>
  dplyr::filter(tech_group == "Green Technology") |>
  dplyr::distinct(docdb_family_id)

cat("PRINZglobal green docdbs: ", nrow(green_prinz), "\n")
cat("Shiny DB green docdbs:    ", nrow(green_shiny), "\n")
cat("Overlap:                  ",
    length(intersect(green_prinz$docdb_family_id, green_shiny$docdb_family_id)), "\n\n")

# In PRINZglobal only (missing from Shiny DB)
only_prinz <- dplyr::anti_join(green_prinz, green_shiny, by = "docdb_family_id")
cat("In PRINZglobal only: ", nrow(only_prinz), "\n")
head(only_prinz)

# In Shiny DB only (missing from PRINZglobal)
only_shiny <- dplyr::anti_join(green_shiny, green_prinz, by = "docdb_family_id")
cat("\nIn Shiny DB only:    ", nrow(only_shiny), "\n")
head(only_shiny)
```

## Compare country-level counts: count_docdb (2009-2018, Green) vs rta_vs_gdp

```{r}
cd_2009_2018 <- count_docdb |>
  dplyr::filter(technology == "Green technology",
                earliest_filing_year >= 2009,
                earliest_filing_year <= 2018) |>
  dplyr::group_by(ctry_code = person_ctry_code) |>
  dplyr::summarise(
    tech_sum = sum(docdb_family_count_tech, na.rm = TRUE),
    all_sum  = sum(docdb_family_count,      na.rm = TRUE),
    .groups = "drop"
  )

comparison <- dplyr::full_join(
    cd_2009_2018,
    rta_vs_gdp |> dplyr::select(ctry_code, innos, Allinnos),
    by = "ctry_code"
  ) |>
  dplyr::mutate(
    country_name = countrycode::countrycode(ctry_code, origin = "iso2c",
                                            destination = "country.name",
                                            warn = FALSE),
    diff_tech = tech_sum - innos,
    diff_all  = all_sum - Allinnos,
    ratio_tech = tech_sum / innos,
    ratio_all  = all_sum  / Allinnos
  ) |>
  dplyr::select(ctry_code, country_name, dplyr::everything()) |>
  dplyr::arrange(dplyr::desc(abs(diff_all)))

cat("Totals across countries:\n")
cat("  count_docdb tech_sum total: ", sum(comparison$tech_sum, na.rm = TRUE), "\n")
cat("  rta_vs_gdp innos total:     ", sum(comparison$innos,    na.rm = TRUE), "\n")
cat("  count_docdb all_sum total:  ", sum(comparison$all_sum,  na.rm = TRUE), "\n")
cat("  rta_vs_gdp Allinnos total:  ", sum(comparison$Allinnos, na.rm = TRUE), "\n\n")

comparison
```

## Green technology: 2009-2018 x income group

```{r}
count_docdb |>
  dplyr::filter(technology == "Green technology",
                earliest_filing_year >= 2009,
                earliest_filing_year <= 2018) |>
  dplyr::mutate(
    income_grp = dplyr::case_when(
      person_ctry_code == "CN"           ~ "China",
      `Income group` == "High income"    ~ "High income",
      TRUE                               ~ "Non-high income (excl. China)"
    )
  ) |>
  dplyr::group_by(income_grp) |>
  dplyr::summarise(
    tech_sum = sum(docdb_family_count_tech, na.rm = TRUE),
    all_sum  = sum(docdb_family_count,      na.rm = TRUE),
    share    = tech_sum / all_sum,
    .groups  = "drop"
  ) |>
  dplyr::arrange(income_grp)
```

## Load data

```{r}
istrax <- read_fst(".bigdata/istraxes/istrax_nationalkey_2009_2018.fst")
country_lookup <- read_parquet("inst/extdata/country_lookup.parquet")

cat("istrax rows:", nrow(istrax), "\n")
cat("country_lookup rows:", nrow(country_lookup), "\n")
```

## Filter to LMICs excluding China

```{r}
lmic_codes <- country_lookup |>
  filter(is_lmic_excl_china) |>
  pull(ctry_code)

istrax_lmic <- istrax |>
  filter(ctry_code %in% lmic_codes)

cat("LMIC (excl. China) rows:", nrow(istrax_lmic), "\n")
cat("LMIC (excl. China) countries in istrax:", n_distinct(istrax_lmic$ctry_code), "\n")
head(istrax_lmic)
```

## Summary by country

```{r}
istrax_lmic |>
  group_by(ctry_code) |>
  summarise(
    n = n(),
    mean_istrax = mean(istrax_nationalkey_2009_2018, na.rm = TRUE),
    median_istrax = median(istrax_nationalkey_2009_2018, na.rm = TRUE)
  ) |>
  arrange(desc(n))
```


## rts vs gdp etc..

```{r}

rta_vs_gdp %>%
  head()

```

## Add 2015 population data

```{r}
library(WDI)

pop_2015 <- WDI(
  indicator = c(pop_2015 = "SP.POP.TOTL"),
  start = 2015, end = 2015,
  extra = FALSE
) |>
  dplyr::select(ctry_code = iso2c, pop_2015)

rta_vs_gdp <- rta_vs_gdp |>
  dplyr::left_join(pop_2015, by = "ctry_code")

cat("Rows with non-missing population:", sum(!is.na(rta_vs_gdp$pop_2015)), "\n")
head(rta_vs_gdp)
```

## Add World Bank income classification

```{r}
wb_income <- as.data.frame(WDI::WDI_data$country) |>
  dplyr::select(ctry_code = iso2c, income_group = income) |>
  dplyr::filter(income_group != "Aggregates") |>
  dplyr::distinct()

rta_vs_gdp <- rta_vs_gdp |>
  dplyr::left_join(wb_income, by = "ctry_code")

table(rta_vs_gdp$income_group, useNA = "ifany")
head(rta_vs_gdp)
```

## Share of innos in Allinnos: High income vs non-High income

```{r}
rta_vs_gdp |>
  dplyr::mutate(
    group = dplyr::if_else(income_group == "High income",
                           "High income", "Non-high income",
                           missing = "Non-high income")
  ) |>
  dplyr::group_by(group) |>
  dplyr::summarise(
    innos_sum    = sum(innos, na.rm = TRUE),
    allinnos_sum = sum(Allinnos, na.rm = TRUE),
    share        = innos_sum / allinnos_sum
  )
```

## Share of innos in Allinnos: High income / China / Non-high income excl. China

```{r}
rta_vs_gdp |>
  dplyr::mutate(
    group = dplyr::case_when(
      ctry_code == "CN"              ~ "China",
      income_group == "High income"  ~ "High income",
      TRUE                           ~ "Non-high income (excl. China)"
    )
  ) |>
  dplyr::group_by(group) |>
  dplyr::summarise(
    innos_sum    = sum(innos, na.rm = TRUE),
    allinnos_sum = sum(Allinnos, na.rm = TRUE),
    share        = innos_sum / allinnos_sum
  )
```

## Scatter: log GDP per capita vs RTA

```{r}
library(ggplot2)
library(ggrepel)

filtered=rta_vs_gdp %>% filter(#pop_2015>0.5*10^6&
                           Allinnos>200
                             #&!(ctry_code %in% c("QA","SA","KW","MO","OM"))
                             )

ggplot(filtered, aes(y = (log_gdp_pc), x = RTA)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE,
              aes(colour = "Unweighted")) +
  geom_smooth(method = "lm", se = TRUE,
              aes(weight = Allinnos, colour = "Weighted by Allinnos")) +
  scale_colour_manual(name = "Trend",
                      values = c("Unweighted" = "steelblue",
                                 "Weighted by Allinnos" = "firebrick")) +
  geom_text_repel(aes(label = ctry_code),
                  size = 3,
                  max.overlaps = Inf,
                  box.padding = 0.3,
                  point.padding = 0.2,
                  segment.alpha = 0.4) +
  labs(
    x = "log GDP per capita",
    y = "RTA",
    title = "RTA vs log GDP per capita"
  ) +  theme_minimal()
```

## EV-weighted Green Technology RTA (EVRTA)

For each country we compute an EV-weighted share based on `pv + ev_global`
(private value plus external spillover value):

- `EVshare_country = sum(pv+ev_global over distinct green docdbs in country) / sum(pv+ev_global over distinct docdbs in country)`
- `EVshare_global  = sum(pv+ev_global over distinct green docdbs globally) / sum(pv+ev_global over distinct docdbs globally)`
- `EVRTA = EVshare_country / (0.5 * (EVshare_country + EVshare_global))`

**Mapping of innovations to countries** uses the UNION of the harmonized
inventor and holder country parquets (`.bigdata/inventor_countries_harm.parquet`
and `.bigdata/holder_countries_harm.parquet`). **Per-family values** `pv` and
`ev` are read from `innos_istraxfield_global_2009_2018.dsv` (the same source
file Step 10a of `01-build-app-parquets.R` uses). **Green families** are
identified from the shiny DB's `patents_x_tech` / `tech_lookup`.

We also compute a count-based RTA in the same query:
`RTA = share_country / (0.5 * (share_country + share_global))`
where `share = green_fams / total_fams`.

Deduplication is enforced at every step so a docdb is never counted twice.

```{r}
library(duckdb)
library(DBI)
library(data.table)

fromWATSON  <- "C:/Users/rmart/Dropbox/patbis2021/data/fromWATSON"
bigdata_dir <- ".bigdata"

con <- dbConnect(duckdb::duckdb())
dbExecute(con, sprintf(
  "CREATE VIEW pxt AS SELECT * FROM read_parquet('%s')",
  normalizePath("inst/extdata/patents_x_tech.parquet")
))
dbExecute(con, sprintf(
  "CREATE VIEW tl  AS SELECT * FROM read_parquet('%s')",
  normalizePath("inst/extdata/tech_lookup.parquet")
))

# Distinct green docdbs (by tech_group = 'Green Technology')
dbExecute(con, "
  CREATE OR REPLACE VIEW green_fams AS
  SELECT DISTINCT pxt.docdb_family_id
  FROM pxt JOIN tl ON pxt.technology = tl.technology
  WHERE tl.tech_group = 'Green Technology'
")

# Per-family pv and ev from the Watson global istrax source file.
pv_ev_df <- fread(
  file.path(fromWATSON, "innos_istraxfield_global_2009_2018.dsv"),
  select = c("docdb_family_id", "pv", "ev")
)
pv_ev_df <- pv_ev_df[!is.na(docdb_family_id),
                      .(pv = sum(pv, na.rm = TRUE),
                        ev_global = sum(ev, na.rm = TRUE)),
                        by = docdb_family_id]
duckdb::duckdb_register(con, "pv_ev", pv_ev_df)

# Family -> country pairs from harmonized inventor + harmonized holder (union).
inv_harm  <- as.data.table(arrow::read_parquet(
  file.path(bigdata_dir, "inventor_countries_harm.parquet")
))
hold_harm <- as.data.table(arrow::read_parquet(
  file.path(bigdata_dir, "holder_countries_harm.parquet")
))
setnames(inv_harm,  "person_ctry_code", "ctry_code")
setnames(hold_harm, "person_ctry_code", "ctry_code")

pairs_df <- unique(rbindlist(list(
  inv_harm [, .(docdb_family_id, ctry_code)],
  hold_harm[, .(docdb_family_id, ctry_code)]
)))
# Restrict to valid ISO2 codes
valid_iso2 <- unique(na.omit(countrycode::codelist$iso2c))
pairs_df   <- pairs_df[ctry_code %in% valid_iso2]
duckdb::duckdb_register(con, "pairs", pairs_df)

cat("pairs (inventor UNION holder, harmonized):", nrow(pairs_df), "rows,",
    data.table::uniqueN(pairs_df$ctry_code), "countries\n\n")

# Country-level sums (both EV-weighted and count-based).
# pairs already has one row per (family, country) after unique(), so we can
# use SUM for EV and COUNT for the count-based measure without further dedup.
country_shares <- dbGetQuery(con, "
  SELECT
    pairs.ctry_code,
    SUM(CASE WHEN pairs.docdb_family_id IN (SELECT docdb_family_id FROM green_fams)
             THEN (COALESCE(v.pv, 0) + COALESCE(v.ev_global, 0))
             ELSE 0 END) AS green_ev,
    SUM(COALESCE(v.pv, 0) + COALESCE(v.ev_global, 0)) AS total_ev,
    SUM(CASE WHEN pairs.docdb_family_id IN (SELECT docdb_family_id FROM green_fams)
             THEN COALESCE(v.pv, 0) ELSE 0 END) AS green_pv,
    SUM(COALESCE(v.pv, 0)) AS total_pv,
    SUM(CASE WHEN pairs.docdb_family_id IN (SELECT docdb_family_id FROM green_fams)
             THEN 1 ELSE 0 END) AS green_n,
    COUNT(*) AS total_n
  FROM pairs
  inner JOIN pv_ev v USING (docdb_family_id)
  GROUP BY pairs.ctry_code
") |>
  dplyr::mutate(EVshare_country = green_ev / total_ev,
                PVshare_country = green_pv / total_pv,
                share_country   = green_n  / total_n)

# Global sums over distinct families that appear somewhere in pairs.
# (Restricting to families in `pairs` so the global reference matches the
# countrymap.)
global <- dbGetQuery(con, "
  WITH fam_universe AS (SELECT DISTINCT docdb_family_id FROM pairs)
  SELECT
    SUM(CASE WHEN f.docdb_family_id IN (SELECT docdb_family_id FROM green_fams)
             THEN (COALESCE(v.pv, 0) + COALESCE(v.ev_global, 0))
             ELSE 0 END) AS green_ev_global,
    SUM(COALESCE(v.pv, 0) + COALESCE(v.ev_global, 0)) AS total_ev_global,
    SUM(CASE WHEN f.docdb_family_id IN (SELECT docdb_family_id FROM green_fams)
             THEN COALESCE(v.pv, 0) ELSE 0 END) AS green_pv_global,
    SUM(COALESCE(v.pv, 0)) AS total_pv_global,
    SUM(CASE WHEN f.docdb_family_id IN (SELECT docdb_family_id FROM green_fams)
             THEN 1 ELSE 0 END) AS green_n_global,
    COUNT(*) AS total_n_global
  FROM fam_universe f
  inner JOIN pv_ev v USING (docdb_family_id)
")
EVshare_global <- global$green_ev_global / global$total_ev_global
PVshare_global <- global$green_pv_global / global$total_pv_global
share_global   <- global$green_n_global  / global$total_n_global
cat("Global green (pv+ev) share:", signif(EVshare_global, 4), "\n")
cat("Global green (pv only) share:", signif(PVshare_global, 4), "\n")
cat("Global green (count) share:", signif(share_global,   4), "\n\n")

# Symmetric RTAs
evrta <- country_shares |>
  dplyr::mutate(
    EVshare_global = EVshare_global,
    PVshare_global = PVshare_global,
    share_global   = share_global,
    EVRTA = EVshare_country / (0.5 * (EVshare_country + EVshare_global)),
    PVRTA = PVshare_country / (0.5 * (PVshare_country + PVshare_global)),
    RTA   = share_country   / (0.5 * (share_country   + share_global))
  ) |>
  dplyr::arrange(dplyr::desc(EVRTA))

dbDisconnect(con, shutdown = TRUE)

head(evrta, 15)
tail(evrta, 15)
```

## Scatter: log GDP per capita vs EVRTA

Same style as the RTA scatter — two trendlines (unweighted and weighted by `innos`),
country-code labels with non-overlapping placement, same population / Allinnos
filters.

```{r}
library(ggplot2)
library(ggrepel)

# Attach log_gdp_pc, pop_2015 and Allinnos from rta_vs_gdp
evrta_gdp <- evrta |>
  dplyr::left_join(
    rta_vs_gdp |> dplyr::select(ctry_code, log_gdp_pc, pop_2015, innos, Allinnos),
    by = "ctry_code"
  )

filtered_ev <- evrta_gdp |>
  dplyr::filter(!is.na(log_gdp_pc),
                #pop_2015 > 1e6,
                Allinnos > 50)

ggplot(filtered_ev, aes(y = log_gdp_pc, x = EVRTA)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE,
              aes(colour = "Unweighted")) +
  geom_smooth(method = "lm", se = TRUE,
              aes(weight = Allinnos, colour = "Weighted by Allinnos")) +
  scale_colour_manual(name = "Trend",
                      values = c("Unweighted" = "steelblue",
                                 "Weighted by Allinnos" = "firebrick")) +
  geom_text_repel(aes(label = ctry_code),
                  size = 3,
                  max.overlaps = Inf,
                  box.padding = 0.3,
                  point.padding = 0.2,
                  segment.alpha = 0.4) +
  labs(
    x = "EVRTA (Green Technology, EV-weighted)",
    y = "log GDP per capita",
    title = "log GDP per capita vs EVRTA"
  ) +
  theme_minimal()
```

## Scatter: log GDP per capita vs count-based RTA

Same filters and styling as the EVRTA scatter, but plotting the count-based
RTA (share of green families in total families per country) on the x-axis.

```{r}
ggplot(filtered_ev, aes(y = log_gdp_pc, x = RTA)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE,
              aes(colour = "Unweighted")) +
  geom_smooth(method = "lm", se = TRUE,
              aes(weight = Allinnos, colour = "Weighted by Allinnos")) +
  scale_colour_manual(name = "Trend",
                      values = c("Unweighted" = "steelblue",
                                 "Weighted by Allinnos" = "firebrick")) +
  geom_text_repel(aes(label = ctry_code),
                  size = 3,
                  max.overlaps = Inf,
                  box.padding = 0.3,
                  point.padding = 0.2,
                  segment.alpha = 0.4) +
  labs(
    x = "RTA (Green Technology, count-based)",
    y = "log GDP per capita",
    title = "log GDP per capita vs count-based RTA"
  ) +
  theme_minimal()
```

## Scatter: log GDP per capita vs PV-weighted RTA (PVRTA)

Weighting by private value only (`pv`), excluding the external spillover
component `ev_global`.

```{r}
ggplot(filtered_ev, aes(y = log_gdp_pc, x = PVRTA)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE,
              aes(colour = "Unweighted")) +
  geom_smooth(method = "lm", se = TRUE,
              aes(weight = Allinnos, colour = "Weighted by Allinnos")) +
  scale_colour_manual(name = "Trend",
                      values = c("Unweighted" = "steelblue",
                                 "Weighted by Allinnos" = "firebrick")) +
  geom_text_repel(aes(label = ctry_code),
                  size = 3,
                  max.overlaps = Inf,
                  box.padding = 0.3,
                  point.padding = 0.2,
                  segment.alpha = 0.4) +
  labs(
    x = "PVRTA (Green Technology, PV-weighted)",
    y = "log GDP per capita",
    title = "log GDP per capita vs PVRTA"
  ) +
  theme_minimal()
```

## Scatter: count-based RTA vs EVRTA

Comparison of the two measures. The 45° reference line (`y = x`) marks perfect
agreement; deviations indicate that EV weighting (by `pv + ev_global`) shifts
the country's apparent green specialization relative to simple counts.

```{r}
ggplot(filtered_ev, aes(x = RTA, y = EVRTA)) +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "grey50") +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE,
              aes(colour = "Unweighted")) +
  geom_smooth(method = "lm", se = TRUE,
              aes(weight = Allinnos, colour = "Weighted by Allinnos")) +
  scale_colour_manual(name = "Trend",
                      values = c("Unweighted" = "steelblue",
                                 "Weighted by Allinnos" = "firebrick")) +
  geom_text_repel(aes(label = ctry_code),
                  size = 3,
                  max.overlaps = Inf,
                  box.padding = 0.3,
                  point.padding = 0.2,
                  segment.alpha = 0.4) +
  labs(
    x = "RTA (count-based)",
    y = "EVRTA (EV-weighted)",
    title = "EVRTA vs count-based RTA"
  ) +
  theme_minimal()
```

## Compare shiny-app RTA (with holders & inventors) vs chunk-computed RTA

Load the shiny-app-exported scatter-data CSV and compare its per-country `RTA`
with the count-based `RTA` we compute in the EVRTA chunk above.

```{r}
library(readr)
library(dplyr)

shiny_rta <- read_csv(
  ".bigdata/rta_vs_gdp_2026-04-18_updated_with_holders_and_innovators.csv"
) |>
  dplyr::select(ctry_code,
                innos_shiny     = innos,
                Allinnos_shiny  = Allinnos,
                share_c_shiny   = share_c,
                share_shiny     = share,
                RTA_shiny       = RTA)

compare <- dplyr::inner_join(
    evrta |>
      dplyr::select(ctry_code,
                    green_n, total_n,
                    share_country, share_global, RTA),
    shiny_rta,
    by = "ctry_code"
  ) |>
  dplyr::mutate(
    diff_greens  = green_n  - innos_shiny,
    diff_totals  = total_n  - Allinnos_shiny,
    diff_sharec  = share_country - share_c_shiny,
    diff_rta     = RTA - RTA_shiny
  ) |>
  dplyr::arrange(dplyr::desc(abs(diff_rta)))

cat("Shape-level check:\n")
cat("  countries in shiny CSV:", nrow(shiny_rta), "\n")
cat("  countries in compare  :", nrow(compare),   "\n\n")

cat("Global 'share' (green / total) chunk vs shiny:\n")
cat("  chunk share_global:", unique(compare$share_global), "\n")
cat("  shiny 'share'     :", unique(compare$share_shiny),  "\n\n")

cat("Largest RTA discrepancies:\n")
print(head(compare, 20))

cat("\nSummary of absolute RTA differences:\n")
print(summary(abs(compare$diff_rta)))
```


```{r}

acompare=compare %>% summarize(sum(Allinnos_shiny),sum(total_n))

```
