---
title: "Test: national istrax for developing countries"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fst)
library(arrow)
library(dplyr)
library(readr)
```

## Load RTA vs GDP data

```{r}
rta_vs_gdp <- read_csv(".bigdata/rta_vs_gdp_2026-04-17.csv")
cat("rta_vs_gdp rows:", nrow(rta_vs_gdp), "\n")
head(rta_vs_gdp)
```

## Load PRINZglobal docdb counts

```{r}
count_docdb <- read_csv("C:/Users/rmart/OneDrive/PRINZglobal/data/count_docdb_2024-11-21.csv")
cat("count_docdb rows:", nrow(count_docdb), "\n")
cat("columns:", paste(names(count_docdb), collapse = ", "), "\n")
head(count_docdb)
count_docdb %>% distinct(technology) %>% pull(technology)
```

## Green technology: pre/post 2010 x income group

```{r}
count_docdb |>
  dplyr::filter(technology == "Green technology") |>
  dplyr::mutate(
    period = dplyr::if_else(earliest_filing_year < 2010, "<2010", ">=2010"),
    income_grp = dplyr::case_when(
      person_ctry_code == "CN"           ~ "China",
      `Income group` == "High income"    ~ "High income",
      TRUE                               ~ "Non-high income (excl. China)"
    )
  ) |>
  dplyr::group_by(period, income_grp) |>
  dplyr::summarise(
    tech_sum = sum(docdb_family_count_tech, na.rm = TRUE),
    all_sum  = sum(docdb_family_count,      na.rm = TRUE),
    share    = tech_sum / all_sum,
    .groups  = "drop"
  ) |>
  dplyr::arrange(period, income_grp)
```

## Compare countries present in count_docdb vs rta_vs_gdp

```{r}
library(countrycode)

code_to_name <- function(codes) {
  nm <- countrycode(codes, origin = "iso2c", destination = "country.name", warn = FALSE)
  paste0(codes, " (", ifelse(is.na(nm), "??", nm), ")")
}

cd_countries  <- sort(unique(na.omit(count_docdb$person_ctry_code)))
rvg_countries <- sort(unique(na.omit(rta_vs_gdp$ctry_code)))

cat("count_docdb countries: ", length(cd_countries), "\n")
cat("rta_vs_gdp countries:  ", length(rvg_countries), "\n")
cat("Overlap:               ", length(intersect(cd_countries, rvg_countries)), "\n\n")

cat("In count_docdb only (not in rta_vs_gdp):\n")
print(code_to_name(setdiff(cd_countries, rvg_countries)))

cat("\nIn rta_vs_gdp only (not in count_docdb):\n")
print(code_to_name(setdiff(rvg_countries, cd_countries)))
```

## Compare Green Technology mapping: PRINZglobal vs Shiny DB

```{r}
# PRINZglobal mapping (all rows are "Green Technology")
green_prinz <- read_csv("C:/Users/rmart/OneDrive/PRINZglobal/data/docdb_green_or_not.csv") |>
  dplyr::filter(green_or_not == "Green Technology") |>
  dplyr::distinct(docdb_family_id)

# Shiny DB mapping: every family in patents_x_tech whose technology belongs
# to tech_group "Green Technology"
patents_x_tech <- arrow::read_parquet("inst/extdata/patents_x_tech.parquet")
tech_lookup    <- arrow::read_parquet("inst/extdata/tech_lookup.parquet")

green_shiny <- patents_x_tech |>
  dplyr::inner_join(tech_lookup, by = "technology") |>
  dplyr::filter(tech_group == "Green Technology") |>
  dplyr::distinct(docdb_family_id)

cat("PRINZglobal green docdbs: ", nrow(green_prinz), "\n")
cat("Shiny DB green docdbs:    ", nrow(green_shiny), "\n")
cat("Overlap:                  ",
    length(intersect(green_prinz$docdb_family_id, green_shiny$docdb_family_id)), "\n\n")

# In PRINZglobal only (missing from Shiny DB)
only_prinz <- dplyr::anti_join(green_prinz, green_shiny, by = "docdb_family_id")
cat("In PRINZglobal only: ", nrow(only_prinz), "\n")
head(only_prinz)

# In Shiny DB only (missing from PRINZglobal)
only_shiny <- dplyr::anti_join(green_shiny, green_prinz, by = "docdb_family_id")
cat("\nIn Shiny DB only:    ", nrow(only_shiny), "\n")
head(only_shiny)
```

## Compare country-level counts: count_docdb (2009-2018, Green) vs rta_vs_gdp

```{r}
cd_2009_2018 <- count_docdb |>
  dplyr::filter(technology == "Green technology",
                earliest_filing_year >= 2009,
                earliest_filing_year <= 2018) |>
  dplyr::group_by(ctry_code = person_ctry_code) |>
  dplyr::summarise(
    tech_sum = sum(docdb_family_count_tech, na.rm = TRUE),
    all_sum  = sum(docdb_family_count,      na.rm = TRUE),
    .groups = "drop"
  )

comparison <- dplyr::full_join(
    cd_2009_2018,
    rta_vs_gdp |> dplyr::select(ctry_code, innos, Allinnos),
    by = "ctry_code"
  ) |>
  dplyr::mutate(
    country_name = countrycode::countrycode(ctry_code, origin = "iso2c",
                                            destination = "country.name",
                                            warn = FALSE),
    diff_tech = tech_sum - innos,
    diff_all  = all_sum - Allinnos,
    ratio_tech = tech_sum / innos,
    ratio_all  = all_sum  / Allinnos
  ) |>
  dplyr::select(ctry_code, country_name, dplyr::everything()) |>
  dplyr::arrange(dplyr::desc(abs(diff_all)))

cat("Totals across countries:\n")
cat("  count_docdb tech_sum total: ", sum(comparison$tech_sum, na.rm = TRUE), "\n")
cat("  rta_vs_gdp innos total:     ", sum(comparison$innos,    na.rm = TRUE), "\n")
cat("  count_docdb all_sum total:  ", sum(comparison$all_sum,  na.rm = TRUE), "\n")
cat("  rta_vs_gdp Allinnos total:  ", sum(comparison$Allinnos, na.rm = TRUE), "\n\n")

comparison
```

## Green technology: 2009-2018 x income group

```{r}
count_docdb |>
  dplyr::filter(technology == "Green technology",
                earliest_filing_year >= 2009,
                earliest_filing_year <= 2018) |>
  dplyr::mutate(
    income_grp = dplyr::case_when(
      person_ctry_code == "CN"           ~ "China",
      `Income group` == "High income"    ~ "High income",
      TRUE                               ~ "Non-high income (excl. China)"
    )
  ) |>
  dplyr::group_by(income_grp) |>
  dplyr::summarise(
    tech_sum = sum(docdb_family_count_tech, na.rm = TRUE),
    all_sum  = sum(docdb_family_count,      na.rm = TRUE),
    share    = tech_sum / all_sum,
    .groups  = "drop"
  ) |>
  dplyr::arrange(income_grp)
```

## Load data

```{r}
istrax <- read_fst(".bigdata/istraxes/istrax_nationalkey_2009_2018.fst")
country_lookup <- read_parquet("inst/extdata/country_lookup.parquet")

cat("istrax rows:", nrow(istrax), "\n")
cat("country_lookup rows:", nrow(country_lookup), "\n")
```

## Filter to LMICs excluding China

```{r}
lmic_codes <- country_lookup |>
  filter(is_lmic_excl_china) |>
  pull(ctry_code)

istrax_lmic <- istrax |>
  filter(ctry_code %in% lmic_codes)

cat("LMIC (excl. China) rows:", nrow(istrax_lmic), "\n")
cat("LMIC (excl. China) countries in istrax:", n_distinct(istrax_lmic$ctry_code), "\n")
head(istrax_lmic)
```

## Summary by country

```{r}
istrax_lmic |>
  group_by(ctry_code) |>
  summarise(
    n = n(),
    mean_istrax = mean(istrax_nationalkey_2009_2018, na.rm = TRUE),
    median_istrax = median(istrax_nationalkey_2009_2018, na.rm = TRUE)
  ) |>
  arrange(desc(n))
```


## rts vs gdp etc..

```{r}

rta_vs_gdp %>%
  head()

```

## Add 2015 population data

```{r}
library(WDI)

pop_2015 <- WDI(
  indicator = c(pop_2015 = "SP.POP.TOTL"),
  start = 2015, end = 2015,
  extra = FALSE
) |>
  dplyr::select(ctry_code = iso2c, pop_2015)

rta_vs_gdp <- rta_vs_gdp |>
  dplyr::left_join(pop_2015, by = "ctry_code")

cat("Rows with non-missing population:", sum(!is.na(rta_vs_gdp$pop_2015)), "\n")
head(rta_vs_gdp)
```

## Add World Bank income classification

```{r}
wb_income <- as.data.frame(WDI::WDI_data$country) |>
  dplyr::select(ctry_code = iso2c, income_group = income) |>
  dplyr::filter(income_group != "Aggregates") |>
  dplyr::distinct()

rta_vs_gdp <- rta_vs_gdp |>
  dplyr::left_join(wb_income, by = "ctry_code")

table(rta_vs_gdp$income_group, useNA = "ifany")
head(rta_vs_gdp)
```

## Share of innos in Allinnos: High income vs non-High income

```{r}
rta_vs_gdp |>
  dplyr::mutate(
    group = dplyr::if_else(income_group == "High income",
                           "High income", "Non-high income",
                           missing = "Non-high income")
  ) |>
  dplyr::group_by(group) |>
  dplyr::summarise(
    innos_sum    = sum(innos, na.rm = TRUE),
    allinnos_sum = sum(Allinnos, na.rm = TRUE),
    share        = innos_sum / allinnos_sum
  )
```

## Share of innos in Allinnos: High income / China / Non-high income excl. China

```{r}
rta_vs_gdp |>
  dplyr::mutate(
    group = dplyr::case_when(
      ctry_code == "CN"              ~ "China",
      income_group == "High income"  ~ "High income",
      TRUE                           ~ "Non-high income (excl. China)"
    )
  ) |>
  dplyr::group_by(group) |>
  dplyr::summarise(
    innos_sum    = sum(innos, na.rm = TRUE),
    allinnos_sum = sum(Allinnos, na.rm = TRUE),
    share        = innos_sum / allinnos_sum
  )
```

## Scatter: log GDP per capita vs RTA

```{r}
library(ggplot2)
library(ggrepel)

filtered=rta_vs_gdp %>% filter(pop_2015>1*10^6& Allinnos>50
                             #&!(ctry_code %in% c("QA","SA","KW","MO","OM"))
                             )

ggplot(filtered, aes(y = (log_gdp_pc), x = RTA)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  geom_text_repel(aes(label = ctry_code),
                  size = 3,
                  max.overlaps = Inf,
                  box.padding = 0.3,
                  point.padding = 0.2,
                  segment.alpha = 0.4) +
  labs(
    x = "log GDP per capita",
    y = "RTA",
    title = "RTA vs log GDP per capita"
  ) +  theme_minimal()
```
