---
title: "Prepare DuckDB Database for ISEApp"
output: html_notebook
---

This notebook reads the source FST/parquet files from the Dropbox `Apps/iseapp` folder
and creates a single DuckDB database file at `{iseapp_local_path}/duck/iseapp.duckdb`.

The resulting database contains all tables needed by the DuckDB-backed Shiny app
(`duck/app.R`), with all data properly normalized and ready for SQL queries.

## Setup

```{r setup}
library(fst)
library(arrow)
library(dplyr)
library(data.table)
library(duckdb)

# Detect Dropbox iseapp path (same logic as app.R)
get_dropbox_path <- function() {
  if (.Platform$OS.type == "windows") {
    info_paths <- c(
      file.path(Sys.getenv("APPDATA"), "Dropbox", "info.json"),
      file.path(Sys.getenv("LOCALAPPDATA"), "Dropbox", "info.json")
    )
  } else {
    info_paths <- file.path(Sys.getenv("HOME"), ".dropbox", "info.json")
  }
  for (p in info_paths) {
    if (file.exists(p)) {
      info <- jsonlite::fromJSON(p)
      if (!is.null(info$business)) return(info$business$path)
      if (!is.null(info$personal)) return(info$personal$path)
    }
  }
  return(NULL)
}

get_local_iseapp_path <- function() {
  env_path <- Sys.getenv("ISEAPP_PATH_LOCAL")
  if (nzchar(env_path) && dir.exists(env_path)) return(env_path)
  dropbox_root <- get_dropbox_path()
  if (!is.null(dropbox_root)) {
    iseapp_path <- file.path(dropbox_root, "Apps", "iseapp")
    if (dir.exists(iseapp_path)) return(iseapp_path)
  }
  return(NULL)
}

iseapp_path <- get_local_iseapp_path()
stopifnot("Cannot find local iseapp path" = !is.null(iseapp_path))
message("Using iseapp data path (Dropbox): ", iseapp_path)

# DuckDB output path — stored in the repo's duck/ folder (not Dropbox)
# to avoid cloud sync corruption issues
duck_dir <- getwd()  # Should be the duck/ directory when knitting this Rmd
message("DuckDB output directory: ", duck_dir)

duck_db_path <- file.path(duck_dir, "iseapp.duckdb")
```

```{r}
# Remove existing database to start fresh
if (file.exists(duck_db_path)) {
  message("Removing existing database: ", duck_db_path)
  file.remove(duck_db_path)
}
```

## Create DuckDB Database

```{r create-db}
con <- dbConnect(duckdb(), dbdir = duck_db_path)
#con <- dbConnect(duckdb(), dbdir = duck_db_path)
message("Created DuckDB database at: ", duck_db_path)
```

## 1. Load and write countrymap

The countrymap contains the patent-country mapping with `appln_id` (application ID).

```{r countrymap}
countrymap <- read_fst(file.path(iseapp_path, "countrymap.fst"))
message("Countrymap: ", nrow(countrymap), " rows, columns: ", paste(names(countrymap), collapse = ", "))

dbWriteTable(con, "countrymap", countrymap, overwrite = TRUE)
message("Written countrymap to DuckDB")
```

## 2. Load and write techmap (with normalization)

The techmap maps patents to technology categories. We normalize category names
and add an "All" category (one row per distinct docdb_family_id in countrymap).

```{r techmap}
techmap_raw <- read_fst(file.path(iseapp_path, "techmap.fst"))
message("Techmap raw: ", nrow(techmap_raw), " rows")

# Add "All" technology from countrymap (every patent belongs to "All")
techmap_all <- countrymap %>%
  select(docdb_family_id) %>%
  distinct() %>%
  mutate(technology = "All")

techmap_processed <- bind_rows(techmap_raw, techmap_all)

# Normalize technology names (matching process_techmap() in app.R)
setDT(techmap_processed)
techmap_processed[, technology := fcase(
  technology == "Any Green technology", "Green Technology",
  technology == "Any battery technology", "Battery Technology",
  technology == "Any Hard to Abate technology", "Hard to Abate Sector Decarbonization",
  default = technology
)]

# Remove duplicates
techmap_processed <- unique(techmap_processed)
message("Techmap processed: ", nrow(techmap_processed), " rows, ",
        length(unique(techmap_processed$technology)), " unique technologies")

dbWriteTable(con, "techmap", as.data.frame(techmap_processed), overwrite = TRUE)
message("Written techmap to DuckDB")
```

## 3. Load and write regionmap

```{r regionmap}
regionmap_path <- file.path(iseapp_path, "regionmap.fst")
if (file.exists(regionmap_path)) {
  regionmap <- read_fst(regionmap_path)
  message("Regionmap: ", nrow(regionmap), " rows, columns: ", paste(names(regionmap), collapse = ", "))
  dbWriteTable(con, "regionmap", regionmap, overwrite = TRUE)
  message("Written regionmap to DuckDB")
} else {
  message("WARNING: regionmap.fst not found, skipping")
}
```

## 4. Load and write firmmap + firmsectormap

```{r firmmap}
firmmap_path <- file.path(iseapp_path, "firmmap.parquet")
if (file.exists(firmmap_path)) {
  firmmap <- arrow::read_parquet(firmmap_path)
  message("Firmmap: ", nrow(firmmap), " rows")
  dbWriteTable(con, "firmmap", as.data.frame(firmmap), overwrite = TRUE)
  message("Written firmmap to DuckDB")
} else {
  message("WARNING: firmmap.parquet not found, skipping")
}

firmsectormap_path <- file.path(iseapp_path, "firmsectormap.parquet")
if (file.exists(firmsectormap_path)) {
  firmsectormap <- arrow::read_parquet(firmsectormap_path)
  message("Firmsectormap: ", nrow(firmsectormap), " rows")
  dbWriteTable(con, "firmsectormap", as.data.frame(firmsectormap), overwrite = TRUE)
  message("Written firmsectormap to DuckDB")
} else {
  message("WARNING: firmsectormap.parquet not found, skipping")
}
```

## 5. Load all istrax flow files into a single long-format table

Each FST file in `istraxes/` has columns: `docdb_family_id, ctry_code, {flow_name}`.
We combine them all into one table: `istrax_values(docdb_family_id, ctry_code, flow_type, value)`.

```{r istrax-values}
istrax_dir <- file.path(iseapp_path, "istraxes")
flow_files <- list.files(istrax_dir, pattern = "\\.fst$", full.names = FALSE)
message("Found ", length(flow_files), " flow files")

# Create the table with the first file, then append the rest
first <- TRUE

for (ff in flow_files) {
  flow_type <- tools::file_path_sans_ext(ff)
  full_path <- file.path(istrax_dir, ff)

  df <- read_fst(full_path)

  # The value column is named after the flow_type
  # Columns are: docdb_family_id, ctry_code, {flow_type}
  value_col <- setdiff(names(df), c("docdb_family_id", "ctry_code"))

  if (length(value_col) != 1) {
    message("WARNING: Unexpected columns in ", ff, ": ", paste(names(df), collapse = ", "))
    next
  }

  # Rename to standard format
  df_long <- data.frame(
    docdb_family_id = df$docdb_family_id,
    ctry_code = df$ctry_code,
    flow_type = flow_type,
    value = df[[value_col]],
    stringsAsFactors = FALSE
  )

  # Replace NAs with 0 (matching current app behavior)
  df_long$value[is.na(df_long$value)] <- 0

  if (first) {
    dbWriteTable(con, "istrax_values", df_long, overwrite = TRUE)
    first <- FALSE
  } else {
    dbAppendTable(con, "istrax_values", df_long)
  }

  message("  Loaded ", ff, ": ", nrow(df_long), " rows (flow_type = ", flow_type, ")")
}

# Check total rows
total <- dbGetQuery(con, "SELECT COUNT(*) AS n FROM istrax_values")
message("Total istrax_values rows: ", total$n)

# Check distinct flow types
flows <- dbGetQuery(con, "SELECT DISTINCT flow_type FROM istrax_values ORDER BY flow_type")
message("Flow types (", nrow(flows), "): ", paste(flows$flow_type, collapse = ", "))
```

## 6. Create country_groups table

Pre-expanded country group definitions for efficient SQL lookups.

```{r country-groups}
all_countries <- sort(unique(na.omit(countrycode::codelist$iso2c)))

lmics <- c("AF","AL","DZ","AO","AR","AM","AZ","BD","BJ","BO","BA","BW","BR","BG",
           "BF","BI","KH","CM","CV","CF","TD","CL","CN","CO","KM","CG","CR","CI",
           "CU","DJ","DM","DO","EC","EG","SV","GQ","ER","ET","FJ","GA","GM","GE",
           "GH","GT","GN","GW","GY","HT","HN","IN","ID","IR","IQ","JM","JO","KZ",
           "KE","KI","KP","KG","LA","LB","LS","LR","LY","MG","MW","MY","MV","ML",
           "MR","MU","MX","MD","MN","ME","MA","MZ","MM","NA","NP","NI","NE","NG",
           "MK","PK","PW","PA","PG","PY","PE","PH","RW","WS","ST","SN","RS","SC",
           "SL","SB","SO","ZA","LK","SD","SR","SY","TJ","TZ","TH","TL","TG","TO",
           "TN","TR","TM","TV","UG","UA","UZ","VU","VE","VN","YE","ZM","ZW")

lmics_excl_china <- setdiff(lmics, "CN")
eu_countries <- c("AT","BE","BG","HR","CY","CZ","DK","EE","FI","FR","DE","GR",
                  "HU","IE","IT","LV","LT","LU","MT","NL","PL","PT","RO","SK",
                  "SI","ES","SE")
hic <- setdiff(all_countries, lmics)

group_definitions <- list(
  "All countries" = all_countries,
  "LMICs" = lmics,
  "LMICs (excl. China)" = lmics_excl_china,
  "EU countries" = eu_countries,
  "High income countries" = hic
)

# Build the long-format table
groups_list <- lapply(names(group_definitions), function(gname) {
  data.frame(
    group_name = gname,
    ctry_code = group_definitions[[gname]],
    stringsAsFactors = FALSE
  )
})
country_groups_df <- do.call(rbind, groups_list)

dbWriteTable(con, "country_groups", country_groups_df, overwrite = TRUE)
message("Written country_groups: ", nrow(country_groups_df), " rows, ",
        length(unique(country_groups_df$group_name)), " groups")
```

## 7. Verify database

```{r verify}
tables <- dbListTables(con)
message("Tables in DuckDB: ", paste(tables, collapse = ", "))

for (tbl in tables) {
  n <- dbGetQuery(con, paste0("SELECT COUNT(*) AS n FROM ", tbl))
  message("  ", tbl, ": ", n$n, " rows")
}

# Spot checks
message("\nSpot checks:")

# Check a sample query matching compute_avstrax_for_techs logic
sample_q <- dbGetQuery(con, "
  SELECT iv.ctry_code,
         AVG(iv.value) AS mean_val,
         COUNT(*) AS n
  FROM istrax_values iv
  WHERE iv.flow_type = 'istrax_global'
    AND iv.ctry_code IN ('US', 'GB', 'DE', 'FR', 'CN')
  GROUP BY iv.ctry_code
  ORDER BY iv.ctry_code
")
message("Sample aggregation (istrax_global, 5 countries):")
print(sample_q)

# Check techmap join
tech_q <- dbGetQuery(con, "
  SELECT t.technology, COUNT(DISTINCT t.docdb_family_id) AS n_patents
  FROM techmap t
  GROUP BY t.technology
  ORDER BY n_patents DESC
  LIMIT 10
")
message("\nTop 10 technology categories by patent count:")
print(tech_q)
```

## 8. Create indexes

Add indexes on key join/filter columns to speed up runtime queries.
DuckDB uses ART (Adaptive Radix Tree) indexes for equality lookups and joins.

```{r indexes}
message("Creating indexes...")

# istrax_values: every query filters on flow_type, then joins on (docdb_family_id, ctry_code)
dbExecute(con, "CREATE INDEX idx_istrax_flow ON istrax_values(flow_type)")
dbExecute(con, "CREATE INDEX idx_istrax_docdb_ctry ON istrax_values(docdb_family_id, ctry_code)")

# countrymap: join key
dbExecute(con, "CREATE INDEX idx_countrymap_docdb_ctry ON countrymap(docdb_family_id, ctry_code)")

# techmap: join on docdb_family_id, filter on technology
dbExecute(con, "CREATE INDEX idx_techmap_docdb ON techmap(docdb_family_id)")
dbExecute(con, "CREATE INDEX idx_techmap_tech ON techmap(technology)")

# regionmap: join key + region filter
if ("regionmap" %in% dbListTables(con)) {
  dbExecute(con, "CREATE INDEX idx_regionmap_docdb_ctry ON regionmap(docdb_family_id, ctry_code)")
  dbExecute(con, "CREATE INDEX idx_regionmap_region ON regionmap(region_code)")
}

# firmmap: subquery lookup
if ("firmmap" %in% dbListTables(con)) {
  dbExecute(con, "CREATE INDEX idx_firmmap_docdb ON firmmap(docdb_family_id)")
  dbExecute(con, "CREATE INDEX idx_firmmap_company ON firmmap(company_raw)")
}

message("Indexes created")
```

## 9. Create pre-joined tables

Pre-join `istrax_values` with `countrymap` and `regionmap` to eliminate the most
expensive runtime joins. The ORDER BY ensures DuckDB's zone maps (min-max indexes
on row groups) are maximally effective for flow_type filtering.

```{r pre-joined-country}
message("Creating pre-joined istrax_country table...")

dbExecute(con, "
  CREATE TABLE istrax_country AS
  SELECT DISTINCT
    iv.docdb_family_id,
    iv.ctry_code,
    cm.appln_id,
    iv.flow_type,
    iv.value
  FROM istrax_values iv
  JOIN countrymap cm
    ON iv.docdb_family_id = cm.docdb_family_id
    AND iv.ctry_code = cm.ctry_code
  ORDER BY iv.flow_type, iv.ctry_code, iv.docdb_family_id
")

n_ic <- dbGetQuery(con, "SELECT COUNT(*) AS n FROM istrax_country")$n
message("istrax_country: ", n_ic, " rows")

# Index the pre-joined table
dbExecute(con, "CREATE INDEX idx_ic_flow ON istrax_country(flow_type)")
dbExecute(con, "CREATE INDEX idx_ic_ctry ON istrax_country(ctry_code)")
dbExecute(con, "CREATE INDEX idx_ic_docdb ON istrax_country(docdb_family_id)")
message("istrax_country indexes created")
```

```{r pre-joined-region}
if ("regionmap" %in% dbListTables(con)) {
  message("Creating pre-joined istrax_region table...")

  dbExecute(con, "
    CREATE TABLE istrax_region AS
    SELECT DISTINCT
      iv.docdb_family_id,
      rm.region_code,
      rm.region_name,
      rm.appln_id,
      iv.flow_type,
      iv.value
    FROM istrax_values iv
    JOIN regionmap rm
      ON iv.docdb_family_id = rm.docdb_family_id
      AND iv.ctry_code = rm.ctry_code
    ORDER BY iv.flow_type, rm.region_code, iv.docdb_family_id
  ")

  n_ir <- dbGetQuery(con, "SELECT COUNT(*) AS n FROM istrax_region")$n
  message("istrax_region: ", n_ir, " rows")

  dbExecute(con, "CREATE INDEX idx_ir_flow ON istrax_region(flow_type)")
  dbExecute(con, "CREATE INDEX idx_ir_region ON istrax_region(region_code)")
  dbExecute(con, "CREATE INDEX idx_ir_docdb ON istrax_region(docdb_family_id)")
  message("istrax_region indexes created")
} else {
  message("Skipping istrax_region (no regionmap)")
}
```

## 10. Pre-compute patent counts per country/region

The RTA calculation needs `Allinnos` (total distinct patents per country, unfiltered
by technology). Pre-computing avoids expensive COUNT DISTINCT queries at runtime.

```{r patent-counts}
message("Creating pre-computed patent count tables...")

dbExecute(con, "
  CREATE TABLE country_patent_counts AS
  SELECT ctry_code, COUNT(DISTINCT docdb_family_id) AS Allinnos
  FROM countrymap
  GROUP BY ctry_code
")
n_cpc <- dbGetQuery(con, "SELECT COUNT(*) AS n FROM country_patent_counts")$n
message("country_patent_counts: ", n_cpc, " rows")

if ("regionmap" %in% dbListTables(con)) {
  dbExecute(con, "
    CREATE TABLE region_patent_counts AS
    SELECT region_code AS ctry_code, COUNT(DISTINCT docdb_family_id) AS Allinnos
    FROM regionmap
    GROUP BY region_code
  ")
  n_rpc <- dbGetQuery(con, "SELECT COUNT(*) AS n FROM region_patent_counts")$n
  message("region_patent_counts: ", n_rpc, " rows")
} else {
  message("Skipping region_patent_counts (no regionmap)")
}
```

## 11. Lookup tables for SQL-complete queries

These small lookup tables enable the Shiny app's SQL queries to return fully
plot-ready dataframes without any R-side post-processing.

### 11a. Technology → greenclass mapping

```{r create-tech-greenclass}
# Technology class definitions (same as app.R)
green_classes <- c("Green Technology","Green Energy", "Green Transport", "Circular Economy",
                   "Green Manufacturing", "Adaptation", "Green Housing", "Green ICT",
                   "Green Agriculture", "GHG Capture")

battery_classes <- c("Battery Technology", "Lithium Extraction & Processing",
                     "Graphite & Carbon Materials", "Cathode Materials", "Anode Materials",
                     "Electrolytes & Additives", "Separators", "Battery Cell Design & Assembly",
                     "Battery Management Systems (BMS)", "Electric Vehicles & Mobility",
                     "Battery Recycling & Recovery")

hard_to_abate_classes <- c("Hard to Abate Sector Decarbonization", "Aviation Decarbonisation",
                           "Cement & Concrete Decarbonisation", "Chemicals & Plastics Decarbonisation",
                           "Shipping Decarbonisation", "Steel & Iron Decarbonisation")

ai_classes <- c("AI", "Machine Learning", "Deep Learning", "Natural Language Processing (NLP)",
                "Computer Vision", "Speech Recognition & Synthesis", "Robotics & Autonomous Systems",
                "Knowledge Representation & Reasoning", "Planning & Decision Making", "Generative AI",
                "Semiconductors", "Cloud & Data Infrastructure", "Data Rettrieval & Processing System",
                "Platform & Frameworks", "Deployment & Support")

cpc_sections <- c("Human Necessities", "Performing Operations; Transporting ",
                  "Chemistry; Metallurgy ", "Textiles; Paper", "Fixed Constructions",
                  "Mechanical Engineering; Lighting; Heating; Weapons; Blasting",
                  "Physics", "Electricity",
                  "General tagging of new or cross-sectional technology")

agrifood_classes <- c("Any Agriculture & Food technology", "Input supply",
                      "Primary food and feed production", "Post-harvest handling & aggregation",
                      "Processing", "Distribution/wholesale", "Retail/consumption", "Crosscutting")

tech_greenclass <- data.frame(
  technology = c(green_classes, battery_classes, hard_to_abate_classes,
                 ai_classes, cpc_sections, agrifood_classes, "All"),
  greenclass = c(rep("green", length(green_classes)),
                 rep("battery", length(battery_classes)),
                 rep("hard to abate", length(hard_to_abate_classes)),
                 rep("AI", length(ai_classes)),
                 rep("CPC Sections", length(cpc_sections)),
                 rep("agrifood", length(agrifood_classes)),
                 "other"),
  stringsAsFactors = FALSE
)

dbWriteTable(con, "tech_greenclass", tech_greenclass, overwrite = TRUE)
message("tech_greenclass: ", nrow(tech_greenclass), " rows")
```

### 11b. Country code → English name mapping

```{r create-country-names}
library(countrycode)

# Get all ISO2 codes from the database
all_codes <- dbGetQuery(con, "SELECT DISTINCT ctry_code FROM countrymap ORDER BY ctry_code")$ctry_code
country_df <- data.frame(
  ctry_code = all_codes,
  country_name = countrycode(all_codes, "iso2c", "country.name.en", warn = FALSE),
  stringsAsFactors = FALSE
)
# Use code itself as fallback for unrecognised codes
country_df$country_name[is.na(country_df$country_name)] <-
  country_df$ctry_code[is.na(country_df$country_name)]

# Add UK NUTS1 region codes (override any existing entries)
uk_regions <- data.frame(
  ctry_code = c("UKC","UKD","UKE","UKF","UKG","UKH","UKI","UKJ","UKK","UKL","UKM","UKN"),
  country_name = c("North East England","North West England","Yorkshire and The Humber",
                    "East Midlands","West Midlands","East of England","London",
                    "South East England","South West England","Wales","Scotland",
                    "Northern Ireland"),
  stringsAsFactors = FALSE
)
country_df <- rbind(country_df[!country_df$ctry_code %in% uk_regions$ctry_code, ], uk_regions)

dbWriteTable(con, "country_names", country_df, overwrite = TRUE)
message("country_names: ", nrow(country_df), " rows")
```

### 11c. GDP per capita data (for scatter plots)

```{r create-gdp-data}
gdp_per_capita_2015 <- data.frame(
  ctry_code = c("AD", "AE", "AF", "AG", "AL", "AM", "AO", "AR", "AT", "AU",
                "AZ", "BA", "BB", "BD", "BE", "BF", "BG", "BH", "BI", "BJ",
                "BN", "BO", "BR", "BS", "BT", "BW", "BY", "BZ", "CA", "CD",
                "CF", "CG", "CH", "CI", "CL", "CM", "CN", "CO", "CR", "CU",
                "CV", "CY", "CZ", "DE", "DJ", "DK", "DM", "DO", "DZ", "EC",
                "EE", "EG", "ER", "ES", "ET", "FI", "FJ", "FM", "FR", "GA",
                "GB", "GD", "GE", "GH", "GM", "GN", "GQ", "GR", "GT", "GW",
                "GY", "HK", "HN", "HR", "HT", "HU", "ID", "IE", "IL", "IN",
                "IQ", "IR", "IS", "IT", "JM", "JO", "JP", "KE", "KG", "KH",
                "KI", "KM", "KN", "KP", "KR", "KW", "KZ", "LA", "LB", "LC",
                "LI", "LK", "LR", "LS", "LT", "LU", "LV", "LY", "MA", "MC",
                "MD", "ME", "MG", "MK", "ML", "MM", "MN", "MO", "MR", "MT",
                "MU", "MV", "MW", "MX", "MY", "MZ", "NA", "NE", "NG", "NI",
                "NL", "NO", "NP", "NZ", "OM", "PA", "PE", "PG", "PH", "PK",
                "PL", "PR", "PS", "PT", "PW", "PY", "QA", "RO", "RS", "RU",
                "RW", "SA", "SB", "SC", "SD", "SE", "SG", "SI", "SK", "SL",
                "SM", "SN", "SO", "SR", "SS", "ST", "SV", "SY", "SZ", "TD",
                "TG", "TH", "TJ", "TL", "TM", "TN", "TO", "TR", "TT", "TV",
                "TW", "TZ", "UA", "UG", "US", "UY", "UZ", "VC", "VE", "VN",
                "VU", "WS", "XK", "YE", "ZA", "ZM", "ZW"),
  gdp_pc_2015 = c(NA, 67127, 1934, 21071, 12387, 11632, 7640, 22536, 52612, 48401,
                  16553, 13679, 15943, 4248, 49145, 2042, 21168, 47384, 758, 2967,
                  76057, 7877, 15065, 31802, 10442, 16687, 18498, 7913, 46780, 1037,
                  868, 5387, 66025, 4789, 24034, 3665, 14439, 14324, 17682, NA,
                  7009, 37898, 34817, 51914, 3744, 52752, 11997, 15784, 13966, 11410,
                  30248, 12035, 1820, 36668, 2086, 45389, 12709, 3587, 43968, 14683,
                  45071, 14790, 12046, 5267, 2331, 2350, 25039, 28037, 8381, 1971,
                  11789, 58573, 5232, 24865, 1795, 27180, 11188, 68883, 37347, 6096,
                  13847, 18765, 50406, 40723, 9025, 10891, 41384, 4373, 3726, 3959,
                  2036, 2914, 23775, NA, 38292, 60566, 25178, 7327, 14073, 14027,
                  NA, 12956, 1350, 3145, 30543, 108797, 26078, 15904, 7780, NA,
                  9614, 18188, 1594, 14949, 2317, 5189, 12456, 107721, 5133, 40160,
                  22534, 17048, 1480, 18934, 27029, 1344, 10622, 1196, 5670, 5678,
                  53955, 64192, 3093, 39935, 32993, 27839, 13009, 4302, 8351, 5174,
                  27963, 37605, 5886, 32871, 16219, 12531, 116582, 24491, 16281, 25218,
                  2178, 49519, 2517, 27632, 4330, 50338, 88155, 35417, 30693, 1622,
                  62644, 3371, NA, 18252, 1879, 3821, 8316, NA, 9895, 2154,
                  1822, 17445, 3280, 5000, 16056, 11604, 6168, 26949, 31251, 4776,
                  NA, 2864, 12270, 2240, 57486, 21928, 6924, 13107, NA, 6826,
                  3008, 6260, 10906, 2591, 13572, 3770, 2413),
  stringsAsFactors = FALSE
)

# Remove NAs before writing
gdp_per_capita_2015 <- gdp_per_capita_2015[!is.na(gdp_per_capita_2015$gdp_pc_2015), ]

dbWriteTable(con, "gdp_data", gdp_per_capita_2015, overwrite = TRUE)
message("gdp_data: ", nrow(gdp_per_capita_2015), " rows")
```

## 12. Final verification

```{r final-verify}
tables <- dbListTables(con)
message("All tables in DuckDB: ", paste(tables, collapse = ", "))

for (tbl in tables) {
  n <- dbGetQuery(con, paste0("SELECT COUNT(*) AS n FROM ", tbl))
  message("  ", tbl, ": ", n$n, " rows")
}

# Verify pre-joined table matches original join
verify_q <- dbGetQuery(con, "
  SELECT ic.ctry_code,
         AVG(ic.value) AS mean_val,
         COUNT(*) AS n
  FROM istrax_country ic
  WHERE ic.flow_type = 'istrax_global'
    AND ic.ctry_code IN ('US', 'GB', 'DE', 'FR', 'CN')
  GROUP BY ic.ctry_code
  ORDER BY ic.ctry_code
")
message("\nVerification query on istrax_country (should match earlier spot check):")
print(verify_q)
```

## 12. Upload to MotherDuck

Upload the completed database to MotherDuck cloud as `iseapp`. Requires the
`MOTHERDUCK_TOKEN` environment variable to be set (get your token from the
MotherDuck UI under Settings → Tokens, then add `MOTHERDUCK_TOKEN=your_token`
to your `~/.Renviron` file and restart R).

```{r motherduck-upload, eval=FALSE}
# Close the local connection first to release the file lock
if (exists("con")) try(dbDisconnect(con, shutdown = TRUE), silent = TRUE)

md_token <- Sys.getenv("MOTHERDUCK_TOKEN")
if (!nzchar(md_token)) {
  message("MOTHERDUCK_TOKEN not set - skipping MotherDuck upload.")
  message("Set it in ~/.Renviron and restart R to enable cloud upload.")
} else {
  db_path_fwd <- gsub("\\\\", "/", duck_db_path)

  # Try R-native approach first (works on Linux/Mac)
  # Note: extension autoloading is OFF in R's duckdb package, so we must
  # use an in-memory connection and explicitly INSTALL/LOAD the extension.
  # This may fail on Windows where the R motherduck extension isn't available.
  upload_ok <- tryCatch({
    message("Connecting to MotherDuck via R duckdb extension...")
    con_md <- dbConnect(duckdb::duckdb())
    dbExecute(con_md, "INSTALL 'motherduck'")
    dbExecute(con_md, "LOAD 'motherduck'")
    dbExecute(con_md, paste0("SET motherduck_token = '", md_token, "'"))
    dbExecute(con_md, "ATTACH 'md:'")

    message("Uploading database to MotherDuck as 'iseapp' (this may take a while)...")
    dbExecute(con_md, paste0("CREATE OR REPLACE DATABASE iseapp FROM '", db_path_fwd, "'"))

    message("Upload complete. Verifying remote database...")
    dbExecute(con_md, "USE iseapp")
    remote_tables <- dbListTables(con_md)
    message("Remote tables: ", paste(remote_tables, collapse = ", "))

    dbDisconnect(con_md, shutdown = TRUE)
    TRUE
  }, error = function(e) {
    message("R duckdb extension failed: ", conditionMessage(e))
    try(dbDisconnect(con_md, shutdown = TRUE), silent = TRUE)
    FALSE
  })

  # Fallback: use DuckDB CLI (works on Windows)
  # Install CLI via: winget install DuckDB.cli
  if (!upload_ok) {
    duckdb_cli <- Sys.which("duckdb")
    if (nzchar(duckdb_cli)) {
      message("Falling back to DuckDB CLI at: ", duckdb_cli)
      # Feed SQL via stdin to avoid shell quoting issues with the JWT token
      sql_lines <- c(
        "INSTALL motherduck;",
        "LOAD motherduck;",
        paste0("SET motherduck_token = '", md_token, "';"),
        "ATTACH 'md:';",
        paste0("CREATE OR REPLACE DATABASE iseapp FROM '", db_path_fwd, "';")
      )
      cli_output <- system2(duckdb_cli, input = sql_lines,
                            stdout = TRUE, stderr = TRUE)
      exit_status <- attr(cli_output, "status")
      if (is.null(exit_status) || exit_status == 0) {
        message("MotherDuck upload via CLI completed successfully.")
      } else {
        warning("DuckDB CLI upload failed. Output:\n", paste(cli_output, collapse = "\n"))
      }
    } else {
      warning("DuckDB CLI not found on PATH.\n",
              "Install it via: winget install DuckDB.cli\n",
              "Or download from: https://duckdb.org/docs/installation/")
    }
  } else {
    message("MotherDuck upload finished successfully.")
  }
}

# Re-open local connection for the cleanup chunk
con <- dbConnect(duckdb(), dbdir = duck_db_path)
```

## 13. Close connection

```{r cleanup}
dbDisconnect(con, shutdown = TRUE)
message("DuckDB database created successfully at: ", duck_db_path)
message("File size: ", round(file.size(duck_db_path) / 1024 / 1024, 1), " MB")
```
