---
title: "Defending Growth"
author: "Innovation Strategy Explorer"
date: "`r format(Sys.Date(), '%-d %B %Y')`"
output:
  html_document:
    self_contained: true
    toc: false
    theme: flatly
    df_print: paged
    css: defending_growth.css
---

```{r setup, include=FALSE}
# svglite is a pure-R SVG backend that does NOT need libX11 / libcairo,
# both of which are missing on this Mac install. The base R `svg()`
# device errors out without them; svglite does not.
knitr::opts_chunk$set(
  echo    = FALSE,
  message = FALSE,
  warning = FALSE,
  dev        = "svglite",   # vector output, no X11/Cairo dependency
  dev.args   = list(),
  fig.width  = 10,
  fig.height = 5.5,
  dpi    = 150
)
suppressPackageStartupMessages({
  library(arrow); library(DBI); library(duckdb)
  library(ggplot2); library(scales); library(svglite)
  library(htmltools)
})
source("R/local_higglobe.R")

# Render a two-panel figure (each panel is a SVG wrapped in a hyperlink
# to the live app) as one rendered HTML string. Built via htmltools so
# the output has no leading-whitespace artefacts that pandoc would
# otherwise interpret as code blocks under `results='asis'`.
panel_pair_figure <- function(title_a, url_a, svg_a,
                              title_b, url_b, svg_b,
                              caption_html) {
  one_panel <- function(letter, title, url, svg) {
    tags$div(
      style = "flex:1 1 380px; min-width:320px;",
      tags$div(
        style = "font-weight:600; text-align:center; margin-bottom:4px; color:#333;",
        HTML(sprintf("<b>(%s)</b> %s", letter, title))
      ),
      tags$a(
        href = url, target = "_blank", rel = "noopener",
        HTML(svg)
      )
    )
  }
  fig <- tags$figure(
    style = "margin:22px 0;",
    tags$div(
      style = paste0("display:flex; flex-wrap:wrap; gap:18px;",
                     " justify-content:center; align-items:flex-start;"),
      one_panel("a", title_a, url_a, svg_a),
      one_panel("b", title_b, url_b, svg_b)
    ),
    tags$figcaption(
      style = paste0("text-align:center; color:#555;",
                     " font-size:0.92rem; margin-top:10px;"),
      HTML(caption_html)
    )
  )
  # htmltools pretty-prints the HTML with indentation, which pandoc
  # then interprets as a markdown code block under `results='asis'`
  # and emits as <pre><code>… literal text. Strip the newlines + any
  # following whitespace so the entire figure becomes a single line.
  out <- as.character(htmltools::renderTags(fig)$html)
  out <- gsub("\n\\s*", "", out)
  cat(out)
}

data_dir <- {
  if      (dir.exists("../inst/extdata")) "../inst/extdata"
  else if (dir.exists("inst/extdata"))    "inst/extdata"
  else    stop("Cannot find inst/extdata/ — knit from project root.")
}
```

The new geopolitical situation we are finding us in will imply more spending
on defence including defence investment. What will be the impact of that on
economic growth overall? A key driver of growth is innovation and a key
driver of innovation are the strength of knowledge spillovers between
different innovations. How does defence compare to other technologies? In
**Figure 1** we estimate the value of knowledge spillovers from different
technologies.

```{r fig1, results='asis'}
# ---- Data ------------------------------------------------------------------
techs <- c("Green Technology", "AI",
           "Any Agriculture & Food technology",
           "Defence Technology", "Biotechnology")

prep_flows <- function(toflow, country = "All countries") {
  fl <- local_value_flows(
    country       = country,
    techs         = c(techs, "All innovations"),
    toflow        = toflow,
    granted_only  = FALSE,
    data_dir      = data_dir
  )
  baseline <- fl$mean[fl$technology == "All innovations"]
  fl <- fl[fl$technology %in% techs, ]
  fl$label <- dplyr::recode(fl$technology,
    "Any Agriculture & Food technology" = "Agri & Food",
    "Green Technology"   = "Green",
    "Defence Technology" = "Defence")
  fl$fill_grp <- dplyr::case_when(
    fl$technology == "Green Technology"                  ~ "green",
    fl$technology == "AI"                                ~ "AI",
    fl$technology == "Any Agriculture & Food technology" ~ "agrifood",
    fl$technology == "Defence Technology"                ~ "defence",
    fl$technology == "Biotechnology"                     ~ "biotech",
    TRUE                                                  ~ "other"
  )
  list(df = fl, baseline = baseline)
}

flows_ev <- prep_flows("ev_global")
flows_is <- prep_flows("is_global")

# Use the SAME y-axis ordering across both panels so the eye can read
# them as a comparison pair. Order by panel A (spillover, $).
ord <- flows_ev$df$label[order(flows_ev$df$mean)]
flows_ev$df$label <- factor(flows_ev$df$label, levels = ord)
flows_is$df$label <- factor(flows_is$df$label, levels = ord)

# is_/av_ flows are stored as decimals; the deployed app multiplies by
# 100 for display. Mirror that here so panel B reads in percent.
flows_is$df$mean <- 100 * flows_is$df$mean
flows_is$df$sem  <- 100 * flows_is$df$sem
flows_is$baseline <- 100 * flows_is$baseline

# ---- Shared scales / theme -------------------------------------------------
fill_pal <- c(green    = "forestgreen", AI       = "orange",
              agrifood = "burlywood",   defence  = "gray35",
              biotech  = "#9467bd")

build_panel <- function(df, baseline, x_lab, value_fmt, baseline_fmt,
                        x_expand_mult = 0.22) {
  ggplot(df, aes(y = label, x = mean)) +
    geom_col(aes(fill = fill_grp), width = 0.6) +
    geom_errorbar(
      aes(xmin = pmax(mean - sem, 0), xmax = mean + sem),
      width = 0.18, colour = "grey30", linewidth = 0.4
    ) +
    geom_point(
      aes(size = innos, fill = fill_grp),
      shape = 21, colour = "grey15", stroke = 0.6, alpha = 0.95
    ) +
    geom_text(
      aes(label = value_fmt(mean)),
      hjust = -0.4, size = 3.4, colour = "grey20"
    ) +
    geom_vline(xintercept = baseline,
               linetype = "dashed", colour = "grey30",
               linewidth = 0.5) +
    annotate("label",
             x = baseline,
             y = nrow(df) + 0.55,
             label = sprintf("All innovations: %s", baseline_fmt(baseline)),
             hjust = -0.05, vjust = 0.5,
             label.size = 0, fill = "white",
             colour = "grey25", size = 3.0,
             fontface = "italic") +
    scale_fill_manual(values = fill_pal, guide = "none") +
    scale_size_area(name = "Innovations",
                    max_size = 12,
                    labels   = scales::comma,
                    breaks   = scales::pretty_breaks(4)) +
    scale_x_continuous(labels = scales::comma,
                       expand = expansion(mult = c(0, x_expand_mult))) +
    scale_y_discrete(expand = expansion(add = c(0.6, 1.1))) +
    labs(y = NULL, x = x_lab) +
    theme_minimal(base_size = 12) +
    theme(panel.grid.major.y = element_blank(),
          legend.position    = "right")
}

p_ev <- build_panel(
  flows_ev$df, flows_ev$baseline,
  x_lab = "Average global spillover per innovation, $ (mean ± 1 SEM)",
  value_fmt    = function(v) scales::dollar(round(v), prefix = "$",
                                            suffix = ""),
  baseline_fmt = function(v) sprintf("$%s", format(round(v, 2),
                                                   nsmall = 2)))
p_is <- build_panel(
  flows_is$df, flows_is$baseline,
  x_lab = "Marginal global return per innovation, % (mean ± 1 SEM)",
  value_fmt    = function(v) sprintf("%s%%", format(round(v),
                                                    big.mark = ",")),
  baseline_fmt = function(v) sprintf("%s%%", format(round(v, 1),
                                                    nsmall = 1)),
  x_expand_mult = 0.18)

# ---- Build URLs for each panel ---------------------------------------------
make_url <- function(toflow) {
  paste0(
    "https://mondpanther-iseapp2.share.connect.posit.cloud/",
    "?_inputs_&country-sidebar=true",
    "&navbar_page=%22Country%20Explorer%22",
    "&country-inner_tabs=%22Value%20flows%20by%20Technology%22",
    "&country-country=%22All%20countries%22",
    "&country-firm=%22No%20firm%20filter%22",
    "&country-techs=%22Green%20Technology%22",
    "&country-city=%22No%20city%20filter%22",
    "&country-toflow=%22", toflow, "%22",
    "&country-tech_categories_plot1=",
    "%5B%22Green%20Technology%22%2C%22AI%22%2C",
    "%22Any%20Agriculture%20%26%20Food%20technology%22%2C",
    "%22Defence%20Technology%22%2C%22Biotechnology%22%5D",
    "&country-widthscale=%22log%22",
    "&country-display_mode=%22confidence%22",
    "&country-granted_only=false",
    "&country-multifam_only=false",
    "&country-include_fallback=false",
    "&country-top_n_ids=10&country-topn=20&country-mininno=10"
  )
}
url_ev <- make_url("ev_global")
url_is <- make_url("is_global")

# ---- Render each panel to its own inline SVG -------------------------------
render_svg <- function(p, w, h) {
  tmp <- tempfile(fileext = ".svg")
  ggsave(tmp, p, device = svglite::svglite, width = w, height = h)
  ln <- readLines(tmp, warn = FALSE)
  ln <- ln[!grepl("^<\\?xml|^<!DOCTYPE", ln)]
  paste(ln, collapse = "\n")
}
svg_ev <- render_svg(p_ev, w = 8.4, h = 5.4)
svg_is <- render_svg(p_is, w = 8.4, h = 5.4)

panel_pair_figure(
  title_a = "Average spillovers", url_a = url_ev, svg_a = svg_ev,
  title_b = "Marginal returns",   url_b = url_is, svg_b = svg_is,
  caption_html = paste0(
    "<b>Figure 1.</b> Per-innovation knowledge-spillover value (panel ",
    "a, Average Global Spillover, $ millions) and per-innovation ",
    "marginal return (panel b, Marginal Global Return, % IRR) by ",
    "technology umbrella. Bars show mean ± 1 SEM; circle area scales ",
    "with the number of patent families behind each estimate; the ",
    "dashed line is the database-wide average across <em>all</em> ",
    "innovations. <em>Click either panel for its interactive, ",
    "drill-down version on the Innovation Strategy Explorer.</em>"
  ))
```

The picture is more mixed than the simple "civilian beats defence"
narrative might suggest, and the two panels tell complementary stories.

**Panel (a) — average spillovers.** This is the dollar value of the
knowledge externality each patent confers on the rest of the world.
**AI is in a class of its own**: the typical AI patent generates roughly
two-and-a-half times the per-innovation spillover that defence does.
Biotechnology comes second. But green technology and agriculture & food
technology, often invoked as defence's growth-relevant counterparts,
actually trail defence on this measure — both sit a touch below the
defence average. The dashed line shows the universe-wide baseline:
defence is well above the typical patent on this measure, AI and biotech
much further still.

**Panel (b) — marginal returns.** This is the implied internal rate of
return on each innovation, scaling the spillover by the patent's alpha
(its private return premium) and dividing through by its R&D cost. The
ranking **flips noticeably**: AI still leads, but green technology and
agriculture & food now sit *above* defence — not below it. The reason
is mechanical: green and agri patents tend to have lower research costs
than defence patents, so each dollar of R&D earns a higher implied
return even when the absolute spillover is smaller. Biotechnology — the
runner-up on spillovers — drops to last on returns, a reflection of its
high R&D cost base.

**Two qualifications.** First, green technology produces roughly **30
times as many patent families as defence does** (the bubble area in
both panels). Even at a similar per-innovation rate, the aggregate
spillover and aggregate-return footprint of green innovation is much
larger — what is similar per-patent and big in aggregate moves the
growth needle more than what is similar per-patent and small in
aggregate. Second, both rankings are sample-sensitive: restricting to
multi-application or LMIC-origin families reshuffles the picture, and
a more careful follow-up should look at a portfolio view (spillover
or return *× volume*) rather than per-innovation alone.

Treated carefully, neither panel says "don't fund defence" —
geopolitical and security cases for defence spending stand on their own
ground. But together they do temper the argument that defence R&D is a
uniquely strong engine of *broader* economic growth: AI is plainly
ahead on both measures, and once research cost is taken into account
(panel b) green technology and agriculture & food look more promising
than the headline spillover comparison alone would suggest.

## Spillovers to the developing world

The "global" measure used above weights every recipient country
equally. From a development-policy angle, however, what often matters
is how much of the spillover lands in low- and middle-income economies
specifically. **Figure 2** repeats the same comparison, but the flow
metric is now narrowed to **LMICs excluding China**: panel (a) is the
average dollar spillover that each source-country patent generates for
those economies, and panel (b) is the implied marginal return of the
same patent measured against the LMIC-bound spillover only.

```{r fig2, results='asis'}
flows_ev_lmic <- prep_flows("ev_emdenocn")
flows_is_lmic <- prep_flows("is_emdenocn")

# Use panel (a) to set the y-axis ordering for both panels.
ord_lmic <- flows_ev_lmic$df$label[order(flows_ev_lmic$df$mean)]
flows_ev_lmic$df$label <- factor(flows_ev_lmic$df$label, levels = ord_lmic)
flows_is_lmic$df$label <- factor(flows_is_lmic$df$label, levels = ord_lmic)

flows_is_lmic$df$mean <- 100 * flows_is_lmic$df$mean
flows_is_lmic$df$sem  <- 100 * flows_is_lmic$df$sem
flows_is_lmic$baseline <- 100 * flows_is_lmic$baseline

p_ev_lmic <- build_panel(
  flows_ev_lmic$df, flows_ev_lmic$baseline,
  x_lab = "Avg spillover to LMICs (excl. China) per innovation, $ (mean ± 1 SEM)",
  value_fmt    = function(v) sprintf("$%s",
                                     format(round(v, 2), nsmall = 2)),
  baseline_fmt = function(v) sprintf("$%s",
                                     format(round(v, 2), nsmall = 2)),
  x_expand_mult = 0.20)

p_is_lmic <- build_panel(
  flows_is_lmic$df, flows_is_lmic$baseline,
  x_lab = "Marginal return to LMICs (excl. China) per innovation, % (mean ± 1 SEM)",
  value_fmt    = function(v) sprintf("%s%%",
                                     format(round(v, 1), nsmall = 1)),
  baseline_fmt = function(v) sprintf("%s%%",
                                     format(round(v, 1), nsmall = 1)),
  x_expand_mult = 0.18)

url_ev_lmic <- make_url("ev_emdenocn")
url_is_lmic <- make_url("is_emdenocn")
svg_ev_lmic <- render_svg(p_ev_lmic, w = 8.4, h = 5.4)
svg_is_lmic <- render_svg(p_is_lmic, w = 8.4, h = 5.4)

panel_pair_figure(
  title_a = "Average spillovers to LMICs (excl. China)",
  url_a   = url_ev_lmic, svg_a = svg_ev_lmic,
  title_b = "Marginal returns to LMICs (excl. China)",
  url_b   = url_is_lmic, svg_b = svg_is_lmic,
  caption_html = paste0(
    "<b>Figure 2.</b> Per-innovation spillover (panel a, $ millions) ",
    "and marginal return (panel b, % IRR) flowing to LMICs ",
    "<em>excluding China</em>, by technology umbrella. Same ",
    "conventions as Figure 1. <em>Click either panel for its ",
    "interactive, drill-down version on the Innovation Strategy ",
    "Explorer.</em>"
  ))
```

The LMIC-focused view tells a different — and politically more
provocative — story. **Defence climbs to second place on both panels**:
its average spillover to LMICs ($0.37m) and its implied marginal return
to LMICs (~480%) sit clearly above what green technology, agriculture &
food, and biotechnology produce on those measures. AI still leads, but
with a wider margin than in the global picture and a defence-shaped
runner-up rather than a biotech-shaped one.

That ordering reflects something real about the geography of citation
flows: defence-tech innovation, like AI, is dominated by a small number
of well-developed innovation hubs (US, UK, France, Israel, Japan,
Korea), and the patents originating in those hubs are disproportionately
likely to be cited by patents elsewhere — including in LMICs that lean
on imported technology rather than producing their own. Green and agri
patenting is more geographically diffuse and tends to be cited by other
patents in similarly diversified domains, so a smaller share of any
given green patent's spillover footprint lands in the LMIC bucket
specifically.

The volume qualification still applies, and arguably hits harder in
this slice. **Green technology produces roughly thirty times as many
patents as defence does**, so even at a quarter the per-innovation rate
the aggregate spillover *into* LMICs from green is comparable to or
larger than defence's. From a "how do we maximise growth-relevant
externalities into the developing world" standpoint, the right
portfolio answer is probably *both* — and certainly not "defence
because of LMIC spillovers". But the per-innovation comparison is still
worth seeing on its own terms.

## What if LMICs do the inventing themselves?

Figures 1 and 2 ranged across all source countries. The development-
policy lens is sharper if we restrict the *source* of the patents to
LMIC inventors as well: how productive is each technology umbrella when
the inventor is themselves resident in an LMIC, and the spillover is
flowing back to other LMICs? **Figure 3** is Figure 2 again, but with
the country filter set to **LMICs excluding China** — so we are looking
at the spillover from LMIC-origin patents to other LMIC-resident
innovation, by umbrella.

```{r fig3, results='asis'}
flows_ev_lmic2 <- prep_flows("ev_emdenocn", country = "LMICs (excl. China)")
flows_is_lmic2 <- prep_flows("is_emdenocn", country = "LMICs (excl. China)")

ord3 <- flows_ev_lmic2$df$label[order(flows_ev_lmic2$df$mean)]
flows_ev_lmic2$df$label <- factor(flows_ev_lmic2$df$label, levels = ord3)
flows_is_lmic2$df$label <- factor(flows_is_lmic2$df$label, levels = ord3)

flows_is_lmic2$df$mean <- 100 * flows_is_lmic2$df$mean
flows_is_lmic2$df$sem  <- 100 * flows_is_lmic2$df$sem
flows_is_lmic2$baseline <- 100 * flows_is_lmic2$baseline

p_ev_lmic2 <- build_panel(
  flows_ev_lmic2$df, flows_ev_lmic2$baseline,
  x_lab = "Avg spillover to LMICs (excl. China) per innovation, $ (mean ± 1 SEM)",
  value_fmt    = function(v) sprintf("$%s",
                                     format(round(v, 2), nsmall = 2)),
  baseline_fmt = function(v) sprintf("$%s",
                                     format(round(v, 2), nsmall = 2)),
  x_expand_mult = 0.20)
p_is_lmic2 <- build_panel(
  flows_is_lmic2$df, flows_is_lmic2$baseline,
  x_lab = "Marginal return to LMICs (excl. China) per innovation, % (mean ± 1 SEM)",
  value_fmt    = function(v) sprintf("%s%%",
                                     format(round(v, 1), nsmall = 1)),
  baseline_fmt = function(v) sprintf("%s%%",
                                     format(round(v, 1), nsmall = 1)),
  x_expand_mult = 0.18)

# Country-restricted live-app URLs (just the country knob differs).
make_url_lmic <- function(toflow) {
  sub("&country-country=%22All%20countries%22",
      "&country-country=%22LMICs%20(excl.%20China)%22",
      make_url(toflow), fixed = TRUE)
}
url_ev_lmic2 <- make_url_lmic("ev_emdenocn")
url_is_lmic2 <- make_url_lmic("is_emdenocn")
svg_ev_lmic2 <- render_svg(p_ev_lmic2, w = 8.4, h = 5.4)
svg_is_lmic2 <- render_svg(p_is_lmic2, w = 8.4, h = 5.4)

panel_pair_figure(
  title_a = "Average spillovers — LMIC source, LMIC destination",
  url_a   = url_ev_lmic2, svg_a = svg_ev_lmic2,
  title_b = "Marginal returns — LMIC source, LMIC destination",
  url_b   = url_is_lmic2, svg_b = svg_is_lmic2,
  caption_html = paste0(
    "<b>Figure 3.</b> Same as Figure 2 but with the country filter ",
    "restricted to <em>LMICs excluding China</em>: the source patents ",
    "now have at least one LMIC-excl-China inventor record. Sample ",
    "sizes drop sharply (defence has roughly 800 LMIC-resident ",
    "families vs. 46,000 globally), so the SEM bars widen. <em>Click ",
    "either panel for its interactive, drill-down version on the ",
    "Innovation Strategy Explorer.</em>"
  ))
```

The picture **flips again** when the source patents are themselves
LMIC-resident. **Defence is now last on spillovers** — it sits *below*
the all-innovations baseline — and only mid-pack on marginal returns.
AI still dominates, with green technology, agriculture & food, and
biotechnology all comfortably above defence on both panels.

Two ways to read this. The benign reading: most LMIC-resident defence
patenting is small, niche, often local-procurement-driven, and doesn't
plug into the global citation network the way LMIC-resident
green-technology or AI work increasingly does. The cautionary reading:
**when LMICs are doing their own inventing, the dollars they put into
defence-related R&D are the dollars least likely to come back to them
as growth-relevant knowledge externalities.** That ought to influence
how LMIC governments frame any tech-and-defence industrial policy.

A note on the sample: with only ~800 defence patents from LMIC
inventors in the database, the defence bar in panel (b) carries a wider
SEM band than the others. The qualitative ranking is robust — defence
trails the four civilian umbrellas on both panels — but the precise
gap to the runners-up should be read with that uncertainty in mind.

## How does spillover-to-LMIC actually travel?

The bar charts above flatten the LMIC-bound spillover into a single
per-innovation number. To see what those flows actually look like as
they propagate through the citation network — and why defence and green
end up in such different places — the maps below trace **seven
generations of citations** from each umbrella's gen-0 patent set,
weighted by `ev_emdenocn` (the per-citation flow into LMICs excluding
China). Both panels start with a random sample of 1,000 patents from
the relevant umbrella; at each generation step we keep a 10% random
sample of citing-docdb candidates so the map stays legible while still
showing the network's reach.

```{r fig4, message=FALSE}
library(leaflet)
panel_green_lmic_flow <- higglobe_local_sample(
  country         = "All countries",
  techs           = "Green Technology",
  toflow          = "ev_emdenocn",
  gen0_n          = 1000L,
  gen0_mode       = "Random",
  n_generations   = 7L,
  edge_sample_pct = 10,
  include_fallback = TRUE,
  seed            = 42,
  data_dir        = data_dir
)
panel_defence_lmic_flow <- higglobe_local_sample(
  country         = "All countries",
  techs           = "Any Defence technology",
  toflow          = "ev_emdenocn",
  gen0_n          = 1000L,
  gen0_mode       = "Random",
  n_generations   = 7L,
  edge_sample_pct = 10,
  include_fallback = TRUE,
  seed            = 42,
  data_dir        = data_dir
)

# Live-app deep-link with the same parameters as the figure builds.
hglobe_live_url <- function(techs) {
  q <- function(x) utils::URLencode(paste0('"', x, '"'), reserved = TRUE)
  paste0(
    "https://mondpanther-iseapp2.share.connect.posit.cloud/",
    "?_inputs_&navbar_page=", q("HiGGlobe"),
    "&hglobe-country=",        q("All countries"),
    "&hglobe-firm=",           q("No firm filter"),
    "&hglobe-techs=",          q(techs),
    "&hglobe-city=",           q("No city filter"),
    "&hglobe-toflow=",         q("ev_emdenocn"),
    "&hglobe-granted_only=false&hglobe-multifam_only=false",
    "&hglobe-include_fallback=true",
    "&hglobe-gen0_select_mode=", q("Random"),
    "&hglobe-gen0_unit_mode=",   q("Number"),
    "&hglobe-gen0_sample_val=1000",
    "&hglobe-edge_select_mode=", q("Random"),
    "&hglobe-edge_unit_mode=",   q("Percent"),
    "&hglobe-edge_sample_val=10",
    "&hglobe-add_generations=7",
    "&hglobe-sidebar=true",
    paste(sprintf("&hglobe-show_gen_%d=true", 0:9), collapse = ""),
    "&step=7"
  )
}
```

```{r fig4-green, fig.cap="(green panel)"}
htmltools::tagList(
  htmltools::tags$figure(
    style = "margin: 22px 0;",
    htmltools::tags$div(
      style = "font-weight: 600; text-align: center; margin-bottom: 6px;",
      "Figure 4. Green-technology citation network, weighted by spillover to LMICs (excl. China)"
    ),
    htmltools::tags$div(
      style = "margin: 4px 0; font-size: 0.85rem; text-align: right;",
      htmltools::tags$a(
        "Open the live HiGGlobe view →",
        href   = hglobe_live_url("Green Technology"),
        target = "_blank", rel = "noopener"
      )
    ),
    higglobe_local_leaflet(panel_green_lmic_flow, height = "520px"),
    htmltools::tags$figcaption(
      style = "text-align:center; color:#555;
               font-size:0.92rem; margin-top:8px;",
      htmltools::HTML(paste0(
        "Seven generations of forward citations from a 1,000-patent ",
        "random gen-0 sample of green-technology patents worldwide; ",
        "edge sampling 10% per step. Marker fill colour encodes ",
        "generation depth. <em>Click any marker</em> for that patent's ",
        "publication number with a direct link to the Espacenet ",
        "search page; click the link above to open the equivalent ",
        "interactive HiGGlobe view on the live app."
      ))
    )
  )
)
```

```{r fig5-defence, fig.cap="(defence panel)"}
htmltools::tagList(
  htmltools::tags$figure(
    style = "margin: 22px 0;",
    htmltools::tags$div(
      style = "font-weight: 600; text-align: center; margin-bottom: 6px;",
      "Figure 5. Defence-technology citation network, weighted by spillover to LMICs (excl. China)"
    ),
    htmltools::tags$div(
      style = "margin: 4px 0; font-size: 0.85rem; text-align: right;",
      htmltools::tags$a(
        "Open the live HiGGlobe view →",
        href   = hglobe_live_url("Any Defence technology"),
        target = "_blank", rel = "noopener"
      )
    ),
    higglobe_local_leaflet(panel_defence_lmic_flow, height = "520px"),
    htmltools::tags$figcaption(
      style = "text-align:center; color:#555;
               font-size:0.92rem; margin-top:8px;",
      htmltools::HTML(paste0(
        "Same construction as Figure 4 with <code>techs = ",
        "\"Any Defence technology\"</code>. The defence network reaches ",
        "deeper (gen 7+) but the geographic spread is much narrower — ",
        "concentrated in a handful of established defence hubs. ",
        "<em>Click any marker</em> for that patent's publication number ",
        "with a direct Espacenet link."
      ))
    )
  )
)
```

The two maps make tangible what the bars in Figure 2 implied. Green
patenting fans out across many continents at every generation. Defence
patenting cascades within a much narrower geography — North America,
Western Europe, Japan, Korea — and only modestly leaks into LMIC
territory at gen 1 / gen 2 before fading. The dollar gap in panel (a)
of Figure 2 (defence > green for spillover *to LMICs*) is delivered by
a small number of high-value defence-source patents whose citation
chains happen to terminate disproportionately in LMIC-resident
descendants — the "defence to LMIC" pipe is strong but narrow.

<small>Source: Innovation Strategy Explorer; full database, granted +
non-granted patent families. Figure 1 covers all source countries;
Figure 2 narrows the *recipient* to LMICs excluding China; Figure 3
narrows both source and recipient to LMICs excluding China; Figures
4–5 trace seven generations of `ev_emdenocn`-weighted citations from
green and defence respectively.
[github.com/mondpanther/iseapp](https://github.com/mondpanther/iseapp).</small>
