---
title: "Spillover flows between HICs and LMICs (excl. China)"
author: "Innovation Strategy Explorer — insights"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 2
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 11, fig.height = 6.5, dpi = 150)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(scales)
library(ggrepel)
```

## What this note shows

How much per-innovation spillover value flows between High-Income Countries (HICs) and LMICs (excl. China) — and how it varies by technology and by where the innovation was filed. All figures use **granted** patent families only.

Three views:

1. **Spillovers to HICs** — LMIC-origin vs HIC-origin innovations.
2. **Spillovers to LMICs (excl. China)** — LMIC-origin vs HIC-origin innovations.
3. **Spillovers to HICs, LMIC-origin only, Marimekko-style** — bar *width* encodes how many innovations LMICs contribute in each technology (so wide bars are categories where LMIC innovation volume is substantial, narrow ones where it is thin), bar *height* is the per-innovation spillover to HICs.

In figures 1 and 2, an overlaid **circle marker** reports the **LMIC share of total innovation in each category** (LMIC-origin family count ÷ all-origin family count) on a secondary axis.

## Data source

Six CSVs from the public Country Explorer tab of <https://mondpanther-iseapp2.share.connect.posit.cloud/>, all with **Granted families only**. Two destination flows × three country filters. Click any "open in app" link to jump to the bookmarked view.

```{r build-app-urls, include=FALSE}
build_app_url <- function(country, toflow, granted_only = TRUE,
                          base = "https://mondpanther-iseapp2.share.connect.posit.cloud/",
                          tech_categories = c(
                            "Green Technology", "AI", "Green Transport",
                            "Green Manufacturing", "GHG Capture",
                            "Green Energy", "Adaptation", "Green Housing",
                            "Circular Economy", "Green ICT",
                            "Green Agriculture",
                            "Any Agriculture & Food technology"
                          )) {
  q <- function(x) utils::URLencode(paste0('"', x, '"'), reserved = TRUE)
  tech_arr <- paste0(
    "%5B",
    paste0(vapply(tech_categories,
                  function(t) paste0("%22", utils::URLencode(t, reserved = TRUE), "%22"),
                  character(1)),
           collapse = "%2C"),
    "%5D"
  )
  paste0(
    base,
    "?_inputs_",
    "&navbar_page=", q("Country Explorer"),
    "&country-inner_tabs=", q("Value flows by Technology"),
    "&country-country=", q(country),
    "&country-firm=", q("No firm filter"),
    "&country-toflow=", q(toflow),
    "&country-tech_categories_plot1=", tech_arr,
    "&country-granted_only=", tolower(as.character(granted_only)),
    "&country-widthscale=", q("log"),
    "&country-display_mode=", q("confidence"),
    "&country-top_n_ids=10"
  )
}
url_L_H <- build_app_url("LMICs (excl. China)",   "ev_hic",      granted_only = TRUE)
url_H_H <- build_app_url("High income countries", "ev_hic",      granted_only = TRUE)
url_A_H <- build_app_url("All countries",         "ev_hic",      granted_only = TRUE)
url_L_L <- build_app_url("LMICs (excl. China)",   "ev_emdenocn", granted_only = TRUE)
url_H_L <- build_app_url("High income countries", "ev_emdenocn", granted_only = TRUE)
url_A_L <- build_app_url("All countries",         "ev_emdenocn", granted_only = TRUE)
```

| CSV in `insights/data/` | Origin filter | Destination flow | App view |
|---|---|---|---|
| `flow_LMIC_to_HIC_granted.csv`  | `LMICs (excl. China)`   | to HICs (`ev_hic`)         | [open in app](`r url_L_H`) |
| `flow_HIC_to_HIC_granted.csv`   | `High income countries` | to HICs (`ev_hic`)         | [open in app](`r url_H_H`) |
| `flow_All_to_HIC_granted.csv`   | `All countries`         | to HICs (`ev_hic`)         | [open in app](`r url_A_H`) |
| `flow_LMIC_to_LMIC_granted.csv` | `LMICs (excl. China)`   | to LMICs (`ev_emdenocn`)   | [open in app](`r url_L_L`) |
| `flow_HIC_to_LMIC_granted.csv`  | `High income countries` | to LMICs (`ev_emdenocn`)   | [open in app](`r url_H_L`) |
| `flow_All_to_LMIC_granted.csv`  | `All countries`         | to LMICs (`ev_emdenocn`)   | [open in app](`r url_A_L`) |

Missing CSVs are fetched automatically via headless Chrome (`chromote`). No direct database access is used.

```{r fetch-csvs, include=FALSE}
source("R/fetch_flow_csv.R")

specs <- list(
  list(name = "LMIC_to_HIC_granted",
       country = "LMICs (excl. China)",
       toflow  = "ev_hic", granted_only = TRUE),
  list(name = "HIC_to_HIC_granted",
       country = "High income countries",
       toflow  = "ev_hic", granted_only = TRUE),
  list(name = "All_to_HIC_granted",
       country = "All countries",
       toflow  = "ev_hic", granted_only = TRUE),
  list(name = "LMIC_to_LMIC_granted",
       country = "LMICs (excl. China)",
       toflow  = "ev_emdenocn", granted_only = TRUE),
  list(name = "HIC_to_LMIC_granted",
       country = "High income countries",
       toflow  = "ev_emdenocn", granted_only = TRUE),
  list(name = "All_to_LMIC_granted",
       country = "All countries",
       toflow  = "ev_emdenocn", granted_only = TRUE)
)

missing <- vapply(specs, function(s)
  !file.exists(file.path("data", paste0("flow_", s$name, ".csv"))),
  logical(1))

if (any(missing)) {
  fetch_flow_csvs(specs[missing], out_dir = "data")
}
```

```{r load-data}
read_flow <- function(file, origin_label) {
  readr::read_csv(file.path("data", file), show_col_types = FALSE) |>
    dplyr::mutate(origin = origin_label)
}

# Helper: assemble dataset for one destination (HIC or LMIC)
build_combined <- function(lmic_csv, hic_csv, all_csv) {
  combined <- dplyr::bind_rows(
    read_flow(lmic_csv, "LMIC-origin (excl. China)"),
    read_flow(hic_csv,  "HIC-origin")
  ) |>
    dplyr::filter(technology != "All") |>
    dplyr::mutate(
      origin = factor(origin,
                      levels = c("LMIC-origin (excl. China)", "HIC-origin"))
    )

  all_countries <- read_flow(all_csv, "All countries") |>
    dplyr::filter(technology != "All") |>
    dplyr::select(technology, innos_all = innos)

  lmic_share <- combined |>
    dplyr::filter(origin == "LMIC-origin (excl. China)") |>
    dplyr::select(technology, innos_lmic = innos) |>
    dplyr::left_join(all_countries, by = "technology") |>
    dplyr::mutate(share = innos_lmic / innos_all)

  list(combined = combined, lmic_share = lmic_share)
}

to_hic  <- build_combined("flow_LMIC_to_HIC_granted.csv",
                          "flow_HIC_to_HIC_granted.csv",
                          "flow_All_to_HIC_granted.csv")
to_lmic <- build_combined("flow_LMIC_to_LMIC_granted.csv",
                          "flow_HIC_to_LMIC_granted.csv",
                          "flow_All_to_LMIC_granted.csv")
```

```{r plot-helper}
# Shared grouped-bar + share-circle renderer
plot_flows <- function(combined, lmic_share, destination_label,
                       flow_code,
                       origin_colors = c(
                         "LMIC-origin (excl. China)" = "#ff7f0e",
                         "HIC-origin"                = "#1f77b4"
                       )) {
  tech_order <- combined |>
    dplyr::filter(origin == "HIC-origin") |>
    dplyr::arrange(dplyr::desc(mean)) |>
    dplyr::pull(technology)

  plot_df  <- combined   |> dplyr::mutate(technology = factor(technology, levels = tech_order))
  share_df <- lmic_share |> dplyr::mutate(technology = factor(technology, levels = tech_order))

  bar_ymax  <- max(plot_df$mean + plot_df$sem, na.rm = TRUE)
  share_max <- max(share_df$share, na.rm = TRUE)
  axis_cap  <- ceiling(share_max * 20) / 20
  scale_k   <- bar_ymax / axis_cap

  ggplot(plot_df, aes(x = technology, y = mean, fill = origin)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.75) +
    geom_errorbar(
      aes(ymin = pmax(mean - sem, 0), ymax = mean + sem),
      position = position_dodge(width = 0.8),
      width = 0.2, colour = "grey30", linewidth = 0.3
    ) +
    geom_point(
      data = share_df,
      aes(x = technology, y = share * scale_k, shape = "LMIC share of innovation"),
      inherit.aes = FALSE,
      colour = "grey15", fill = "white", size = 4, stroke = 1.2
    ) +
    scale_shape_manual(name = NULL, values = c("LMIC share of innovation" = 21)) +
    scale_y_continuous(
      labels   = scales::comma,
      sec.axis = sec_axis(
        transform = ~ . / scale_k,
        name      = paste0("LMIC share of total innovation in category (axis capped at ",
                           scales::percent(axis_cap, accuracy = 0.1), ")"),
        labels    = scales::percent_format(accuracy = 0.1),
        breaks    = scales::pretty_breaks(n = 5)
      )
    ) +
    scale_fill_manual(values = origin_colors) +
    labs(
      title    = sprintf("Average spillover value to %s, per innovation (granted families only)",
                         destination_label),
      subtitle = sprintf("By technology and origin — destination flow: %s · bars: spillover \u00b1 SEM · circles: LMIC share of total innovation",
                         flow_code),
      x = NULL, y = sprintf("Spillover to %s per innovation (\u00b1 1 SEM)", destination_label),
      fill = NULL,
      caption = "Source: Innovation Strategy Explorer, Country Explorer CSV downloads"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x        = element_text(angle = 35, hjust = 1),
      legend.position    = "top",
      panel.grid.major.x = element_blank()
    )
}
```

## Figure 1 — Spillovers to HICs

```{r fig-hic}
plot_flows(to_hic$combined, to_hic$lmic_share,
           destination_label = "HICs",
           flow_code         = "ev_hic")
```

## Figure 2 — Spillovers to LMICs (excl. China)

```{r fig-lmic}
plot_flows(to_lmic$combined, to_lmic$lmic_share,
           destination_label = "LMICs (excl. China)",
           flow_code         = "ev_emdenocn")
```

## Figure 3 — LMIC-origin only, bar width ∝ number of innovations

Same LMIC-origin bars as in Figure 1 (spillovers to HICs) but rendered **Marimekko-style**: bar *width* is proportional to the count of LMIC-origin innovations in that technology category (so you immediately see which technologies contribute meaningful volume, not just meaningful per-innovation value). Bar *height* is the average spillover to HICs per innovation.

```{r fig-hic-lmiconly-marimekko, fig.height=6}
mm_df <- to_hic$combined |>
  dplyr::filter(origin == "LMIC-origin (excl. China)") |>
  dplyr::arrange(dplyr::desc(mean))

mm_df <- mm_df |>
  dplyr::mutate(
    w           = innos / sum(innos),
    xmax        = cumsum(w),
    xmin        = xmax - w,
    xmid        = (xmin + xmax) / 2,
    total_value = mean * innos,                                   # in $M
    total_label = sprintf("$%s M",
                          formatC(round(total_value), format = "d", big.mark = ","))
  )

ggplot(mm_df) +
  geom_rect(
    aes(xmin = xmin, xmax = xmax,
        ymin = 0,    ymax = mean),
    fill = "#ff7f0e", colour = "white", linewidth = 0.3
  ) +
  geom_errorbar(
    aes(x = xmid, ymin = pmax(mean - sem, 0), ymax = mean + sem),
    width = 0.005, colour = "grey30", linewidth = 0.4
  ) +
  ggrepel::geom_label_repel(
    aes(x = xmid, y = mean / 2, label = total_label),
    colour        = "grey20",
    fill          = "white",
    label.size    = 0.25,
    label.padding = grid::unit(0.2, "lines"),
    label.r       = grid::unit(0.08, "lines"),
    size          = 3,
    min.segment.length = 0,
    segment.colour     = "grey40",
    segment.size       = 0.3,
    box.padding        = 0.4,
    force              = 2,
    max.overlaps       = Inf,
    direction          = "both",
    seed               = 1
  ) +
  scale_x_continuous(
    breaks = mm_df$xmid,
    labels = mm_df$technology,
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title    = "Spillovers to HICs per LMIC-origin innovation — bar width \u221d N innovations",
    subtitle = "Wider bars = more LMIC-origin innovations in that category · height = mean spillover",
    x = NULL, y = "Spillover to HICs per innovation (\u00b1 1 SEM)",
    caption = "Source: Innovation Strategy Explorer, Country Explorer CSV downloads"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x        = element_text(angle = 35, hjust = 1),
    panel.grid.major.x = element_blank()
  )
```

## Reading the charts

- **Figures 1 & 2** — Within each technology group, the **blue** bar is the average destination-bound spillover per HIC-origin innovation; the **orange** bar is the same figure for LMIC-origin innovations (excl. China). The **open circle** is the LMIC share of total innovation in the category (right axis). Error bars are $\pm$ 1 SEM. Technologies are sorted left-to-right by HIC-origin spillover magnitude within each figure.
- **Figure 3** — Only LMIC-origin bars, spillover *to HICs*. Width proportional to the number of LMIC-origin innovations in the category: this deliberately combines intensity (height) and volume (width), so the total area of each bar is proportional to the **aggregate spillover value flowing to HICs from LMIC innovation in that technology**. Each white-boxed label reports that aggregate (per-innovation mean × number of innovations) in millions of USD; labels are offset by `ggrepel` with a leader line pointing to the bar they describe.

## Summary table

```{r table}
bind_tbl <- function(x, destination_label) {
  x$combined |>
    dplyr::select(technology, origin, mean, sem, innos) |>
    tidyr::pivot_wider(
      id_cols     = technology,
      names_from  = origin,
      values_from = c(mean, sem, innos)
    ) |>
    dplyr::mutate(destination = destination_label, .before = 1)
}

summary_tbl <- dplyr::bind_rows(
  bind_tbl(to_hic,  "HICs"),
  bind_tbl(to_lmic, "LMICs (excl. China)")
) |>
  dplyr::mutate(
    ratio_LMIC_over_HIC = `mean_LMIC-origin (excl. China)` / `mean_HIC-origin`
  ) |>
  dplyr::arrange(destination, dplyr::desc(`mean_HIC-origin`))

knitr::kable(
  summary_tbl,
  digits = 2,
  col.names = c("Destination", "Technology",
                "Mean LMIC\u2192dest", "Mean HIC\u2192dest",
                "SEM LMIC\u2192dest",  "SEM HIC\u2192dest",
                "N LMIC-origin",        "N HIC-origin",
                "Ratio LMIC / HIC"),
  caption = "Per-innovation spillover means by destination, technology, and origin (granted families only)."
)
```

## Notes & caveats

- Bars in Figures 1 and 2 use the same destination flow within each figure, so the levels are directly comparable within a technology.
- The sample size column `innos` is **not** the same across origins: LMIC innovator pools are smaller, so their SEMs are typically larger. Treat any orange/blue gap inside a large SEM with caution.
- Only **granted** patent families are included (the app's default).
- The origin-country filter selects patents filed from the relevant country group; downstream-citing patents can come from anywhere in the world, but only destination-country citations count toward the `ev_*` flow.
- Data acquisition uses only the app's public CSV-download endpoint (via `chromote`) — the patent database is never accessed directly.

---

```{r version-stamp, results='asis'}
local({
  util <- "../R/utils_version.R"
  if (file.exists(util)) source(util, local = TRUE)
})
v <- tryCatch(app_version(), error = function(e) "unknown")
cat(sprintf(
  "<p style='font-size:0.85em; color:#666;'>Rendered %s &middot; app version <code>%s</code></p>",
  format(Sys.time(), "%Y-%m-%d %H:%M %Z"), v
))
```
