---
title: "Spillovers to High-Income Countries — LMIC-origin vs HIC-origin innovations"
author: "Innovation Strategy Explorer — insights"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    self_contained: true
    toc: false
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 11, fig.height = 6.5, dpi = 150)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(scales)
```

## What this note shows

How much per-innovation spillover value do different technologies generate **for High-Income Countries (HICs)** — and does this depend on where the innovation was filed and on whether we restrict to *granted* families? The faceted grouped bar chart below compares three populations of innovations, all scored on the *same* destination flow (*Average Spillovers to HICs* = `ev_hic`):

- **LMIC-origin** innovations, excluding China
- **China-origin** innovations
- **HIC-origin** innovations

It shows each origin group twice — once restricted to *granted* patent families (the app's default) and once including *all* families regardless of grant status. Technologies are on the x-axis.

## Data source

Six downloads from the public Country Explorer tab of <https://mondpanther-iseapp2.share.connect.posit.cloud/>, with **Value flow = *Average Spillovers to HICs*** in all six cases. Three origins × two grant statuses. Click any "open in app" link below to jump directly to the bookmarked view used for that row:

```{r build-app-urls, include=FALSE}
build_app_url <- function(country, toflow, granted_only = TRUE,
                          base = "https://mondpanther-iseapp2.share.connect.posit.cloud/",
                          tech_categories = c(
                            "Green Technology", "AI", "Green Transport",
                            "Green Manufacturing", "GHG Capture",
                            "Green Energy", "Adaptation", "Green Housing",
                            "Circular Economy", "Green ICT",
                            "Green Agriculture",
                            "Any Agriculture & Food technology"
                          )) {
  q <- function(x) utils::URLencode(paste0('"', x, '"'), reserved = TRUE)
  tech_arr <- paste0(
    "%5B",
    paste0(vapply(tech_categories,
                  function(t) paste0("%22", utils::URLencode(t, reserved = TRUE), "%22"),
                  character(1)),
           collapse = "%2C"),
    "%5D"
  )
  paste0(
    base,
    "?_inputs_",
    "&navbar_page=", q("Country Explorer"),
    "&country-inner_tabs=", q("Value flows by Technology"),
    "&country-country=", q(country),
    "&country-firm=", q("No firm filter"),
    "&country-toflow=", q(toflow),
    "&country-tech_categories_plot1=", tech_arr,
    "&country-granted_only=", tolower(as.character(granted_only)),
    "&country-widthscale=", q("log"),
    "&country-display_mode=", q("confidence"),
    "&country-top_n_ids=10"
  )
}
url_lmic_g  <- build_app_url("LMICs (excl. China)",   "ev_hic", granted_only = TRUE)
url_china_g <- build_app_url("CN",                     "ev_hic", granted_only = TRUE)
url_hic_g   <- build_app_url("High income countries", "ev_hic", granted_only = TRUE)
url_lmic_a  <- build_app_url("LMICs (excl. China)",   "ev_hic", granted_only = FALSE)
url_china_a <- build_app_url("CN",                     "ev_hic", granted_only = FALSE)
url_hic_a   <- build_app_url("High income countries", "ev_hic", granted_only = FALSE)
```

| File in `insights/data/` | `country-country` filter | Granted only? | App view |
|---|---|---|---|
| `flow_LMIC_to_HIC_granted.csv`  | `LMICs (excl. China)`     | yes | [open in app](`r url_lmic_g`)  |
| `flow_China_to_HIC_granted.csv` | `CN` (China)              | yes | [open in app](`r url_china_g`) |
| `flow_HIC_to_HIC_granted.csv`   | `High income countries`   | yes | [open in app](`r url_hic_g`)   |
| `flow_LMIC_to_HIC_all.csv`      | `LMICs (excl. China)`     | **no** | [open in app](`r url_lmic_a`)  |
| `flow_China_to_HIC_all.csv`     | `CN` (China)              | **no** | [open in app](`r url_china_a`) |
| `flow_HIC_to_HIC_all.csv`       | `High income countries`   | **no** | [open in app](`r url_hic_a`)   |

If any CSV is missing, the chunk below fetches it via headless Chrome using the [`chromote`](https://rstudio.github.io/chromote/) R package. No access to the underlying patent database is used at any point.

```{r fetch-csvs, include=FALSE}
source("R/fetch_flow_csv.R")

specs <- list(
  list(name = "LMIC_to_HIC_granted",
       country = "LMICs (excl. China)",
       toflow  = "ev_hic", granted_only = TRUE),
  list(name = "China_to_HIC_granted",
       country = "CN",
       toflow  = "ev_hic", granted_only = TRUE),
  list(name = "HIC_to_HIC_granted",
       country = "High income countries",
       toflow  = "ev_hic", granted_only = TRUE),
  list(name = "LMIC_to_HIC_all",
       country = "LMICs (excl. China)",
       toflow  = "ev_hic", granted_only = FALSE),
  list(name = "China_to_HIC_all",
       country = "CN",
       toflow  = "ev_hic", granted_only = FALSE),
  list(name = "HIC_to_HIC_all",
       country = "High income countries",
       toflow  = "ev_hic", granted_only = FALSE)
)

missing <- vapply(specs, function(s)
  !file.exists(file.path("data", paste0("flow_", s$name, ".csv"))),
  logical(1))

if (any(missing)) {
  fetch_flow_csvs(specs[missing], out_dir = "data")
}
```

```{r load-data}
read_flow <- function(file, origin_label, grant_label) {
  readr::read_csv(file.path("data", file), show_col_types = FALSE) |>
    dplyr::mutate(origin = origin_label, grant_status = grant_label)
}

combined <- dplyr::bind_rows(
  read_flow("flow_LMIC_to_HIC_granted.csv", "LMIC-origin (excl. China)", "Granted families only"),
  read_flow("flow_China_to_HIC_granted.csv","China-origin",              "Granted families only"),
  read_flow("flow_HIC_to_HIC_granted.csv",  "HIC-origin",                "Granted families only"),
  read_flow("flow_LMIC_to_HIC_all.csv",     "LMIC-origin (excl. China)", "All families"),
  read_flow("flow_China_to_HIC_all.csv",    "China-origin",              "All families"),
  read_flow("flow_HIC_to_HIC_all.csv",      "HIC-origin",                "All families")
) |>
  dplyr::filter(technology != "All") |>
  dplyr::mutate(
    origin = factor(origin,
                    levels = c("LMIC-origin (excl. China)",
                               "China-origin",
                               "HIC-origin")),
    grant_status = factor(grant_status,
                          levels = c("Granted families only", "All families"))
  )
```

## Figure — average spillover value to HICs per innovation, by technology and origin

```{r plot, fig.height=8}
tech_order <- combined |>
  dplyr::filter(origin == "HIC-origin",
                grant_status == "Granted families only") |>
  dplyr::arrange(dplyr::desc(mean)) |>
  dplyr::pull(technology)

plot_df <- combined |>
  dplyr::mutate(technology = factor(technology, levels = tech_order))

ggplot(plot_df, aes(x = technology, y = mean, fill = origin)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  geom_errorbar(
    aes(ymin = pmax(mean - sem, 0), ymax = mean + sem),
    position = position_dodge(width = 0.8),
    width = 0.2, colour = "grey30", linewidth = 0.3
  ) +
  facet_wrap(~ grant_status, ncol = 1, scales = "fixed") +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(values = c(
    "LMIC-origin (excl. China)" = "#ff7f0e",
    "China-origin"              = "#d62728",
    "HIC-origin"                = "#1f77b4"
  )) +
  labs(
    title    = "Average spillover value to HICs, per innovation",
    subtitle = "By technology, origin, and grant filter — same destination flow (ev_hic) throughout",
    x = NULL, y = "Spillover to HICs per innovation (\u00b1 1 SEM)",
    fill = NULL,
    caption = "Source: Innovation Strategy Explorer, Country Explorer CSV downloads"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x        = element_text(angle = 35, hjust = 1),
    legend.position    = "top",
    strip.text         = element_text(face = "bold"),
    panel.grid.major.x = element_blank()
  )
```

## Reading the chart

- Within each technology group: **orange** = LMIC-origin (excl. China), **red** = China-origin, **blue** = HIC-origin — all measuring the average downstream spillover value per innovation that accrues to HICs.
- Error bars are $\pm$ one standard error of the mean.
- Technologies are sorted (consistently across both facets) by HIC-origin magnitude in the *Granted families only* panel.
- The top facet keeps only **granted** families (the app's default). The bottom facet uses **all** families regardless of grant status — so it includes pending or abandoned applications. Differences between the facets show how much of the gap between origins is driven by grant selectivity versus by innovation quality itself.

China sits between the two comparison groups analytically: it's the world's largest patent-filer and accounts for the bulk of LMIC innovation volume, but its patents typically face weaker examiner citation links into HIC markets than HIC-origin patents do. Seeing where the red bar tracks orange (weak HIC absorption, like the broader LMIC pool) versus where it tracks or exceeds blue (HIC-comparable absorption) helps identify technology fields where Chinese innovation has already crossed into the HIC citation mainstream. Moving from the *Granted* facet to the *All families* facet typically shrinks HIC-origin and LMIC-origin averages unevenly — a useful test for how sensitive the ordering is to grant selection.

## Summary table

```{r table}
summary_tbl <- combined |>
  dplyr::select(technology, origin, grant_status, mean, innos) |>
  tidyr::pivot_wider(
    id_cols     = c(technology, grant_status),
    names_from  = origin,
    values_from = c(mean, innos)
  ) |>
  dplyr::arrange(grant_status, dplyr::desc(`mean_HIC-origin`))

knitr::kable(
  summary_tbl,
  digits = 2,
  col.names = c("Technology", "Grant filter",
                "Mean LMIC\u2192HIC", "Mean China\u2192HIC", "Mean HIC\u2192HIC",
                "N LMIC-origin",      "N China-origin",       "N HIC-origin"),
  caption = "Per-innovation HIC-spillover means by technology, origin, and grant filter."
)
```

## Notes & caveats

- All six bars use the same destination flow `ev_hic` — *Average Spillovers to HICs* — so the levels are directly comparable within a technology facet.
- The sample size column `innos` is **not** the same across origins: LMIC innovator pools are smaller, so their SEMs are typically larger. Treat any orange/blue gap inside a large SEM with caution.
- Grant rates differ by origin (HIC applicants are granted at higher rates than LMIC applicants on average). Mean changes between the *Granted only* and *All families* facets therefore mix two effects — average innovation quality in the population and the selectivity of the grant filter. Use the two panels together, not in isolation.
- The origin-country filter selects patents filed from the relevant country group; downstream-citing patents can come from anywhere in the world, but only HIC-located citations count toward `ev_hic`.
- Data acquisition uses only the app's public CSV-download endpoint (via `chromote`) — the patent database is never accessed directly.
