---
title: "Spillover flows between HICs and LMICs (excl. China)"
author: "Innovation Strategy Explorer — insights"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 2
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 11, fig.height = 6.5, dpi = 150)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(scales)
library(ggrepel)
```

## What this note shows

How much per-innovation spillover value flows between High-Income Countries (HICs) and LMICs (excl. China) — and how it varies by technology and by where the innovation was filed.

We run the same three views three times:

1. **No restriction** — every patent family the app's database knows about, granted or not, single- or multi-application.
2. **Granted families only** — the same sample as the original public-facing version of this note.
3. **`docdb_family_size` > 1** — restricting to multi-application patent families (PATSTAT `docdb_family_size >= 2`), which filters out one-off applications and tends to focus the sample on technologies firms commit further investment to.

Within each sample we show:

1. **Spillovers to HICs** — LMIC-origin vs HIC-origin innovations.
2. **Spillovers to LMICs (excl. China)** — LMIC-origin vs HIC-origin innovations.
3. **Spillovers to HICs, LMIC-origin only, Marimekko-style** — bar *width* encodes how many innovations LMICs contribute in each technology (so wide bars are categories where LMIC innovation volume is substantial, narrow ones where it is thin), bar *height* is the per-innovation spillover to HICs.

In figures 1 and 2, an overlaid **circle marker** reports the **LMIC share of total innovation in each category** (LMIC-origin family count ÷ all-origin family count) on a secondary axis.

## Data source

CSVs from the public Country Explorer tab of <https://mondpanther-iseapp2.share.connect.posit.cloud/>: per sample, two destination flows × three country filters = 6 CSVs. Missing CSVs are fetched automatically via headless Chrome (`chromote`) — the patent database is never accessed directly.

```{r helpers, include=FALSE}
source("R/fetch_flow_csv.R")

build_app_url <- function(country, toflow, granted_only = TRUE,
                          multifam_only = FALSE,
                          base = "https://mondpanther-iseapp2.share.connect.posit.cloud/",
                          tech_categories = c(
                            "Green Technology", "AI", "Green Transport",
                            "Green Manufacturing", "GHG Capture",
                            "Green Energy", "Adaptation", "Green Housing",
                            "Circular Economy", "Green ICT",
                            "Green Agriculture",
                            "Any Agriculture & Food technology"
                          )) {
  q <- function(x) utils::URLencode(paste0('"', x, '"'), reserved = TRUE)
  tech_arr <- paste0(
    "%5B",
    paste0(vapply(tech_categories,
                  function(t) paste0("%22", utils::URLencode(t, reserved = TRUE), "%22"),
                  character(1)),
           collapse = "%2C"),
    "%5D"
  )
  paste0(
    base,
    "?_inputs_",
    "&navbar_page=", q("Country Explorer"),
    "&country-inner_tabs=", q("Value flows by Technology"),
    "&country-country=", q(country),
    "&country-firm=", q("No firm filter"),
    "&country-toflow=", q(toflow),
    "&country-tech_categories_plot1=", tech_arr,
    "&country-granted_only=",  tolower(as.character(granted_only)),
    "&country-multifam_only=", tolower(as.character(multifam_only)),
    "&country-widthscale=", q("log"),
    "&country-display_mode=", q("confidence"),
    "&country-top_n_ids=10"
  )
}

read_flow <- function(file, origin_label) {
  readr::read_csv(file.path("data", file), show_col_types = FALSE) |>
    dplyr::mutate(origin = origin_label)
}

build_combined <- function(lmic_csv, hic_csv, all_csv) {
  combined <- dplyr::bind_rows(
    read_flow(lmic_csv, "LMIC-origin (excl. China)"),
    read_flow(hic_csv,  "HIC-origin")
  ) |>
    dplyr::filter(technology != "All") |>
    dplyr::mutate(
      origin = factor(origin,
                      levels = c("LMIC-origin (excl. China)", "HIC-origin"))
    )

  all_countries <- read_flow(all_csv, "All countries") |>
    dplyr::filter(technology != "All") |>
    dplyr::select(technology, innos_all = innos)

  lmic_share <- combined |>
    dplyr::filter(origin == "LMIC-origin (excl. China)") |>
    dplyr::select(technology, innos_lmic = innos) |>
    dplyr::left_join(all_countries, by = "technology") |>
    dplyr::mutate(share = innos_lmic / innos_all)

  list(combined = combined, lmic_share = lmic_share)
}

# Shared grouped-bar + share-circle renderer (figures 1 & 2)
plot_flows <- function(combined, lmic_share, destination_label,
                       flow_code, sample_label,
                       origin_colors = c(
                         "LMIC-origin (excl. China)" = "#ff7f0e",
                         "HIC-origin"                = "#1f77b4"
                       )) {
  tech_order <- combined |>
    dplyr::filter(origin == "HIC-origin") |>
    dplyr::arrange(dplyr::desc(mean)) |>
    dplyr::pull(technology)

  plot_df  <- combined   |> dplyr::mutate(technology = factor(technology, levels = tech_order))
  share_df <- lmic_share |> dplyr::mutate(technology = factor(technology, levels = tech_order))

  bar_ymax  <- max(plot_df$mean + plot_df$sem, na.rm = TRUE)
  share_max <- max(share_df$share, na.rm = TRUE)
  axis_cap  <- ceiling(share_max * 20) / 20
  scale_k   <- bar_ymax / axis_cap

  ggplot(plot_df, aes(x = technology, y = mean, fill = origin)) +
    geom_col(position = position_dodge(width = 0.8), width = 0.75) +
    geom_errorbar(
      aes(ymin = pmax(mean - sem, 0), ymax = mean + sem),
      position = position_dodge(width = 0.8),
      width = 0.2, colour = "grey30", linewidth = 0.3
    ) +
    geom_point(
      data = share_df,
      aes(x = technology, y = share * scale_k, shape = "LMIC share of innovation"),
      inherit.aes = FALSE,
      colour = "grey15", fill = "white", size = 4, stroke = 1.2
    ) +
    scale_shape_manual(name = NULL, values = c("LMIC share of innovation" = 21)) +
    scale_y_continuous(
      labels   = scales::comma,
      sec.axis = sec_axis(
        transform = ~ . / scale_k,
        name      = paste0("LMIC share of total innovation in category (axis capped at ",
                           scales::percent(axis_cap, accuracy = 0.1), ")"),
        labels    = scales::percent_format(accuracy = 0.1),
        breaks    = scales::pretty_breaks(n = 5)
      )
    ) +
    scale_fill_manual(values = origin_colors) +
    labs(
      title    = sprintf("Average spillover value to %s, per innovation (%s)",
                         destination_label, sample_label),
      subtitle = sprintf("By technology and origin - destination flow: %s | bars: spillover +/- SEM | circles: LMIC share of total innovation",
                         flow_code),
      x = NULL, y = sprintf("Spillover to %s per innovation (+/- 1 SEM)", destination_label),
      fill = NULL,
      caption = "Source: Innovation Strategy Explorer, Country Explorer CSV downloads"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x        = element_text(angle = 35, hjust = 1),
      legend.position    = "top",
      panel.grid.major.x = element_blank()
    )
}

# Marimekko-style LMIC-origin -> HIC plot (figure 3)
plot_marimekko <- function(combined, sample_label) {
  mm_df <- combined |>
    dplyr::filter(origin == "LMIC-origin (excl. China)") |>
    dplyr::arrange(dplyr::desc(mean)) |>
    dplyr::mutate(
      w           = innos / sum(innos),
      xmax        = cumsum(w),
      xmin        = xmax - w,
      xmid        = (xmin + xmax) / 2,
      total_value = mean * innos,
      total_label = sprintf("$%s M",
                            formatC(round(total_value), format = "d", big.mark = ","))
    )

  ggplot(mm_df) +
    geom_rect(
      aes(xmin = xmin, xmax = xmax, ymin = 0, ymax = mean),
      fill = "#ff7f0e", colour = "white", linewidth = 0.3
    ) +
    geom_errorbar(
      aes(x = xmid, ymin = pmax(mean - sem, 0), ymax = mean + sem),
      width = 0.005, colour = "grey30", linewidth = 0.4
    ) +
    ggrepel::geom_label_repel(
      aes(x = xmid, y = mean / 2, label = total_label),
      colour = "grey20", fill = "white",
      label.size = 0.25,
      label.padding = grid::unit(0.2, "lines"),
      label.r       = grid::unit(0.08, "lines"),
      size = 3, min.segment.length = 0,
      segment.colour = "grey40", segment.size = 0.3,
      box.padding = 0.4, force = 2,
      max.overlaps = Inf, direction = "both", seed = 1
    ) +
    scale_x_continuous(
      breaks = mm_df$xmid,
      labels = mm_df$technology,
      expand = expansion(mult = c(0.01, 0.01))
    ) +
    scale_y_continuous(labels = scales::comma) +
    labs(
      title    = sprintf("Spillovers to HICs per LMIC-origin innovation - bar width prop. to N innovations (%s)",
                         sample_label),
      subtitle = "Wider bars = more LMIC-origin innovations in that category | height = mean spillover",
      x = NULL, y = "Spillover to HICs per innovation (+/- 1 SEM)",
      caption = "Source: Innovation Strategy Explorer, Country Explorer CSV downloads"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x        = element_text(angle = 35, hjust = 1),
      panel.grid.major.x = element_blank()
    )
}

# Pivoted summary table (one row per technology x destination)
build_summary_table <- function(to_hic, to_lmic, caption_extra = "") {
  bind_tbl <- function(x, destination_label) {
    x$combined |>
      dplyr::select(technology, origin, mean, sem, innos) |>
      tidyr::pivot_wider(
        id_cols     = technology,
        names_from  = origin,
        values_from = c(mean, sem, innos)
      ) |>
      dplyr::mutate(destination = destination_label, .before = 1)
  }

  summary_tbl <- dplyr::bind_rows(
    bind_tbl(to_hic,  "HICs"),
    bind_tbl(to_lmic, "LMICs (excl. China)")
  ) |>
    dplyr::mutate(
      ratio_LMIC_over_HIC = `mean_LMIC-origin (excl. China)` / `mean_HIC-origin`
    ) |>
    dplyr::arrange(destination, dplyr::desc(`mean_HIC-origin`))

  knitr::kable(
    summary_tbl,
    digits = 2,
    col.names = c("Destination", "Technology",
                  "Mean LMIC->dest", "Mean HIC->dest",
                  "SEM LMIC->dest",  "SEM HIC->dest",
                  "N LMIC-origin",   "N HIC-origin",
                  "Ratio LMIC / HIC"),
    caption = paste0("Per-innovation spillover means by destination, technology, and origin",
                     caption_extra, ".")
  )
}

# Top-level driver: builds 6 specs, fetches missing CSVs, returns the
# combined datasets needed by the figures and the summary table.
run_sample <- function(suffix, granted_only, multifam_only,
                       data_dir = "data") {
  stopifnot(is.character(suffix), nchar(suffix) > 0)

  specs <- list(
    list(country = "LMICs (excl. China)",   toflow = "ev_hic"),
    list(country = "High income countries", toflow = "ev_hic"),
    list(country = "All countries",         toflow = "ev_hic"),
    list(country = "LMICs (excl. China)",   toflow = "ev_emdenocn"),
    list(country = "High income countries", toflow = "ev_emdenocn"),
    list(country = "All countries",         toflow = "ev_emdenocn")
  )
  short <- c("LMIC", "HIC", "All", "LMIC", "HIC", "All")
  dest  <- c("HIC",  "HIC", "HIC", "LMIC", "LMIC", "LMIC")
  for (i in seq_along(specs)) {
    specs[[i]]$name          <- sprintf("%s_to_%s_%s", short[i], dest[i], suffix)
    specs[[i]]$granted_only  <- granted_only
    specs[[i]]$multifam_only <- multifam_only
  }

  missing <- vapply(specs, function(s)
    !file.exists(file.path(data_dir, paste0("flow_", s$name, ".csv"))),
    logical(1))
  if (any(missing)) fetch_flow_csvs(specs[missing], out_dir = data_dir)

  list(
    to_hic = build_combined(
      sprintf("flow_LMIC_to_HIC_%s.csv", suffix),
      sprintf("flow_HIC_to_HIC_%s.csv",  suffix),
      sprintf("flow_All_to_HIC_%s.csv",  suffix)),
    to_lmic = build_combined(
      sprintf("flow_LMIC_to_LMIC_%s.csv", suffix),
      sprintf("flow_HIC_to_LMIC_%s.csv",  suffix),
      sprintf("flow_All_to_LMIC_%s.csv",  suffix))
  )
}
```

---

# Sample 1: no restriction

Every family in the app's database, regardless of grant status or family
size. This is the broadest universe — useful as a baseline before
applying the granted / multi-application filters in the next two
sections.

```{r build-all, results='asis'}
a <- run_sample("all",
                granted_only  = FALSE,
                multifam_only = FALSE)
cat(sprintf(
  "App-view URL templates: [LMIC->HIC](%s) | [HIC->HIC](%s) | [LMIC->LMIC](%s) | [HIC->LMIC](%s)\n\n",
  build_app_url("LMICs (excl. China)",   "ev_hic",      FALSE, FALSE),
  build_app_url("High income countries", "ev_hic",      FALSE, FALSE),
  build_app_url("LMICs (excl. China)",   "ev_emdenocn", FALSE, FALSE),
  build_app_url("High income countries", "ev_emdenocn", FALSE, FALSE)
))
```

## Figure 1 (no restriction) — Spillovers to HICs

```{r fig-hic-all}
plot_flows(a$to_hic$combined, a$to_hic$lmic_share,
           destination_label = "HICs",
           flow_code         = "ev_hic",
           sample_label      = "no restriction")
```

## Figure 2 (no restriction) — Spillovers to LMICs (excl. China)

```{r fig-lmic-all}
plot_flows(a$to_lmic$combined, a$to_lmic$lmic_share,
           destination_label = "LMICs (excl. China)",
           flow_code         = "ev_emdenocn",
           sample_label      = "no restriction")
```

## Figure 3 (no restriction) — LMIC-origin only, bar width prop. to innovation count

```{r fig-marimekko-all, fig.height=6}
plot_marimekko(a$to_hic$combined, sample_label = "no restriction")
```

## Summary table (no restriction)

```{r table-all}
build_summary_table(a$to_hic, a$to_lmic,
                    caption_extra = " (no restriction)")
```

---

# Sample 2: granted families only

```{r build-granted, results='asis'}
g <- run_sample("granted",
                granted_only  = TRUE,
                multifam_only = FALSE)
cat(sprintf(
  "App-view URL templates: [LMIC->HIC](%s) | [HIC->HIC](%s) | [LMIC->LMIC](%s) | [HIC->LMIC](%s)\n\n",
  build_app_url("LMICs (excl. China)",   "ev_hic",      TRUE, FALSE),
  build_app_url("High income countries", "ev_hic",      TRUE, FALSE),
  build_app_url("LMICs (excl. China)",   "ev_emdenocn", TRUE, FALSE),
  build_app_url("High income countries", "ev_emdenocn", TRUE, FALSE)
))
```

## Figure 1 (granted) — Spillovers to HICs

```{r fig-hic-granted}
plot_flows(g$to_hic$combined, g$to_hic$lmic_share,
           destination_label = "HICs",
           flow_code         = "ev_hic",
           sample_label      = "granted families only")
```

## Figure 2 (granted) — Spillovers to LMICs (excl. China)

```{r fig-lmic-granted}
plot_flows(g$to_lmic$combined, g$to_lmic$lmic_share,
           destination_label = "LMICs (excl. China)",
           flow_code         = "ev_emdenocn",
           sample_label      = "granted families only")
```

## Figure 3 (granted) — LMIC-origin only, bar width prop. to innovation count

```{r fig-marimekko-granted, fig.height=6}
plot_marimekko(g$to_hic$combined, sample_label = "granted families only")
```

## Summary table (granted)

```{r table-granted}
build_summary_table(g$to_hic, g$to_lmic,
                    caption_extra = " (granted families only)")
```

---

# Sample 3: `docdb_family_size` > 1

Restricts to families with at least two applications in PATSTAT
(`docdb_family_size >= 2`). The granted filter is left **off** here so the
multi-application restriction is the only difference versus the granted
sample above.

```{r build-multifam, results='asis'}
m <- run_sample("multifam",
                granted_only  = FALSE,
                multifam_only = TRUE)
cat(sprintf(
  "App-view URL templates: [LMIC->HIC](%s) | [HIC->HIC](%s) | [LMIC->LMIC](%s) | [HIC->LMIC](%s)\n\n",
  build_app_url("LMICs (excl. China)",   "ev_hic",      FALSE, TRUE),
  build_app_url("High income countries", "ev_hic",      FALSE, TRUE),
  build_app_url("LMICs (excl. China)",   "ev_emdenocn", FALSE, TRUE),
  build_app_url("High income countries", "ev_emdenocn", FALSE, TRUE)
))
```

## Figure 1 (fam_size > 1) — Spillovers to HICs

```{r fig-hic-multifam}
plot_flows(m$to_hic$combined, m$to_hic$lmic_share,
           destination_label = "HICs",
           flow_code         = "ev_hic",
           sample_label      = "fam_size >= 2")
```

## Figure 2 (fam_size > 1) — Spillovers to LMICs (excl. China)

```{r fig-lmic-multifam}
plot_flows(m$to_lmic$combined, m$to_lmic$lmic_share,
           destination_label = "LMICs (excl. China)",
           flow_code         = "ev_emdenocn",
           sample_label      = "fam_size >= 2")
```

## Figure 3 (fam_size > 1) — LMIC-origin only, bar width prop. to innovation count

```{r fig-marimekko-multifam, fig.height=6}
plot_marimekko(m$to_hic$combined, sample_label = "fam_size >= 2")
```

## Summary table (fam_size > 1)

```{r table-multifam}
build_summary_table(m$to_hic, m$to_lmic,
                    caption_extra = " (fam_size >= 2)")
```

---

## Reading the charts

- **Figures 1 & 2** — Within each technology group, the **blue** bar is the average destination-bound spillover per HIC-origin innovation; the **orange** bar is the same figure for LMIC-origin innovations (excl. China). The **open circle** is the LMIC share of total innovation in the category (right axis). Error bars are ± 1 SEM. Technologies are sorted left-to-right by HIC-origin spillover magnitude within each figure.
- **Figure 3** — Only LMIC-origin bars, spillover *to HICs*. Width proportional to the number of LMIC-origin innovations in the category: this deliberately combines intensity (height) and volume (width), so the total area of each bar is proportional to the **aggregate spillover value flowing to HICs from LMIC innovation in that technology**. Each white-boxed label reports that aggregate (per-innovation mean × number of innovations) in millions of USD; labels are offset by `ggrepel` with a leader line pointing to the bar they describe.

## Notes & caveats

- Bars in Figures 1 and 2 use the same destination flow within each figure, so the levels are directly comparable within a technology.
- The sample size column `innos` is **not** the same across origins: LMIC innovator pools are smaller, so their SEMs are typically larger. Treat any orange/blue gap inside a large SEM with caution.
- The three top-level samples are: (1) no restriction (the underlying universe), (2) "granted families only" UI checkbox, (3) "multi-application families only" checkbox (i.e. PATSTAT `docdb_family_size >= 2`). Samples 2 and 3 are **not** subsets of each other — a granted family can be a single application, and a multi-application family need not be granted.
- The origin-country filter selects patents filed from the relevant country group; downstream-citing patents can come from anywhere in the world, but only destination-country citations count toward the `ev_*` flow.
- Data acquisition uses only the app's public CSV-download endpoint (via `chromote`) — the patent database is never accessed directly.

---

```{r version-stamp, results='asis'}
local({
  util <- "../R/utils_version.R"
  if (file.exists(util)) source(util, local = TRUE)
})
v <- tryCatch(app_version(), error = function(e) "unknown")
cat(sprintf(
  "<p style='font-size:0.85em; color:#666;'>Rendered %s &middot; app version <code>%s</code></p>",
  format(Sys.time(), "%Y-%m-%d %H:%M %Z"), v
))
```
