---
title: "Spillovers to LMICs (excl. China) — LMIC-origin vs HIC-origin innovations"
author: "Innovation Strategy Explorer — insights"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    self_contained: true
    toc: false
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 11, fig.height = 6.5, dpi = 150)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(scales)
```

## What this note shows

How much per-innovation spillover value do different technologies generate **for LMICs excluding China** — and does this depend on where the innovation was filed? The grouped bar chart below compares two populations of innovations, both scored on the *same* destination flow (*Average Spillovers to LMICs (excl. China)* = `ev_emdenocn`) and both restricted to **granted** patent families (the app's default):

- **LMIC-origin** innovations, excluding China
- **HIC-origin** innovations

Technologies are on the x-axis; within each technology the two bars show the average spillover value per innovation that flows to LMICs (excl. China), for the two origin groups. An overlaid **circle marker** reports the **LMIC share of total innovation in that category** (LMIC-origin family count ÷ all-origin family count), read off a secondary axis.

## Data source

Three downloads from the public Country Explorer tab of <https://mondpanther-iseapp2.share.connect.posit.cloud/>, with **Value flow = *Average Spillovers to LMICs (excl. China)*** and **Granted families only** in all three cases, and only the **Country / Group** filter changing. Click any "open in app" link below to jump directly to the bookmarked view used for that row:

```{r build-app-urls, include=FALSE}
build_app_url <- function(country, toflow, granted_only = TRUE,
                          base = "https://mondpanther-iseapp2.share.connect.posit.cloud/",
                          tech_categories = c(
                            "Green Technology", "AI", "Green Transport",
                            "Green Manufacturing", "GHG Capture",
                            "Green Energy", "Adaptation", "Green Housing",
                            "Circular Economy", "Green ICT",
                            "Green Agriculture",
                            "Any Agriculture & Food technology"
                          )) {
  q <- function(x) utils::URLencode(paste0('"', x, '"'), reserved = TRUE)
  tech_arr <- paste0(
    "%5B",
    paste0(vapply(tech_categories,
                  function(t) paste0("%22", utils::URLencode(t, reserved = TRUE), "%22"),
                  character(1)),
           collapse = "%2C"),
    "%5D"
  )
  paste0(
    base,
    "?_inputs_",
    "&navbar_page=", q("Country Explorer"),
    "&country-inner_tabs=", q("Value flows by Technology"),
    "&country-country=", q(country),
    "&country-firm=", q("No firm filter"),
    "&country-toflow=", q(toflow),
    "&country-tech_categories_plot1=", tech_arr,
    "&country-granted_only=", tolower(as.character(granted_only)),
    "&country-widthscale=", q("log"),
    "&country-display_mode=", q("confidence"),
    "&country-top_n_ids=10"
  )
}
url_lmic <- build_app_url("LMICs (excl. China)",   "ev_emdenocn", granted_only = TRUE)
url_hic  <- build_app_url("High income countries", "ev_emdenocn", granted_only = TRUE)
url_all  <- build_app_url("All countries",         "ev_emdenocn", granted_only = TRUE)
```

| File in `insights/data/` | `country-country` filter | Granted only? | App view |
|---|---|---|---|
| `flow_LMIC_to_LMIC_granted.csv` | `LMICs (excl. China)`                 | yes | [open in app](`r url_lmic`) |
| `flow_HIC_to_LMIC_granted.csv`  | `High income countries`               | yes | [open in app](`r url_hic`)  |
| `flow_All_to_LMIC_granted.csv`  | `All countries` (for share denominator) | yes | [open in app](`r url_all`)  |

If any CSV is missing, the chunk below fetches it via headless Chrome using the [`chromote`](https://rstudio.github.io/chromote/) R package. No access to the underlying patent database is used at any point.

```{r fetch-csvs, include=FALSE}
source("R/fetch_flow_csv.R")

specs <- list(
  list(name = "LMIC_to_LMIC_granted",
       country = "LMICs (excl. China)",
       toflow  = "ev_emdenocn", granted_only = TRUE),
  list(name = "HIC_to_LMIC_granted",
       country = "High income countries",
       toflow  = "ev_emdenocn", granted_only = TRUE),
  list(name = "All_to_LMIC_granted",
       country = "All countries",
       toflow  = "ev_emdenocn", granted_only = TRUE)
)

missing <- vapply(specs, function(s)
  !file.exists(file.path("data", paste0("flow_", s$name, ".csv"))),
  logical(1))

if (any(missing)) {
  fetch_flow_csvs(specs[missing], out_dir = "data")
}
```

```{r load-data}
read_flow <- function(file, origin_label) {
  readr::read_csv(file.path("data", file), show_col_types = FALSE) |>
    dplyr::mutate(origin = origin_label)
}

combined <- dplyr::bind_rows(
  read_flow("flow_LMIC_to_LMIC_granted.csv", "LMIC-origin (excl. China)"),
  read_flow("flow_HIC_to_LMIC_granted.csv",  "HIC-origin")
) |>
  dplyr::filter(technology != "All") |>
  dplyr::mutate(
    origin = factor(origin,
                    levels = c("LMIC-origin (excl. China)", "HIC-origin"))
  )

# LMIC share of total innovation per technology: innos_LMIC / innos_AllCountries
all_countries <- read_flow("flow_All_to_LMIC_granted.csv", "All countries") |>
  dplyr::filter(technology != "All") |>
  dplyr::select(technology, innos_all = innos)

lmic_share <- combined |>
  dplyr::filter(origin == "LMIC-origin (excl. China)") |>
  dplyr::select(technology, innos_lmic = innos) |>
  dplyr::left_join(all_countries, by = "technology") |>
  dplyr::mutate(share = innos_lmic / innos_all)
```

## Figure — average spillover value to LMICs (excl. China) per innovation, by technology and origin

```{r plot}
tech_order <- combined |>
  dplyr::filter(origin == "HIC-origin") |>
  dplyr::arrange(dplyr::desc(mean)) |>
  dplyr::pull(technology)

plot_df <- combined |>
  dplyr::mutate(technology = factor(technology, levels = tech_order))

share_df <- lmic_share |>
  dplyr::mutate(technology = factor(technology, levels = tech_order))

# Scale factor: map share (0..1) onto the bar-value axis so both can coexist.
# Tune scale_k so the largest observed share lands near the top of the plot —
# without this, shares of a few percent would be crushed against the x-axis.
bar_ymax   <- max(plot_df$mean + plot_df$sem, na.rm = TRUE)
share_max  <- max(share_df$share, na.rm = TRUE)
# round the sec-axis cap up to a nice break above the largest observed share
axis_cap   <- ceiling(share_max * 20) / 20
scale_k    <- bar_ymax / axis_cap

ggplot(plot_df, aes(x = technology, y = mean, fill = origin)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  geom_errorbar(
    aes(ymin = pmax(mean - sem, 0), ymax = mean + sem),
    position = position_dodge(width = 0.8),
    width = 0.2, colour = "grey30", linewidth = 0.3
  ) +
  geom_point(
    data = share_df,
    aes(x = technology, y = share * scale_k, shape = "LMIC share of innovation"),
    inherit.aes = FALSE,
    colour = "grey15", fill = "white", size = 4, stroke = 1.2
  ) +
  scale_shape_manual(name = NULL, values = c("LMIC share of innovation" = 21)) +
  scale_y_continuous(
    labels   = scales::comma,
    sec.axis = sec_axis(
      transform = ~ . / scale_k,
      name      = paste0("LMIC share of total innovation in category (axis capped at ",
                         scales::percent(axis_cap, accuracy = 0.1), ")"),
      labels    = scales::percent_format(accuracy = 0.1),
      breaks    = scales::pretty_breaks(n = 5)
    )
  ) +
  scale_fill_manual(values = c(
    "LMIC-origin (excl. China)" = "#ff7f0e",
    "HIC-origin"                = "#1f77b4"
  )) +
  labs(
    title    = "Average spillover value to LMICs (excl. China), per innovation (granted families only)",
    subtitle = "By technology and origin — bars: spillover \u00b1 SEM · circles: LMIC share of total innovation in the category",
    x = NULL, y = "Spillover to LMICs per innovation (\u00b1 1 SEM)",
    fill = NULL,
    caption = "Source: Innovation Strategy Explorer, Country Explorer CSV downloads"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x        = element_text(angle = 35, hjust = 1),
    legend.position    = "top",
    panel.grid.major.x = element_blank()
  )
```

## Reading the chart

- Within each technology group, the **blue** bar is the average LMIC-bound spillover per HIC-origin innovation; the **orange** bar is the same figure for LMIC-origin innovations (excl. China). Read against the **left axis** (spillover per innovation).
- The **open circle** above each group is the LMIC share of total innovation in that technology category — read against the **right axis**.
- Error bars are $\pm$ one standard error of the mean.
- Technologies are sorted left-to-right by HIC-origin magnitude.

If orange sits *above* blue for a technology, LMIC-filed innovations in that field inject more downstream value into other LMICs per innovation than HIC-filed innovations do. A blue-above-orange pattern, conversely, suggests HIC-origin innovations are more strongly absorbed into LMIC citation networks (either because they set the technological frontier or because LMIC examiners cite them disproportionately).

## Summary table

```{r table}
summary_tbl <- combined |>
  dplyr::select(technology, origin, mean, sem, innos) |>
  tidyr::pivot_wider(
    id_cols     = technology,
    names_from  = origin,
    values_from = c(mean, sem, innos)
  ) |>
  dplyr::mutate(
    ratio_LMIC_over_HIC = `mean_LMIC-origin (excl. China)` / `mean_HIC-origin`
  ) |>
  dplyr::arrange(dplyr::desc(`mean_HIC-origin`))

knitr::kable(
  summary_tbl,
  digits = 2,
  col.names = c("Technology",
                "Mean LMIC\u2192LMIC", "Mean HIC\u2192LMIC",
                "SEM LMIC\u2192LMIC",  "SEM HIC\u2192LMIC",
                "N LMIC-origin",       "N HIC-origin",
                "Ratio LMIC / HIC"),
  caption = "Per-innovation LMIC-spillover means by technology and origin (granted families only)."
)
```

## Notes & caveats

- Both bars use the same destination flow `ev_emdenocn` — *Average Spillovers to LMICs (excl. China)* — so the levels are directly comparable within a technology.
- The sample size column `innos` is **not** the same across origins: LMIC innovator pools are smaller, so their SEMs are typically larger. Treat any orange/blue gap inside a large SEM with caution.
- Only **granted** patent families are included (the app's default).
- The origin-country filter selects patents filed from the relevant country group; downstream-citing patents can come from anywhere in the world, but only LMIC (excl. China) citations count toward `ev_emdenocn`.
- Data acquisition uses only the app's public CSV-download endpoint (via `chromote`) — the patent database is never accessed directly.

---

```{r version-stamp, results='asis'}
# Reuse the same app_version() function the Shiny "About" page uses.
local({
  util <- "../R/utils_version.R"
  if (file.exists(util)) source(util, local = TRUE)
})
v <- tryCatch(app_version(), error = function(e) "unknown")
cat(sprintf(
  "<p style='font-size:0.85em; color:#666;'>Rendered %s &middot; app version <code>%s</code></p>",
  format(Sys.time(), "%Y-%m-%d %H:%M %Z"), v
))
```
