---
title: "City-level innovation spillovers: global vs. LMICs (excl. China)"
author: "Innovation Strategy Explorer — insights"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 2
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 12, fig.height = 7.5, dpi = 150)

library(dplyr)
library(ggplot2)
library(scales)
library(ggrepel)
library(DBI)
library(duckdb)
library(htmltools)
```

## What this note shows

For every city in the app's `countrymap`, two per-city averages are
computed:

* **Average Global Spillover** — the mean of `ev_global` (Watson average
  global spillover per patent family) across all docdbs mapped to the
  city.
* **Average Spillover to LMICs (excl. China)** — the mean of
  `ev_emdenocn` across the same docdbs, i.e. the share of those
  spillovers that land in low- and middle-income countries other than
  China.

The scatter plot below shows one dot per city. A linear trend line
summarises the overall relationship; the largest cities (by patent
count) are highlighted, and a hand-picked list of LMIC / mid-tier
cities is always labelled regardless of size.

```{r data}
# Locate inst/extdata regardless of where the Rmd is rendered from.
find_extdata <- function(file) {
  cand <- c(
    system.file("extdata", file, package = "innovationStrategyExplorer"),
    file.path("..", "inst", "extdata", file),
    file.path("inst", "extdata", file)
  )
  hit <- cand[nzchar(cand) & file.exists(cand)]
  if (!length(hit)) stop("Could not locate ", file)
  normalizePath(hit[[1]], winslash = "/")
}
pdb_path <- find_extdata("patent_database.parquet")
cm_path  <- find_extdata("countrymap.parquet")
pxt_path <- find_extdata("patents_x_tech.parquet")
tl_path  <- find_extdata("tech_lookup.parquet")

# Pull the per-city averages directly from DuckDB so we never hold the
# 28 M-row patent_database in R memory. We:
#   1. Collapse patent_database to one row per docdb_family_id (ev_*
#      values don't vary across countries within a family).
#   2. Tag each docdb as Green Technology if any of its
#      patents_x_tech rows resolve to either tech_group = "Green
#      Technology" OR technology = "Green Technology" — same logic
#      the Country Explorer's tech filter uses, so the share matches
#      what HiGGlo / Country Explorer "Green Technology" filtering
#      would produce.
#   3. Join countrymap to attach the city + capital-fallback flag.
#   4. Drop capital-fallback rows (geocode_missing = TRUE) — those
#      entries are positioned at the country's capital simply because
#      no real address survived geocoding, so they would inflate the
#      patent count of the wrong city.
#   5. Aggregate per city: n_docdbs, n_green, avg ev_global overall,
#      avg ev_global on the green subset, avg ev_emdenocn overall.
con <- dbConnect(duckdb::duckdb())
city_stats <- dbGetQuery(con, sprintf("
  WITH
    p AS (SELECT * FROM read_parquet('%s')),
    cm AS (
      SELECT * FROM read_parquet('%s')
      WHERE geocode_missing = FALSE
        AND city IS NOT NULL
        AND city <> ''
    ),
    pxt AS (SELECT * FROM read_parquet('%s')),
    tl  AS (SELECT * FROM read_parquet('%s')),
    -- Distinct docdbs flagged green via tech_lookup. tech_group OR
    -- technology so e.g. 'Battery Technology' (a sub-tech under the
    -- Green Technology umbrella) also flips the flag.
    green AS (
      SELECT DISTINCT pxt.docdb_family_id
      FROM pxt
      JOIN tl ON tl.technology = pxt.technology
      WHERE tl.tech_group = 'Green Technology'
         OR pxt.technology = 'Green Technology'
    ),
    docdb AS (
      SELECT
        p.docdb_family_id,
        ANY_VALUE(p.ev_global)   AS ev_global,
        ANY_VALUE(p.ev_emdenocn) AS ev_emdenocn,
        CASE WHEN MAX(g.docdb_family_id) IS NOT NULL THEN 1 ELSE 0 END
                                  AS is_green
      FROM p
      LEFT JOIN green g ON g.docdb_family_id = p.docdb_family_id
      WHERE p.ev_global IS NOT NULL
      GROUP BY p.docdb_family_id
    ),
    docdb_city AS (
      -- Each (docdb, ctry) row in countrymap gives one city to credit.
      -- A single docdb may show up under multiple cities (multi-country
      -- families); each city sees the docdb's full ev value, which is
      -- the natural per-city interpretation.
      SELECT DISTINCT cm.city, d.docdb_family_id,
             d.ev_global, d.ev_emdenocn, d.is_green
      FROM cm
      JOIN docdb d ON d.docdb_family_id = cm.docdb_family_id
    )
  SELECT
    city,
    COUNT(DISTINCT docdb_family_id)                       AS n_docdbs,
    SUM(is_green)                                         AS n_green,
    AVG(ev_global)                                        AS avg_ev_global,
    AVG(ev_emdenocn)                                      AS avg_ev_emdenocn,
    AVG(CASE WHEN is_green = 1 THEN ev_global END)        AS avg_ev_global_green
  FROM docdb_city
  GROUP BY city
", pdb_path, cm_path, pxt_path, tl_path))
dbDisconnect(con, shutdown = TRUE)

# Derived city-level features.
#   share_green  = fraction of city's docdbs flagged Green Technology
#   green_premium = avg_ev_global_green / avg_ev_global
#                   - 1 means green docdbs match the city average
#                   > 1 means green docdbs over-perform vs the city's
#                     overall innovation footprint
#                   NA when there are no green docdbs OR avg is 0
city_stats <- city_stats |>
  dplyr::mutate(
    share_green   = ifelse(n_docdbs > 0L, n_green / n_docdbs, NA_real_),
    green_premium = dplyr::if_else(
      !is.na(avg_ev_global_green) &
        !is.na(avg_ev_global) &
        avg_ev_global != 0,
      avg_ev_global_green / avg_ev_global,
      NA_real_
    )
  )

cat(sprintf(
  "cities seen: %s   (median patents/city: %s, max: %s, share_green median: %.1f%%)\n",
  format(nrow(city_stats),                                    big.mark = ","),
  format(stats::median(city_stats$n_docdbs),                  big.mark = ","),
  format(max(city_stats$n_docdbs),                            big.mark = ","),
  100 * stats::median(city_stats$share_green, na.rm = TRUE)
))
```

## City selection

To stop the scatter from being dominated by tiny cities with one
outlier patent, we keep only those with at least `min_patents`
distinct patent families. **The hand-picked label set is always kept,
regardless of size** — otherwise small LMIC cities (Belém has only ~14
families, Luang Prabang fewer still) would be dropped from the plot
and from the abstract-summary section before they can be flagged.

The patent universe here is the **"All innovations"** view: every
docdb in `patent_database` with a non-null `ev_global`, no tech
filter and no granted-only filter. This matches the HiGGlo "All
innovations" tab counts exactly — e.g. Belém: 14 families total,
8 granted.

```{r labels}
min_patents <- 10L

# Hand-picked cities. Each entry is a regex (case-insensitive) so we
# can match common spelling variants — Belém vs Belem, Antalya vs
# Antaya, etc.
required_patterns <- c(
  "Bogot[aá]",                       # Bogotá, Colombia
  "Bel[eé]m",                        # Belém, Brazil
  "Ibadan",                          # Ibadan, Nigeria
  "Konya",                           # Konya, Turkey
  "Antalya|Antaya",                  # Antalya, Turkey
  "Sofia|Sofiya",                    # Sofia, Bulgaria
  "Luang ?Prabang|Luangprabang|Luanpraban"  # Luang Prabang, Laos
)
required_re <- paste(required_patterns, collapse = "|")

cities <- city_stats |>
  dplyr::mutate(
    is_required = grepl(required_re, city, ignore.case = TRUE)
  ) |>
  # Keep cities that EITHER pass the threshold OR are in the
  # required label set. Without the OR-required clause, small LMIC
  # cities get dropped before `is_required` even gets a chance to
  # tag them.
  dplyr::filter((n_docdbs >= min_patents) | is_required) |>
  dplyr::filter(!is.na(avg_ev_global), !is.na(avg_ev_emdenocn)) |>
  dplyr::mutate(
    city_lc      = tolower(city),
    rank_by_size = dplyr::min_rank(dplyr::desc(n_docdbs))
  )

cat(sprintf(
  "after filter (>= %d patents OR in required list): %s cities kept (%.1f%% of all)\n",
  min_patents,
  format(nrow(cities),  big.mark = ","),
  100 * nrow(cities) / nrow(city_stats)))

big_n  <- 25L
cities <- cities |>
  dplyr::mutate(
    is_big   = rank_by_size <= big_n,
    label_me = is_required | is_big
  )
```

```{r helper_links}
# Helpers shared by the HiGGlo capture chunk, the HiGGlo PNG render
# chunk, and the per-city summary render chunk. Defined once here so
# all three sections produce the same URL / file paths and so the
# render chunks can hyperlink even when `chromote` isn't installed.

# Live app endpoint. Single source of truth — change here if the
# Posit Connect URL ever moves.
app_root <- "https://mondpanther-iseapp2.share.connect.posit.cloud/"

# Build the bookmarked HiGGlo URL for a given city. Same shape Shiny
# native bookmarks emit (each value is JSON, then URL-encoded). The
# city literal is URL-encoded then wrapped in %22...%22 so accented
# names like "Belém" survive UTF-8 round-tripping.
build_higglo_url <- function(city) {
  city_enc <- utils::URLencode(city, reserved = TRUE)
  paste0(
    app_root,
    "?_inputs_",
    "&hglobe-gen0_select_mode=%22Random%22",
    "&hglobe-gen0_unit_mode=%22Number%22",
    "&hglobe-edge_select_mode=%22Random%22",
    "&hglobe-edge_unit_mode=%22Percent%22",
    "&hglobe-render_map=1",
    "&hglobe-next_step=1",
    "&hglobe-sidebar=true",
    "&navbar_page=%22HiGGlobe%22",
    "&hglobe-country=%22All%20countries%22",
    "&hglobe-firm=%22No%20firm%20filter%22",
    "&hglobe-techs=%22All%20innovations%22",
    "&hglobe-city=%22", city_enc, "%22",
    "&hglobe-toflow=%22ev_global%22",
    "&hglobe-granted_only=false",
    "&hglobe-multifam_only=false",
    "&hglobe-include_fallback=false",
    "&hglobe-gen0_sample_val=100",
    "&hglobe-edge_sample_val=100",
    "&hglobe-add_generations=2",
    "&hglobe-map_zoom=2",
    "&hglobe-map_center=%7B%22lng%22%3A10%2C%22lat%22%3A20%7D"
  )
}

# Filesystem-safe slug for the cached PNG.
city_png_path <- function(city) {
  slug <- gsub("[^A-Za-z0-9]+", "_", city)
  slug <- gsub("^_|_$", "", slug)
  file.path(cache_dir, sprintf("higglo_%s.png", slug))
}

# Markdown anchor pointing to the live HiGGlo view for a city.
# `target="_blank"` so clicking doesn't navigate the user away from
# the rendered insight HTML; arrow glyph signals "external link".
higglo_link_md <- function(city) {
  sprintf(
    "<a href='%s' target='_blank' rel='noopener'>Open %s on HiGGlo &rarr;</a>",
    build_higglo_url(city), city
  )
}

# Quick visibility check: report whether each required city was found.
required_check <- tibble::tibble(
  pattern = required_patterns,
  matches = vapply(required_patterns, function(p) {
    paste(cities$city[grepl(p, cities$city, ignore.case = TRUE)],
          collapse = ", ")
  }, character(1))
)
required_check
```

## Scatter plot

```{r scatter, fig.width=12, fig.height=8}
# Background dots = every city. Highlighted dots overlay the big-city
# anchors and the required-label set so they pop out.
base_size <- 1.4
hi_size   <- 3.2

p <- ggplot(cities,
            aes(x = avg_ev_global, y = avg_ev_emdenocn)) +
  # Background cloud
  geom_point(aes(size = n_docdbs),
             colour = "#a8a8a8", alpha = 0.35) +
  # Linear trend across all kept cities
  geom_smooth(method = "lm", formula = y ~ x,
              colour = "#1f77b4", fill = "#1f77b4", alpha = 0.15,
              linewidth = 0.7) +
  # Big-city anchors (orange)
  geom_point(data = subset(cities, is_big & !is_required),
             colour = "#ff7f0e", size = hi_size, alpha = 0.9) +
  # Required label set (red)
  geom_point(data = subset(cities, is_required),
             colour = "#d62728", size = hi_size + 0.6, alpha = 0.95) +
  # Repel labels for both sets, drawn together so the layout solver
  # avoids the highlighted dots regardless of category.
  ggrepel::geom_text_repel(
    data            = subset(cities, label_me),
    aes(label = city,
        colour      = is_required),
    size            = 3.4,
    fontface        = "bold",
    box.padding     = 0.45,
    point.padding   = 0.35,
    min.segment.length = 0,
    segment.colour  = "grey50",
    segment.size    = 0.3,
    max.overlaps    = Inf,
    show.legend     = FALSE
  ) +
  scale_colour_manual(values = c(`TRUE` = "#d62728", `FALSE` = "#333333")) +
  scale_size_continuous(
    name   = "Patent families per city",
    range  = c(0.6, 8),
    breaks = c(50, 500, 5000, 50000),
    labels = scales::label_comma()
  ) +
  scale_x_continuous(labels = scales::label_number(accuracy = 0.01)) +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.01)) +
  labs(
    title    = "City-level innovation spillovers",
    subtitle = sprintf(
      "All cities with >=%d patent families. Labelled: top %d by size + hand-picked LMIC / mid-tier cities.",
      min_patents, big_n),
    x        = "Average Global Spillover  (mean ev_global per city)",
    y        = "Average Spillover to LMICs excl. China  (mean ev_emdenocn per city)",
    caption  = "Each dot is one city. Trend line: OLS of LMICs-excl-China spillover on global spillover."
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title       = element_text(face = "bold"),
    plot.subtitle    = element_text(colour = "#555"),
    panel.grid.minor = element_blank(),
    legend.position  = "right"
  )

print(p)
```

## Trend-line fit

```{r fit}
fit <- stats::lm(avg_ev_emdenocn ~ avg_ev_global, data = cities)
fit_s <- summary(fit)
cat(sprintf(
  "OLS:  avg_ev_emdenocn = %.4f + %.4f * avg_ev_global\n  R^2 = %.3f, n = %d cities\n",
  stats::coef(fit)[1L],
  stats::coef(fit)[2L],
  fit_s$r.squared,
  nrow(cities)))
```

## Scatter: green-tech spillover vs overall spillover

For every city we now have **two** average global-spillover numbers:
the overall mean across all of its docdbs, and the mean restricted to
docdbs flagged as Green Technology. Plotting them against each other
shows whether a city's green innovations punch above or below its
overall innovation-spillover footprint. The dashed grey line is the
1:1 reference — points above it are cities whose green innovations
generate higher-than-average global spillover for that city; points
below sit on the other side of the gap.

Cities with zero green docdbs drop out of this view automatically
(no green mean to plot). The same big-city anchors and hand-picked
LMIC cities are highlighted as in the first scatter.

```{r scatter_green, fig.width=12, fig.height=8}
cities_green <- cities |>
  dplyr::filter(!is.na(avg_ev_global_green))

axis_lim <- range(c(cities_green$avg_ev_global,
                    cities_green$avg_ev_global_green),
                  na.rm = TRUE, finite = TRUE)

p_green <- ggplot(cities_green,
                  aes(x = avg_ev_global, y = avg_ev_global_green)) +
  # 1:1 reference line; anything above means green spillover beats
  # the city's overall mean, anything below means it lags.
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", colour = "#888888", linewidth = 0.5) +
  # Background cloud, sized by the green-only docdb count so a city
  # with 1 green family doesn't visually outweigh one with hundreds.
  geom_point(aes(size = n_green),
             colour = "#a8a8a8", alpha = 0.35) +
  # OLS trend line over the city set
  geom_smooth(method = "lm", formula = y ~ x,
              colour = "#2ca02c", fill = "#2ca02c", alpha = 0.15,
              linewidth = 0.7) +
  # Big-city anchors (orange)
  geom_point(data = subset(cities_green, is_big & !is_required),
             colour = "#ff7f0e", size = hi_size, alpha = 0.9) +
  # Required label set (red)
  geom_point(data = subset(cities_green, is_required),
             colour = "#d62728", size = hi_size + 0.6, alpha = 0.95) +
  ggrepel::geom_text_repel(
    data            = subset(cities_green, label_me),
    aes(label = city, colour = is_required),
    size            = 3.4,
    fontface        = "bold",
    box.padding     = 0.45,
    point.padding   = 0.35,
    min.segment.length = 0,
    segment.colour  = "grey50",
    segment.size    = 0.3,
    max.overlaps    = Inf,
    show.legend     = FALSE
  ) +
  scale_colour_manual(values = c(`TRUE` = "#d62728", `FALSE` = "#333333")) +
  scale_size_continuous(
    name   = "Green patent families per city",
    range  = c(0.6, 8),
    breaks = c(5, 50, 500, 5000),
    labels = scales::label_comma()
  ) +
  scale_x_continuous(labels = scales::label_number(accuracy = 0.01),
                     limits = axis_lim) +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.01),
                     limits = axis_lim) +
  coord_fixed() +
  labs(
    title    = "Green-tech spillover vs overall spillover, per city",
    subtitle = sprintf(
      "All cities with >=%d patent families and at least one Green Technology family. Dashed line: 1:1.",
      min_patents),
    x        = "Average Global Spillover  (mean ev_global, all docdbs)",
    y        = "Average Global Spillover  (mean ev_global, Green Tech only)",
    caption  = "Each dot is one city. Trend line: OLS of green-only on overall city mean."
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title       = element_text(face = "bold"),
    plot.subtitle    = element_text(colour = "#555"),
    panel.grid.minor = element_blank(),
    legend.position  = "right"
  )

print(p_green)
```

```{r fit_green}
fit_green <- stats::lm(avg_ev_global_green ~ avg_ev_global,
                       data = cities_green)
fit_green_s <- summary(fit_green)
cat(sprintf(
  "OLS:  avg_ev_global_green = %.4f + %.4f * avg_ev_global\n  R^2 = %.3f, n = %d cities\n",
  stats::coef(fit_green)[1L],
  stats::coef(fit_green)[2L],
  fit_green_s$r.squared,
  nrow(cities_green)))

cat(sprintf(
  "Cities where green > overall: %s of %s (%.1f%%)\n",
  format(sum(cities_green$avg_ev_global_green > cities_green$avg_ev_global,
             na.rm = TRUE),
         big.mark = ","),
  format(nrow(cities_green), big.mark = ","),
  100 * mean(cities_green$avg_ev_global_green > cities_green$avg_ev_global,
             na.rm = TRUE)
))
```

## Top-of-table view

```{r toptable}
fmt_pct <- function(x, digits = 1) {
  ifelse(is.na(x), NA_character_,
         paste0(formatC(100 * x, format = "f", digits = digits), "%"))
}
fmt_num <- function(x, digits = 3) {
  ifelse(is.na(x), NA_character_,
         formatC(x, format = "f", digits = digits))
}

top_for_table <- cities |>
  dplyr::arrange(dplyr::desc(n_docdbs)) |>
  dplyr::slice_head(n = 30L) |>
  dplyr::transmute(
    city,
    `# patents`               = n_docdbs,
    `# green`                 = n_green,
    `Share green`             = fmt_pct(share_green),
    `Avg ev_global`           = fmt_num(avg_ev_global,       3),
    `Avg ev_global (green)`   = fmt_num(avg_ev_global_green, 3),
    `Green premium`           = fmt_num(green_premium,       2),
    `Avg ev_emdenocn (LMICs)` = fmt_num(avg_ev_emdenocn,     3)
  )

required_for_table <- cities |>
  dplyr::filter(is_required) |>
  dplyr::arrange(dplyr::desc(n_docdbs)) |>
  dplyr::transmute(
    city,
    `# patents`               = n_docdbs,
    `# green`                 = n_green,
    `Share green`             = fmt_pct(share_green),
    `Avg ev_global`           = fmt_num(avg_ev_global,       3),
    `Avg ev_global (green)`   = fmt_num(avg_ev_global_green, 3),
    `Green premium`           = fmt_num(green_premium,       2),
    `Avg ev_emdenocn (LMICs)` = fmt_num(avg_ev_emdenocn,     3)
  )
```

### 30 largest cities

`Share green` is the fraction of the city's `# patents` that are
flagged as Green Technology (via `tech_lookup.tech_group = "Green
Technology"` or `technology = "Green Technology"`). `Green premium`
is `Avg ev_global (green) / Avg ev_global` — values **greater than
1** mean the city's green innovations have higher-than-average global
spillover for that city.

```{r toptable_render}
knitr::kable(top_for_table, format = "html",
             format.args = list(big.mark = ","),
             align = c("l", rep("r", ncol(top_for_table) - 1L))) |>
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  )
```

### Hand-picked label set

```{r requiredtable_render}
knitr::kable(required_for_table, format = "html",
             format.args = list(big.mark = ","),
             align = c("l", rep("r", ncol(required_for_table) - 1L))) |>
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  )
```

## Patent abstracts and OpenAI summaries

For each highlighted city, we retrieve the top-N patent applications
(ranked by `ev_global`) via the EPO Open Patent Services (OPS) API,
then ask `gpt-4o-mini` to summarise the technological themes that
emerge from the titles + abstracts.

This section reuses the EPO/OPS plumbing from
`insights/patent_cluster_compare.Rmd`. Two API keys are required as
environment variables; if either is missing the section short-circuits
to a warning rather than failing the knit.

```{r summary_setup, results='asis'}
have_ops <- nchar(Sys.getenv("OPS_CONSUMER_KEY",    unset = "")) > 0 &&
            nchar(Sys.getenv("OPS_CONSUMER_SECRET", unset = "")) > 0
have_oai <- nchar(Sys.getenv("OPENAI_API_KEY",      unset = "")) > 0

if (!have_ops || !have_oai) {
  cat(
    "<div style='padding:12px;border:1px solid #d62728;",
    "background:#fff3f3;border-radius:6px;color:#600;'>",
    "<b>Skipping abstract fetch + summarisation.</b> ",
    "Set the following environment variables before knitting to enable ",
    "this section: <code>OPS_CONSUMER_KEY</code>, ",
    "<code>OPS_CONSUMER_SECRET</code>, <code>OPENAI_API_KEY</code>.",
    "</div>",
    sep = "")
}
```

```{r summary_helpers, eval = have_ops && have_oai}
# --- Configuration --------------------------------------------------------
# Which cities get full abstract fetch + summary. Defaults to just the
# hand-picked LMIC / mid-tier list (those have a few hundred patents at
# most — practical for OPS). Add `cities$city[cities$is_big]` here if
# you also want summaries for the top-25 by size; expect the OPS
# fetch to take much longer in that case.
cities_to_summarise <- cities |>
  dplyr::filter(is_required) |>
  dplyr::pull(city)

# Cap per city. ev_global ranks docdbs from highest to lowest; we ask
# OPS for at most `max_per_city` of each city's patents. Keeps the
# fetch budget bounded for huge cities and prevents Luang Prabang's
# tiny universe from being silently extended past what it actually
# has.
max_per_city <- 75L

cache_dir <- ".bigdata/innocity"
if (!dir.exists(cache_dir)) dir.create(cache_dir, recursive = TRUE)

# --- Libraries (only loaded for this section) -----------------------------
library(httr2)
library(xml2)
library(jsonlite)
library(stringr)
library(digest)
library(purrr)
library(tibble)

# --- EPO OPS helpers (ported from patent_cluster_compare.Rmd) -------------
ns_epo <- c(ex = "http://www.epo.org/exchange", ops = "http://ops.epo.org")

extract_en_abstract <- function(doc) {
  abs_nodes <- xml2::xml_find_all(doc, ".//ex:abstract", ns_epo)
  if (length(abs_nodes) == 0) return(NA_character_)
  langs <- xml2::xml_attr(abs_nodes, "lang")
  en <- abs_nodes[!is.na(langs) & langs == "en"]
  node <- if (length(en) > 0) en[[1]] else abs_nodes[[1]]
  p_nodes <- xml2::xml_find_all(node, ".//ex:p", ns_epo)
  text <- if (length(p_nodes) > 0)
    paste(xml2::xml_text(p_nodes), collapse = " ")
  else xml2::xml_text(node)
  text <- stringr::str_squish(text)
  if (nchar(text) < 10) NA_character_ else text
}

extract_en_title <- function(doc) {
  title_nodes <- xml2::xml_find_all(doc, ".//ex:invention-title", ns_epo)
  if (length(title_nodes) == 0) return(NA_character_)
  langs <- xml2::xml_attr(title_nodes, "lang")
  en <- title_nodes[!is.na(langs) & langs == "en"]
  node <- if (length(en) > 0) en[[1]] else title_nodes[[1]]
  stringr::str_squish(xml2::xml_text(node))
}

parse_app_number <- function(x) {
  m <- stringr::str_match(x, "^([A-Z]{2,4})(\\d.*)$")
  tibble::tibble(raw = x, country = m[, 2], docnumber = m[, 3])
}

make_token_manager <- function(ops_key, ops_secret, initial_token) {
  env <- new.env()
  env$token      <- initial_token
  env$expires_at <- Sys.time() + 15 * 60
  env$refresh <- function() {
    resp <- httr2::request("https://ops.epo.org/3.2/auth/accesstoken") |>
      httr2::req_auth_basic(ops_key, ops_secret) |>
      httr2::req_body_form(grant_type = "client_credentials") |>
      httr2::req_perform()
    env$token      <- httr2::resp_body_json(resp)$access_token
    env$expires_at <- Sys.time() + 15 * 60
  }
  env$get <- function() {
    if (Sys.time() > env$expires_at) env$refresh()
    env$token
  }
  env
}

make_ops_request <- function(token_mgr, max_retries = 15) {
  function(url) {
    for (attempt in seq_len(max_retries + 1)) {
      resp <- httr2::request(url) |>
        httr2::req_headers(Authorization = paste("Bearer", token_mgr$get()),
                           Accept        = "application/xml") |>
        httr2::req_error(is_error = \(resp) FALSE) |>
        httr2::req_perform()
      status <- httr2::resp_status(resp)
      throttle_hdr <- httr2::resp_header(resp, "X-Throttling-Control")
      if (!is.null(throttle_hdr) &&
          grepl("overloaded", throttle_hdr, ignore.case = TRUE)) {
        Sys.sleep(60); next
      }
      if (status %in% c(401, 403)) { token_mgr$refresh(); Sys.sleep(5); next }
      if (status %in% c(429, 500, 502, 503, 504)) { Sys.sleep(60); next }
      if (status >= 400) return(NULL)
      return(httr2::resp_body_xml(resp))
    }
    NULL
  }
}

# Pull applicant names out of an OPS biblio XML doc. Mirrors the
# patent_cluster_compare.Rmd `extract_applicants()` but returns just a
# `;`-joined string of names (we'll explode + harmonise later); we
# don't need the per-applicant types here.
extract_applicants <- function(doc) {
  app_nodes <- xml2::xml_find_all(doc,
    ".//ex:applicant[@data-format='docdb']", ns_epo)
  if (length(app_nodes) == 0L) {
    app_nodes <- xml2::xml_find_all(doc, ".//ex:applicant", ns_epo)
  }
  if (length(app_nodes) == 0L) return(NA_character_)
  names_vec <- vapply(app_nodes, function(node) {
    name_nodes <- xml2::xml_find_all(node, ".//ex:applicant-name", ns_epo)
    if (length(name_nodes) == 0L) return(NA_character_)
    langs <- xml2::xml_attr(name_nodes, "lang")
    en <- name_nodes[!is.na(langs) & langs == "en"]
    nm <- if (length(en) > 0L) en[[1]] else name_nodes[[1]]
    stringr::str_squish(xml2::xml_text(nm))
  }, character(1))
  names_vec <- unique(names_vec[!is.na(names_vec) & nzchar(names_vec)])
  if (length(names_vec) == 0L) return(NA_character_)
  paste(names_vec, collapse = "; ")
}

fetch_one_patent <- function(raw_id, ops_request) {
  search_url <- paste0(
    "https://ops.epo.org/3.2/rest-services/published-data/search/biblio?q=pn%3D",
    raw_id
  )
  doc <- tryCatch(ops_request(search_url), error = function(e) NULL)
  if (is.null(doc)) return(tibble::tibble(title = NA_character_,
                                          abstract = NA_character_,
                                          applicant_names = NA_character_))
  tibble::tibble(
    title           = extract_en_title(doc),
    abstract        = extract_en_abstract(doc),
    applicant_names = extract_applicants(doc)
  )
}

fetch_for_city <- function(city_label, appln_ids, token_mgr,
                            cache_file, sleep_between = 1.5) {
  if (file.exists(cache_file)) {
    return(readRDS(cache_file))
  }
  ops_request <- make_ops_request(token_mgr)
  out <- vector("list", length(appln_ids))
  for (i in seq_along(appln_ids)) {
    if (i %% 10 == 0) message("    [", city_label, "] ", i, "/",
                              length(appln_ids))
    out[[i]] <- tryCatch(
      fetch_one_patent(appln_ids[i], ops_request),
      error = function(e) tibble::tibble(title = NA_character_,
                                         abstract = NA_character_))
    Sys.sleep(sleep_between)
  }
  res <- dplyr::bind_cols(
    tibble::tibble(appln_id = appln_ids),
    dplyr::bind_rows(out)
  )
  saveRDS(res, cache_file)
  res
}

# --- OpenAI chat helper ---------------------------------------------------
chat_completion <- function(prompt, api_key, model = "gpt-4o-mini",
                            temperature = 0.2, max_tokens = 700) {
  resp <- httr2::request("https://api.openai.com/v1/chat/completions") |>
    httr2::req_headers(Authorization = paste("Bearer", api_key),
                       `Content-Type` = "application/json") |>
    httr2::req_body_json(list(
      model = model,
      messages = list(list(role = "user", content = prompt)),
      temperature = temperature,
      max_tokens = max_tokens
    )) |>
    httr2::req_perform()
  httr2::resp_body_json(resp)$choices[[1]]$message$content
}
```

```{r summary_get_appln_ids, eval = have_ops && have_oai}
# For each city to summarise, query DuckDB for the top-N patents by
# ev_global. We rejoin patent_database (for ev_global + appln_id) with
# countrymap (filtered to the chosen city, geocode_missing = FALSE).
# DISTINCT ON keeps one row per docdb_family_id — multi-country
# families would otherwise show up multiple times.
con <- dbConnect(duckdb::duckdb())

city_top_apps <- purrr::map_dfr(cities_to_summarise, function(target_city) {
  q <- sprintf("
    WITH
      p  AS (SELECT * FROM read_parquet('%s')),
      cm AS (
        SELECT * FROM read_parquet('%s')
        WHERE city = '%s' AND geocode_missing = FALSE
      ),
      docdb AS (
        SELECT
          p.docdb_family_id,
          ANY_VALUE(p.appln_id)  AS appln_id,
          ANY_VALUE(p.ev_global) AS ev_global
        FROM p
        WHERE p.ev_global IS NOT NULL
          AND p.docdb_family_id IN (SELECT docdb_family_id FROM cm)
        GROUP BY p.docdb_family_id
      )
    SELECT appln_id, ev_global
    FROM docdb
    WHERE appln_id IS NOT NULL
    ORDER BY ev_global DESC NULLS LAST
    LIMIT %d
  ",
    pdb_path, cm_path,
    gsub("'", "''", target_city, fixed = TRUE),
    max_per_city)
  res <- dbGetQuery(con, q)
  res$city <- target_city
  res
})

dbDisconnect(con, shutdown = TRUE)

cat(sprintf("Will fetch up to %d patents for each of %d cities (total: %d).\n",
            max_per_city,
            length(cities_to_summarise),
            nrow(city_top_apps)))
```

```{r summary_fetch_abstracts, eval = have_ops && have_oai}
# Authenticate once; the token manager auto-refreshes during the run.
ops_key    <- Sys.getenv("OPS_CONSUMER_KEY")
ops_secret <- Sys.getenv("OPS_CONSUMER_SECRET")
token_resp <- httr2::request("https://ops.epo.org/3.2/auth/accesstoken") |>
  httr2::req_auth_basic(ops_key, ops_secret) |>
  httr2::req_body_form(grant_type = "client_credentials") |>
  httr2::req_perform()
initial_token <- httr2::resp_body_json(token_resp)$access_token
token_mgr <- make_token_manager(ops_key, ops_secret, initial_token)

city_abstracts <- city_top_apps |>
  dplyr::group_by(city) |>
  dplyr::group_modify(~{
    cache_key <- digest::digest(
      list(.y$city, sort(.x$appln_id), max_per_city), algo = "md5")
    cache_file <- file.path(
      cache_dir,
      sprintf("ops_%s_%s.rds",
              gsub("[^A-Za-z0-9]", "_", .y$city),
              substr(cache_key, 1L, 8L))
    )
    fetched <- fetch_for_city(.y$city, .x$appln_id, token_mgr, cache_file)
    dplyr::left_join(.x, fetched, by = "appln_id")
  }) |>
  dplyr::ungroup()

# Coverage report: how many abstracts came back per city.
city_abstracts |>
  dplyr::group_by(city) |>
  dplyr::summarise(
    requested = dplyr::n(),
    titles    = sum(!is.na(title)),
    abstracts = sum(!is.na(abstract)),
    .groups   = "drop"
  )
```

```{r summary_call_openai, eval = have_ops && have_oai}
# Two summaries per city are produced:
#   * "all"  — every fetched patent that returned usable text from
#              OPS (titles and / or abstracts).
#   * "top"  — the subset whose ev_global is strictly above the
#              MEDIAN of the fetched set. With max_per_city = 75 this
#              is roughly the top 35-37 patents per city, i.e. the
#              very best of the best by global spillover.
# Both summaries are cached separately, so adding the "top" cut here
# does not invalidate the existing all-fetched cache.
openai_key <- Sys.getenv("OPENAI_API_KEY")

build_prompt <- function(city, df, scope_label = "all") {
  texts <- df |>
    dplyr::filter(!is.na(title) | !is.na(abstract)) |>
    dplyr::mutate(
      tt = dplyr::if_else(is.na(title),    "(no title)",    title),
      ab = dplyr::if_else(is.na(abstract),
                          "(no abstract)",
                          stringr::str_trunc(abstract, 1200))
    ) |>
    dplyr::transmute(line = sprintf("- %s :: %s", tt, ab))

  scope_blurb <- if (identical(scope_label, "top")) {
    paste0(
      "These are the patents whose `ev_global` Watson spillover is ",
      "ABOVE the median of the fetched set for this city — i.e. the ",
      "top half by global spillover."
    )
  } else {
    paste0(
      "These are the up to ", max_per_city, " highest-ev_global ",
      "patents fetched for this city."
    )
  }

  paste0(
    "You are summarising the technological focus of patents associated ",
    "with one city in the Innovation Strategy Explorer database. ",
    scope_blurb, " The inventor address resolved to ", city, ".\n\n",
    "Patents:\n",
    paste(texts$line, collapse = "\n"),
    "\n\n",
    "Write a concise summary (~150-200 words) for ", city, " covering:\n",
    "  1. The two or three dominant technological themes,\n",
    "  2. Anything distinctive about the mix (e.g. heavy concentration ",
    "in one area, or surprising breadth),\n",
    "  3. Notable patent applicants or institutions if any are inferable ",
    "from the wording.\n",
    "Use plain English. Do not list patents one by one. Do not invent ",
    "facts that aren't supported by the abstracts."
  )
}

# Helper: get-or-fetch a single OpenAI completion, keyed by a digest
# of (city, scope, appln_ids, abstracts). Means changing the scope
# does NOT invalidate the other scope's cache.
get_or_fetch_summary <- function(city_label, scope_label, df) {
  cache_file <- file.path(
    cache_dir,
    sprintf("oai_%s_%s_%s.rds",
            gsub("[^A-Za-z0-9]", "_", city_label),
            scope_label,
            substr(digest::digest(list(df$appln_id, df$abstract,
                                        max_per_city, scope_label),
                                  algo = "md5"),
                   1L, 8L))
  )
  if (file.exists(cache_file)) return(readRDS(cache_file))
  out <- tryCatch(
    chat_completion(build_prompt(city_label, df, scope_label),
                    openai_key),
    error = function(e) sprintf("OpenAI call failed: %s",
                                conditionMessage(e))
  )
  saveRDS(out, cache_file)
  out
}

city_summaries <- city_abstracts |>
  dplyr::group_by(city) |>
  dplyr::group_split() |>
  purrr::map_dfr(function(df) {
    city_label  <- df$city[[1L]]
    has_text    <- !is.na(df$title) | !is.na(df$abstract)
    n_with_text <- sum(has_text)

    if (n_with_text < 3L) {
      return(tibble::tibble(
        city          = city_label,
        n_used_all    = n_with_text,
        n_used_top    = 0L,
        median_ev     = NA_real_,
        summary_all   = sprintf(
          "Only %d patent(s) returned usable text from OPS — too few for a meaningful summary.",
          n_with_text),
        summary_top   = "Top-innovations summary skipped (insufficient text returned from OPS)."
      ))
    }

    df_text <- df[has_text, , drop = FALSE]

    # Median of the fetched set's ev_global — the universe we have
    # text for. `> median` means strictly above (drops the median
    # case so the two halves can't both include the same patent).
    med_ev <- stats::median(df_text$ev_global, na.rm = TRUE)
    df_top <- df_text |>
      dplyr::filter(!is.na(ev_global) & ev_global > med_ev)

    summary_all <- get_or_fetch_summary(city_label, "all", df_text)
    summary_top <- if (nrow(df_top) < 3L) {
      sprintf(
        "Only %d patent(s) sit above the median ev_global for this city — too few to summarise separately.",
        nrow(df_top))
    } else {
      get_or_fetch_summary(city_label, "top", df_top)
    }

    tibble::tibble(
      city        = city_label,
      n_used_all  = n_with_text,
      n_used_top  = nrow(df_top),
      median_ev   = med_ev,
      summary_all = summary_all,
      summary_top = summary_top
    )
  })

city_summaries
```

```{r summary_render, eval = have_ops && have_oai, results='asis'}
for (i in seq_len(nrow(city_summaries))) {
  city_label  <- city_summaries$city[i]
  n_all       <- city_summaries$n_used_all[i]
  n_top       <- city_summaries$n_used_top[i]
  med_ev      <- city_summaries$median_ev[i]
  summary_all <- city_summaries$summary_all[i]
  summary_top <- city_summaries$summary_top[i]

  cat(sprintf("\n### %s\n\n", city_label))
  cat(higglo_link_md(city_label), "\n\n", sep = "")

  cat(sprintf("**All fetched patents** &mdash; %d with usable title/abstract text from OPS.\n\n",
              n_all))
  cat(summary_all, "\n\n")

  med_str <- if (!is.na(med_ev))
    sprintf(" (median ev_global = %.3f)", med_ev) else ""
  cat(sprintf(
    "**Top innovations only** &mdash; %d patents with `ev_global` above the median%s.\n\n",
    n_top, med_str))
  cat(summary_top, "\n\n")
}
```

## Firm word cloud per city

These wordclouds use the **harmonised firm data shipped with the
Innovation Strategy Explorer** (`inst/extdata/patents_x_firm.parquet`)
rather than EPO OPS applicants. That avoids the need for OPS API
calls and a GPT harmonisation pass, and the names are already
canonical. The trade-off: `patents_x_firm.parquet` is restricted to
the **top-100 firms by patent count globally** (see
`data-raw-2025/01-build-app-parquets.R`). Patents whose owners are
not in that top-100 list contribute nothing to a city's wordcloud,
which leaves smaller LMIC cities sparse or empty. Where that happens,
the OPS-based wordclouds (in earlier git history) are the more
inclusive view.

Each wordcloud word is sized by how many patent families in the city
are mapped to the firm. Colour encodes `firm_sector` from
`firm_lookup.parquet` when available.

```{r firms_local_query}
library(ggwordcloud)

# Resolve the bundle paths the same way the data chunk did. The
# helper was defined inside the `data` chunk above, so it's already
# in scope here, but redefining the path lookups locally keeps this
# section readable on its own.
pxf_path <- find_extdata("patents_x_firm.parquet")
fl_path  <- find_extdata("firm_lookup.parquet")

# All highlighted cities — same set used by the abstract summaries
# and the HiGGlo PNG capture.
cities_for_wc <- cities |>
  dplyr::filter(is_required) |>
  dplyr::arrange(dplyr::desc(n_docdbs)) |>
  dplyr::pull(city)

# One row per (city, firm) with patent-family count. We query
# DuckDB directly so we don't have to materialise the full
# patents_x_firm bridge in R memory. Firms are joined to
# firm_lookup for the sector colour; LEFT JOIN so firms missing
# from the lookup still render (with `firm_sector = NA`).
con <- dbConnect(duckdb::duckdb())
city_firms <- dbGetQuery(con, sprintf("
  WITH
    cm  AS (
      SELECT * FROM read_parquet('%s')
      WHERE geocode_missing = FALSE
        AND city IS NOT NULL
        AND city <> ''
    ),
    pxf AS (SELECT * FROM read_parquet('%s')),
    fl  AS (SELECT * FROM read_parquet('%s'))
  SELECT
    cm.city,
    pxf.firm,
    fl.firm_sector,
    COUNT(DISTINCT pxf.docdb_family_id) AS n
  FROM pxf
  JOIN cm ON cm.docdb_family_id = pxf.docdb_family_id
  LEFT JOIN fl ON fl.firm = pxf.firm
  WHERE cm.city IN (%s)
  GROUP BY cm.city, pxf.firm, fl.firm_sector
", cm_path, pxf_path, fl_path,
   paste0("'", gsub("'", "''", cities_for_wc, fixed = TRUE), "'",
          collapse = ", ")))
dbDisconnect(con, shutdown = TRUE)

# Top firms per city for the wordcloud.
max_firms_per_city <- 30L
city_firms <- city_firms |>
  dplyr::group_by(city) |>
  dplyr::slice_max(n, n = max_firms_per_city, with_ties = FALSE) |>
  dplyr::ungroup() |>
  dplyr::mutate(
    firm_sector = dplyr::if_else(is.na(firm_sector) | firm_sector == "",
                                 "Unclassified", firm_sector)
  )

cat(sprintf("city x firm rows kept: %s across %d cities (%d distinct sectors)\n",
            format(nrow(city_firms), big.mark = ","),
            dplyr::n_distinct(city_firms$city),
            dplyr::n_distinct(city_firms$firm_sector)))
```

```{r firms_wordcloud, fig.width=10, fig.height=4.5}
# Distinct sector colours. Sectors not in this palette fall back to
# a neutral grey via the named-vector default.
sector_palette <- c(
  "Pharmaceuticals"      = "#377eb8",
  "Chemicals"            = "#e41a1c",
  "Electronics"          = "#4daf4a",
  "Automotive"           = "#984ea3",
  "Energy"               = "#ff7f00",
  "Aerospace & Defense"  = "#a65628",
  "Industrial"           = "#1b9e77",
  "Consumer Goods"       = "#d95f02",
  "Telecommunications"   = "#7570b3",
  "Software"             = "#66a61e",
  "Materials"            = "#e7298a",
  "Unclassified"         = "#999999"
)
fallback_color <- "#777777"

color_for_sector <- function(s) {
  out <- sector_palette[s]
  out[is.na(out)] <- fallback_color
  unname(out)
}

# Loop city-by-city. One plot per city so individual labels can use
# a much larger max_size; empty cities (no top-100 firms in the
# patent set) get an italic placeholder. Cities are ordered by
# patent count, biggest first.
for (target_city in cities_for_wc) {
  cnt <- city_firms |>
    dplyr::filter(city == target_city)

  if (nrow(cnt) == 0L) {
    p <- ggplot() +
      annotate(
        "text", x = 0.5, y = 0.5,
        label = paste0(
          "No top-100 firms identified for ", target_city, ".\n",
          "patents_x_firm.parquet only carries the global top 100 ",
          "firms by patent count;\nLMIC cities often sit below that ",
          "threshold."
        ),
        size = 4, fontface = "italic", colour = "#666"
      ) +
      labs(title = target_city) +
      theme_void() +
      theme(
        plot.title  = element_text(size = 16, face = "bold",
                                   hjust = 0.5,
                                   margin = margin(b = 4)),
        plot.margin = margin(2, 2, 2, 2)
      ) +
      xlim(0, 1) + ylim(0, 1)
  } else {
    cnt$colour <- color_for_sector(cnt$firm_sector)
    p <- ggplot(cnt,
                aes(label = firm, size = n, colour = I(colour))) +
      ggwordcloud::geom_text_wordcloud_area(
        rm_outside   = TRUE,
        shape        = "square",
        eccentricity = 1.4
      ) +
      scale_size_area(max_size = 50) +
      labs(title = target_city) +
      theme_void() +
      theme(
        plot.title  = element_text(size = 16, face = "bold",
                                   hjust = 0.5,
                                   margin = margin(b = 4)),
        plot.margin = margin(2, 2, 2, 2)
      )
  }
  print(p)
}
```

```{r firms_wordcloud_legend, fig.width=10, fig.height=0.6}
# Single colour-by-sector legend at the end of the section, showing
# only the sectors that actually appear in any of the rendered
# wordclouds.
present_sectors <- intersect(names(sector_palette),
                              unique(city_firms$firm_sector))
if (length(present_sectors) > 0L) {
  legend_df <- data.frame(
    firm_sector = factor(present_sectors, levels = present_sectors),
    x           = seq_along(present_sectors),
    y           = 1
  )
  ggplot(legend_df,
         aes(x, y, colour = firm_sector, label = firm_sector)) +
    geom_point(size = 6) +
    geom_text(nudge_x = 0.15, hjust = 0, size = 5, colour = "#222") +
    scale_colour_manual(values = sector_palette[present_sectors],
                        guide  = "none") +
    coord_cartesian(xlim = c(0.5, length(present_sectors) + 1.5),
                    ylim = c(0.5, 1.5)) +
    theme_void() +
    theme(plot.margin = margin(2, 2, 2, 2))
}
```

```{r firms_table_render, results='asis'}
# Top-10 firms per city in table form. Mirrors the previous
# OPS-based table but uses the local firm names directly.
top_firms_table <- city_firms |>
  dplyr::group_by(city) |>
  dplyr::arrange(dplyr::desc(n), firm, .by_group = TRUE) |>
  dplyr::slice_head(n = 10L) |>
  dplyr::mutate(rank = dplyr::row_number()) |>
  dplyr::ungroup() |>
  dplyr::transmute(
    City        = city,
    Rank        = rank,
    Firm        = firm,
    Sector      = firm_sector,
    `# patents` = n
  )

if (nrow(top_firms_table) > 0L) {
  knitr::kable(top_firms_table, format = "html",
               align = c("l", "r", "l", "l", "r")) |>
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = FALSE
    )
} else {
  cat("<p><em>No firms identified for any of the highlighted cities ",
      "in the top-100 firm subset.</em></p>", sep = "")
}
```

```{r firms_legacy_block_dropped, eval = FALSE}
# The previous OPS-derived wordcloud machinery
# (assignee_helpers / assignee_pipeline / assignee_wordcloud /
# assignee_wordcloud_legend / assignee_table_render) lived here and
# was retired in favour of the local-firm path above. Kept as an
# `eval = FALSE` placeholder so the chunk-level diff is contained;
# the body remains in version control.
library(ggwordcloud)

# Heuristic regexes — same as patent_cluster_compare.Rmd. Pre-classify
# common organisation patterns before sending the unknowns to GPT;
# saves tokens and avoids accidental cross-type merges (e.g. a
# university variant collapsing onto a company).
org_suffix_regex <- stringr::regex(
  "\\b(inc|corp|ltd|llc|gmbh|ag|sa|sas|bv|nv|co|plc|pty|ab|oy|as|se|spa|srl|kk)\\b",
  ignore_case = TRUE
)
university_regex <- stringr::regex(
  paste0("\\b(univ|univers|universit|college|polytechnic|",
         "institute|institut|academy|academia|school|schule|hochschule|",
         "ecole|escuela|scuola|faculty)\\b"),
  ignore_case = TRUE
)
research_regex <- stringr::regex(
  paste0("\\b(cnrs|csic|csiro|csir|fraunhofer|max planck|helmholtz|riken|",
         "national lab|laboratory|laboratoire|laboratorio|research|forschung|",
         "instituto|agency|ministry|ministr|commission|consejo|conseil|",
         "council|foundation|fund|fonds|fundacion|stiftung)\\b"),
  ignore_case = TRUE
)

# One row per (city, patent, raw assignee). Strips the EPO-style `[CC]`
# country suffix and applies the heuristic classification used to
# split GPT prompts later.
explode_assignees <- function(df) {
  df |>
    dplyr::filter(!is.na(applicant_names)) |>
    dplyr::select(city, appln_id, applicant_names) |>
    dplyr::mutate(assignee = stringr::str_split(applicant_names, ";\\s*")) |>
    tidyr::unnest(assignee) |>
    dplyr::mutate(
      assignee = stringr::str_squish(stringr::str_remove(
        assignee, "\\s*\\[[A-Z]{2,3}\\]\\s*$"))
    ) |>
    dplyr::filter(nchar(assignee) > 0L) |>
    dplyr::mutate(
      is_university_h = stringr::str_detect(assignee, university_regex),
      is_research_h   = stringr::str_detect(assignee, research_regex),
      is_company_h    = stringr::str_detect(assignee, org_suffix_regex),
      heuristic_type  = dplyr::case_when(
        is_university_h ~ "university",
        is_research_h   ~ "research_institute",
        is_company_h    ~ "company",
        TRUE            ~ "unknown"
      )
    )
}

# Ask GPT to: (1) drop individual person names; (2) harmonise variants
# to a short canonical name; (3) return entity_type. Sent in batches
# split BY heuristic type so a university variant can't accidentally
# merge with a company. Caches the full mapping in
# `.bigdata/innocity/oai_assignee_map_*.rds` so re-knits skip the GPT
# call.
harmonize_assignees <- function(assignees_df, openai_key,
                                batch_size = 150,
                                cache_file = NULL) {
  if (!is.null(cache_file) && file.exists(cache_file)) {
    return(readRDS(cache_file))
  }

  uniq <- assignees_df |>
    dplyr::group_by(assignee) |>
    dplyr::summarise(freq           = dplyr::n(),
                     heuristic_type = dplyr::first(heuristic_type),
                     .groups        = "drop") |>
    dplyr::arrange(dplyr::desc(freq))
  if (nrow(uniq) == 0L) {
    out <- tibble::tibble(from = character(), to = character(),
                          entity_type = character())
    if (!is.null(cache_file)) saveRDS(out, cache_file)
    return(out)
  }

  call_gpt_batch <- function(batch, prompt_prefix) {
    entries <- paste0("- ", batch$assignee, " (n=", batch$freq, ")",
                      collapse = "\n")
    prompt <- paste0(prompt_prefix, "\n\n", entries, "\n\n---\n",
      "Return ONLY a JSON array of objects with keys \"from\" (original name), ",
      "\"to\" (canonical short name, Title Case, 1-4 words, null for persons), ",
      "and \"entity_type\" (one of: \"company\", \"university\", ",
      "\"research_institute\", \"government\", \"other\", \"person\"). ",
      "EVERY input name MUST appear in the output. No other text."
    )
    raw <- chat_completion(prompt, openai_key,
                           temperature = 0.1, max_tokens = 6000)
    raw <- stringr::str_remove_all(raw, "```json\\s*|```\\s*")
    tryCatch(
      jsonlite::fromJSON(raw) |>
        tibble::as_tibble() |>
        dplyr::mutate(
          from = stringr::str_squish(from),
          to   = if (is.character(to)) stringr::str_squish(to)
                 else as.character(to),
          to   = dplyr::na_if(to, "NA"),
          to   = dplyr::na_if(to, ""),
          entity_type = stringr::str_to_lower(
            stringr::str_squish(entity_type))
        ),
      error = function(e) {
        warning("Failed to parse assignee batch: ", e$message)
        tibble::tibble(from = character(), to = character(),
                       entity_type = character())
      }
    )
  }

  run_for_type <- function(df, label, type_hint) {
    if (nrow(df) == 0L) return(tibble::tibble(
      from = character(), to = character(), entity_type = character()))
    batches <- split(df, ceiling(seq_len(nrow(df)) / batch_size))
    prefix <- if (label == "unknown") {
      paste0(
        "Below is a list of raw patent assignee names (with frequencies). ",
        "These could not be pre-classified. For EACH name:\n",
        "1. DECIDE if it is a PERSON or an organization.\n",
        "2. For organizations, HARMONISE variants to a SHORT canonical name.\n",
        "3. ASSIGN an entity_type."
      )
    } else {
      paste0(
        "Below is a list of raw patent assignee names (with frequencies) ",
        "all pre-classified as ", type_hint, ". HARMONISE variants to a ",
        "SHORT canonical name (Title Case, 1-4 words; drop generic ",
        "suffixes like Inc/Corp/Ltd/GmbH). Set entity_type = \"",
        type_hint, "\" unless the name is clearly something else. ",
        "A name is only a person if clearly an individual human."
      )
    }
    lapply(seq_along(batches), function(bi) {
      message("    [assignees:", label, "] batch ", bi, "/", length(batches),
              " (", nrow(batches[[bi]]), " names)")
      call_gpt_batch(batches[[bi]], prefix)
    }) |> dplyr::bind_rows()
  }

  groups <- split(uniq, uniq$heuristic_type)
  empty  <- uniq[0L, , drop = FALSE]
  out <- dplyr::bind_rows(
    run_for_type(groups[["company"]]            %||% empty, "company",            "company"),
    run_for_type(groups[["university"]]         %||% empty, "university",         "university"),
    run_for_type(groups[["research_institute"]] %||% empty, "research_institute", "research_institute"),
    run_for_type(groups[["unknown"]]            %||% empty, "unknown",            NA_character_)
  )
  if (!is.null(cache_file)) saveRDS(out, cache_file)
  out
}

# Tiny coalesce helper — patent_cluster_compare.Rmd uses %||% from
# rlang; avoid the dependency by defining one locally.
`%||%` <- function(a, b) if (is.null(a) || !length(a)) b else a
```

```{r assignee_pipeline, eval = FALSE}
# Retired alongside the OPS-derived wordcloud (see
# firms_legacy_block_dropped above). Kept under `eval = FALSE` so
# the source remains in version control without being executed.
# Defensive coercion: OPS cache files written before applicant
# extraction was wired in (`extract_applicants` was added later in
# the development of this Rmd) don't carry an `applicant_names`
# column at all. Without this guard, `explode_assignees` would error
# on the missing column AND any downstream chunk that touches
# `city_abstracts$applicant_names` would silently produce zero rows.
# Guarantee the column exists, even if all-NA, so the rest of the
# pipeline degrades gracefully (cities with stale caches end up with
# empty wordclouds rather than disappearing entirely).
if (!"applicant_names" %in% names(city_abstracts)) {
  city_abstracts$applicant_names <- NA_character_
}

# Step 1 — explode raw assignees and apply heuristic classification.
assignees_long <- explode_assignees(city_abstracts)

n_unique_raw <- dplyr::n_distinct(assignees_long$assignee)
cat(sprintf("raw assignees: %s rows, %s unique names across %d cities\n",
            format(nrow(assignees_long), big.mark = ","),
            format(n_unique_raw,         big.mark = ","),
            length(unique(assignees_long$city))))

# Step 2 — GPT harmonisation, cached against the unique-name set.
assignee_cache <- file.path(
  cache_dir,
  sprintf("oai_assignee_map_%s.rds",
          substr(digest::digest(
            list(sort(unique(assignees_long$assignee)),
                 max_per_city,
                 cities_to_summarise),
            algo = "md5"), 1L, 8L))
)
assignee_map <- harmonize_assignees(assignees_long, openai_key,
                                    cache_file = assignee_cache)

# Step 3 — fold canonical names + entity_type back onto each row.
assignees_clean <- assignees_long |>
  dplyr::left_join(assignee_map, by = c("assignee" = "from")) |>
  dplyr::mutate(
    canonical   = dplyr::if_else(!is.na(to), to, assignee),
    entity_type = dplyr::if_else(!is.na(entity_type),
                                 entity_type, heuristic_type)
  )

# Step 4 — collapse to one row per (city, canonical, patent) so a
# patent with two variants of the same firm doesn't double-count.
assignees_dedup <- assignees_clean |>
  dplyr::filter(!is.na(canonical),
                !entity_type %in% c("person", NA_character_)) |>
  dplyr::distinct(city, appln_id, canonical, entity_type)
```

```{r assignee_wordcloud, eval = FALSE}
# Retired (see firms_wordcloud chunk for the active local-firm
# version).
entity_type_colors <- c(
  "company"            = "#377eb8",
  "university"         = "#e41a1c",
  "research_institute" = "#4daf4a",
  "government"         = "#984ea3",
  "other"              = "#999999"
)

# Top 30 firms per city. Tied counts get arbitrary order; with at most
# 75 patents per city the head of the distribution is dominated by
# 1-2 high-frequency assignees, with everything else tied at small
# values, so 30 is plenty for visual inspection without crowding the
# panel.
max_firms_per_city <- 30L

counts_by_city <- assignees_dedup |>
  dplyr::group_by(city, canonical) |>
  dplyr::summarise(
    n           = dplyr::n(),
    entity_type = names(sort(table(entity_type), decreasing = TRUE))[1L],
    .groups     = "drop"
  ) |>
  dplyr::filter(entity_type %in% names(entity_type_colors)) |>
  dplyr::group_by(city) |>
  dplyr::slice_max(n, n = max_firms_per_city, with_ties = FALSE) |>
  dplyr::ungroup()

# Loop over EVERY required city, not just the ones with data. That
# way readers can tell whether a missing wordcloud means "no firms
# survived the filter" vs "this city wasn't even checked". Each
# city's plot is its own figure (no faceting) so individual labels
# can use a much larger max_size without crowding adjacent panels.
# Empty cities get a small placeholder ggplot rather than silently
# disappearing.
cities_for_wc <- cities |>
  dplyr::filter(is_required) |>
  dplyr::arrange(dplyr::desc(n_docdbs)) |>
  dplyr::pull(city)

for (target_city in cities_for_wc) {
  cnt <- counts_by_city |>
    dplyr::filter(city == target_city)

  if (nrow(cnt) == 0L) {
    p <- ggplot() +
      annotate(
        "text", x = 0.5, y = 0.5,
        label = paste0(
          "No firm assignees identified for ", target_city, ".\n",
          "If this is unexpected, the OPS cache for this city was ",
          "likely written before applicant\nextraction was added. ",
          "Delete .bigdata/innocity/ops_*.rds and re-knit to refetch."
        ),
        size = 4, fontface = "italic", colour = "#666"
      ) +
      labs(title = target_city) +
      theme_void() +
      theme(
        plot.title  = element_text(size = 16, face = "bold",
                                   hjust = 0.5,
                                   margin = margin(b = 4)),
        plot.margin = margin(2, 2, 2, 2)
      ) +
      xlim(0, 1) + ylim(0, 1)
  } else {
    p <- ggplot(cnt,
                aes(label = canonical, size = n, colour = entity_type)) +
      ggwordcloud::geom_text_wordcloud_area(
        rm_outside   = TRUE,
        shape        = "square",
        eccentricity = 1.4
      ) +
      # Per-plot wordcloud has the full figure area to itself. Step
      # max_size up from the old facet-friendly 22 to 50 so the
      # smallest labels are still legible and the largest aren't
      # squashed.
      scale_size_area(max_size = 50) +
      scale_colour_manual(values = entity_type_colors,
                          name   = "Entity type",
                          guide  = "none") +
      labs(title = target_city) +
      theme_void() +
      theme(
        plot.title  = element_text(size = 16, face = "bold",
                                   hjust = 0.5,
                                   margin = margin(b = 4)),
        plot.margin = margin(2, 2, 2, 2)
      )
  }
  print(p)
}
```

```{r assignee_wordcloud_legend, eval = FALSE}
# Retired (see firms_wordcloud_legend for the active version).
# Colour legend rendered separately so it appears once at the end of
# the section instead of repeated under every per-city wordcloud.
# Pulled into its own short chunk so the figure-height knob doesn't
# affect the wordclouds above.
present_types <- intersect(names(entity_type_colors),
                            unique(counts_by_city$entity_type))
if (length(present_types) > 0L) {
  legend_df <- data.frame(
    entity_type = factor(present_types, levels = present_types),
    x           = seq_along(present_types),
    y           = 1
  )
  ggplot(legend_df, aes(x, y, colour = entity_type, label = entity_type)) +
    geom_point(size = 6) +
    geom_text(nudge_x = 0.15, hjust = 0, size = 5, colour = "#222") +
    scale_colour_manual(values = entity_type_colors[present_types],
                        guide  = "none") +
    coord_cartesian(xlim = c(0.5, length(present_types) + 1.5),
                    ylim = c(0.5, 1.5)) +
    theme_void() +
    theme(plot.margin = margin(2, 2, 2, 2))
}
```

### Top firms per city (table form)

```{r assignee_table_render, eval = FALSE}
# Retired (see firms_table_render for the active version).
top_table <- counts_by_city |>
  dplyr::group_by(city) |>
  dplyr::arrange(dplyr::desc(n), canonical, .by_group = TRUE) |>
  dplyr::slice_head(n = 10L) |>
  dplyr::mutate(rank = dplyr::row_number()) |>
  dplyr::ungroup() |>
  dplyr::transmute(
    City = city, Rank = rank, Firm = canonical,
    Type = entity_type, `# patents` = n
  )

knitr::kable(top_table, format = "html",
             align = c("l", "r", "l", "l", "r")) |>
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  )
```

## HiGGlo network: per highlighted city

For each of the hand-picked / highlighted cities we capture the
HiGGlo citation-network view directly from the live app via headless
Chrome (`chromote`). Same parameters as the URL the user provided
for Sofia:

* City filter: the hand-picked city
* `toflow = ev_global`, no firm/tech filter, **All countries**
* Gen-0 sampling: Random / Number / 100
* Edge sampling: Random / Percent / 100
* `add_generations = 2` (one Generate click → gens 1 and 2)

PNGs are cached under `.bigdata/innocity/higglo_<city>.png`. To force
any one of them to refresh, delete the corresponding file and
re-knit. The first run captures all of them sequentially in a
single chromote session (re-using the warm Shiny server between
navigations); budget ~50-70 s per city for a cold first knit, near
zero on subsequent knits.

```{r higglo_capture, eval = requireNamespace("chromote", quietly = TRUE)}
# All cities tagged is_required by the regex earlier in the document.
# Each gets its own captured screenshot. `build_higglo_url()` and
# `city_png_path()` come from the `helper_links` chunk above so all
# three downstream chunks (capture, PNG render, summary render) share
# the same URL shape.
cities_for_higglo <- cities |>
  dplyr::filter(is_required) |>
  dplyr::pull(city)

# Decide which cities still need fetching.
todo <- cities_for_higglo[
  !vapply(cities_for_higglo,
          function(c) file.exists(city_png_path(c)),
          logical(1))
]

if (length(todo) > 0L) {
  if (!dir.exists(cache_dir)) dir.create(cache_dir, recursive = TRUE)
  message(sprintf("Capturing HiGGlo PNGs for %d cities (cached: %d)...",
                  length(todo),
                  length(cities_for_higglo) - length(todo)))

  # Bump chromote's default CDP wait substantially. Posit Connect on
  # the share.connect.posit.cloud free tier can take well over a minute
  # to cold-start a Shiny container; if Page.navigate doesn't get a
  # CDP response in that window the synchronize() in chromote throws
  # `timed out waiting for response to command Page.navigate`. 600 s
  # is the chromote ceiling and matches what we use elsewhere
  # (insights/R/fetch_flow_csv.R bumps to 120 s for the simpler CSV
  # fetch — screenshots need more headroom).
  prev_to <- tryCatch(chromote::default_timeout(),
                      error = function(e) NULL)
  try(chromote::default_timeout(600), silent = TRUE)

  # One chromote session for the whole batch. Navigating between
  # URLs is much cheaper than spinning up a new headless Chrome per
  # city (Shiny container stays warm too, so cities 2..N typically
  # render in 30-40 s instead of 60 s).
  b <- chromote::ChromoteSession$new()
  on.exit({
    try(b$close(), silent = TRUE)
    if (!is.null(prev_to))
      try(chromote::default_timeout(prev_to), silent = TRUE)
  }, add = TRUE)

  # Higher DPR for sharper PNGs. The leaflet base tiles at the chosen
  # zoom level (2) are themselves rasters, so deviceScaleFactor only
  # helps for the markers, arcs, legend, and labels — but those are
  # the visually important bits, and stepping from 2x to 3x removes
  # most of the graininess on text and circle outlines without making
  # the file size unmanageable. Width/height kept at 1280x800 so the
  # output PNG aspect ratio matches what readers see in the browser
  # at the canonical map zoom.
  b$Emulation$setDeviceMetricsOverride(
    width             = 1280,
    height            = 800,
    deviceScaleFactor = 3,
    mobile            = FALSE
  )

  # ---------------------------------------------------------------------
  # PRE-WARM the deployed Shiny app. Hitting the URL once and waiting
  # before the real loop means the FIRST per-city Page.navigate call
  # doesn't have to wait for a cold-start container — that's the
  # navigation that was hitting the Page.navigate timeout. The warm-up
  # navigates to the app root with no query params so it isn't tied to
  # the fate of any one city's URL.
  app_root <- "https://mondpanther-iseapp2.share.connect.posit.cloud/"
  message("  [warm-up] visiting app root to wake the container...")
  warmup_ok <- tryCatch({
    b$Page$navigate(app_root)
    Sys.sleep(45)
    TRUE
  }, error = function(e) {
    message("    [warm-up failed: ", conditionMessage(e),
            "] continuing anyway — first city will retry.")
    FALSE
  })

  for (i in seq_along(todo)) {
    city <- todo[i]
    url  <- build_higglo_url(city)
    out  <- city_png_path(city)
    # Generous post-navigate sleep budgets. Cold starts that survive
    # the warm-up still need ~20 s of init before render_map fires;
    # gen 0 + 2 follow-on generations then need another ~30 s.
    sleep_s <- if (i == 1L && !warmup_ok) 90L else 55L

    message(sprintf("  [%d/%d] %s (sleep %ds)",
                    i, length(todo), city, sleep_s))

    # Wrap each city in tryCatch so a single failure (timeout, network
    # blip, malformed URL) does NOT kill the rest of the batch. We
    # log the error and move on; the per-city render fallback will
    # show a warning box for any cities that didn't produce a PNG.
    tryCatch({
      b$Page$navigate(url)
      Sys.sleep(sleep_s)
      # Hide the leaflet +/- zoom controls in the captured PNG. The
      # Shiny app exposes a `body.higglobe-capture-mode` CSS hook
      # (defined in R/module_hglobe.R) that suppresses
      # `.leaflet-control-zoom` whenever the body has that class.
      # Setting it via Runtime.evaluate just before the screenshot
      # keeps the interactive UI unchanged — the zoom controls are
      # only hidden for the duration of the snapshot.
      b$Runtime$evaluate(
        "document.body.classList.add('higglobe-capture-mode')"
      )
      b$screenshot(filename = out, selector = "#hglobe-map")
      message(sprintf("    -> wrote %s", out))
    }, error = function(e) {
      message("    [city failed: ", conditionMessage(e), "] skipping.")
    })
  }
}
```

```{r higglo_render, results='asis', echo=FALSE}
cities_for_higglo <- cities |>
  dplyr::filter(is_required) |>
  dplyr::arrange(dplyr::desc(n_docdbs)) |>   # biggest first
  dplyr::pull(city)

for (city in cities_for_higglo) {
  png_path <- city_png_path(city)
  cat(sprintf("\n### %s\n\n", city))
  cat(higglo_link_md(city), "\n\n", sep = "")
  if (file.exists(png_path)) {
    # Use raw markdown image syntax so knitr embeds a relative file
    # reference rather than the absolute one `include_graphics()`
    # produces inside a loop.
    cat(sprintf("![](%s){width=100%%}\n\n",
                normalizePath(png_path, winslash = "/", mustWork = FALSE)))
  } else {
    cat(
      "<div style='padding:12px;border:1px solid #d62728;",
      "background:#fff3f3;border-radius:6px;color:#600;'>",
      "<b>HiGGlo screenshot not available.</b> Either ",
      "<code>chromote</code> isn't installed, the live app URL was ",
      "unreachable, or the screenshot timed out. ",
      "Install chromote and re-knit, or pre-populate the file at ",
      sprintf("<code>%s</code>", png_path),
      " by hand using the Save PNG button on the live app.",
      "</div>\n\n",
      sep = "")
  }
}
```

---

*Notes.*

* `ev_global` is the Watson per-family expected spillover to the whole
  world; `ev_emdenocn` is the share of that spillover that lands in
  LMICs excluding China. Both are docdb-level (one value per family),
  so per-city averages are unweighted means across the families
  associated with the city through `countrymap`.
* Capital-fallback rows (`geocode_missing = TRUE`) are dropped before
  aggregation. Without that, every patent for which no real address
  survived geocoding would inflate the patent count of the country's
  capital.
* A single docdb may be mapped to multiple cities (e.g. multinational
  inventor teams). Each city it touches gets credit for the family —
  there is no proportional split.
* The abstract-fetch + OpenAI summary section caches both OPS responses
  and OpenAI completions under `.bigdata/innocity/`, keyed on the
  cities and the per-city patent set. Re-knitting the document is
  free as long as the inputs don't change. Delete files in that
  directory to force a refetch / re-summarise.
* Default scope is the seven hand-picked cities only. To also
  summarise the 25 biggest cities, edit `cities_to_summarise` in the
  `summary_helpers` chunk — but expect a very long OPS fetch the
  first time round, since each big city sends `max_per_city`
  HTTP calls.
