---
title: "Are developing country innovators hidden giants?"
author: "Innovation Strategy Explorer — insights"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 2
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 11, fig.height = 6.5, dpi = 150)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(scales)
library(ggrepel)
library(DBI)
library(duckdb)
library(htmltools)

# Light, theme-proof table styling. RStudio's inline notebook preview
# renders chunk output inside an iframe whose CSS inherits the IDE colour
# scheme — including dark backgrounds that make ordinary kable output
# unreadable. Inline `style="..."` with `!important` on every cell beats
# both the iframe stylesheet and any document-level <style> block.
nice_kable <- function(df, digits = NULL,
                       format.args = list(big.mark = ",")) {
  cols  <- names(df)
  ndig  <- function(j) {
    if (is.null(digits)) NA_integer_
    else if (length(digits) == 1L) digits
    else digits[j]
  }
  fmt   <- function(v, j) {
    d <- ndig(j)
    if (is.numeric(v)) {
      v2 <- if (is.na(d)) v else round(v, d)
      do.call(format, c(list(v2), format.args, list(scientific = FALSE)))
    } else as.character(v)
  }

  th_s  <- "background:#e8e8e8 !important;color:#111 !important;border:1px solid #bbb;padding:4px 10px;text-align:center;font-weight:600;"
  td_s  <- "background:#ffffff !important;color:#222 !important;border:1px solid #ddd;padding:4px 10px;text-align:right;"
  td_z  <- "background:#f7f7f7 !important;color:#222 !important;border:1px solid #ddd;padding:4px 10px;text-align:right;"
  ltxt  <- "text-align:left;"
  tbl_s <- "background:#ffffff !important;color:#222 !important;border-collapse:collapse;font-family:-apple-system,Segoe UI,Roboto,sans-serif;font-size:13px;margin:8px 0;"

  thead <- paste0(
    "<tr>",
    paste0(sprintf('<th style="%s">%s</th>', th_s, cols), collapse = ""),
    "</tr>"
  )
  rows <- vapply(seq_len(nrow(df)), function(i) {
    cells <- vapply(seq_along(cols), function(j) {
      base <- if (i %% 2 == 0) td_z else td_s
      sty  <- if (j == 1) paste0(base, ltxt) else base
      sprintf('<td style="%s">%s</td>', sty, fmt(df[[j]][i], j))
    }, character(1))
    paste0("<tr>", paste(cells, collapse = ""), "</tr>")
  }, character(1))

  htmltools::HTML(paste0(
    sprintf('<table style="%s">', tbl_s),
    "<thead>", thead, "</thead>",
    "<tbody>", paste(rows, collapse = ""), "</tbody>",
    "</table>"
  ))
}
```

## What this note shows

## Citation, EV, and AV ranks at the docdb level

Collapse the app's main database to one row per `docdb_family_id`, attach
flags telling us whether the family is mapped to **China**, any **HIC**, or
any **non-HIC country (excluding China)**, then rank docdbs (1 = highest)
by:

* **Citation count** (`cit_count`) — number of times a docdb is cited by
  another docdb in the app database.
* **EV global** (`ev_global`) — Watson average global spillover.
* **AV global** (`av_global`) — Watson average global return.

Country-class flags are not mutually exclusive (a family can be mapped to
e.g. both the US and China). For colouring the scatter we collapse to a
single primary class with priority **China → HIC → non-HIC excl. China**.

```{r build_docdb_ranks}
# Locate inst/extdata regardless of where the Rmd is rendered from.
find_extdata <- function(file) {
  cand <- c(
    system.file("extdata", file, package = "innovationStrategyExplorer"),
    file.path("..", "inst", "extdata", file),
    file.path("inst", "extdata", file)
  )
  hit <- cand[nzchar(cand) & file.exists(cand)]
  if (!length(hit)) stop("Could not locate ", file)
  normalizePath(hit[[1]], winslash = "/")
}
pdb_path <- find_extdata("patent_database.parquet")
clu_path <- find_extdata("country_lookup.parquet")

# Pull a per-docdb table directly from DuckDB so we never materialise the
# 28 M-row patent_database in R memory. ANY_VALUE on the family-constant
# columns is safe because cit_count, ev_global, av_global don't vary
# across countries within a docdb.
con <- dbConnect(duckdb::duckdb())
# Note: av_global is no longer materialised in patent_database.parquet —
# it is derived by the runtime view in R/runAppPackage.R from
# (pv + ev_global) / cost. We reproduce the same expression here so this
# Rmd can read the raw parquet without going through the package view.
docdb <- dbGetQuery(con, sprintf("
  WITH p   AS (SELECT * FROM read_parquet('%s')),
       cl  AS (SELECT * FROM read_parquet('%s'))
  SELECT
    p.docdb_family_id,
    ANY_VALUE(p.cit_count)  AS cit_count,
    ANY_VALUE(p.ev_global)  AS ev_global,
    ANY_VALUE(p.granted)    AS granted,
    ANY_VALUE(
      CASE WHEN p.cost IS NULL OR p.cost = 0 THEN 0
           ELSE (p.pv + p.ev_global) / p.cost
      END
    ) AS av_global,
    MAX(CASE WHEN cl.is_hic = TRUE THEN 1 ELSE 0 END)               AS any_hic,
    MAX(CASE WHEN p.ctry_code = 'CN' THEN 1 ELSE 0 END)             AS any_cn,
    MAX(CASE WHEN cl.is_hic = FALSE AND p.ctry_code <> 'CN' THEN 1
             ELSE 0 END)                                            AS any_lmic_excl_cn
  FROM p
  LEFT JOIN cl ON cl.ctry_code = p.ctry_code
  GROUP BY p.docdb_family_id
", pdb_path, clu_path))
dbDisconnect(con, shutdown = TRUE)

docdb <- docdb %>%
  mutate(
    cit_rank      = min_rank(desc(cit_count)),
    ev_rank       = min_rank(desc(ev_global)),
    av_rank       = min_rank(desc(av_global)),
    is_hic        = any_hic == 1,
    is_china      = any_cn  == 1,
    is_lmic_xcn   = any_lmic_excl_cn == 1,
    region_class  = factor(
      case_when(
        is_china    ~ "China",
        is_hic      ~ "HIC",
        is_lmic_xcn ~ "non-HIC excl China",
        TRUE        ~ "unmapped"
      ),
      levels = c("China", "HIC", "non-HIC excl China", "unmapped")
    )
  )

cat(sprintf("docdbs total: %s   (HIC: %s, China: %s, non-HIC excl China: %s)\n",
            format(nrow(docdb), big.mark = ","),
            format(sum(docdb$is_hic),      big.mark = ","),
            format(sum(docdb$is_china),    big.mark = ","),
            format(sum(docdb$is_lmic_xcn), big.mark = ",")))
```

### Hidden-giant indices (citation rank vs value rank)

* `higx_ev = cit_rank − ev_rank`
* `higx_av = cit_rank − av_rank`

Both ranks have **1 = highest**, so a *negative* index means a docdb is
more highly cited than its value rank would predict — a candidate "hidden
giant". A *positive* value means the opposite (high value, few citations).

```{r higx_indices}
docdb <- docdb %>%
  mutate(
    higx_ev = ((cit_rank) - (ev_rank))/max(cit_rank)*100,
    higx_av = ((cit_rank) - (av_rank))/max(cit_rank)*100
  )

docdb %>%
  filter(region_class != "unmapped") %>%
  group_by(region_class) %>%
  summarise(
    n              = dplyr::n(),
    mean_higx_ev   = mean(higx_ev),
    median_higx_ev = stats::median(higx_ev),
    p75_higx_ev    = stats::quantile(higx_ev, 0.75),
    p90_higx_ev    = stats::quantile(higx_ev, 0.90),
    p95_higx_ev    = stats::quantile(higx_ev, 0.95),
    .groups        = "drop"
  ) %>%
  nice_kable(digits = 2, format.args = list(big.mark = ","))


docdb %>% distinct(docdb_family_id) %>% nrow()
docdb %>% nrow()
```

### Country-class flag overlap

Since `is_hic`, `is_china`, and `is_lmic_xcn` are not mutually exclusive,
the next regressions ask three overlapping yes/no questions. The table
below tabulates all 2^3 = 8 combinations of the flags so the overlap is
explicit. The `region_class` column shows where each combination ends up
under the priority rule **China → HIC → non-HIC excl China** that drives
the scatter.

```{r flag_overlap}
priority <- function(hic, cn, lmic) {
  ifelse(cn,  "China",
  ifelse(hic, "HIC",
  ifelse(lmic, "non-HIC excl China",
         "unmapped")))
}

docdb %>%
  count(is_hic, is_china, is_lmic_xcn, name = "n") %>%
  mutate(
    share        = n / sum(n),
    region_class = priority(is_hic, is_china, is_lmic_xcn)
  ) %>%
  select(is_hic, is_china, is_lmic_xcn, region_class, n, share) %>%
  arrange(desc(n)) %>%
  nice_kable(digits = c(NA, NA, NA, NA, 0, 4),
             format.args = list(big.mark = ","))
```

#### Granted families only

Same overlap table restricted to docdbs with at least one granted
application — useful as a check on whether the flag distribution shifts
when applications that never made it through examination are excluded.

```{r flag_overlap_granted}
docdb %>%
  filter(granted) %>%
  count(is_hic, is_china, is_lmic_xcn, name = "n") %>%
  mutate(
    share        = n / sum(n),
    region_class = priority(is_hic, is_china, is_lmic_xcn)
  ) %>%
  select(is_hic, is_china, is_lmic_xcn, region_class, n, share) %>%
  arrange(desc(n)) %>%
  nice_kable(digits = c(NA, NA, NA, NA, 0, 4),
             format.args = list(big.mark = ","))
```

#### `docdb_family_size` > 1

Same overlap table restricted to docdbs whose PATSTAT
`docdb_family_size` is at least 2 (multi-application families). The
eligible-family list is read from
`.bigdata/fam_size_min2.parquet`, the BigQuery cache that
`01-build-app-parquets.R` produces when the build-time family-size
filter is enabled.

```{r flag_overlap_multifam}
fam2_path <- local({
  cand <- c(file.path("..", ".bigdata", "fam_size_min2.parquet"),
            file.path(".bigdata", "fam_size_min2.parquet"))
  hit  <- cand[file.exists(cand)]
  if (!length(hit)) NA_character_ else normalizePath(hit[[1]], winslash = "/")
})

if (is.na(fam2_path)) {
  cat("fam_size_min2.parquet not found — rerun the build pipeline (or set\n",
      "FAMILY_SIZE_MIN <- 2L) to materialise the eligible-family cache.\n")
} else {
  con_fs <- dbConnect(duckdb::duckdb())
  fams2 <- dbGetQuery(con_fs, sprintf(
    "SELECT DISTINCT docdb_family_id FROM read_parquet('%s')",
    fam2_path))
  dbDisconnect(con_fs, shutdown = TRUE)

  docdb %>%
    semi_join(fams2, by = "docdb_family_id") %>%
    count(is_hic, is_china, is_lmic_xcn, name = "n") %>%
    mutate(
      share        = n / sum(n),
      region_class = priority(is_hic, is_china, is_lmic_xcn)
    ) %>%
    select(is_hic, is_china, is_lmic_xcn, region_class, n, share) %>%
    arrange(desc(n)) %>%
    nice_kable(digits = c(NA, NA, NA, NA, 0, 4),
               format.args = list(big.mark = ","))
}
```

### Regressions on country-class flags

Both flag triples are non-mutually-exclusive — a docdb mapped to both the
US and China contributes to `is_hic = TRUE` *and* `is_china = TRUE`. The
intercept therefore represents a docdb with all three flags `FALSE`
(none); each coefficient is the conditional change in the index when that
particular flag flips on.

```{r reg_higx}
library(broom)
library(fixest)

mapped <- docdb %>%
  filter(is_hic | is_china | is_lmic_xcn)   # drop unmapped families

# fixest::feols beats stats::lm on memory and wall time for 28 M-row data
# even without fixed effects (multi-threaded C++ path, no full model
# matrix allocation). Default SEs are HC1 (heteroskedasticity-robust);
# pass vcov = "iid" if you want classical OLS SEs to match lm exactly.
fit_ev <- fixest::feols(higx_ev ~  is_china + is_lmic_xcn, data = mapped)

tidy(fit_ev, conf.int = TRUE) %>%
  mutate(outcome = "higx_ev") %>%
  select(outcome, term, estimate, std.error, statistic, p.value,
         conf.low, conf.high) %>%
  nice_kable(digits = c(NA, NA, 3, 3, 2, 4, 3, 3),
             format.args = list(big.mark = ","))

cat(sprintf(
  "n (mapped docdbs): %s   |   R^2 (higx_ev): %.4f\n",
  format(nrow(mapped), big.mark = ","),
  fixest::r2(fit_ev, type = "r2")
))
```

#### Granted families only

Re-run on the subset where `granted = TRUE` (the docdb has at least one
granted application). This filters out applications that never made it
through examination, which can carry citation patterns very different
from granted patents.

```{r reg_higx_granted}
mapped_g <- mapped %>% filter(granted)

fit_ev_g <- fixest::feols(higx_ev ~ is_hic + is_china + is_lmic_xcn,
                          data = mapped_g)

tidy(fit_ev_g, conf.int = TRUE) %>%
  mutate(outcome = "higx_ev") %>%
  select(outcome, term, estimate, std.error, statistic, p.value,
         conf.low, conf.high) %>%
  nice_kable(digits = c(NA, NA, 3, 3, 2, 4, 3, 3),
             format.args = list(big.mark = ","))

cat(sprintf(
  "n (granted, mapped docdbs): %s   |   R^2 (higx_ev): %.4f\n",
  format(nrow(mapped_g), big.mark = ","),
  fixest::r2(fit_ev_g, type = "r2")
))
```

#### `docdb_family_size` > 1

Re-run on the multi-application subset (PATSTAT `docdb_family_size >= 2`),
using the same `.bigdata/fam_size_min2.parquet` cache that the overlap
table above reads.

```{r reg_higx_multifam, results='asis'}
fam2_path <- local({
  cand <- c(file.path("..", ".bigdata", "fam_size_min2.parquet"),
            file.path(".bigdata", "fam_size_min2.parquet"))
  hit  <- cand[file.exists(cand)]
  if (!length(hit)) NA_character_ else normalizePath(hit[[1]], winslash = "/")
})

if (is.na(fam2_path)) {
  cat("fam_size_min2.parquet not found — skipping multi-family regressions.\n")
} else {
  con_fs <- dbConnect(duckdb::duckdb())
  fams2  <- dbGetQuery(con_fs, sprintf(
    "SELECT DISTINCT docdb_family_id FROM read_parquet('%s')", fam2_path))
  dbDisconnect(con_fs, shutdown = TRUE)

  mapped_m <- mapped %>% semi_join(fams2, by = "docdb_family_id")

  fit_ev_m <- fixest::feols(higx_ev ~ is_hic + is_china + is_lmic_xcn,
                            data = mapped_m)

  # results='asis' lets us emit the raw HTML kable from inside this
  # if-block. `print()` would have wrapped the htmltools::HTML object as
  # plain-text output instead of rendering it.
  cat(as.character(
    tidy(fit_ev_m, conf.int = TRUE) %>%
      mutate(outcome = "higx_ev") %>%
      select(outcome, term, estimate, std.error, statistic, p.value,
             conf.low, conf.high) %>%
      nice_kable(digits = c(NA, NA, 3, 3, 2, 4, 3, 3),
                 format.args = list(big.mark = ","))
  ))

  cat(sprintf(
    "<pre style='margin-top:8px;background:#f5f5f5;color:#222;padding:6px 10px;font-size:12px;'>n (fam_size>=2, mapped docdbs): %s   |   R^2 (higx_ev): %.4f</pre>\n",
    format(nrow(mapped_m), big.mark = ","),
    fixest::r2(fit_ev_m, type = "r2")
  ))
}
```

### Coefficient comparison across specifications

Stacks the three sub-samples (all mapped docdbs, granted only,
`fam_size >= 2`) into one figure so the coefficient on each country-class
flag is comparable across them. Bars are coefficient point estimates with
95 % confidence intervals; the dashed zero line is the no-effect
reference.

```{r reg_higx_compare}
collect_tidy <- function(fit, spec) {
  if (!exists(deparse(substitute(fit)), inherits = FALSE) || is.null(fit))
    return(NULL)
  tidy(fit, conf.int = TRUE) %>% mutate(spec = spec)
}

reg_long <- bind_rows(
  collect_tidy(fit_ev,   "all"),
  collect_tidy(fit_ev_g, "granted"),
  if (exists("fit_ev_m")) collect_tidy(fit_ev_m, "fam_size>=2")
) %>%
  filter(term != "(Intercept)") %>%
  mutate(spec = factor(spec, levels = c("all", "granted", "fam_size>=2")))

ggplot(reg_long, aes(x = term, y = estimate, fill = spec)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "#888") +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                position = position_dodge(width = 0.8),
                width = 0.25) +
  scale_fill_manual(values = c(
    "all"          = "#1f77b4",
    "granted"      = "#ff7f0e",
    "fam_size>=2"  = "#2ca02c"
  )) +
  labs(
    title = "higx_ev country-class coefficients across specifications",
    x     = NULL,
    y     = "Coefficient estimate (with 95% CI)",
    fill  = "Sample"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
```

### EV-rank vs citation-rank scatter (1 % sample)

```{r scatter_helper, include=FALSE}
# Reusable scatter so the three subsamples (full / granted / fam_size>1)
# share identical aesthetics. `subset_df` is a pre-filtered docdb table;
# the helper handles the unmapped drop + Bernoulli sample + plot.
scatter_ev_cit <- function(subset_df, prop = 0.05, subtitle_extra = NULL,
                           seed = 42) {
  set.seed(seed)
  plot_df <- subset_df %>%
    filter(region_class != "unmapped") %>%
    # Drop China-only docdbs (is_china = TRUE with neither HIC nor non-HIC
    # excl-China overlap) — the user's request was to exclude families
    # mapped exclusively to China across all three scatters.
    filter(!(is_china & !is_hic & !is_lmic_xcn)) %>%
    # Convert ranks to percentiles within the eligible (post-exclusion)
    # universe so the two axes share the same 0-100 scale and the y=x
    # diagonal is interpretable.
    mutate(
      ev_pct  = ev_rank  / max(ev_rank,  na.rm = TRUE) * 100,
      cit_pct = cit_rank / max(cit_rank, na.rm = TRUE) * 100
    ) %>%
    slice_sample(prop = prop)

  # Cap each axis at the 99th percentile of the *displayed* sample so the
  # bulk of the cloud is visible. Use the joint max across both axes so
  # the 45° line stays meaningful.
  cap <- max(stats::quantile(plot_df$ev_pct,  0.99, na.rm = TRUE),
             stats::quantile(plot_df$cit_pct, 0.99, na.rm = TRUE))

  sub <- sprintf("n = %s sampled docdbs; axes are percentiles (1 = top of distribution)",
                 format(nrow(plot_df), big.mark = ","))
  if (!is.null(subtitle_extra)) sub <- paste0(subtitle_extra, " — ", sub)

  ggplot(plot_df, aes(ev_pct, cit_pct, colour = region_class)) +
    geom_abline(slope = 1, intercept = 0,
                colour = "#888", linetype = "dashed") +
    geom_point(alpha = 0.35, size = 0.6) +
    # Per-group 2D kernel-density contours over the same scatter — same
    # colour scale as the dots so each region_class is identifiable. The
    # `group = region_class` makes density estimates separate per class
    # rather than a single global surface.
    geom_density_2d(aes(group = region_class),
                    linewidth = 0.55, alpha = 0.9) +
    coord_cartesian(xlim = c(0, cap), ylim = c(0, cap)) +
    scale_x_continuous(labels = label_number(accuracy = 0.1)) +
    scale_y_continuous(labels = label_number(accuracy = 0.1)) +
    scale_colour_manual(values = c(
      "China"              = "#d62728",
      "HIC"                = "#1f77b4",
      "non-HIC excl China" = "#2ca02c"
    )) +
    labs(
      title    = "EV-global percentile vs citation-count percentile",
      subtitle = sub,
      x        = "EV global percentile (rank / max × 100)",
      y        = "Citation count percentile",
      colour   = "Country class"
    ) +
    guides(colour = guide_legend(override.aes = list(alpha = 1, size = 2))) +
    theme_minimal(base_size = 12)
}
```

```{r scatter_ev_vs_cit}
scatter_ev_cit(docdb, prop = 0.05, subtitle_extra = "all docdbs")
```

#### Granted families only

```{r scatter_ev_vs_cit_granted}
scatter_ev_cit(docdb %>% filter(granted),
               prop = 0.05,
               subtitle_extra = "granted families only")
```

#### `docdb_family_size` > 1

```{r scatter_ev_vs_cit_multifam}
fam2_path <- local({
  cand <- c(file.path("..", ".bigdata", "fam_size_min2.parquet"),
            file.path(".bigdata", "fam_size_min2.parquet"))
  hit  <- cand[file.exists(cand)]
  if (!length(hit)) NA_character_ else normalizePath(hit[[1]], winslash = "/")
})

if (is.na(fam2_path)) {
  cat("fam_size_min2.parquet not found — skipping fam_size>1 scatter.\n")
} else {
  con_fs <- dbConnect(duckdb::duckdb())
  fams2  <- dbGetQuery(con_fs, sprintf(
    "SELECT DISTINCT docdb_family_id FROM read_parquet('%s')", fam2_path))
  dbDisconnect(con_fs, shutdown = TRUE)
  print(scatter_ev_cit(
    docdb %>% semi_join(fams2, by = "docdb_family_id"),
    prop = 0.05,
    subtitle_extra = "fam_size >= 2"
  ))
}
```

