---
title: "Are developing country innovators hidden giants?"
author: "Innovation Strategy Explorer — insights"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 2
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 11, fig.height = 6.5, dpi = 150)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(scales)
library(ggrepel)
library(DBI)
library(duckdb)
library(htmltools)

# Light, theme-proof table styling. RStudio's inline notebook preview
# renders chunk output inside an iframe whose CSS inherits the IDE colour
# scheme — including dark backgrounds that make ordinary kable output
# unreadable. Inline `style="..."` with `!important` on every cell beats
# both the iframe stylesheet and any document-level <style> block.
nice_kable <- function(df, digits = NULL,
                       format.args = list(big.mark = ",")) {
  cols  <- names(df)
  ndig  <- function(j) {
    if (is.null(digits)) NA_integer_
    else if (length(digits) == 1L) digits
    else digits[j]
  }
  fmt   <- function(v, j) {
    d <- ndig(j)
    if (is.numeric(v)) {
      v2 <- if (is.na(d)) v else round(v, d)
      do.call(format, c(list(v2), format.args, list(scientific = FALSE)))
    } else as.character(v)
  }

  th_s  <- "background:#e8e8e8 !important;color:#111 !important;border:1px solid #bbb;padding:4px 10px;text-align:center;font-weight:600;"
  td_s  <- "background:#ffffff !important;color:#222 !important;border:1px solid #ddd;padding:4px 10px;text-align:right;"
  td_z  <- "background:#f7f7f7 !important;color:#222 !important;border:1px solid #ddd;padding:4px 10px;text-align:right;"
  ltxt  <- "text-align:left;"
  tbl_s <- "background:#ffffff !important;color:#222 !important;border-collapse:collapse;font-family:-apple-system,Segoe UI,Roboto,sans-serif;font-size:13px;margin:8px 0;"

  thead <- paste0(
    "<tr>",
    paste0(sprintf('<th style="%s">%s</th>', th_s, cols), collapse = ""),
    "</tr>"
  )
  rows <- vapply(seq_len(nrow(df)), function(i) {
    cells <- vapply(seq_along(cols), function(j) {
      base <- if (i %% 2 == 0) td_z else td_s
      sty  <- if (j == 1) paste0(base, ltxt) else base
      sprintf('<td style="%s">%s</td>', sty, fmt(df[[j]][i], j))
    }, character(1))
    paste0("<tr>", paste(cells, collapse = ""), "</tr>")
  }, character(1))

  htmltools::HTML(paste0(
    sprintf('<table style="%s">', tbl_s),
    "<thead>", thead, "</thead>",
    "<tbody>", paste(rows, collapse = ""), "</tbody>",
    "</table>"
  ))
}
```

## What this note shows

## Citation, EV, and AV ranks at the docdb level

Collapse the app's main database to one row per `docdb_family_id`, attach
flags telling us whether the family is mapped to **China**, any **HIC**, or
any **non-HIC country (excluding China)**, then rank docdbs (1 = highest)
by:

* **Citation count** (`cit_count`) — number of times a docdb is cited by
  another docdb in the app database.
* **EV global** (`ev_global`) — Watson average global spillover.
* **AV global** (`av_global`) — Watson average global return.

Country-class flags are not mutually exclusive (a family can be mapped to
e.g. both the US and China). For colouring the scatter we collapse to a
single primary class with priority **China → HIC → non-HIC excl. China**.

```{r build_docdb_ranks}
# Locate inst/extdata regardless of where the Rmd is rendered from.
find_extdata <- function(file) {
  cand <- c(
    system.file("extdata", file, package = "innovationStrategyExplorer"),
    file.path("..", "inst", "extdata", file),
    file.path("inst", "extdata", file)
  )
  hit <- cand[nzchar(cand) & file.exists(cand)]
  if (!length(hit)) stop("Could not locate ", file)
  normalizePath(hit[[1]], winslash = "/")
}
pdb_path <- find_extdata("patent_database.parquet")
clu_path <- find_extdata("country_lookup.parquet")

# Pull a per-docdb table directly from DuckDB so we never materialise the
# 28 M-row patent_database in R memory. ANY_VALUE on the family-constant
# columns is safe because cit_count, ev_global, av_global don't vary
# across countries within a docdb.
con <- dbConnect(duckdb::duckdb())
# Note: av_global is no longer materialised in patent_database.parquet —
# it is derived by the runtime view in R/runAppPackage.R from
# (pv + ev_global) / cost. We reproduce the same expression here so this
# Rmd can read the raw parquet without going through the package view.
docdb <- dbGetQuery(con, sprintf("
  WITH p   AS (SELECT * FROM read_parquet('%s')),
       cl  AS (SELECT * FROM read_parquet('%s'))
  SELECT
    p.docdb_family_id,
    ANY_VALUE(p.cit_count)  AS cit_count,
    ANY_VALUE(p.ev_global)  AS ev_global,
    ANY_VALUE(p.granted)    AS granted,
    ANY_VALUE(
      CASE WHEN p.cost IS NULL OR p.cost = 0 THEN 0
           ELSE (p.pv + p.ev_global) / p.cost
      END
    ) AS av_global,
    MAX(CASE WHEN cl.is_hic = TRUE THEN 1 ELSE 0 END)               AS any_hic,
    MAX(CASE WHEN p.ctry_code = 'CN' THEN 1 ELSE 0 END)             AS any_cn,
    MAX(CASE WHEN cl.is_hic = FALSE AND p.ctry_code <> 'CN' THEN 1
             ELSE 0 END)                                            AS any_lmic_excl_cn
  FROM p
  LEFT JOIN cl ON cl.ctry_code = p.ctry_code
  GROUP BY p.docdb_family_id
", pdb_path, clu_path))
dbDisconnect(con, shutdown = TRUE)

# Restrict to docdbs that have at least one incoming citation. Without
# this filter the bottom ~half of the universe (cit_count = 0) all share
# the same lowest citation rank, which collapses the lower-left of every
# scatter plot into a single point and pulls every regression's mean
# index toward zero. Filtering before rank computation reshapes ranks to
# range over the cited subset only.
docdb <- docdb %>%
  filter(cit_count > 0) %>%
  mutate(
    cit_rank      = min_rank(desc(cit_count)),
    ev_rank       = min_rank(desc(ev_global)),
    av_rank       = min_rank(desc(av_global)),
    is_hic        = any_hic == 1,
    is_china      = any_cn  == 1,
    is_lmic_xcn   = any_lmic_excl_cn == 1,
    region_class  = factor(
      case_when(
        is_china    ~ "China",
        is_hic      ~ "HIC",
        is_lmic_xcn ~ "non-HIC excl China",
        TRUE        ~ "unmapped"
      ),
      levels = c("China", "HIC", "non-HIC excl China", "unmapped")
    )
  )

cat(sprintf("cited docdbs: %s   (HIC: %s, China: %s, non-HIC excl China: %s)\n",
            format(nrow(docdb), big.mark = ","),
            format(sum(docdb$is_hic),      big.mark = ","),
            format(sum(docdb$is_china),    big.mark = ","),
            format(sum(docdb$is_lmic_xcn), big.mark = ",")))
```

### Hidden-giant indices (citation rank vs value rank)

* `higx_ev = cit_rank − ev_rank`
* `higx_av = cit_rank − av_rank`

Both ranks have **1 = highest**, so a *negative* index means a docdb is
more highly cited than its value rank would predict — a candidate "hidden
giant". A *positive* value means the opposite (high value, few citations).

```{r higx_indices}
docdb <- docdb %>%
  mutate(
    higx_ev = ((cit_rank) - (ev_rank))/max(cit_rank)*100,
    higx_av = ((cit_rank) - (av_rank))/max(cit_rank)*100
  )

docdb %>%
  filter(region_class != "unmapped") %>%
  group_by(region_class) %>%
  summarise(
    n              = dplyr::n(),
    mean_higx_ev   = mean(higx_ev),
    median_higx_ev = stats::median(higx_ev),
    p75_higx_ev    = stats::quantile(higx_ev, 0.75),
    p90_higx_ev    = stats::quantile(higx_ev, 0.90),
    p95_higx_ev    = stats::quantile(higx_ev, 0.95),
    .groups        = "drop"
  ) %>%
  nice_kable(digits = 2, format.args = list(big.mark = ","))


docdb %>% distinct(docdb_family_id) %>% nrow()
docdb %>% nrow()
```

### Country-class flag overlap

Since `is_hic`, `is_china`, and `is_lmic_xcn` are not mutually exclusive,
the next regressions ask three overlapping yes/no questions. The table
below tabulates all 2^3 = 8 combinations of the flags so the overlap is
explicit. The `region_class` column shows where each combination ends up
under the priority rule **China → HIC → non-HIC excl China** that drives
the scatter.

```{r flag_overlap}
priority <- function(hic, cn, lmic) {
  ifelse(cn,  "China",
  ifelse(hic, "HIC",
  ifelse(lmic, "non-HIC excl China",
         "unmapped")))
}

docdb %>%
  count(is_hic, is_china, is_lmic_xcn, name = "n") %>%
  mutate(
    share        = n / sum(n),
    region_class = priority(is_hic, is_china, is_lmic_xcn)
  ) %>%
  select(is_hic, is_china, is_lmic_xcn, region_class, n, share) %>%
  arrange(desc(n)) %>%
  nice_kable(digits = c(NA, NA, NA, NA, 0, 4),
             format.args = list(big.mark = ","))
```

#### Granted families only

Same overlap table restricted to docdbs with at least one granted
application — useful as a check on whether the flag distribution shifts
when applications that never made it through examination are excluded.

```{r flag_overlap_granted}
docdb %>%
  filter(granted) %>%
  count(is_hic, is_china, is_lmic_xcn, name = "n") %>%
  mutate(
    share        = n / sum(n),
    region_class = priority(is_hic, is_china, is_lmic_xcn)
  ) %>%
  select(is_hic, is_china, is_lmic_xcn, region_class, n, share) %>%
  arrange(desc(n)) %>%
  nice_kable(digits = c(NA, NA, NA, NA, 0, 4),
             format.args = list(big.mark = ","))
```

#### `docdb_family_size` > 1

Same overlap table restricted to docdbs whose PATSTAT
`docdb_family_size` is at least 2 (multi-application families). The
eligible-family list is read from
`.bigdata/fam_size_min2.parquet`, the BigQuery cache that
`01-build-app-parquets.R` produces when the build-time family-size
filter is enabled.

```{r flag_overlap_multifam}
fam2_path <- local({
  cand <- c(file.path("..", ".bigdata", "fam_size_min2.parquet"),
            file.path(".bigdata", "fam_size_min2.parquet"))
  hit  <- cand[file.exists(cand)]
  if (!length(hit)) NA_character_ else normalizePath(hit[[1]], winslash = "/")
})

if (is.na(fam2_path)) {
  cat("fam_size_min2.parquet not found — rerun the build pipeline (or set\n",
      "FAMILY_SIZE_MIN <- 2L) to materialise the eligible-family cache.\n")
} else {
  con_fs <- dbConnect(duckdb::duckdb())
  fams2 <- dbGetQuery(con_fs, sprintf(
    "SELECT DISTINCT docdb_family_id FROM read_parquet('%s')",
    fam2_path))
  dbDisconnect(con_fs, shutdown = TRUE)

  docdb %>%
    semi_join(fams2, by = "docdb_family_id") %>%
    count(is_hic, is_china, is_lmic_xcn, name = "n") %>%
    mutate(
      share        = n / sum(n),
      region_class = priority(is_hic, is_china, is_lmic_xcn)
    ) %>%
    select(is_hic, is_china, is_lmic_xcn, region_class, n, share) %>%
    arrange(desc(n)) %>%
    nice_kable(digits = c(NA, NA, NA, NA, 0, 4),
               format.args = list(big.mark = ","))
}
```

### Regressions on country-class flags

Both flag triples are non-mutually-exclusive — a docdb mapped to both the
US and China contributes to `is_hic = TRUE` *and* `is_china = TRUE`. The
intercept therefore represents a docdb with all three flags `FALSE`
(none); each coefficient is the conditional change in the index when that
particular flag flips on.

```{r reg_higx}
library(broom)
library(fixest)

mapped <- docdb %>%
  filter(is_hic | is_china | is_lmic_xcn)   # drop unmapped families

# fixest::feols beats stats::lm on memory and wall time for 28 M-row data
# even without fixed effects (multi-threaded C++ path, no full model
# matrix allocation). Default SEs are HC1 (heteroskedasticity-robust);
# pass vcov = "iid" if you want classical OLS SEs to match lm exactly.
fit_ev <- fixest::feols(higx_ev ~   is_china + is_lmic_xcn, data = mapped)

tidy(fit_ev, conf.int = TRUE) %>%
  mutate(outcome = "higx_ev") %>%
  select(outcome, term, estimate, std.error, statistic, p.value,
         conf.low, conf.high) %>%
  nice_kable(digits = c(NA, NA, 3, 3, 2, 4, 3, 3),
             format.args = list(big.mark = ","))

cat(sprintf(
  "n (mapped docdbs): %s   |   R^2 (higx_ev): %.4f\n",
  format(nrow(mapped), big.mark = ","),
  fixest::r2(fit_ev, type = "r2")
))
```

#### Granted families only

Re-run on the subset where `granted = TRUE` (the docdb has at least one
granted application). This filters out applications that never made it
through examination, which can carry citation patterns very different
from granted patents.

```{r reg_higx_granted}
mapped_g <- mapped %>% filter(granted)

fit_ev_g <- fixest::feols(higx_ev ~   is_china + is_lmic_xcn,
                          data = mapped_g)

tidy(fit_ev_g, conf.int = TRUE) %>%
  mutate(outcome = "higx_ev") %>%
  select(outcome, term, estimate, std.error, statistic, p.value,
         conf.low, conf.high) %>%
  nice_kable(digits = c(NA, NA, 3, 3, 2, 4, 3, 3),
             format.args = list(big.mark = ","))

cat(sprintf(
  "n (granted, mapped docdbs): %s   |   R^2 (higx_ev): %.4f\n",
  format(nrow(mapped_g), big.mark = ","),
  fixest::r2(fit_ev_g, type = "r2")
))
```

#### `docdb_family_size` > 1

Re-run on the multi-application subset (PATSTAT `docdb_family_size >= 2`),
using the same `.bigdata/fam_size_min2.parquet` cache that the overlap
table above reads.

```{r reg_higx_multifam, results='asis'}
fam2_path <- local({
  cand <- c(file.path("..", ".bigdata", "fam_size_min2.parquet"),
            file.path(".bigdata", "fam_size_min2.parquet"))
  hit  <- cand[file.exists(cand)]
  if (!length(hit)) NA_character_ else normalizePath(hit[[1]], winslash = "/")
})

if (is.na(fam2_path)) {
  cat("fam_size_min2.parquet not found — skipping multi-family regressions.\n")
} else {
  con_fs <- dbConnect(duckdb::duckdb())
  fams2  <- dbGetQuery(con_fs, sprintf(
    "SELECT DISTINCT docdb_family_id FROM read_parquet('%s')", fam2_path))
  dbDisconnect(con_fs, shutdown = TRUE)

  mapped_m <- mapped %>% semi_join(fams2, by = "docdb_family_id")

  fit_ev_m <- fixest::feols(higx_ev ~  is_china + is_lmic_xcn,
                            data = mapped_m)

  # results='asis' lets us emit the raw HTML kable from inside this
  # if-block. `print()` would have wrapped the htmltools::HTML object as
  # plain-text output instead of rendering it.
  cat(as.character(
    tidy(fit_ev_m, conf.int = TRUE) %>%
      mutate(outcome = "higx_ev") %>%
      select(outcome, term, estimate, std.error, statistic, p.value,
             conf.low, conf.high) %>%
      nice_kable(digits = c(NA, NA, 3, 3, 2, 4, 3, 3),
                 format.args = list(big.mark = ","))
  ))

  cat(sprintf(
    "<pre style='margin-top:8px;background:#f5f5f5;color:#222;padding:6px 10px;font-size:12px;'>n (fam_size>=2, mapped docdbs): %s   |   R^2 (higx_ev): %.4f</pre>\n",
    format(nrow(mapped_m), big.mark = ","),
    fixest::r2(fit_ev_m, type = "r2")
  ))
}
```

### Coefficient comparison across specifications

Stacks the three sub-samples (all mapped docdbs, granted only,
`fam_size >= 2`) into one figure so the coefficient on each country-class
flag is comparable across them. Bars are coefficient point estimates with
95 % confidence intervals; the dashed zero line is the no-effect
reference.

```{r reg_higx_compare}
collect_tidy <- function(fit, spec) {
  if (is.null(fit)) return(NULL)
  tidy(fit, conf.int = TRUE) %>% mutate(spec = spec)
}

reg_long <- bind_rows(
  collect_tidy(fit_ev,   "all"),
  collect_tidy(fit_ev_g, "granted"),
  if (exists("fit_ev_m")) collect_tidy(fit_ev_m, "fam_size>=2")
) %>%
  filter(term != "(Intercept)") %>%
  mutate(spec = factor(spec, levels = c("all", "granted", "fam_size>=2")))

ggplot(reg_long, aes(x = term, y = estimate, fill = spec)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "#888") +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
                position = position_dodge(width = 0.8),
                width = 0.25) +
  scale_fill_manual(values = c(
    "all"          = "#1f77b4",
    "granted"      = "#ff7f0e",
    "fam_size>=2"  = "#2ca02c"
  )) +
  labs(
    title = "higx_ev country-class coefficients across specifications",
    x     = NULL,
    y     = "Coefficient estimate (with 95% CI)",
    fill  = "Sample"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
```

### EV-rank vs citation-rank scatter (1 % sample)

```{r scatter_helper, include=FALSE}
# Reusable scatter so the three subsamples (full / granted / fam_size>1)
# share identical aesthetics. `subset_df` is a pre-filtered docdb table;
# the helper handles the unmapped drop + Bernoulli sample + plot.
scatter_ev_cit <- function(subset_df, prop = 0.05, subtitle_extra = NULL,
                           seed = 42) {
  set.seed(seed)
  plot_df <- subset_df %>%
    filter(region_class != "unmapped") %>%
    # Drop China-only docdbs (is_china = TRUE with neither HIC nor non-HIC
    # excl-China overlap) — the user's request was to exclude families
    # mapped exclusively to China across all three scatters.
    filter(!(is_china & !is_hic & !is_lmic_xcn)) %>%
    # Convert ranks to percentiles within the eligible (post-exclusion)
    # universe so the two axes share the same 0-100 scale and the y=x
    # diagonal is interpretable.
    mutate(
      ev_pct  = ev_rank  / max(ev_rank,  na.rm = TRUE) * 100,
      cit_pct = cit_rank / max(cit_rank, na.rm = TRUE) * 100
    ) %>%
    slice_sample(prop = prop)

  # Cap each axis at the 25th percentile of the *displayed* sample so we
  # zoom in on the top-ranked end of the distribution (where the hidden-
  # giant signal lives). Joint max across both axes keeps the 45-degree
  # line interpretable.
  cap <- max(stats::quantile(plot_df$ev_pct,  0.25, na.rm = TRUE),
             stats::quantile(plot_df$cit_pct, 0.25, na.rm = TRUE))

  sub <- sprintf("n = %s sampled docdbs; axes are percentiles (1 = top of distribution)",
                 format(nrow(plot_df), big.mark = ","))
  if (!is.null(subtitle_extra)) sub <- paste0(subtitle_extra, " — ", sub)

  ggplot(plot_df, aes(ev_pct, cit_pct, colour = region_class)) +
    geom_abline(slope = 1, intercept = 0,
                colour = "#888", linetype = "dashed") +
    geom_point(alpha = 0.35, size = 0.6) +
    # Per-group 2D kernel-density contours over the same scatter — same
    # colour scale as the dots so each region_class is identifiable. The
    # `group = region_class` makes density estimates separate per class
    # rather than a single global surface.
    geom_density_2d(aes(group = region_class),
                    linewidth = 0.55, alpha = 0.9) +
    coord_cartesian(xlim = c(0, cap), ylim = c(0, cap)) +
    scale_x_continuous(labels = label_number(accuracy = 0.1)) +
    scale_y_continuous(labels = label_number(accuracy = 0.1)) +
    scale_colour_manual(values = c(
      "China"              = "#d62728",
      "HIC"                = "#1f77b4",
      "non-HIC excl China" = "#2ca02c"
    )) +
    labs(
      title    = "EV-global percentile vs citation-count percentile",
      subtitle = sub,
      x        = "EV global percentile (rank / max * 100)",
      y        = "Citation count percentile",
      colour   = "Country class"
    ) +
    guides(colour = guide_legend(override.aes = list(alpha = 1, size = 2))) +
    theme_minimal(base_size = 12)
}
```

```{r scatter_ev_vs_cit}
scatter_ev_cit(docdb, prop = 0.05, subtitle_extra = "all docdbs")
```

#### Granted families only

```{r scatter_ev_vs_cit_granted}
scatter_ev_cit(docdb %>% filter(granted),
               prop = 0.05,
               subtitle_extra = "granted families only")
```

#### `docdb_family_size` > 1

```{r scatter_ev_vs_cit_multifam}
fam2_path <- local({
  cand <- c(file.path("..", ".bigdata", "fam_size_min2.parquet"),
            file.path(".bigdata", "fam_size_min2.parquet"))
  hit  <- cand[file.exists(cand)]
  if (!length(hit)) NA_character_ else normalizePath(hit[[1]], winslash = "/")
})

if (is.na(fam2_path)) {
  cat("fam_size_min2.parquet not found — skipping fam_size>1 scatter.\n")
} else {
  con_fs <- dbConnect(duckdb::duckdb())
  fams2  <- dbGetQuery(con_fs, sprintf(
    "SELECT DISTINCT docdb_family_id FROM read_parquet('%s')", fam2_path))
  dbDisconnect(con_fs, shutdown = TRUE)
  print(scatter_ev_cit(
    docdb %>% semi_join(fams2, by = "docdb_family_id"),
    prop = 0.05,
    subtitle_extra = "fam_size >= 2"
  ))
}
```

### Quintile grid: share by country class

Bin both axes into five equal-width percentile bins (0-20, 20-40, 40-60,
60-80, 80-100) and assign every docdb to one of the resulting 5 x 5 = 25
cells. Within each country-class group (China / HIC / non-HIC excl
China) compute the share of the group's docdbs falling in each cell.
Each cell shows three coloured dots — one per group — with the dot's
size proportional to that group's share of its own population in that
cell, and a white-boxed label showing the share value. Because we
collapse to a 25-cell summary, this can be computed on the full data,
no Bernoulli sample needed.

```{r quintile_grid_helper, include=FALSE}
# Helper: compute and plot the per-cell share grid for a docdb subset.
quintile_grid <- function(subset_df, subtitle_extra = NULL) {
  pdf <- subset_df %>%
    filter(region_class != "unmapped") %>%
    filter(!(is_china & !is_hic & !is_lmic_xcn)) %>%
    mutate(
      ev_pct  = ev_rank  / max(ev_rank,  na.rm = TRUE) * 100,
      cit_pct = cit_rank / max(cit_rank, na.rm = TRUE) * 100,
      ev_q    = cut(ev_pct,  breaks = seq(0, 100, by = 20),
                    include.lowest = TRUE, labels = 1:5),
      cit_q   = cut(cit_pct, breaks = seq(0, 100, by = 20),
                    include.lowest = TRUE, labels = 1:5)
    )

  shares <- pdf %>%
    count(region_class, ev_q, cit_q, name = "n") %>%
    group_by(region_class) %>%
    mutate(share = n / sum(n)) %>%
    ungroup() %>%
    mutate(
      ev_q  = as.integer(as.character(ev_q)),
      cit_q = as.integer(as.character(cit_q))
    )

  sub <- sprintf("n = %s docdbs (full sample, no Bernoulli sub-sampling)",
                 format(nrow(pdf), big.mark = ","))
  if (!is.null(subtitle_extra)) sub <- paste0(subtitle_extra, " - ", sub)

  ggplot(shares, aes(x = ev_q, y = cit_q,
                     size = share, colour = region_class)) +
    geom_point(position = position_dodge(width = 0.65), alpha = 0.85) +
    # Label sits on top of (in front of) the circle in a white box so
    # the percentage stays legible regardless of circle colour or size.
    # `colour` inherits region_class -> coloured text + matching border.
    geom_label(aes(label = scales::label_percent(accuracy = 0.1)(share)),
               position = position_dodge(width = 0.65),
               size = 2.4, show.legend = FALSE,
               fill = "white", alpha = 0.95,
               label.size = 0.25,
               label.padding = unit(0.10, "lines"),
               label.r = unit(0.12, "lines")) +
    scale_size_area(max_size = 16,
                    labels = scales::label_percent(accuracy = 0.1)) +
    scale_colour_manual(values = c(
      "China"              = "#d62728",
      "HIC"                = "#1f77b4",
      "non-HIC excl China" = "#2ca02c"
    )) +
    scale_x_continuous(breaks = 1:5,
                       labels = c("Q1\n0-20", "Q2\n20-40", "Q3\n40-60",
                                  "Q4\n60-80", "Q5\n80-100")) +
    scale_y_continuous(breaks = 1:5,
                       labels = c("Q1\n0-20", "Q2\n20-40", "Q3\n40-60",
                                  "Q4\n60-80", "Q5\n80-100")) +
    labs(
      title    = "Quintile grid - share of country-class docdbs per cell",
      subtitle = sub,
      x        = "EV global percentile bin (rank, 1 = top)",
      y        = "Citation count percentile bin (rank, 1 = top)",
      colour   = "Country class",
      size     = "Share within group"
    ) +
    guides(colour = guide_legend(override.aes = list(alpha = 1, size = 4))) +
    theme_minimal(base_size = 12) +
    theme(panel.grid.minor = element_blank(),
          legend.position = "right")
}
```

```{r quintile_grid_all}
quintile_grid(docdb, subtitle_extra = "all docdbs")
```

#### Granted families only

```{r quintile_grid_granted}
quintile_grid(docdb %>% filter(granted),
              subtitle_extra = "granted families only")
```

#### `docdb_family_size` > 1

```{r quintile_grid_multifam}
fam2_path <- local({
  cand <- c(file.path("..", ".bigdata", "fam_size_min2.parquet"),
            file.path(".bigdata", "fam_size_min2.parquet"))
  hit  <- cand[file.exists(cand)]
  if (!length(hit)) NA_character_ else normalizePath(hit[[1]], winslash = "/")
})

if (is.na(fam2_path)) {
  cat("fam_size_min2.parquet not found - skipping fam_size>1 quintile grid.\n")
} else {
  con_fs <- dbConnect(duckdb::duckdb())
  fams2  <- dbGetQuery(con_fs, sprintf(
    "SELECT DISTINCT docdb_family_id FROM read_parquet('%s')", fam2_path))
  dbDisconnect(con_fs, shutdown = TRUE)
  print(quintile_grid(
    docdb %>% semi_join(fams2, by = "docdb_family_id"),
    subtitle_extra = "fam_size >= 2"
  ))
}
```

### `higx_ev` distribution by country class

`higx_ev` is `cit_rank - ev_rank` rescaled to percentile points (divided
by `max(cit_rank)` and multiplied by 100), so it ranges from roughly -100
(low citation rank, high EV rank) to +100 (high citation rank, low EV
rank). Negative values are the "hidden giant" direction.

Each diagram is a single side-by-side boxplot — one box per country class.
Three diagrams cover the three samples (all / granted / `fam_size >= 2`).
Outlier dots are suppressed because each subset has millions of rows;
the median, IQR, and whiskers carry the same information.

```{r boxplot_higx_helper, include=FALSE}
# Side-by-side boxplots of higx_ev for the three country classes within
# `subset_df`. The unmapped and China-only filters mirror the scatter and
# quartile-grid helpers so all three views share an identical universe.
boxplot_higx <- function(subset_df, subtitle_extra = NULL) {
  pdf <- subset_df %>%
    filter(region_class != "unmapped") %>%
    filter(!(is_china & !is_hic & !is_lmic_xcn))

  sub <- sprintf("n = %s docdbs", format(nrow(pdf), big.mark = ","))
  if (!is.null(subtitle_extra)) sub <- paste0(subtitle_extra, " - ", sub)

  ggplot(pdf, aes(x = region_class, y = higx_ev, fill = region_class)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "#888") +
    geom_boxplot(outlier.shape = NA, alpha = 0.85, width = 0.55) +
    scale_fill_manual(values = c(
      "China"              = "#d62728",
      "HIC"                = "#1f77b4",
      "non-HIC excl China" = "#2ca02c"
    )) +
    labs(
      title    = "higx_ev distribution by country class",
      subtitle = sub,
      x        = NULL,
      y        = "higx_ev = (cit_rank - ev_rank) / max(cit_rank) * 100"
    ) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none")
}
```

```{r boxplot_higx_all}
boxplot_higx(docdb, subtitle_extra = "all docdbs")
```

#### Granted families only

```{r boxplot_higx_granted}
boxplot_higx(docdb %>% filter(granted),
             subtitle_extra = "granted families only")
```

#### `docdb_family_size` > 1

```{r boxplot_higx_multifam}
fam2_path <- local({
  cand <- c(file.path("..", ".bigdata", "fam_size_min2.parquet"),
            file.path(".bigdata", "fam_size_min2.parquet"))
  hit  <- cand[file.exists(cand)]
  if (!length(hit)) NA_character_ else normalizePath(hit[[1]], winslash = "/")
})

if (is.na(fam2_path)) {
  cat("fam_size_min2.parquet not found - skipping fam_size>1 boxplot.\n")
} else {
  con_fs <- dbConnect(duckdb::duckdb())
  fams2  <- dbGetQuery(con_fs, sprintf(
    "SELECT DISTINCT docdb_family_id FROM read_parquet('%s')", fam2_path))
  dbDisconnect(con_fs, shutdown = TRUE)
  print(boxplot_higx(
    docdb %>% semi_join(fams2, by = "docdb_family_id"),
    subtitle_extra = "fam_size >= 2"
  ))
}
```

### `higx_ev` density by country class

Same data, same exclusions, but now as overlaid kernel-density curves so
the full shape of each group's distribution is visible — modes, skew,
heavy tails — not just the five-number summary the boxplots show. Each
group is normalised to integrate to 1, so the heights are directly
comparable regardless of group size.

```{r density_higx_helper, include=FALSE}
density_higx <- function(subset_df, subtitle_extra = NULL) {
  pdf <- subset_df %>%
    filter(region_class != "unmapped") %>%
    filter(!(is_china & !is_hic & !is_lmic_xcn))

  sub <- sprintf("n = %s docdbs", format(nrow(pdf), big.mark = ","))
  if (!is.null(subtitle_extra)) sub <- paste0(subtitle_extra, " - ", sub)

  ggplot(pdf, aes(x = higx_ev, colour = region_class)) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "#888") +
    # Lines only - filled densities make the three overlapping curves
    # hard to read in the tails. `fill = NA` keeps the geom but turns
    # off the polygon fill.
    geom_density(fill = NA, linewidth = 0.8) +
    scale_colour_manual(values = c(
      "China"              = "#d62728",
      "HIC"                = "#1f77b4",
      "non-HIC excl China" = "#2ca02c"
    )) +
    labs(
      title    = "higx_ev density by country class",
      subtitle = sub,
      x        = "higx_ev = (cit_rank - ev_rank) / max(cit_rank) * 100",
      y        = "Density",
      colour   = "Country class"
    ) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "right")
}
```

```{r density_higx_all}
density_higx(docdb, subtitle_extra = "all docdbs")
```

#### Granted families only

```{r density_higx_granted}
density_higx(docdb %>% filter(granted),
             subtitle_extra = "granted families only")
```

#### `docdb_family_size` > 1

```{r density_higx_multifam}
fam2_path <- local({
  cand <- c(file.path("..", ".bigdata", "fam_size_min2.parquet"),
            file.path(".bigdata", "fam_size_min2.parquet"))
  hit  <- cand[file.exists(cand)]
  if (!length(hit)) NA_character_ else normalizePath(hit[[1]], winslash = "/")
})

if (is.na(fam2_path)) {
  cat("fam_size_min2.parquet not found - skipping fam_size>1 density.\n")
} else {
  con_fs <- dbConnect(duckdb::duckdb())
  fams2  <- dbGetQuery(con_fs, sprintf(
    "SELECT DISTINCT docdb_family_id FROM read_parquet('%s')", fam2_path))
  dbDisconnect(con_fs, shutdown = TRUE)
  print(density_higx(
    docdb %>% semi_join(fams2, by = "docdb_family_id"),
    subtitle_extra = "fam_size >= 2"
  ))
}
```

### `higx_ev` cumulative distribution by country class

Empirical CDFs of the same three groups, overlaid as coloured lines.
Where one group's curve sits above another, that group has more mass at
or below the corresponding `higx_ev` value — i.e. is more skewed toward
the "EV-rank ahead of citation-rank" side of the index.

```{r ecdf_higx_helper, include=FALSE}
ecdf_higx <- function(subset_df, subtitle_extra = NULL) {
  pdf <- subset_df %>%
    filter(region_class != "unmapped") %>%
    filter(!(is_china & !is_hic & !is_lmic_xcn))

  sub <- sprintf("n = %s docdbs", format(nrow(pdf), big.mark = ","))
  if (!is.null(subtitle_extra)) sub <- paste0(subtitle_extra, " - ", sub)

  ggplot(pdf, aes(x = higx_ev, colour = region_class)) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "#888") +
    geom_hline(yintercept = 0.5, linetype = "dotted", colour = "#bbb") +
    stat_ecdf(geom = "step", linewidth = 0.8, pad = FALSE) +
    scale_colour_manual(values = c(
      "China"              = "#d62728",
      "HIC"                = "#1f77b4",
      "non-HIC excl China" = "#2ca02c"
    )) +
    scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
    labs(
      title    = "higx_ev cumulative distribution by country class",
      subtitle = sub,
      x        = "higx_ev = (cit_rank - ev_rank) / max(cit_rank) * 100",
      y        = "Empirical CDF (within group)",
      colour   = "Country class"
    ) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "right")
}
```

```{r ecdf_higx_all}
ecdf_higx(docdb, subtitle_extra = "all docdbs")
```

#### Granted families only

```{r ecdf_higx_granted}
ecdf_higx(docdb %>% filter(granted),
          subtitle_extra = "granted families only")
```

#### `docdb_family_size` > 1

```{r ecdf_higx_multifam}
fam2_path <- local({
  cand <- c(file.path("..", ".bigdata", "fam_size_min2.parquet"),
            file.path(".bigdata", "fam_size_min2.parquet"))
  hit  <- cand[file.exists(cand)]
  if (!length(hit)) NA_character_ else normalizePath(hit[[1]], winslash = "/")
})

if (is.na(fam2_path)) {
  cat("fam_size_min2.parquet not found - skipping fam_size>1 ECDF.\n")
} else {
  con_fs <- dbConnect(duckdb::duckdb())
  fams2  <- dbGetQuery(con_fs, sprintf(
    "SELECT DISTINCT docdb_family_id FROM read_parquet('%s')", fam2_path))
  dbDisconnect(con_fs, shutdown = TRUE)
  print(ecdf_higx(
    docdb %>% semi_join(fams2, by = "docdb_family_id"),
    subtitle_extra = "fam_size >= 2"
  ))
}
```

