---
title: "Patent Cluster Comparison — LMICs vs HICs"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: true
    toc_float: true
---

```{r knitr-defaults, include=FALSE}
# Keep code in the output but collapsed behind a "Show code" toggle (via
# `code_folding: hide` in the YAML header). Suppress messages, warnings,
# and text output by default so the rendered HTML shows only figures and
# tables. Chunks that intentionally print a table or progress text
# override locally with `results='asis'` / `results='markup'` (see the
# per-technology display chunk).
knitr::opts_chunk$set(
  message   = FALSE,
  warning   = FALSE,
  results   = "hide",
  fig.align = "center"
)
```

This notebook compares two sets of patents (LMICs vs HICs) for each technology
in a configurable list, by embedding their titles and abstracts, clustering
them together, and visualizing with distinct styling for each group.

To analyse a new technology, just add it to `tech_selections` in Section 2.

## 0 — Setup

```{r setup, message=FALSE, warning=FALSE}
pkgs <- c("httr2", "xml2", "dplyr", "tidyr", "tidytext", "stringr",
          "ggplot2", "ggwordcloud", "purrr", "tibble", "digest", "umap",
          "jsonlite", "ggrepel", "DBI", "duckdb", "glue")
install.packages(setdiff(pkgs, rownames(installed.packages())), quiet = TRUE)

library(httr2)
library(xml2)
library(dplyr)
library(tidyr)
library(tidytext)
library(stringr)
library(ggplot2)
library(ggwordcloud)
library(purrr)
library(tibble)
library(umap)
library(jsonlite)
library(ggrepel)
library(DBI)
library(duckdb)
library(glue)
```

## 1 — Helper functions

All reusable machinery lives here. Section 3 defines the pipeline function
that calls these; Section 4 runs the pipeline for each technology.

```{r helpers}
# ---- DuckDB connection & patent query ----------------------------------------

open_patent_db <- function(parquet_dir = file.path("..", "inst", "extdata")) {
  con <- DBI::dbConnect(duckdb::duckdb())
  DBI::dbExecute(con, glue::glue(
    "CREATE OR REPLACE VIEW full_patent_database AS
     SELECT * FROM read_parquet('{file.path(parquet_dir, \"patent_database.parquet\")}')"
  ))
  DBI::dbExecute(con, glue::glue(
    "CREATE OR REPLACE TABLE patents_x_tech AS
     SELECT * FROM read_parquet('{file.path(parquet_dir, \"patents_x_tech.parquet\")}')"
  ))
  DBI::dbExecute(con, glue::glue(
    "CREATE OR REPLACE TABLE tech_lookup AS
     SELECT * FROM read_parquet('{file.path(parquet_dir, \"tech_lookup.parquet\")}')"
  ))
  con
}

get_top_patent_ids <- function(con, country_group, tech, toflow, top_n,
                               group_definitions, granted_only = TRUE) {
  country_codes <- if (country_group %in% names(group_definitions)) {
    group_definitions[[country_group]]
  } else {
    country_group
  }
  country_sql <- paste0("'", paste(country_codes, collapse = "', '"), "'")
  granted_clause <- if (isTRUE(granted_only)) "AND p.granted = TRUE" else ""

  sql <- glue::glue("
    WITH filtered_tech AS (
      SELECT DISTINCT t.docdb_family_id, '{tech}' AS technology
      FROM patents_x_tech t
      JOIN tech_lookup tl ON t.technology = tl.technology
      WHERE tl.tech_group = '{tech}' OR t.technology = '{tech}'
    ),

    deduped AS (
      SELECT DISTINCT ON (ft.technology, p.docdb_family_id)
        ft.technology, p.docdb_family_id, p.appln_id, p.{toflow}
      FROM full_patent_database p
      INNER JOIN filtered_tech ft ON p.docdb_family_id = ft.docdb_family_id
      WHERE p.ctry_code IN ({country_sql})
        AND p.{toflow} IS NOT NULL
        {granted_clause}
    ),

    ranked AS (
      SELECT appln_id, {toflow},
        ROW_NUMBER() OVER (ORDER BY {toflow} DESC) AS rnk
      FROM deduped
    )

    SELECT appln_id FROM ranked
    WHERE rnk <= {top_n}
    ORDER BY {toflow} DESC
  ")
  DBI::dbGetQuery(con, sql)$appln_id
}

parse_app_number <- function(x) {
  m <- str_match(x, "^([A-Z]{2,4})(\\d.*)$")
  tibble(raw = x, country = m[, 2], docnumber = m[, 3])
}

# ---- EPO OPS: XML parsing -----------------------------------------------------

ns_epo <- c(ex = "http://www.epo.org/exchange", ops = "http://ops.epo.org")

extract_en_abstract <- function(doc) {
  abs_nodes <- xml_find_all(doc, ".//ex:abstract", ns_epo)
  if (length(abs_nodes) == 0) return(NA_character_)
  langs <- xml_attr(abs_nodes, "lang")
  en <- abs_nodes[!is.na(langs) & langs == "en"]
  node <- if (length(en) > 0) en[[1]] else abs_nodes[[1]]

  p_nodes <- xml_find_all(node, ".//ex:p", ns_epo)
  text <- if (length(p_nodes) > 0) paste(xml_text(p_nodes), collapse = " ") else xml_text(node)
  text <- str_squish(text)
  if (nchar(text) < 10) NA_character_ else text
}

extract_en_title <- function(doc) {
  title_nodes <- xml_find_all(doc, ".//ex:invention-title", ns_epo)
  if (length(title_nodes) == 0) return(NA_character_)
  langs <- xml_attr(title_nodes, "lang")
  en <- title_nodes[!is.na(langs) & langs == "en"]
  node <- if (length(en) > 0) en[[1]] else title_nodes[[1]]
  str_squish(xml_text(node))
}

extract_applicants <- function(doc) {
  app_nodes <- xml_find_all(doc, ".//ex:applicant[@data-format='docdb']", ns_epo)
  if (length(app_nodes) == 0) {
    app_nodes <- xml_find_all(doc, ".//ex:applicant", ns_epo)
  }
  empty <- tibble(applicant_names = NA_character_,
                  applicant_types = NA_character_,
                  is_company = NA)
  if (length(app_nodes) == 0) return(empty)

  applicants <- lapply(app_nodes, function(node) {
    name_nodes <- xml_find_all(node, ".//ex:applicant-name", ns_epo)
    name_langs <- xml_attr(name_nodes, "lang")
    en_name <- name_nodes[!is.na(name_langs) & name_langs == "en"]
    name_node <- if (length(en_name) > 0) en_name[[1]]
                 else if (length(name_nodes) > 0) name_nodes[[1]]
                 else NULL
    name <- if (!is.null(name_node)) str_squish(xml_text(name_node)) else NA_character_

    has_orgname   <- length(xml_find_all(name_node, ".//ex:orgname", ns_epo)) > 0
    has_firstname <- length(xml_find_all(name_node, ".//ex:first-name", ns_epo)) > 0
    app_type <- xml_attr(node, "app-type")

    is_corp <- has_orgname || (!has_firstname && !is.na(name) &&
      str_detect(name, regex("\\b(inc|corp|ltd|llc|gmbh|ag|sa|sas|bv|nv|co|plc|pty|ab|oy|as|se|spa|srl|kk|univ)\\b",
                             ignore_case = TRUE)))

    list(name = name, app_type = app_type %||% NA_character_, is_company = is_corp)
  })

  names_vec <- sapply(applicants, \(a) a$name)
  unique_idx <- !duplicated(names_vec) & !is.na(names_vec)
  applicants <- applicants[unique_idx]
  if (length(applicants) == 0) return(empty)

  tibble(
    applicant_names = paste(sapply(applicants, \(a) a$name), collapse = "; "),
    applicant_types = paste(sapply(applicants, \(a) a$app_type), collapse = "; "),
    is_company      = any(sapply(applicants, \(a) a$is_company))
  )
}

# ---- EPO OPS: token management & HTTP with retries ---------------------------

make_token_manager <- function(ops_key, ops_secret, initial_token) {
  env <- new.env()
  env$token      <- initial_token
  env$expires_at <- Sys.time() + 15 * 60

  env$refresh <- function() {
    message("    [TOKEN] Refreshing OPS access token...")
    resp <- request("https://ops.epo.org/3.2/auth/accesstoken") |>
      req_auth_basic(ops_key, ops_secret) |>
      req_body_form(grant_type = "client_credentials") |>
      req_perform()
    env$token      <- resp_body_json(resp)$access_token
    env$expires_at <- Sys.time() + 15 * 60
    message("    [TOKEN] Refreshed, valid until ",
            format(env$expires_at, "%H:%M:%S"))
  }

  env$get <- function() {
    if (Sys.time() > env$expires_at) env$refresh()
    env$token
  }
  env
}

make_ops_request <- function(token_mgr, max_retries = 15) {
  function(url) {
    for (attempt in seq_len(max_retries + 1)) {
      resp <- request(url) |>
        req_headers(Authorization = paste("Bearer", token_mgr$get()),
                    Accept        = "application/xml") |>
        req_error(is_error = \(resp) FALSE) |>
        req_perform()

      status <- resp_status(resp)
      throttle_hdr <- resp_header(resp, "X-Throttling-Control")

      if (!is.null(throttle_hdr) && grepl("overloaded", throttle_hdr, ignore.case = TRUE)) {
        message("    [THROTTLED] waiting 60s (attempt ", attempt, "/", max_retries, ")...")
        Sys.sleep(60); next
      }
      if (status %in% c(401, 403)) {
        message("    [", status, "] possibly expired token, refreshing...")
        token_mgr$refresh(); Sys.sleep(5); next
      }
      if (status %in% c(429, 500, 502, 503, 504)) {
        message("    [", status, "] waiting 60s (attempt ", attempt, "/", max_retries, ")...")
        Sys.sleep(60); next
      }
      if (status >= 400) {
        # Log non-retryable errors so we can see what went wrong
        message("    [HTTP ", status, "] giving up: ", url)
        return(NULL)
      }
      return(resp_body_xml(resp))
    }
    NULL
  }
}

fetch_one_patent <- function(raw_id, ops_request) {
  abstract <- NA_character_
  title    <- NA_character_
  applicant_info <- tibble(applicant_names = NA_character_,
                           applicant_types = NA_character_,
                           is_company = NA)

  # Use `biblio` as the search constant — lightweight, returns the biblio
  # exchange document which includes title, abstract, and applicants.
  # NOTE: the search endpoint accepts ONE constant only (not comma-separated).
  # `raw_id` is the publication number (publn_auth || publn_nr, e.g.
  # "EP1234567"), so we query with pn= (publication number), NOT ap=
  # (application number — a different PATSTAT identifier entirely).
  search_url <- paste0(
    "https://ops.epo.org/3.2/rest-services/published-data/search/biblio?q=pn%3D",
    raw_id
  )
  doc <- tryCatch(ops_request(search_url), error = function(e) NULL)

  if (!is.null(doc)) {
    abstract       <- extract_en_abstract(doc)
    title          <- extract_en_title(doc)
    applicant_info <- extract_applicants(doc)
  }

  src <- if (!is.na(abstract) || !is.na(title)) "OK" else "miss"
  message("  [", src, "] ", raw_id,
          if (!is.na(title)) paste0(" - ", str_trunc(title, 50)) else "",
          if (!is.na(applicant_info$applicant_names))
            paste0(" | ", str_trunc(applicant_info$applicant_names, 40)) else "")

  bind_cols(tibble(title = title, abstract = abstract), applicant_info)
}

fetch_patents <- function(parsed, cache_file, token_mgr) {
  if (file.exists(cache_file)) {
    cat("Loading cached EPO data from", cache_file, "\n")
    results <- readRDS(cache_file)
    cat("Loaded", sum(!is.na(results$abstract)), "abstracts from cache\n")
    return(results)
  }

  cat("No cache found, fetching from EPO API...\n")
  ops_request <- make_ops_request(token_mgr)
  partial_file <- paste0(cache_file, ".partial.rds")

  if (file.exists(partial_file)) {
    done <- readRDS(partial_file)
    done_keys <- done$raw
    cat("Resuming from partial cache:", nrow(done), "already fetched\n")
  } else {
    done <- tibble(); done_keys <- character(0)
  }

  todo <- parsed |> filter(!raw %in% done_keys)
  n_total <- nrow(parsed); n_done <- nrow(done)
  cat("Fetching", nrow(todo), "remaining patents (", n_done, "already done)...\n")

  for (i in seq_len(nrow(todo))) {
    row <- todo[i, ]
    Sys.sleep(2)
    if ((n_done + i) %% 50 == 0)
      message("--- Progress: ", n_done + i, " / ", n_total, " ---")

    fetched <- tryCatch(
      fetch_one_patent(row$raw, ops_request),
      error = function(e) {
        message("  [ERROR] ", row$raw, ": ", e$message)
        bind_cols(
          tibble(title = NA_character_, abstract = NA_character_),
          tibble(applicant_names = NA_character_,
                 applicant_types = NA_character_, is_company = NA)
        )
      }
    )
    done <- bind_rows(done, bind_cols(row, fetched))
    if (i %% 100 == 0) {
      saveRDS(done, partial_file)
      message("  [CHECKPOINT] Saved ", nrow(done), " / ", n_total)
    }
  }

  results <- done
  cat("\nFetched", sum(!is.na(results$abstract)), "abstracts,",
      sum(!is.na(results$title)), "titles, and",
      sum(!is.na(results$applicant_names)), "applicant records\n")
  cat("  of which", sum(results$is_company, na.rm = TRUE), "have a company assignee\n")

  saveRDS(results, cache_file)
  if (file.exists(partial_file)) file.remove(partial_file)
  results
}

# ---- OpenAI: embeddings & chat ----------------------------------------------

get_embeddings <- function(texts, api_key) {
  resp <- request("https://api.openai.com/v1/embeddings") |>
    req_headers(Authorization = paste("Bearer", api_key),
                `Content-Type` = "application/json") |>
    req_body_json(list(model = "text-embedding-3-small", input = texts)) |>
    req_perform()
  result <- resp_body_json(resp)
  do.call(rbind, lapply(result$data, function(x) unlist(x$embedding)))
}

chat_completion <- function(prompt, api_key, model = "gpt-4o-mini",
                            temperature = 0.2, max_tokens = 500) {
  resp <- request("https://api.openai.com/v1/chat/completions") |>
    req_headers(Authorization = paste("Bearer", api_key),
                `Content-Type` = "application/json") |>
    req_body_json(list(
      model = model,
      messages = list(list(role = "user", content = prompt)),
      temperature = temperature,
      max_tokens = max_tokens
    )) |>
    req_perform()
  raw <- resp_body_json(resp)$choices[[1]]$message$content
  str_remove_all(raw, "```json\\s*|```\\s*")
}

label_clusters <- function(embed_data, openai_key) {
  cluster_summaries <- embed_data |>
    group_by(cluster) |>
    summarise(
      patents = paste(
        if_else(!is.na(title), title, str_trunc(text, 100)),
        collapse = "\n- "),
      n = n(), .groups = "drop"
    ) |>
    mutate(entry = paste0("Cluster ", cluster, " (", n, " patents):\n- ", patents))

  prompt <- paste0(
    "Below are clusters of patents grouped by semantic similarity.\n\n",
    paste(cluster_summaries$entry, collapse = "\n\n"),
    "\n\n---\n",
    "For EACH cluster, provide a short descriptive label (2-5 words) that captures ",
    "the common technological theme. Be specific and technical.\n\n",
    "Return ONLY a JSON array of objects with keys \"cluster\" (the number) ",
    "and \"label\" (the short description). No other text."
  )

  raw_json <- chat_completion(prompt, openai_key, max_tokens = 500)
  jsonlite::fromJSON(raw_json) |>
    as_tibble() |>
    mutate(cluster = as.character(cluster))
}

consolidate_titles <- function(titles_df, openai_key, batch_size = 120) {
  counts <- titles_df |>
    filter(!is.na(simple_title)) |>
    count(simple_title, sort = TRUE, name = "freq")

  if (nrow(counts) <= 1) {
    return(titles_df |> mutate(simple_title_raw = simple_title))
  }

  consolidation_prompt <- paste0(
    "Your job: consolidate this list. Many labels are synonyms or near-duplicates ",
    "of the same broad theme — they should be merged into a single canonical label. ",
    "Canonical labels should be ONE or TWO words — use one word when it naturally ",
    "captures the theme, and two words when they give a more meaningful description.\n\n",
    "Examples of merges: 'solar panel' + 'photovoltaic' + 'pv module' → 'solar'; ",
    "'cancer drug' + 'oncology drug' + 'tumor therapy' → 'cancer drug'; ",
    "'lithium battery' + 'battery cell' + 'energy storage' → 'battery'; ",
    "'water filter' + 'desalination' + 'water treatment' → 'water purification'.\n\n",
    "Aim for roughly 10-25 broad canonical themes across the full dataset. ",
    "Every input label MUST appear in the output mapping.\n\n",
    "Return ONLY a JSON array of objects with keys \"from\" (the original label) ",
    "and \"to\" (the canonical label, lowercase, 1-2 words). No other text."
  )

  # Batch the labels to avoid output truncation
  batches <- split(counts, ceiling(seq_len(nrow(counts)) / batch_size))

  all_mappings <- lapply(seq_along(batches), function(bi) {
    batch <- batches[[bi]]
    entries <- paste0("- ", batch$simple_title, " (n=", batch$freq, ")",
                      collapse = "\n")
    prompt <- paste0(
      "Below is a list of short patent theme labels with their frequencies.\n\n",
      entries, "\n\n---\n", consolidation_prompt
    )
    message("    [consolidate] batch ", bi, "/", length(batches),
            " (", nrow(batch), " labels)")

    raw_json <- chat_completion(prompt, openai_key,
                                temperature = 0.1, max_tokens = 6000)
    tryCatch(
      jsonlite::fromJSON(raw_json) |>
        as_tibble() |>
        mutate(from = str_to_lower(str_squish(from)),
               to   = str_to_lower(str_squish(to))),
      error = function(e) {
        warning("Failed to parse consolidation batch ", bi, ": ", e$message)
        tibble(from = character(), to = character())
      }
    )
  }) |> bind_rows()

  if (nrow(all_mappings) == 0) {
    return(titles_df |> mutate(simple_title_raw = simple_title))
  }

  # If there were multiple batches, do a final cross-batch harmonization pass
  # on just the canonical labels to unify across batches
  canon_labels <- all_mappings |> distinct(to) |> pull(to)
  if (length(batches) > 1 && length(canon_labels) > 25) {
    message("    [consolidate] cross-batch pass on ", length(canon_labels),
            " canonical labels")
    cross_entries <- paste0("- ", canon_labels, collapse = "\n")
    cross_prompt <- paste0(
      "Below are canonical patent theme labels from separate batches.\n\n",
      cross_entries, "\n\n---\n",
      "Merge any remaining synonyms or near-duplicates into a single canonical label. ",
      "Keep labels that are already distinct. Aim for 10-25 themes total.\n\n",
      "Return ONLY a JSON array of objects with keys \"from\" and \"to\". ",
      "Labels that stay unchanged should still appear (from = to). No other text."
    )
    raw_json <- chat_completion(cross_prompt, openai_key,
                                temperature = 0.1, max_tokens = 4000)
    cross_map <- tryCatch(
      jsonlite::fromJSON(raw_json) |>
        as_tibble() |>
        mutate(from = str_to_lower(str_squish(from)),
               to   = str_to_lower(str_squish(to))),
      error = function(e) {
        warning("Cross-batch consolidation failed: ", e$message, " — skipping")
        tibble(from = character(), to = character())
      }
    )
    if (nrow(cross_map) > 0) {
      cross_vec <- setNames(cross_map$to, cross_map$from)
      all_mappings <- all_mappings |>
        mutate(to = unname(ifelse(to %in% names(cross_vec), cross_vec[to], to)))
    }
  }

  map_vec <- setNames(all_mappings$to, all_mappings$from)

  titles_df |>
    mutate(
      simple_title_raw = simple_title,
      simple_title = unname(ifelse(
        simple_title %in% names(map_vec),
        map_vec[simple_title],
        simple_title
      ))
    )
}

generate_simple_titles <- function(embed_data, openai_key, batch_size = 100) {
  prep <- embed_data |>
    mutate(global_id = row_number()) |>
    mutate(entry_text = paste0(
      if_else(!is.na(title), paste0("Title: ", title, ". "), ""),
      if_else(!is.na(abstract), paste0("Abstract: ", str_trunc(abstract, 300)), "")
    ))

  n <- nrow(prep)
  batches <- split(prep, ceiling(seq_len(n) / batch_size))

  results <- lapply(seq_along(batches), function(bi) {
    batch <- batches[[bi]]
    # Use local indices [1..batch_size] in the prompt, map back afterwards
    entries <- paste0("[", seq_len(nrow(batch)), "] ", batch$entry_text) |>
      paste(collapse = "\n\n")

    prompt <- paste0(
      "Below are patents with their titles and/or abstracts.\n\n",
      entries, "\n\n---\n",
      "For EACH patent, create a very short label — ONE or TWO words — ",
      "that a non-expert can understand. Use plain language, not technical jargon. ",
      "Prefer one word when it naturally captures the theme, but two words are ",
      "fine when they give a more meaningful description.\n\n",
      "CRITICAL: labels MUST repeat across similar patents. The goal is to group ",
      "patents into broad themes, NOT to describe each patent uniquely. ",
      "Think of categories, not descriptions. ",
      "Good examples: 'vaccine', 'battery', 'catalyst', 'solar', 'fertilizer', ",
      "'turbine', 'cancer drug', 'solar cell', 'water purification', 'crop protection'.\n\n",
      "Many patents SHOULD share the same label. If you find yourself writing ",
      "many distinct labels, you are being too specific — generalize further.\n\n",
      "Return ONLY a JSON array of objects with keys \"id\" (the bracket number) ",
      "and \"simple_title\" (the label, lowercase, 1-2 words). No other text."
    )

    message("    [titles] batch ", bi, "/", length(batches),
            " (", nrow(batch), " patents)")

    raw_json <- chat_completion(prompt, openai_key,
                                temperature = 0.3, max_tokens = 4000)

    parsed <- tryCatch(
      jsonlite::fromJSON(raw_json) |> as_tibble(),
      error = function(e) {
        warning("Failed to parse batch ", bi, ": ", e$message)
        tibble(id = integer(), simple_title = character())
      }
    )
    if (nrow(parsed) == 0) return(parsed)

    parsed |>
      mutate(
        id = as.integer(id),
        simple_title = str_to_lower(str_squish(simple_title)),
        # Map local batch id back to global row_number id
        id = batch$global_id[id]
      )
  })

  bind_rows(results)
}

# ---- Assignee harmonization --------------------------------------------------

# Regex patterns for per-assignee heuristic classification
# Organizations: corporate suffixes, universities, research institutions
org_suffix_regex <- regex(
  "\\b(inc|corp|ltd|llc|gmbh|ag|sa|sas|bv|nv|co|plc|pty|ab|oy|as|se|spa|srl|kk)\\b",
  ignore_case = TRUE
)
# University/academia keywords (multilingual)
university_regex <- regex(
  paste0("\\b(univ|univers|universit|college|polytechnic|",
         "institute|institut|academy|academia|school|schule|hochschule|",
         "ecole|ecole polytechnique|escuela|scuola|faculty)\\b"),
  ignore_case = TRUE
)
# Research institute / government keywords
research_regex <- regex(
  paste0("\\b(cnrs|csic|csiro|csir|fraunhofer|max planck|helmholtz|riken|",
         "national lab|laboratory|laboratoire|laboratorio|research|forschung|",
         "instituto|agency|ministry|ministr|commission|consejo|conseil|",
         "council|foundation|fund|fonds|fundacion|stiftung)\\b"),
  ignore_case = TRUE
)

# Explode the `applicant_names` column (semicolon-separated) into one row per
# (patent, raw assignee). Strips EPO-style "[CC]" country suffixes and applies
# per-assignee heuristic classification (company / university / research / person).
explode_assignees <- function(embed_data) {
  embed_data |>
    filter(!is.na(applicant_names)) |>
    select(raw, source, applicant_names) |>
    mutate(assignee = str_split(applicant_names, ";\\s*")) |>
    tidyr::unnest(assignee) |>
    mutate(assignee = str_squish(str_remove(assignee, "\\s*\\[[A-Z]{2,3}\\]\\s*$"))) |>
    filter(nchar(assignee) > 0) |>
    mutate(
      # Per-assignee heuristic classification
      is_university_h = str_detect(assignee, university_regex),
      is_research_h   = str_detect(assignee, research_regex),
      is_company_h    = str_detect(assignee, org_suffix_regex),
      is_org_h        = is_university_h | is_research_h | is_company_h,
      heuristic_type  = case_when(
        is_university_h ~ "university",
        is_research_h   ~ "research_institute",
        is_company_h    ~ "company",
        TRUE            ~ "unknown"  # needs GPT to decide
      )
    )
}

# Ask GPT to: (1) drop individual person names, (2) harmonise variants to a
# canonical firm/institution name, (3) classify as company/university/etc.
# Returns a mapping: from -> (to, entity_type).  to == NA for persons.
# Only names the heuristic couldn't classify confidently are sent to GPT;
# names pre-classified by the heuristic are harmonised in a separate GPT pass
# WITHIN each type to avoid cross-type merges.
harmonize_assignees <- function(assignees_df, openai_key, batch_size = 150) {
  # Get unique names with frequency AND their heuristic classification
  uniq <- assignees_df |>
    group_by(assignee) |>
    summarise(
      freq           = n(),
      heuristic_type = first(heuristic_type),
      .groups        = "drop"
    ) |>
    arrange(desc(freq))

  if (nrow(uniq) == 0) {
    return(tibble(from = character(), to = character(), entity_type = character()))
  }

  call_gpt_batch <- function(batch, prompt_prefix) {
    entries <- paste0("- ", batch$assignee, " (n=", batch$freq, ")",
                      collapse = "\n")
    prompt <- paste0(prompt_prefix, "\n\n", entries, "\n\n---\n",
      "Return ONLY a JSON array of objects with keys \"from\" (original name), ",
      "\"to\" (canonical short name, Title Case, 1-4 words, null for persons), ",
      "and \"entity_type\" (one of: \"company\", \"university\", ",
      "\"research_institute\", \"government\", \"other\", \"person\"). ",
      "EVERY input name MUST appear in the output. No other text."
    )

    raw_json <- chat_completion(prompt, openai_key,
                                temperature = 0.1, max_tokens = 6000)

    tryCatch(
      jsonlite::fromJSON(raw_json) |>
        as_tibble() |>
        mutate(
          from = str_squish(from),
          to   = if (is.character(to)) str_squish(to) else as.character(to),
          to   = na_if(to, "NA"),
          to   = na_if(to, ""),
          entity_type = str_to_lower(str_squish(entity_type))
        ),
      error = function(e) {
        warning("Failed to parse assignee batch: ", e$message)
        tibble(from = character(), to = character(), entity_type = character())
      }
    )
  }

  # Split by heuristic type so GPT harmonises WITHIN each type
  # (prevents e.g. a university variant accidentally merging with a company)
  run_for_type <- function(df, label, type_hint) {
    if (nrow(df) == 0) return(tibble(from = character(), to = character(),
                                     entity_type = character()))
    batches <- split(df, ceiling(seq_len(nrow(df)) / batch_size))
    prefix <- switch(label,
      "unknown" = paste0(
        "Below is a list of raw patent assignee names (with frequencies). ",
        "These could not be pre-classified. For EACH name:\n",
        "1. DECIDE if it is a PERSON or an organization (company, ",
        "university, research institute, government). ",
        "Persons look like 'SMITH JOHN', 'WANG WEI', 'GARCIA MARIA [US]'.\n",
        "2. For organizations, HARMONISE variants to a SHORT canonical name. ",
        "Examples: 'MERCK SHARP & DOHME CORP' + 'MERCK & CO INC' → 'Merck'; ",
        "'UNIV CALIFORNIA' + 'REGENTS OF THE UNIVERSITY OF CALIFORNIA' → 'UC System'.\n",
        "3. ASSIGN an entity_type."
      ),
      paste0(
        "Below is a list of raw patent assignee names (with frequencies) that ",
        "are all pre-classified as ", type_hint, ". ",
        "HARMONISE variants to a SHORT canonical name (Title Case, 1-4 words, ",
        "drop generic suffixes like Inc/Corp/Ltd/GmbH). ",
        "Examples: 'MERCK SHARP & DOHME CORP' + 'MERCK & CO INC' → 'Merck'. ",
        "Set entity_type = \"", type_hint, "\" for every entry unless a name ",
        "is clearly something else (in which case correct it). ",
        "A name is only a person if it is clearly an individual human — be strict, ",
        "since these were pre-classified as organizations."
      )
    )

    lapply(seq_along(batches), function(bi) {
      message("    [assignees:", label, "] batch ", bi, "/", length(batches),
              " (", nrow(batches[[bi]]), " names)")
      call_gpt_batch(batches[[bi]], prefix)
    }) |> bind_rows()
  }

  groups <- split(uniq, uniq$heuristic_type)

  bind_rows(
    run_for_type(groups$company            %||% uniq[0,], "company",            "company"),
    run_for_type(groups$university         %||% uniq[0,], "university",         "university"),
    run_for_type(groups$research_institute %||% uniq[0,], "research_institute", "research_institute"),
    run_for_type(groups$unknown            %||% uniq[0,], "unknown",            NA)
  )
}

# ---- Plotting functions ------------------------------------------------------

cluster_colors <- c(
  "#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00",
  "#a65628", "#f781bf", "#1b9e77", "#d95f02", "#7570b3"
)

make_umap_plot <- function(embed_data, k, tech_name) {
  # Two side-by-side panels (LMIC / HIC) sharing the UMAP coordinate system.
  # Clustering and UMAP positioning were computed on the combined data
  # upstream, so umap_x / umap_y and cluster_label are already comparable
  # across the two panels; facet_wrap with fixed (default) scales keeps the
  # axes aligned. Ensures each patent appears only in its own panel.
  n_lmic <- sum(embed_data$source == "LMIC")
  n_hic  <- sum(embed_data$source == "HIC")

  # Force panel order: LMIC on the left, HIC on the right
  embed_data <- embed_data |>
    mutate(source = factor(source, levels = c("LMIC", "HIC")))

  ggplot(embed_data,
         aes(x = umap_x, y = umap_y, colour = cluster_label)) +
    geom_point(size = 3.5, alpha = 0.85, shape = 16) +
    geom_text_repel(aes(label = plot_label),
                    size = 2.3, alpha = 0.9,
                    max.overlaps = 15, show.legend = FALSE) +
    scale_colour_manual(values = cluster_colors) +
    facet_wrap(~source, ncol = 2) +
    theme_minimal() +
    labs(
      title    = paste0("Patent Embedding Clusters — ", tech_name),
      subtitle = paste0(n_lmic, " LMIC patents  |  ", n_hic,
                        " HIC patents  |  ", k, " shared clusters"),
      x = "UMAP 1", y = "UMAP 2", colour = "Theme"
    ) +
    theme(
      legend.position  = "right",
      legend.text      = element_text(size = 9),
      strip.text       = element_text(size = 13, face = "bold"),
      panel.spacing.x  = unit(1.2, "lines")
    )
}

make_wordcloud_plot <- function(embed_data, tech_name, max_per_source = 60) {
  counts <- embed_data |>
    filter(!is.na(simple_title)) |>
    count(source, simple_title, sort = TRUE, name = "n") |>
    group_by(source) |>
    slice_max(n, n = max_per_source, with_ties = FALSE) |>
    ungroup()

  if (nrow(counts) == 0) {
    return(ggplot() + theme_void() +
           labs(title = paste0("No themes to display — ", tech_name)))
  }

  ggplot(counts, aes(label = simple_title, size = n, colour = source)) +
    geom_text_wordcloud_area(rm_outside = TRUE, shape = "square",
                             eccentricity = 1.4) +
    scale_size_area(max_size = 44) +
    scale_colour_manual(values = c("LMIC" = "#e41a1c", "HIC" = "#377eb8")) +
    facet_wrap(~source, ncol = 2) +
    theme_void() +
    labs(
      title = paste0("Patent Innovation Themes — ", tech_name),
      subtitle = "Top canonical themes by source"
    ) +
    theme(strip.text        = element_text(size = 14, face = "bold"),
          plot.title        = element_text(size = 16, face = "bold",
                                            margin = margin(0, 0, 2, 0)),
          plot.subtitle     = element_text(size = 11,
                                            margin = margin(0, 0, 4, 0)),
          plot.margin       = margin(2, 2, 2, 2),
          panel.spacing     = unit(0.15, "lines"),
          legend.position   = "none")
}

make_top_titles_bar <- function(title_counts, tech_name) {
  title_counts |>
    slice_max(n, n = 25, with_ties = FALSE) |>
    mutate(simple_title = reorder(simple_title, n)) |>
    ggplot(aes(n, simple_title, fill = n)) +
    geom_col(show.legend = FALSE) +
    scale_fill_gradient(low = "#2c7bb6", high = "#d7191c") +
    labs(x = "Number of Patents", y = NULL,
         title = paste0("Top Innovation Themes — ", tech_name)) +
    theme_minimal()
}

entity_type_colors <- c(
  "company"            = "#377eb8",
  "university"         = "#e41a1c",
  "research_institute" = "#4daf4a",
  "government"         = "#984ea3",
  "other"              = "#999999"
)

make_assignee_wordcloud <- function(assignees_df, tech_name, max_per_source = 60) {
  # Drop persons and unclassified; keep most common entity_type per canonical
  counts <- assignees_df |>
    filter(!is.na(canonical),
           !entity_type %in% c("person", NA)) |>
    group_by(source, canonical) |>
    summarise(n = n(),
              entity_type = names(sort(table(entity_type), decreasing = TRUE))[1],
              .groups = "drop") |>
    filter(entity_type %in% names(entity_type_colors)) |>
    group_by(source) |>
    slice_max(n, n = max_per_source, with_ties = FALSE) |>
    ungroup()

  if (nrow(counts) == 0) {
    return(ggplot() + theme_void() +
           labs(title = paste0("No assignees to display — ", tech_name)))
  }

  ggplot(counts, aes(label = canonical, size = n, colour = entity_type)) +
    geom_text_wordcloud_area(rm_outside = TRUE, shape = "square",
                             eccentricity = 1.4) +
    scale_size_area(max_size = 38) +
    scale_colour_manual(values = entity_type_colors, name = "Entity type") +
    facet_wrap(~source, ncol = 2) +
    theme_void() +
    labs(
      title = paste0("Harmonized Assignees — ", tech_name),
      subtitle = "Firms and institutions only (persons filtered out); colored by entity type"
    ) +
    theme(strip.text = element_text(size = 14, face = "bold"),
          plot.title        = element_text(size = 16, face = "bold",
                                            margin = margin(0, 0, 2, 0)),
          plot.subtitle     = element_text(size = 11,
                                            margin = margin(0, 0, 4, 0)),
          legend.position = "bottom",
          plot.margin = margin(2, 2, 2, 2),
          panel.spacing = unit(0.5, "lines"))
}

make_source_comparison_bar <- function(embed_data, tech_name) {
  wide <- embed_data |>
    filter(!is.na(simple_title)) |>
    count(simple_title, source) |>
    tidyr::pivot_wider(names_from = source, values_from = n, values_fill = 0)
  if (!"LMIC" %in% names(wide)) wide$LMIC <- 0L
  if (!"HIC"  %in% names(wide)) wide$HIC  <- 0L

  long <- wide |>
    mutate(Total = LMIC + HIC) |>
    slice_max(Total, n = 20, with_ties = FALSE) |>
    tidyr::pivot_longer(cols = c(LMIC, HIC), names_to = "source", values_to = "n")

  ggplot(long, aes(x = n, y = reorder(simple_title, Total), fill = source)) +
    geom_col(position = "dodge") +
    scale_fill_manual(values = c("LMIC" = "#e41a1c", "HIC" = "#377eb8")) +
    labs(x = "Number of Patents", y = NULL,
         title = paste0("Top Innovation Themes by Source — ", tech_name),
         fill = "Source") +
    theme_minimal() +
    theme(legend.position = "top")
}
```

## 2 — Configuration

Add or remove technology names here to change what gets analysed. Each entry
produces its own full set of plots in Section 5.

```{r config}
tech_selections <- c("Any Agriculture & Food technology" )#, "Green Technology")

toflow         <- "av_global"
top_n_ids      <- 250
country_group1 <- "LMICs (excl. China)"
country_group2 <- "High income countries"

# Load country group definitions from the package's sysdata
load(file.path("..", "R", "sysdata.rda"))

# Open shared DuckDB connection
con <- open_patent_db()

# OPS authentication (done once; token_mgr will auto-refresh)
ops_key    <- Sys.getenv("OPS_CONSUMER_KEY",    unset = "YOUR_KEY_HERE")
ops_secret <- Sys.getenv("OPS_CONSUMER_SECRET", unset = "YOUR_SECRET_HERE")
stopifnot("Set your OPS credentials via env vars OPS_CONSUMER_KEY / OPS_CONSUMER_SECRET" =
            ops_key != "YOUR_KEY_HERE")

token_resp <- request("https://ops.epo.org/3.2/auth/accesstoken") |>
  req_auth_basic(ops_key, ops_secret) |>
  req_body_form(grant_type = "client_credentials") |>
  req_perform()
initial_token <- resp_body_json(token_resp)$access_token
token_mgr <- make_token_manager(ops_key, ops_secret, initial_token)

# OpenAI key
openai_key <- Sys.getenv("OPENAI_API_KEY", unset = "")
stopifnot("Set OPENAI_API_KEY in your .Renviron" = nchar(openai_key) > 0)

cache_dir <- ".bigdata"
if (!dir.exists(cache_dir)) dir.create(cache_dir, recursive = TRUE)
```

## 3 — Pipeline function

`analyze_tech()` runs the full pipeline for a single technology and returns a
list containing the processed data and the plots.

```{r pipeline}
analyze_tech <- function(tech_selection) {
  cat("\n========== ", tech_selection, " ==========\n")

  # --- 1. Query patent IDs ---
  apps_lmic <- get_top_patent_ids(con, country_group1, tech_selection,
                                  toflow, top_n_ids, group_definitions)
  apps_hic  <- get_top_patent_ids(con, country_group2, tech_selection,
                                  toflow, top_n_ids, group_definitions)
  cat("LMICs:", length(apps_lmic), "  HICs:", length(apps_hic), "\n")

  # Keep ALL rows (a patent in both lists appears twice, once per source)
  all_apps <- tibble(
    raw = c(apps_lmic, apps_hic),
    source = c(rep("LMIC", length(apps_lmic)), rep("HIC", length(apps_hic)))
  )
  n_overlap <- sum(duplicated(all_apps$raw))
  cat("Union:", nrow(all_apps), "rows (", n_overlap, "patents in both groups)\n")

  # --- 2. Parse & fetch abstracts (deduplicated to avoid double-fetching) ---
  unique_apps <- all_apps |> distinct(raw)
  parsed <- unique_apps |>
    left_join(bind_rows(lapply(unique_apps$raw, parse_app_number)), by = "raw")

  cache_file <- file.path(cache_dir, paste0(
    "epo_compare_", digest::digest(sort(unique_apps$raw), algo = "md5"), ".rds"
  ))
  fetched <- fetch_patents(parsed, cache_file, token_mgr)

  # Rejoin fetched data back to ALL rows (including both sources for overlaps)
  results <- all_apps |>
    left_join(fetched |> select(-any_of("source")), by = "raw")

  # --- 3. Build embedding dataset ---
  embed_data <- results |>
    filter(!is.na(abstract) | !is.na(title)) |>
    mutate(
      text = paste(
        if_else(!is.na(title), title, ""),
        if_else(!is.na(abstract), abstract, "")
      ) |> str_squish()
    ) |>
    filter(nchar(text) > 10)

  if (nrow(embed_data) < 3) {
    warning("Too few patents with text for ", tech_selection, " — skipping.")
    return(NULL)
  }

  # Compute embeddings on UNIQUE patents only (avoid embedding overlaps twice)
  unique_embed <- embed_data |> distinct(raw, .keep_all = TRUE)
  cat("Creating embeddings for", nrow(unique_embed), "unique patents...\n")
  embedding_matrix <- get_embeddings(unique_embed$text, openai_key)

  # --- 4. Cluster (on unique patents) ---
  set.seed(42)
  k <- min(10, nrow(embedding_matrix))
  clusters <- kmeans(embedding_matrix, centers = k, nstart = 25)
  unique_embed$cluster <- as.factor(clusters$cluster)

  # --- 5. AI cluster labels ---
  cluster_labels <- label_clusters(unique_embed, openai_key)
  label_map <- setNames(cluster_labels$label, cluster_labels$cluster)
  unique_embed <- unique_embed |>
    mutate(cluster_label = factor(label_map[as.character(cluster)], levels = label_map))

  # --- 6. UMAP projection (on unique patents) ---
  umap_config <- umap.defaults
  umap_config$n_neighbors <- min(15, nrow(embedding_matrix) - 1)
  umap_config$min_dist <- 0.1
  umap_result <- umap(embedding_matrix, config = umap_config)
  unique_embed$umap_x <- umap_result$layout[, 1]
  unique_embed$umap_y <- umap_result$layout[, 2]
  unique_embed <- unique_embed |>
    mutate(plot_label = if_else(!is.na(title), str_trunc(title, 30), raw))

  # --- 7. Simple titles (two-pass: label, then consolidate synonyms) ---
  #     Run on unique patents only (avoid labelling the same text twice)
  simple_titles_df <- generate_simple_titles(unique_embed, openai_key)
  simple_titles_df <- consolidate_titles(simple_titles_df, openai_key)

  n_raw    <- dplyr::n_distinct(simple_titles_df$simple_title_raw, na.rm = TRUE)
  n_canon  <- dplyr::n_distinct(simple_titles_df$simple_title,     na.rm = TRUE)
  cat("Simple titles: ", n_raw, " raw → ", n_canon, " canonical\n", sep = "")

  unique_embed <- unique_embed |>
    mutate(id = row_number()) |>
    left_join(simple_titles_df, by = "id")

  # Rejoin all computed fields back to the full embed_data (both LMIC & HIC rows)
  computed_cols <- c("raw", "cluster", "cluster_label", "umap_x", "umap_y",
                     "plot_label", "simple_title", "simple_title_raw", "id")
  computed_cols <- intersect(computed_cols, names(unique_embed))
  embed_data <- embed_data |>
    select(-any_of(setdiff(computed_cols, "raw"))) |>
    left_join(unique_embed |> select(all_of(computed_cols)), by = "raw")

  # --- 8. Assignee harmonization (drop persons, merge variants, classify) ---
  assignees_long <- explode_assignees(embed_data)
  assignee_map   <- harmonize_assignees(assignees_long, openai_key)

  assignees_clean <- assignees_long |>
    left_join(assignee_map, by = c("assignee" = "from")) |>
    rename(canonical = to)

  n_raw_assignees   <- dplyr::n_distinct(assignees_long$assignee)
  n_canon_assignees <- dplyr::n_distinct(assignees_clean$canonical[!is.na(assignees_clean$canonical)])
  n_persons_dropped <- sum(is.na(assignees_clean$canonical) |
                           assignees_clean$entity_type == "person", na.rm = TRUE)
  cat("Assignees: ", n_raw_assignees, " raw → ", n_canon_assignees,
      " canonical firms/institutions (",
      n_persons_dropped, " person-rows dropped)\n", sep = "")
  cat("  Entity types: ",
      paste(names(table(assignees_clean$entity_type)), "=",
            as.integer(table(assignees_clean$entity_type)), collapse = ", "),
      "\n", sep = "")

  # --- 9. Build plots ---
  title_counts <- embed_data |>
    filter(!is.na(simple_title)) |>
    count(simple_title, sort = TRUE)

  src_summary_wide <- embed_data |>
    count(source, cluster_label) |>
    tidyr::pivot_wider(names_from = source, values_from = n, values_fill = 0)
  if (!"LMIC" %in% names(src_summary_wide)) src_summary_wide$LMIC <- 0L
  if (!"HIC"  %in% names(src_summary_wide)) src_summary_wide$HIC  <- 0L
  src_summary <- src_summary_wide |>
    mutate(Total = LMIC + HIC) |>
    arrange(desc(Total))

  list(
    tech              = tech_selection,
    embed_data        = embed_data,
    cluster_labels    = cluster_labels,
    k                 = k,
    title_counts      = title_counts,
    src_summary       = src_summary,
    assignees_clean   = assignees_clean,
    umap_plot         = make_umap_plot(embed_data, k, tech_selection),
    wordcloud         = make_wordcloud_plot(embed_data, tech_selection),
    top_titles_bar    = make_top_titles_bar(title_counts, tech_selection),
    source_bar        = make_source_comparison_bar(embed_data, tech_selection),
    assignee_wordcloud = make_assignee_wordcloud(assignees_clean, tech_selection)
  )
}
```

## 4 — Run pipeline for each technology

```{r run_pipelines, message=TRUE, warning=TRUE}
results_by_tech <- lapply(tech_selections, analyze_tech)
names(results_by_tech) <- tech_selections

# Drop any that were skipped (NULL)
results_by_tech <- results_by_tech[!sapply(results_by_tech, is.null)]
```

## 5 — Results per technology

```{r results_display, results='asis', fig.width=14, fig.height=11, dpi=150, out.width='100%'}
for (tech in names(results_by_tech)) {
  res <- results_by_tech[[tech]]

  cat("\n\n### ", tech, "\n\n")

  cat("**Cluster labels:**\n\n")
  print(knitr::kable(res$cluster_labels))
  cat("\n\n")

  cat("**Patents per cluster × source:**\n\n")
  print(knitr::kable(res$src_summary))
  cat("\n\n")

  print(res$umap_plot)
  print(res$wordcloud)
  print(res$top_titles_bar)
  print(res$source_bar)
  print(res$assignee_wordcloud)
}
```

## 6 — Clean up

```{r cleanup}
DBI::dbDisconnect(con, shutdown = TRUE)
```
