---
title: "Tanks to Turbines"
subtitle: "Green innovation creates broader spillovers than defence — across origin economies"
author: "Innovation Strategy Explorer — insights"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 2
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 13, fig.height = 5.2, dpi = 150)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
library(scales)
library(htmltools)
suppressPackageStartupMessages({
  library(arrow); library(DBI); library(duckdb); library(leaflet)
})
source("R/local_higglobe.R")

# Resolve `inst/extdata/` whether the rmd is being knitted from the project
# root or from the insights/ directory.
data_dir <- {
  if      (dir.exists("../inst/extdata")) "../inst/extdata"
  else if (dir.exists("inst/extdata"))    "inst/extdata"
  else    stop("Cannot find inst/extdata/ — knit from project root.")
}
```

## The argument in one sentence

Defence R&D is routinely justified with the spillover argument — "DARPA gave
us the internet" — but applying the same yardstick to green-technology
innovation reveals that green tech generates broader and deeper citation
footprints than defence does, both at the umbrella level (per-innovation
spillovers, three origin economies) and at the citation-network level
(the citation graphs of the world's top-1,000 patents in each domain).

```{r autofetch, include=FALSE}
# ----------------------------------------------------------------------------
# Everything this note needs is computed LOCALLY from the parquets in
# inst/extdata/. No chromote, no HTTP, no live-app dependency.
#
#  * Bar charts: local_value_flows() aggregates patent_database directly,
#    matching the live Country-Explorer "Value flows by Technology" CSV
#    output (technology, mean, sem, innos). `granted_only = FALSE` =
#    full database, both granted and non-granted families.
#  * HiGGlobe maps: higglobe_local_sample() builds the citation-network
#    samples; higglobe_local_leaflet() renders them as interactive
#    leaflet widgets inline in the HTML output below.
#
# Both helpers live in insights/R/local_higglobe.R.
# ----------------------------------------------------------------------------

# Origin economies we contrast in the bar charts.
origins <- list(
  list(label = "LMICs (excl. China)",   local = "LMICs (excl. China)",
       ctry  = "LMICs (excl. China)",  tag   = "lmic"),
  list(label = "China",                 local = "CN",
       ctry  = "CN",                   tag   = "cn"),
  list(label = "High-Income Countries", local = "High income countries",
       ctry  = "High income countries", tag  = "hic")
)

tech_categories_full <- c(
  "All innovations",
  "Green Technology", "Green Energy", "Green Transport",
  "Green Manufacturing", "GHG Capture", "Adaptation",
  "Green Housing", "Circular Economy", "Green ICT",
  "Green Agriculture",
  "AI", "Any Agriculture & Food technology",
  "Defence Technology", "Any Defence technology",
  "Small Arms & Ordnance", "Non-firearm Weapons",
  "Projectile & Missile Launching", "Weapon Sights & Aiming",
  "Missile Propulsion", "Armour & Armoured Vehicles",
  "Targets & Training", "Ammunition", "Fuzes & Ammunition Safety",
  "Blasting & Demolition", "Naval Warfare",
  "Military Aviation Equipment", "Defence Radar & Sonar"
)

# Build a list of FOUR per-origin aggregates, one per sample variant:
#   full     — granted + non-granted, all family sizes (the default story)
#   granted  — granted families only
#   multifam — `docdb_family_size >= 2` only (one-off filings stripped out)
#   both     — granted AND `docdb_family_size >= 2`
# The same 28 tech categories and the same three origins are aggregated
# under each filter. About 3-4 s per variant.
flows_variants <- list(
  full     = list(label = "Full database",
                  desc  = "Granted + non-granted, all family sizes",
                  granted = FALSE, multifam = FALSE),
  granted  = list(label = "Granted only",
                  desc  = "Restricted to families flagged granted",
                  granted = TRUE,  multifam = FALSE),
  multifam = list(label = "Multi-application families",
                  desc  = "Restricted to docdb_family_size >= 2",
                  granted = FALSE, multifam = TRUE),
  both     = list(label = "Granted & multi-application",
                  desc  = "Both filters applied",
                  granted = TRUE,  multifam = TRUE)
)

flows_list <- lapply(flows_variants, function(v) {
  do.call(rbind, lapply(origins, function(o) {
    df <- local_value_flows(
      country       = o$local,
      techs         = tech_categories_full,
      toflow        = "ev_global",
      granted_only  = v$granted,
      multifam_only = v$multifam,
      data_dir      = data_dir
    )
    df$origin <- factor(o$label, levels = vapply(origins, `[[`, "", "label"))
    df
  }))
})

# Default `flows` for any code that still references it directly: the
# full-DB variant. Most charts now consume `flows_list[[variant]]`.
flows <- flows_list$full
have_data <- nrow(flows) > 0

# Build TWO large global HiGGlobe samples locally — one for green, one
# for defence. Both seeded with the top 1,000 patents worldwide by
# Average Global Spillover (`ev_global`); 1% edge sampling at each
# generation step keeps the map readable while following 4 generations
# of forward citations. Mirrors the live HiGGlobe params on
# https://mondpanther-iseapp2.share.connect.posit.cloud/?navbar_page=...
# (gen0_select_mode=Top, gen0_unit_mode=Number, gen0_sample_val=1000,
# edge_sample_val=1, add_generations=4).
panel_green <- higglobe_local_sample(
  country          = "All countries",
  techs            = "Green Technology",
  toflow           = "ev_global",
  gen0_n           = 1000L,
  gen0_mode        = "Top",
  n_generations    = 4L,
  edge_sample_pct  = 1,
  include_fallback = FALSE,
  seed             = 42,
  data_dir         = data_dir
)
panel_defence <- higglobe_local_sample(
  country          = "All countries",
  techs            = "Any Defence technology",
  toflow           = "ev_global",
  gen0_n           = 1000L,
  gen0_mode        = "Top",
  n_generations    = 4L,
  edge_sample_pct  = 1,
  include_fallback = FALSE,
  seed             = 42,
  data_dir         = data_dir
)
panel_ai <- higglobe_local_sample(
  country          = "All countries",
  techs            = "AI",
  toflow           = "ev_global",
  gen0_n           = 1000L,
  gen0_mode        = "Top",
  n_generations    = 4L,
  edge_sample_pct  = 1,
  include_fallback = FALSE,
  seed             = 42,
  data_dir         = data_dir
)
```

```{r helpers, include=FALSE}

# Build a HiGGlobe deep-link URL on the public app for any (techs, country)
# pair so each panel caption can link straight to the live, pannable view.
higglobe_url <- function(country, techs,
                         base = "https://mondpanther-iseapp2.share.connect.posit.cloud/") {
  # Scalar quoted-string format matches Shiny's own bookmark serialiser
  # output for multi-select selectize inputs, and round-trips correctly
  # through Shiny's bookmark-restore handler. JSON-array form
  # (`["…"]`) was tried and silently dropped on restore.
  q <- function(x) utils::URLencode(paste0('"', x, '"'), reserved = TRUE)
  paste0(
    base, "?_inputs_",
    "&navbar_page=", q("HiGGlobe"),
    "&hglobe-country=", q(country),
    "&hglobe-firm=",    q("No firm filter"),
    "&hglobe-techs=",   q(techs),
    "&hglobe-city=",    q("No city filter"),
    "&hglobe-toflow=",  q("ev_global"),
    "&hglobe-granted_only=false&hglobe-multifam_only=false",
    "&hglobe-include_fallback=false",
    # Match the local panels: top 1000 by ev_global, 1% edge sampling,
    # 4 follow-on generations.
    "&hglobe-gen0_select_mode=", q("Top"),
    "&hglobe-gen0_unit_mode=",   q("Number"),
    "&hglobe-gen0_sample_val=1000",
    "&hglobe-edge_select_mode=", q("Random"),
    "&hglobe-edge_unit_mode=",   q("Percent"),
    "&hglobe-edge_sample_val=1",
    "&hglobe-add_generations=4",
    "&hglobe-sidebar=true",
    "&hglobe-show_gen_0=true&hglobe-show_gen_1=true",
    "&hglobe-show_gen_2=true&hglobe-show_gen_3=true",
    "&hglobe-show_gen_4=true",
    # `step=N` flips the URL-driven auto-init switch in module_hglobe.R:
    # the module clicks Initiate Innovation 1.2 s after its observers
    # register, then auto-clicks Generate once gen 0 is seeded. Without
    # this the link only RESTORES the input values; nothing renders, so
    # the dropdown ends up looking empty even when its value is set.
    "&step=4"
  )
}
```

## How knowledge travels: green technology

Forward citations radiating out from the **top 1,000 green-tech innovations
worldwide** (ranked by Average Global Spillover, `ev_global`). `gen 0` (blue)
are the source patents; `gen 1` (red) is direct citations; `gen 2-4`
(green / orange / purple) are citations of citations. Edges are a 1%
random sample at each step so the map stays legible. Knowledge fans out
across every continent.

```{r green-globe}
htmltools::tagList(
  htmltools::tags$div(
    style = "margin: 6px 0 4px 0; font-size: 0.85rem; text-align: right;",
    htmltools::tags$a(
      "Open the live HiGGlobe view →",
      href   = higglobe_url("All countries", "Green Technology"),
      target = "_blank", rel = "noopener"
    )
  ),
  higglobe_local_leaflet(panel_green, height = "560px")
)
```

## Contrast: defence technology

Same construction, same parameters — top 1,000 defence-tech innovations
worldwide, four generations of forward citations, 1% edge sample.

```{r defence-globe}
htmltools::tagList(
  htmltools::tags$div(
    style = "margin: 6px 0 4px 0; font-size: 0.85rem; text-align: right;",
    htmltools::tags$a(
      "Open the live HiGGlobe view →",
      href   = higglobe_url("All countries", "Any Defence technology"),
      target = "_blank", rel = "noopener"
    )
  ),
  higglobe_local_leaflet(panel_defence, height = "560px")
)
```

Defence citations stay closer to their home cluster, fade to `gen 1` /
`gen 2`, and rarely reach the deeper generations — even though the gen 0
seeds are the top 1,000 defence patents the database knows about.

## For comparison: AI technology

The same construction applied to the **top 1,000 AI patents worldwide** —
where the spillover-driven justification for public investment is
arguably even louder than for defence. AI's citation network is denser
than green's at every generation, but is overwhelmingly concentrated in a
narrower band of countries (US west coast, China east coast, Korea,
Western Europe).

```{r ai-globe}
htmltools::tagList(
  htmltools::tags$div(
    style = "margin: 6px 0 4px 0; font-size: 0.85rem; text-align: right;",
    htmltools::tags$a(
      "Open the live HiGGlobe view →",
      href   = higglobe_url("All countries", "AI"),
      target = "_blank", rel = "noopener"
    )
  ),
  higglobe_local_leaflet(panel_ai, height = "560px")
)
```

## Cross-tech comparisons by sample {.tabset .tabset-pills}

The four pills below toggle the **patent sample** behind every chart in
this section — the headline four-umbrella bars, the subcategory drilldown,
and the share-by-subcategory donuts all redraw together. Every chart
re-runs against `patent_database.parquet` locally, so flipping between
samples is instantaneous after the first render.

* **Full database** — granted *and* non-granted, all family sizes (the
  default thesis: every patent the database knows about).
* **Granted only** — restricts to families flagged `granted = TRUE`.
* **Multi-application families** — `docdb_family_size >= 2`, i.e. only
  families that the original applicant cared enough to file in more than
  one office. A common signal of "real" innovation vs. one-off filings.
* **Granted & multi-application** — both filters at once. Strictest cut.

### Full database

```{r umb-full, fig.height=4.2, fig.width=13}
t2t_make_umbrella_bar(flows_list$full,
                      sample_label = flows_variants$full$desc)
```

[Open this view in the live Country Explorer →](https://mondpanther-iseapp2.share.connect.posit.cloud/?_inputs_&country-sidebar=true&navbar_page=%22Country%20Explorer%22&country-inner_tabs=%22Value%20flows%20by%20Technology%22&country-country=%22LMICs%20(excl.%20China)%22&country-firm=%22No%20firm%20filter%22&country-toflow=%22ev_global%22&country-tech_categories_plot1=%5B%22Green%20Technology%22%2C%22AI%22%2C%22Any%20Agriculture%20%26%20Food%20technology%22%2C%22Defence%20Technology%22%5D&country-widthscale=%22log%22&country-display_mode=%22confidence%22&country-granted_only=false&country-multifam_only=false)

```{r drill-full, fig.height=9, fig.width=13}
t2t_make_drilldown_bar(flows_list$full,
                       sample_label = flows_variants$full$desc)
```

```{r donut-full, fig.height=8.6, fig.width=13}
t2t_make_donut_grid(flows_list$full, origins)
```

### Granted only

```{r umb-granted, fig.height=4.2, fig.width=13}
t2t_make_umbrella_bar(flows_list$granted,
                      sample_label = flows_variants$granted$desc)
```

```{r drill-granted, fig.height=9, fig.width=13}
t2t_make_drilldown_bar(flows_list$granted,
                       sample_label = flows_variants$granted$desc)
```

```{r donut-granted, fig.height=8.6, fig.width=13}
t2t_make_donut_grid(flows_list$granted, origins)
```

### Multi-application families

```{r umb-multifam, fig.height=4.2, fig.width=13}
t2t_make_umbrella_bar(flows_list$multifam,
                      sample_label = flows_variants$multifam$desc)
```

```{r drill-multifam, fig.height=9, fig.width=13}
t2t_make_drilldown_bar(flows_list$multifam,
                       sample_label = flows_variants$multifam$desc)
```

```{r donut-multifam, fig.height=8.6, fig.width=13}
t2t_make_donut_grid(flows_list$multifam, origins)
```

### Granted & multi-application

```{r umb-both, fig.height=4.2, fig.width=13}
t2t_make_umbrella_bar(flows_list$both,
                      sample_label = flows_variants$both$desc)
```

```{r drill-both, fig.height=9, fig.width=13}
t2t_make_drilldown_bar(flows_list$both,
                       sample_label = flows_variants$both$desc)
```

```{r donut-both, fig.height=8.6, fig.width=13}
t2t_make_donut_grid(flows_list$both, origins)
```

## {-}

Whichever sample you flip to, the qualitative ranking — green and
agriculture above defence, on a per-innovation basis — is robust. The
"Granted & multi-application" pill is the most conservative cut, where
defence's relative position is highest; the headline gap survives.

## Why this matters

If the spillover argument is what justifies public R&D — DARPA, Horizon
Europe's defence pillar, the UK clean-tech budget — then the comparison
above ought to inform how the next budget cycle gets carved up. Two
findings stand out:

1. **Green-tech innovation travels further than defence** in every origin
   economy we look at — LMICs, China, and the High-Income countries.
2. **The per-innovation spillover gap is large** even for the origin
   economies where defence patenting is dense (HICs); the gap widens for
   LMICs where defence sample sizes are tiny.

Neither result tells us defence R&D has zero externalities — classified
filings don't show up in PATSTAT, and dual-use families that are tagged
civilian-first won't appear in the F41/F42 + cross-cutting CPC scope used
here. But on the evidence that *is* published, the developing-world
clean-tech innovator delivers the broader public good.

## Caveats

* **Defence patent coverage is partial.** Some military innovation is
  classified or filed under cover names. The CPC scope follows the academic
  patent-economics literature
  ([Acosta et al. 2013](https://link.springer.com/article/10.1007/s11192-012-0857-8)).
* **`gen 0` sample sizes differ across the HiGGlobe panels.** All three
  globes draw the same fixed-N random sample (`gen0_sample_val=100`) but
  the universe they're drawn from varies — China's defence universe is
  small; LMIC defence is smaller still.
* **`ev_global` is a model-derived value**, the shadow value implied by the
  citation diffusion model that powers the Innovation Strategy Explorer;
  the original methodology is in [Efficient industrial policy for
  innovation: standing on the shoulders of hidden
  giants](https://cep.lse.ac.uk/_NEW/publications/abstract.asp?index=8614).

## How to reproduce

1. The note reads from `inst/extdata/{patent_database, citenet,
   countrymap, patents_x_tech, tech_lookup, country_lookup}.parquet`. If
   those exist (you've run `data-raw-2025/whole_databuild_pipeline.R`),
   step 2 just works.
2. From the project root:
   `rmarkdown::render("insights/tanks2turbines.rmd")`.
   * Both the bar charts and the HiGGlobe samples are now computed
     locally via DuckDB over those parquets. See
     [`insights/R/local_higglobe.R`](R/local_higglobe.R) for both
     `local_value_flows()` (the bar-chart aggregator) and
     `higglobe_local_sample()` / `higglobe_local_leaflet()` (the
     citation-network sampler + renderer). No chromote, no live-app
     dependency, no per-render fetch.
3. Toggling `granted_only = TRUE / FALSE` in the autofetch chunk is
   instant — the aggregates rebuild from the parquet in a few seconds.
4. `source("insights/render_all.R")` to bundle the HTML for deployment.
