---
title: "Prepare istraxes"
output: html_notebook
---

C:/Users/rmartin7/AppData/Local/Temp/image.png--- title: "Prepare ISTRAX
tool" output: #word_document: default html_document: default date:
"2025-09-04" ---

# Pulling data together

This pulls the data together we already have in our directory.
Ultimately we want to merge this into a more dynamic data pipeline that
ideally access the the PATSTAT API.

Note that all directory links are relative so you should be able to find
all underlying datasets.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE,fig.width = 9,fig.height = 10,dpi=2200)
source("../code/helpers.r")
library(dplyr)
library(readr)
library(readr)
library(dplyr)
library(wbstats)
library(arrow)
library(readr)
library(data.table)

```

I am pulling various datasets we have in the one drive to create three
datastructures. Eventually we should establish a datapipeline that pulls
that data from the PATSTAT API pipeline. All data should be in the
shared drive and paths are relative so it should all work on your
machines.

1.  A dataset at the patent families (docdb level) with various family
    charactistics that we might want to show (patchar)

2.  A dataset that maps patents to countrys (countrymap). Currently
    that's done via inventor locations. Note because the same inno can
    have innovators in multiple countries. The countrymap dataset
    includes - at least for now - also a national istrax variable. This
    is a variable which computes istrax such that only spillovers are
    considered that benefit the source country (i.e. for innovations
    linked to India say it is flows to other innovators in India). The
    long term goal - somebody; e.g. Denis, needs to re-run the algorithm
    for all countries - to have in the patchar dataset a to flow istrax
    column for every country and country group of interest. Once we have
    that national istrax is contained in this information.

3.  A dataset that maps patent families to technology classes (techmap,
    again a family can be in multiple). (Maybe in the medium run we want
    to replace that with a mapping of all patents to CPC subclasses
    combined with a separate - and much smaller - mapping from
    subclasses to specific technology tags)

4.  We combine 1 & 2 to create patchar_countrymap, which is now more
    convenient for simply computing bar charts with national vs global
    spillovers istrax.

## Build mapping from patents to countries (countrymap)







```{r}
url0="../../data/patent_data/patbis2021/data/fromWATSON/innos_ev_nationalkey_2009_2018.dsv"
test0= read_tsv(url0)
atest=test0 %>% group_by(ctry_code) %>% summarise(n())


url="../../data/patent_data/patbis2021/data/fromWATSON/innos_ctry_indicators.dsv"
#df=read_tsv("https://www.dropbox.com/scl/fi/84hj4f81d9ups0z3v7gcu/innos_ctry_indicators.dsv?rlkey=5t8qnhib02e44aqpo48yu3m9b&st=0zcqzokk&dl=1")

#url="../../data/patent_data/patbis2021/data/fromWATSON/innos_istraxfield_nationalkey_2009_2018.dsv"

countrymap <- read_tsv(url)
#countrymap=countrymap %>% distinct(docdb_family_id,ctry_code,istrax) %>% rename(istrax_nationalkey_2009_2018=istrax)
countrymap=countrymap %>% distinct(docdb_family_id,ctry_code,istrax_nationalkey_2009_2018)

test=countrymap %>% group_by(ctry_code) %>% summarise(nn=n())

nrow(countrymap %>% filter(ctry_code=="CN"))


```

## Build patchar

```{r}

url2="../../data/patent_data/patbis2021/data/fromWATSON/innos_istraxfield_global_2009_2018.dsv"
patchar=read_tsv(url2) %>%  select(docdb_family_id,istrax,alphapvyear_2009_2018,costpvyear_2009_2018,pv,ev) %>% 
  rename(istrax_global=istrax) %>% 
   rename(cost=costpvyear_2009_2018,
          alpha=alphapvyear_2009_2018,
          ev_global=ev) %>% 
   mutate(avstrax_global=(ev_global+pv)/cost)

#library(fixest)
#reg=fepois(pv~ev,patchar) 
#reg %>% summary()






```

## Build patchar_countrymap

```{r}




patchar_countrymap = read_tsv(url) %>% select(docdb_family_id,ctry_code,
                                              ev_nationalkey_2009_2018,pv,
                            earliest_filing_date,earliest_filing_year,
                            docdb_family_size,
                            istrax_nationalkey_2009_2018,
                            cost) %>% 
                            mutate(avstrax_nationalkey_2009_2018=(pv+ev_nationalkey_2009_2018)/cost
                            ) %>% select(-cost)#%>% rename(istrax_nationalkey_2009_2018=istrax)

#...assuming this covers national we need global....

#test=
```
### trying building up country stuff from person file...


```{r}

library(arrow)

localbigwb="C:\\Users\\rmartin7\\OneDrive - WBG\\projects\\inglobe"
dfnew=read_parquet(paste0(localbigwb,"\\data\\dfnew.parquet"))

inventors=paste0(localbigwb,"\\data\\inventors.parquet") %>% 
  read_parquet()

holders=paste0(localbigwb,"\\data\\holders.parquet") %>% 
  read_parquet()


holder_countries=paste0(localbigwb,"\\data\\holder_countries.parquet") %>% 
  read_parquet()

inventor_countries=paste0(localbigwb,"\\data\\holder_countries.parquet") %>% 
  read_parquet()


```


## 😊 build country map from scracth

```{r}
dfnew %>% names()
inventors %>% names()
holders %>% names()

dfnew=dfnew %>% rename(ctry_code=person_ctry_code)

income_classification <- tibble::tribble(
  ~income_group, ~ctry_code,
  "High income", c("AD", "AE", "AG", "AR", "AU", "AT", "AW", "BE", "BH", "BM", "BS", "BW", "CA", "CH", "CL", "CN", "CW", "CY", "CZ", "DE", "DK", "EE", "ES", "FI", "FR", "GB", "GI", "GR", "GU", "HK", "HR", "HU", "IE", "IL", "IM", "IS", "IT", "JP", "KR", "KW", "KY", "LI", "LT", "LU", "LV", "MC", "MP", "MT", "MO", "NL", "NO", "NZ", "OM", "PA", "PL", "PR", "PT", "QA", "RO", "SA", "SC", "SG", "SI", "SK", "SM", "SE", "TC", "TT", "TW", "UY", "US", "VI", "VG"),
  "Upper middle income", c("AL", "AM", "AZ", "BY", "BZ", "BR", "BG", "CO", "CR", "CU", "DM", "DO", "EC", "FJ", "GA", "GD", "GE", "GQ", "GT", "GY", "ID", "IQ", "IR", "JM", "JO", "KZ", "LB", "LC", "LY", "MK", "MU", "MV", "MX", "MY", "NA", "PE", "PY", "RU", "RS", "SR", "SV", "TH", "TM", "TO", "TR", "VE", "ZA", "ME", "XK", "ST", "VC", "WS"),
  "Lower middle income", c("AO", "BD", "BJ", "BO", "BT", "CI", "CM", "CG", "CV", "DZ", "EG", "GH", "HN", "HT", "IN", "KE", "KG", "KH", "KI", "LA", "LK", "LS", "MA", "MD", "MN", "MM", "MR", "NG", "NI", "NP", "PK", "PG", "PH", "PS", "SB", "SN", "TJ", "TL", "TN", "TZ", "UA", "UZ", "VN", "VU", "ZM", "ZW"),
  "Low income", c("AF", "BF", "BI", "CD", "CF", "ER", "ET", "GM", "GN", "GW", "KM", "LR", "MG", "ML", "MW", "MZ", "NE", "RW", "SD", "SL", "SO", "SS", "SY", "TD", "TG", "UG", "YE")
) %>%
  tidyr::unnest(ctry_code)
dfnew=dfnew %>% inner_join(income_classification)


head(dfnew)
#xcountrymap=bind_rows(inventors,holders) %>% distinct() %>% inner_join(dfnew) %>% 
#  distinct(person_ctry_code,docdb_family_id)

#nrow(dfnew %>% filter(is.na(person_ctry_code)))
test=dfnew %>% group_by(ctry_code) %>% 
  summarize(n())



test2=inventors %>% inner_join(patchar)
rm(test2)

xcountrymap_i = inventors   %>% distinct() %>% inner_join(dfnew %>% select(person_id,ctry_code))
xcountrymap_i = xcountrymap_i %>% distinct(ctry_code,docdb_family_id)
rm(xcountrymap)


xcountrymap_h = holders   %>% distinct() %>% inner_join(dfnew %>% select(person_id,ctry_code))
xcountrymap_h = xcountrymap_h %>% distinct(ctry_code,docdb_family_id)

xcountrymap=xcountrymap_h %>% bind_rows(xcountrymap_i) %>% distinct(ctry_code,docdb_family_id)


nrow(countrymap %>% filter(ctry_code=="CN"))
nrow(xcountrymap %>% filter(ctry_code=="CN"))

nrow(countrymap)
```

### .....add patchar

```{r}
xcountrymap_patchar=xcountrymap %>% inner_join(patchar)

xcountrymap_patchar %>% inner_join(income_classification) %>% 
  filter(ctry_code=="CN" #|income_group=="High income"
         ) %>% group_by(income_group) %>% 
  summarize(mean(ev_global))



```



### let's do that again from the inventor country dataset...


```{r}

library(arrow)
# Define list of tables to read
tables_to_read <- c( "inventor_countries", "holder_countries")

# Loop through each table and read into a dataframe
for (table in tables_to_read) {
  # Create filename
  local_file <- paste0(localbigwb, "\\data\\", table, ".parquet")
  
  # Read parquet file and assign to variable with table name
  if (file.exists(local_file)) {
    assign(table, read_parquet(local_file))
    message("Loaded ", table, " (", nrow(get(table)), " rows)")
  } else {
    warning("File not found: ", local_file)
  }
}



holderORinventor_famXcountry = holder_countries %>% 
  mutate(type="holder") %>%
  bind_rows(inventor_countries %>% 
              mutate(type="inventor")) %>% 
  rename(ctry_code=person_ctry_code) %>% 
  #filter(type=="inventor") %>% 
  distinct(docdb_family_id,ctry_code,type) %>% inner_join(patchar) 
```


## just holders
```{r}
holderORinventor_famXcountry %>% 
  filter(ctry_code=="CN" & type=="holder" #|income_group=="High income"
         ) %>%  
  summarize(n(),mean(ev_global))

```

## just inventors
```{r}
holderORinventor_famXcountry %>% 
  filter(ctry_code=="CN" & type=="inventor" #|income_group=="High income"
         ) %>%  
  summarize(n(),mean(ev_global))
```

## combined
```{r}
holderORinventor_famXcountry %>% 
  distinct(docdb_family_id,ctry_code,ev_global) %>% 
  filter(ctry_code=="CN" 
         ) %>%  
  summarize(n(),mean(ev_global))


```

## Vietnam?
```{r}

holderORinventor_famXcountry %>% filter(ctry_code=="VN") %>% distinct(docdb_family_id) %>% 
  nrow()
```

## HICs?

```{r,eval=FALSE}
holderORinventor_famXcountry %>% 
  distinct(docdb_family_id,ctry_code,ev_global,HIC) %>% group_by(HIC) %>% 
  summarise(n(),mean(ev_global))

```

##👌###We should try here with all...rather than patchar...i.e. get all from bigquery..




### check pv vs ev

```{r}
library(fixest)


test0=patchar_countrymap %>% select(docdb_family_id,ctry_code,docdb_family_size) %>% 
   inner_join(patchar %>% select(docdb_family_id,ev_global,pv,cost,alpha)) %>% 
   mutate(lowid=pv<2*cost,
          thresdist=(pv-2*cost)/(2*cost),
          pvOcost=pv/cost,
          istraxid=(alpha+1)/cost*ev_global,
          evOcost=ev_global/cost) 


atest0=test0  %>% group_by(ctry_code) %>% summarize(mean(ev_global))
atest0 %>% filter(ctry_code=="CN")
test=test0 %>%
  select(-ctry_code) %>% distinct()


```

```{r}
   
#feols(ev ~ pv,test)
#feols(ev ~  lowid, test)


#feols(ev ~  thresdist, test)

#feols(istraxid ~  thresdist, test)

#feols(evOcost ~  pvOcost, test)




#################################################

eu_countries <- c("AT", "BE", "BG", "HR", "CY", "CZ", "DK", "EE", 
                  "FI", "FR", "DE", "GR", "HU", "IE", "IT", "LV", 
                  "LT", "LU", "MT", "NL", "PL", "PT", "RO", "SK", 
                  "SI", "ES", "SE")

# Create a logical variable identifying EU countries
test0$is_eu <- test0$ctry_code %in% eu_countries
test0$ctry_group <- ifelse(test0$ctry_code %in% eu_countries, "EU", test0$ctry_code)


library(wbstats)
library(dplyr)

# Get latest population data from World Bank
pop_data <- wb_data(
  indicator = "SP.POP.TOTL",
  start_date = 2023,
  end_date = 2023
) %>%
  select(iso2c, population = SP.POP.TOTL) %>%
  filter(!is.na(population))

# Merge with your dataframe
test0 <- test0 %>%
  left_join(pop_data, by = c("ctry_code" = "iso2c"))



test0a=test0 %>% group_by(ctry_code) %>% summarize(innos=n(),population=mean(population),innosFL1=sum(docdb_family_size>1))
test0a$ctry_group <- ifelse(test0a$ctry_code %in% eu_countries, "EU", test0a$ctry_code)
test0af=test0a %>% mutate(innosOpop=innos/population, innosFL1Opop=innosFL1/population)%>% filter(ctry_group %in% c("EU","CN","US","JP","GB"))
print(test0af)

library(ggplot2)
library(dplyr)
library(ggrepel)



plot_country_variable <- function(data, 
                                  var, 
                                  title = NULL,
                                  y_label = NULL,
                                  top_n = 15,
                                  quantile_threshold = 0.9) {
  
  # Use tidy evaluation to handle variable argument
  var_quo <- enquo(var)
  var_name <- rlang::as_label(var_quo)
  
  # Set default title and y_label if not provided
  if (is.null(title)) {
    title <- paste(var_name, "by Country")
  }
  if (is.null(y_label)) {
    y_label <- var_name
  }
  
  # Prepare data with labels
  data_plot <- data %>%
    mutate(
      pop_rank = rank(-population, na.last = "keep"),
      label = if_else(
        pop_rank <= top_n | {{var}} > quantile({{var}}, quantile_threshold, na.rm = TRUE), 
        ctry_code, 
        ""
      )
    ) %>%
    arrange({{var}})
  
  # Create plot
  ggplot(data_plot, aes(x = reorder(ctry_code, {{var}}), y = {{var}})) +
    geom_segment(aes(xend = ctry_code, y = 0, yend = {{var}}, 
                     color = ctry_group, size = population), alpha = 0.7) +
    geom_point(aes(color = ctry_group, size = population)) +
    geom_text_repel(aes(label = label), size = 3, max.overlaps = 20) +
    scale_size_continuous(name = "Population", labels = scales::comma, 
                          range = c(0.5, 4)) +
    scale_color_manual(values = c("EU" = "steelblue", "Other" = "coral"),
                       na.value = "gray60") +
    coord_flip() +
    labs(title = title,
         subtitle = paste0("Point size = population; Top ", top_n, 
                          " countries + high performers labeled"),
         x = "Country",
         y = y_label) +
    theme_minimal() +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank())
}

# Usage examples:




plot_country_variable(test0af, innosOpop, 
                      title = "Innovation per Population by Country",
                      y_label = "Innovations per Population")



plot_country_variable(test0af,  innosFL1Opop, 
                      title = "Multi country Innovations per Population by Country",
                      y_label = "Innovations per Population")





```

```{r}
test0 %>% filter(ctry_group %in% c("EU","CN","US","JP","GB")) %>% 
  ggplot( aes(x = ctry_group, y = cost)) +
  geom_boxplot() +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3, 
               fill = "red", color = "red") +
  geom_hline(yintercept = mean(test0$cost, na.rm = TRUE), 
             linetype = "dashed", color = "blue", alpha = 0.5) +
  labs(title = "Cost Distribution by Country Group",
       x = "Country Group",
       y = "Cost",
       caption = "Red diamonds = group means; Blue dashed line = overall mean") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))







```

#### nice figures

```{r}
#test %>% ggplot(aes(x=pvOcost,y=evOcost)) + geom_point()




library(ggplot2)
library(dplyr)

# Create quintile bins and calculate averages
test_binned <- test %>%
  mutate(evOcost_quintile = ntile(evOcost, 10)) %>%
  group_by(evOcost_quintile) %>%
  summarise(
    mean_evOcost = mean(evOcost, na.rm = TRUE),
    mean_pvOcost = mean(pvOcost, na.rm = TRUE)
  )

# Create the plot
ggplot(test_binned, aes(x = mean_evOcost, y = mean_pvOcost)) +
  geom_point(size = 3) +
  geom_line() +
  labs(
    x = "evOcost (bin averages)",
    y = "Average pvOcost",
    title = "Average pvOcost by evOcost Quintiles"
  ) +
  theme_minimal()

```

#### now for thresdist

```{r}

library(ggplot2)
library(dplyr)

# Create quintile bins and calculate averages
test_binned <- test %>%
  mutate(thresdist_quintile = ntile(thresdist, 10)) %>%
  group_by(thresdist_quintile) %>%
  summarise(
    mean_thresdist = mean(thresdist, na.rm = TRUE),
    mean_ev = mean(ev_global, na.rm = TRUE)
  )

# Create the plot
ggplot(test_binned, aes(x = mean_thresdist*100, y = mean_ev)) +
  geom_point(size = 3) +
  geom_line() +
  labs(
    x = "% distance from cutoff threshold",
    y = "Average external value ($ millions)",
    title = "External vs private value"
  ) +
  theme_minimal()+
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") 


```

#### with horizontal lines...

```{r}
library(ggplot2)
library(dplyr)

# Calculate average ev for both lowid groups
avg_ev_lowid1 <- test %>%
  filter(lowid == 1) %>%
  summarise(mean_ev = mean(ev, na.rm = TRUE)) %>%
  pull(mean_ev)

avg_ev_lowid0 <- test %>%
  filter(lowid == 0) %>%
  summarise(mean_ev = mean(ev, na.rm = TRUE)) %>%
  pull(mean_ev)

# Print the averages
cat("Average EV for lowid==1:", avg_ev_lowid1, "\n")
cat("Average EV for lowid==0:", avg_ev_lowid0, "\n")

# Create quintile bins and calculate averages
test_binned <- test %>%
  mutate(thresdist_quintile = ntile(thresdist, 10)) %>%
  group_by(thresdist_quintile) %>%
  summarise(
    mean_thresdist = mean(thresdist, na.rm = TRUE),
    mean_ev = mean(ev, na.rm = TRUE)
  )

# Create the plot
ggplot(test_binned, aes(x = mean_thresdist*100, y = mean_ev)) +
  geom_point(size = 3) +
  geom_line() +
  # Horizontal line for lowid==1 (from x=-100 to x=0)
  geom_segment(aes(x = -100, xend = 0, 
                   y = avg_ev_lowid1, yend = avg_ev_lowid1),
               color = "blue", linetype = "dashed", linewidth = 1) +
  # Horizontal line for lowid==0 (from x=0 to x=125)
  geom_segment(aes(x = 0, xend = 125, 
                   y = avg_ev_lowid0, yend = avg_ev_lowid0),
               color = "darkgreen", linetype = "dashed", linewidth = 1) +
  # Vertical line at cutoff
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "% distance from cutoff threshold",
    y = "Average external value ($ millions)",
    title = "External vs private value"
  ) +
  theme_minimal()



#########################################
library(ggplot2)
library(dplyr)

# Calculate average ev for both lowid groups
avg_ev_lowid1 <- test %>%
  filter(lowid == 1) %>%
  summarise(mean_ev = mean(ev, na.rm = TRUE)) %>%
  pull(mean_ev)

avg_ev_lowid0 <- test %>%
  filter(lowid == 0) %>%
  summarise(mean_ev = mean(ev, na.rm = TRUE)) %>%
  pull(mean_ev)

# Print the averages
cat("Average EV for lowid==1:", avg_ev_lowid1, "\n")
cat("Average EV for lowid==0:", avg_ev_lowid0, "\n")

# Create quintile bins and calculate averages
test_binned <- test %>%
  mutate(thresdist_quintile = ntile(thresdist, 10)) %>%
  group_by(thresdist_quintile) %>%
  summarise(
    mean_thresdist = mean(thresdist, na.rm = TRUE),
    mean_ev = mean(ev, na.rm = TRUE)
  )

# Create the plot
ggplot(test_binned, aes(x = mean_thresdist*100, y = mean_ev)) +
  geom_point(size = 3) +
  geom_line() +
  # Horizontal line for lowid==1 (from x=-100 to x=0)
  geom_segment(aes(x = -100, xend = 0, 
                   y = avg_ev_lowid1, yend = avg_ev_lowid1),
               color = "blue", linetype = "solid", linewidth = 1.5) +
  # Horizontal line for lowid==0 (from x=0 to x=125)
  geom_segment(aes(x = 0, xend = 125, 
                   y = avg_ev_lowid0, yend = avg_ev_lowid0),
               color = "darkgreen", linetype = "solid", linewidth = 1.5) +
  # Vertical line at cutoff
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "% distance from cutoff threshold",
    y = "Average external value ($ millions)",
    title = "External vs private value"
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 16, face = "bold"),
    axis.title.y = element_text(size = 16, face = "bold"),
    plot.title = element_text(size = 18, face = "bold"),
    axis.text = element_text(size = 12)
  )

  
```

## Build More ISTRAX (we only have EV on file so still need to apply the istrax formula)

The formula is

$$\frac{1}{c}\times EV \times \left(\alpha-\alpha \times I\{PV>2 c\}+I\{PV<2 c\}\right) $$

$$=\frac{1}{c}\times EV \times \left(\alpha-\alpha \times I\{PV>2 c\}+\left(1-I\{PV>2 c\}\right)\right) $$

$$=\frac{1}{c}\times EV \times \left(1+\alpha-\left(\alpha+1\right) \times I\{PV>2 c\}\right) $$

$$ISTRAX_i=\frac{\alpha+1}{c}\times EV_i \times\left(1-  I\{PV_i>2 c\}\right) $$

## combine EVs

```{r}





library(rlang)




flist = c( "CN","emde","eu","euplusuk","g7","IN","emdenocn","emdenocnin","hic","AT","GB","DE","FR","US" )
#flist = c( "FR","US" )
#flist=("VN")
for(ff in flist){
  #ff="CN"
  ff=ff %>% toupper()
  newev=paste0("ev_",ff)
  jurl=paste0("../../data/patent_data/patbis2021/data/fromWATSON/innos_ev_",ff,"_2009_2018.dsv")
  newdf=read_tsv(jurl)
  #test=
  ev_var_name <- names(newdf)[startsWith(names(newdf), "ev")]
  newdf=newdf %>% select(docdb_family_id,starts_with("ev")) %>% rename( !!newev:=!!sym(ev_var_name) )
  
  patchar=patchar %>% left_join( newdf )
}



patchar=patchar   %>% distinct() #%>% 
  #mutate(ev_EMDE=ev_EMDE,
         #ev_EMDEexCN=ev_EMDENOCN  ,
         #ev_EMDEexCNIN=ev_EMDENOCNIN,
         #ev_HIC=ev_HIC) 
  
  

#test=patchar  %>% filter(ev_EMDEexCNIN<0)
#test2=patchar %>% filter(ev_EMDEexCN<0)
#test3=patchar %>% filter(ev_test<0)
#test4=patchar %>% filter(ev_EMDE>ev)
```

# Making Istraxes

```{r}


library(dplyr)
library(stringr)

# Optionally guard against division by zero:
# patchar <- patchar %>% mutate(cost = na_if(cost, 0))

patchar <- patchar %>%
  # Create istrax_* columns for every ev_* column
  mutate(
    across(
      .cols = starts_with("ev_"),
      .fns  = ~ ((alpha + 1) / cost) * .x * as.integer(pv <= 2 * cost),
      .names = "istrax_{.col}"      # temporary names: istrax_ev_XXX
    )
    
  ) %>%
  # Drop the extra "ev_" in the new names: istrax_ev_XXX -> istrax_XXX
  select(-istrax_global) %>% 
  rename_with(~ str_replace(., "^istrax_ev_", "istrax_"), starts_with("istrax_ev_")) 

  
patchar <- patchar %>%
  mutate(
    across(
      .cols = starts_with("ev_"),
      .fns  = ~ (1 / cost) * (.x + pv)  ,
      .names = "avstrax_{.col}"      # temporary names: istrax_ev_XXX
    )
    
  ) %>%
  # Drop the extra "ev_" in the new names: istrax_ev_XXX -> istrax_XXX
  select(-avstrax_global) %>% 
  rename_with(~ str_replace(., "^avstrax_ev_", "avstrax_"), starts_with("avstrax_ev_"))
  




```

## Build patchar_countrymap

```{r}




# Add global istrax to patchar_countrympa
patchar_countrymap = patchar_countrymap %>% left_join(patchar) %>% 
  select(#-starts_with("ev_"),
    ,-starts_with("earliest"),-alpha,-cost,-pv,-docdb_family_size,-istrax_G7)

# Note: we have loads of innovations which are not mapped to countries...

#countrymap=patchar_countrymap %>% distinct(docdb_family_id,ctry_code)

#istraxes=patchar_countrymap %>% select(-ctry_code) %>% distinct()

#test2=test %>% select(docdb_family_id)


```


## ### test the global hypothesis...

```{r}

#patchar_countrymap=patchar_countrymap %>% mutate(testglobal=ev_hic+ev_emde)

patchar_countrymap %>% 
   mutate(testglobal=ev_HIC+ev_EMDE) %>% 

  filter(ctry_code %in% c("CN", "US", "VN","DE","JP","KR")) %>% 
  group_by(ctry_code) %>% 
  summarise(across(c(testglobal, ev_global), mean))




```
correcting this issue for now by adding hic to emde....

```{r}


prefixes <- patchar_countrymap %>% 
  select(ends_with("_global")) %>% 
  names() %>% 
  str_remove("_global$")

for(vvv in prefixes){
  global_col <- paste0(vvv, "_global")
  hic_col <- paste0(vvv, "_HIC")
  emde_col <- paste0(vvv, "_EMDE")
  
  patchar_countrymap[[global_col]] <- 
    patchar_countrymap[[hic_col]] + patchar_countrymap[[emde_col]]
}


```


## Build techmap

```{r}
url="../patent analysis/Data/Other data/docdb_tech_class.csv"

techmap=fread(url)
techmap=techmap %>% arrange(docdb_family_id) %>% filter(technology!="")
#head(classes)
#classes %>% distinct(technology)

green_classes=c("Green Energy",
                "Green Transport",
                "Circular Economy",
                "Green Manufacturing",
                "Adaptation",
                "Green Housing",
                "Green ICT",
                "Green Agriculture",
                "GHG Capture",
                "Any Green")

#classes = classes %>% mutate(green_id=technology %in%  green_classes)

#classes %>% distinct(technology)



#url="../patent analysis/Data/Other data/docdb_green_or_not.csv"


# Add a mapping to green.... Sam, double check if you agree...

url="../../data/patent_data/Y02_docdb_family_id_patstat2021.dsv"
greenornot=fread(url) %>% mutate(technology="Any Green") %>% select(docdb_family_id, technology)
#head(greenornot)
techmap=techmap %>% bind_rows(greenornot) %>% distinct()


```


##### We should integrate this with extrasectors.RMD which creates some extra sectors....

## Save stuff

```{r}

library(arrow)
library(fst)
# Save patchar_countrymap as patchar_countrymap.parquet
write_fst(patchar_countrymap, "patchar_countrymap.fst",compress=100)

# Save techmap as techmap.parquet
write_parquet(techmap, "techmap.parquet")

write_fst(patchar, "patchar.fst",compress=100)
rm(patchar)
#write_parquet(countrymap, "countrymap.parquet")
#write_parquet(istraxes, "istraxes.parquet")

```

### save istraxes

```{r}
#library(qs)
library(fst)
# save istraxes in separate files so we can more easily upload to github
#names(istraxes)
# Create output directory if it doesn't exist
if (!dir.exists("istraxes")) {
  dir.create("istraxes")
}

# Identify istrax_* columns
istrax_cols <- grep("^istrax_", names(patchar_countrymap), value = TRUE)
# Loop through each istrax_* column and save with docdb_family_id and ctry_code
for (col in istrax_cols) {
  #col="istraxi_HIC" 
  #col="istrax_global"
  #avcol=gsub("^i", "av", col)
  subset_df <- patchar_countrymap[, c("docdb_family_id","ctry_code", col)]
  #write_parquet(subset_df, paste0("istraxes/", col, ".parquet"))
  write_fst(subset_df, paste0("istraxes/", col, ".fst"),compress=100)
}


# Av strax columns
istrax_cols <- grep("^avstrax_", names(patchar_countrymap), value = TRUE)
# Loop through each istrax_* column and save with docdb_family_id and ctry_code
for (col in istrax_cols) {
  #col="istraxi_HIC" 
  #col="istrax_global"
  #avcol=gsub("^i", "av", col)
  subset_df <- patchar_countrymap[, c("docdb_family_id","ctry_code", col)]
  #write_parquet(subset_df, paste0("istraxes/", col, ".parquet"))
  write_fst(subset_df, paste0("istraxes/", col, ".fst"),compress=100)
}



# EV columns
istrax_cols <- grep("^ev_", names(patchar_countrymap), value = TRUE)
# Loop through each istrax_* column and save with docdb_family_id and ctry_code
for (col in istrax_cols) {
  #col="istraxi_HIC" 
  #col="istrax_global"
  #avcol=gsub("^i", "av", col)
  subset_df <- patchar_countrymap[, c("docdb_family_id","ctry_code", col)]
  #write_parquet(subset_df, paste0("istraxes/", col, ".parquet"))
  write_fst(subset_df, paste0("istraxes/", col, ".fst"),compress=100)
}




```

## copy stuff to ISEapp github repo as well

```{r}

if(Sys.info()["nodename"]=="PCACL-9RVSQ74" ){
  
  githubdir="C:/Users/rmartin7/GitHub/"  
  
}



# List of relative file paths
files1 <- list.files(path="istraxes", pattern = "fst$", full.names = TRUE)
file_list <- c(files1,"techmap.parquet")
#file_list=files1


# Destination root directory
dest_root1 <- paste0(githubdir,"iseapp/")
dest_root2 <- paste0(githubdir,"KEVCE/data/")



# Create destination directories and copy files
for (file in file_list) {
  # iseapp directory
  dest_path <- file.path(dest_root1, file)
  dest_dir <- dirname(dest_path)
  
  if (!dir.exists(dest_dir)) {
    dir.create(dest_dir, recursive = TRUE)
  }
  
  file.copy(from = file, to = dest_path, overwrite = TRUE)
  #########################################################
  
  # kevce directory
  dest_path <- file.path(dest_root2, file)
  dest_dir <- dirname(dest_path)
  
  if (!dir.exists(dest_dir)) {
    dir.create(dest_dir, recursive = TRUE)
  }
  
  file.copy(from = file, to = dest_path, overwrite = TRUE)


  
    
  
  
}

```

# Analysis

```{r}

source("istraxfunctions.r")

```

## Global vs National ISTRAX

Here we compare returns from further innovation at global vs national
spillovers. Note: going forward we want to add further scopes; e.g. LMIC
istrax where only spillovers to LMICs are considered.

## Vietnam

toflowlabels: Global Returns, National Returns, Returns within LMICs
(excl. China), Returns within HICs

toflowcatnames: istrax_global, istrax_nationalkey_2009_2018,
istrax_EMDEexCN, istrax_HIC

```{r}





plot_avstrax_by_country(patchar_countrymap, 
                        techmap, 
                        green_classes, 
                        country_code="VN", 
                        toflow="istrax_global", 
                        custom_colors)+ggtitle("Global Returns")

plot_avstrax_by_country(patchar_countrymap,
                        techmap,
                        green_classes, 
                        country_code="VN", 
                        toflow="istrax_nationalkey_2009_2018", 
                        custom_colors)+ggtitle("National Returns")

plot_avstrax_by_country(patchar_countrymap,
                        techmap,
                        green_classes, 
                        country_code="VN", 
                        toflow="istrax_EMDEexCN", 
                        custom_colors)+ggtitle("Returns within LMICs (excl. China)")



```

## All dvelopping (ex China)

```{r}

lmics=c("AF","AL","DZ","AO","AR","AM","AZ","BD","BJ","BO","BA","BW","BR","BG","BF","BI","KH","CM","CV","CF","TD","CL","CO","KM","CG","CR","CI","CU","DJ","DM","DO","EC","EG","SV","GQ","ER","ET","FJ","GA","GM","GE","GH","GT","GN","GW","GY","HT","HN","IN","ID","IR","IQ","JM","JO","KZ","KE","KI","KP","KG","LA","LB","LS","LR","LY","MG","MW","MY","MV","ML","MR","MU","MX","MD","MN","ME","MA","MZ","MM","NA","NP","NI","NE","NG","MK","PK","PW","PA","PG","PY","PE","PH","RW","WS","ST","SN","RS","SC","SL","SB","SO","ZA","LK","SD","SR","SY","TJ","TZ","TH","TL","TG","TO","TN","TR","TM","TV","UG","UA","UZ","VU","VE","VN","YE","ZM","ZW")


plot_avstrax_by_country(patchar_countrymap, 
                        techmap,
                        green_classes, 
                        country_code=lmics, 
                        toflow="istrax_global", 
                        custom_colors)+ggtitle("Global Returns")

plot_avstrax_by_country(patchar_countrymap, 
                        techmap,
                         green_classes, 
                        country_code=lmics, 
                        toflow="istrax_nationalkey_2009_2018", 
                        custom_colors)+ggtitle("National Returns")
```

## China

```{r}



plot_avstrax_by_country(patchar_countrymap, 
                        techmap,
                        green_classes, 
                        country_code="CN", 
                        toflow="istrax_global", 
                        custom_colors)+ggtitle("Global Returns")

plot_avstrax_by_country( patchar_countrymap, 
                        techmap,
                        green_classes, 
                        country_code="CN", 
                        toflow="istrax_nationalkey_2009_2018", 
                        custom_colors)+ggtitle("National Returns")




```

Africa only?

```{r}

lmics=c("TD","DZ","DJ","MZ","GM","GN","UG","GH","SN","ER",
        "CG","MA","TN","CM","MW","SD","NG","SO","EG","SL",
        "BF","BJ","ML","NE","MG","RW","CD","EH","CV","GW",
        "LR","ZA","SZ","ZM","KE","MR","TZ","LY","ZW","TG",
        "CF","ET","LS","GQ","KM","GA","CI","BI")


plot_avstrax_by_country(patchar_countrymap, 
                        techmap,
                        green_classes, 
                        country_code=lmics, 
                        toflow="istrax_global", 
                        custom_colors)+ggtitle("Global Returns")

plot_avstrax_by_country(patchar_countrymap, 
                        techmap, 
                         green_classes, 
                        country_code=lmics, 
                        toflow="istrax_nationalkey_2009_2018", 
                        custom_colors)+ggtitle("National Returns")



```

# Turkey

```{r}

plot_avstrax_by_country(patchar_countrymap, 
                        techmap,
                        green_classes, 
                        country_code="TR", 
                        toflow="istrax_global", 
                        custom_colors)+ggtitle("Global Returns")

plot_avstrax_by_country(patchar_countrymap, 
                        techmap, green_classes, 
                        country_code="TR", 
                        toflow="istrax_nationalkey_2009_2018", 
                        custom_colors)+ggtitle("National Returns")



```

# India

```{r}

plot_avstrax_by_country(patchar_countrymap, 
                        techmap,
                        green_classes, 
                        country_code="IN", 
                        toflow="istrax_global", 
                        custom_colors)+ggtitle("Global Returns")

plot_avstrax_by_country(patchar_countrymap, 
                        techmap, green_classes, 
                        country_code="IN", 
                        toflow="istrax_nationalkey_2009_2018", 
                        custom_colors)+ggtitle("National Returns")



```

# South Africa

```{r}

test=patchar_countrymap %>% filter(ctry_code=="ZA")

plot_avstrax_by_country(patchar_countrymap, 
                        techmap,green_classes ,
                        country_code="ZA", 
                        toflow="istrax_global", 
                        custom_colors)+ggtitle("Global Returns")


plot_avstrax_by_country(patchar_countrymap, 
                        techmap, green_classes, 
                        country_code="ZA", 
                        toflow="istrax_nationalkey_2009_2018", 
                        custom_colors)+ggtitle("National Returns")








```

# UK

```{r}

plot_avstrax_by_country(patchar_countrymap, 
                        techmap, 
                        green_classes, 
                        country_code="GB", 
                        toflow="istrax_global", 
                        custom_colors)+ggtitle("Global Returns")

plot_avstrax_by_country(patchar_countrymap, 
                        techmap,
                        green_classes ,
                        country_code="GB", 
                        toflow="istrax_nationalkey_2009_2018", 
                        custom_colors)+ggtitle("National Returns")



```

# Germany

```{r}

plot_avstrax_by_country(patchar_countrymap, 
                        techmap,
                        green_classes , 
                        country_code="DE", 
                        toflow="istrax_global", 
                        custom_colors)+ggtitle("Global Returns")

plot_avstrax_by_country(patchar_countrymap, 
                        techmap,
                        green_classes ,
                        country_code="DE", 
                        toflow="istrax_nationalkey_2009_2018", 
                        custom_colors)+ggtitle("National Returns")



```






# A check...do we really not have patents from Argentina (or namibia)
```{r}
if (Sys.info()["nodename"] == "PCACL-9RVSQ74") {
  
  inglobe="C:\\Users\\rmartin7\\OneDrive - Imperial College London\\inglobe"
  inglobewb="C:\\Users\\rmartin7\\OneDrive - WBG\\projects\\inglobe"
}



personfile=paste0(inglobewb,"\\data\\person_address.parquet")


personfile=read_parquet(personfile)

head(personfile)




```
